### 2019-09-14 leadership_across_cultures_contexts.Rmd
Adding Markdown file for manusript and R sript file for analyses.

---
title: "Leadership across cultures and contexts"
author: "Zachary Garfield, Kristen Syme, and Edward Hagen"
date: "4/25/2019"
output: 
  bookdown::html_document2
bibliography: Leader_contexts.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- A thematic note: One of our primary goals is to explore how an evolutionary framework can illuminate when we expect to find cultural, behavioral, physiological, etc. variation and then explain (some of) that variation. Part of going "Beyond WEIRD" is getting to theories that can account for population-level variation and predict invariant universals (or, at least, less variable phenotypes). We want to push things beyond the 'butterfly collecting' of simply documenting population-level variation. -->

# Abstract

Leader-follower dynamics are central to much of social life. Processes of leadership and followership observed in traditional societies, especially among small-scale societies, are often quite distinct from those observed among WEIRD societies and their complex organizations. Many researchers have turned to evolutionary theory to integrate diverse perspectives on human leadership and to better understand the diversity of leadership and followership observed across human cultures and across social contexts. Evolutionary theories of leadership, in turn, draw on ethnographic case studies from traditional societies that are thought to more closely resemble the small, face-to-face societies in which humans evolved. Currently, though, there is limited systematic data on the nature of leadership in such societies.

We build on our previous cross-cultural findings to systematically review the nature of leadership across the full range of human cultural diversity. Over 1200 ethnographic texts from the Human Relations Area Files (HRAF) were coded on leader qualities, their functions, and the costs and benefits for both leaders and followers across diverse social contexts. We discovered evidence for both cultural universals in leadership, as well as important variation, especially when compared to leadership in WEIRD societies. Across all contexts, leaders in the ethnographic record are most commonly described as intelligent and knowledgeable, of high social status, accomplished, wealthy, and prosocial. The primary benefits leaders offer to followers are resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources. The primary benefits leaders receive from followers include non-food resources, increased social status, and social services from followers. The most frequent costs of leadership for leaders include loss of social status and resources, and increased risk of harm or conflict. For followers, the most frequent costs include loss of resources, costs in providing social services, and increased risk of harm or conflict.

Leadership in WEIRD settings often differs substantially from non-WEIRD settings. In WEIRD settings influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face. A detailed view of leaders, their roles, and the costs and benefits of leadership across traditional societies and across contexts stands to make a significant empirical contribution for the development of more integrative evolutionary theories of leadership and social hierarchy.

# Introduction

>It's good to be king, if just for a while. - Tom Petty

Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes. Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, as well as follower preferences can greatly constrain a leader's agenda. This interaction between the strategies of individual leaders with cultural forces and group dynamics they likely play an out-sized role in contributing to organizational variation including social and economic trajectories of populations [@wiessner_risk_1982], and cultural evolutionary processes [@henrich_big_2015]. Leadership and followership are critical in understanding group structure and have been the focus of much scienfitic attention across psychology, sociology, political science, and managerial studies. Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology. Under a unifying evolutionary framework, social and biological scientits have now produced to a sizble body of theory on the evolution and cross-culural patterning of leadership and followership [e.g., CITE], accomanyed by a growing body of empirical results [e.g., CITE]. This approach as been informative. Drawing on the study of leadership across diverse, small-scale and tradiational societies can help develop more generalizable theories and offer applications to shape ogranizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015]. Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains. Moreover, while the benefits leaders acrue for their services have recieved substantial attention the costs of leadership for both leaders and followers have also been relatively underinvestigated.

Leaders are individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be. Leadership roles are also often associated with specific rights and responsibilities. Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the challenges specific groups face. Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor certain types of leaders, however there can be substantial variation in the qualities of leaders within these groups as well as and between them.

<!-- Reference a historical example of a short term leader implimenting change with lasting effects, good and/or bad. -->

## The qualities of leaders across cultures and contexts

## The functions of leadership across cultures and contexts

## The costs of leadership across cultures and contexts

## The benefits of leadership across cultures and contexts

The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.'

# Methods

# Results

# Discussion

# References

### 2019-09-14 leader_contexts_analyses.R

@@ -1,0 +1,121 @@
+#Temporary analyses script for Leadership across context and cultures paper
+#Eventually this will be added to the leadershipdata package, or probably it's own package. 
+
+
+# Load data library -------------------------------------------------------
+library(leadershipdata)
+
+# Load libraries ----------------------------------------------------------
+library(tidyverse)
+library(EGAnet)
+library(glasso)
+library(qgraph)
+
+
+# Create named lists of variables by group --------------------------------
+
+# Functions
+function_vars = c("function_bestow.mate","function_organize.cooperation","function_political.appointments",
+                  "function_resolve.conflcit","functions_construction.infrastructure","functions_control.calendar",
+                  "functions_control.economics","functions_control.immigration","functions_council.member",
+                  "functions_distribute.resources","functions_group.determination","functions_group.representative",
+                  "functions_hospitality","functions_moral.authority","functions_new.settlement",
+                  "functions_policymaking","functions_prosocial.investment","functions_protection",
+                  "functions_provide.counsel","functions_provide.subsistence","functions_punishment",
+                  "functions_ritual","functions_serve.leader","functions_social.functions","functions_strategic.planning")
+
+# Qualities
+quality_vars = c("qualities_artistic.performance","qualities_drug.consumption","qualities_exp.accomplished",
+                 "qualities_generous", "qualities_high.status","qualities_wealthy","qualities.age","qualities.aggressive",
+                 "qualities.ambition","qualities.attractive","qualities.bravery","qualities.charisma","qualities.coercive.authority",
+                 "qualities.confident","qualities.culturally.conservative","qualities.culturally.progressive","qualities.decisive",
+                 "qualities.fairness","qualities.favorable.personality","qualities.feared","qualities.high.quality.spouse",
+                 "qualities.honest","qualities.humble","qualities.industriousness","qualities.ingroup.member","qualities.innovative",
+                 "qualities.interpersonal.skills", "qualities.killer", "qualities.knowlageable.intellect", "qualities.loyalty",
+                 "qualities.many.children","qualities.oratory.skill","qualities.physical.health","qualities.physically.strong",
[…]

### 2019-09-25 leadership_across_cultures_contexts.Rmd
Added costs and benefits to variable support script.
BUT, added variable support plots to main script.

Working on MS.

@@
   bookdown::html_document2
-bibliography: Leader_contexts.bib
+bibliography: Leader2.bib
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+load("Leader2.Rdata")
+library(pander)
 ```
@@
 Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes.
-Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, as well as follower preferences can greatly constrain a leader's agenda.
-This interaction between the strategies of individual leaders with cultural forces and group dynamics they likely play an out-sized role in contributing to organizational variation including social and economic trajectories of populations [@wiessner_risk_1982], and cultural evolutionary processes [@henrich_big_2015].
-Leadership and followership are critical in understanding group structure and have been the focus of much scienfitic attention across psychology, sociology, political science, and managerial studies.
+Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda.
+The interactions between the strategies of individual leaders, pressures of followership, and cultural forces and group dynamics likely play an out-sized role in contributing to organizational variation across human populations including social and economic trajectories of populations [@wiessner_risk_1982], and cultural evolutionary processes [@henrich_big_2015].
+Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies.
 Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientits have now produced to a sizble body of theory on the evolution and cross-culural patterning of leadership and followership [e.g., CITE], accomanyed by a growing body of empirical results [e.g., CITE].
+Under a unifying evolutionary framework, social and biological scientists have now produced to a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., CITE], accompanied by a growing body of empirical results [e.g., CITE].
 This approach as been informative.
-Drawing on the study of leadership across diverse, small-scale and tradiational societies can help develop more generalizable theories and offer applications to shape ogranizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015].
+Drawing on the study of leadership across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015].
 Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains.
-Moreover, while the benefits leaders acrue for their services have recieved substantial attention the costs of leadership for both leaders and followers have also been relatively underinvestigated.
+Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated.
 
@@
 
+## The functions of leadership across cultures and contexts
+
+The emergence of leadership and the functions leaders serve across cultures is closely tied to social and political structure.
+Among ethnographically described small-scale societies which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
+Leaders in such contexts often draw on persuasive capacities and leverage community-based respect, their greater knowledge and social support to express their influence.
+The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of conflict resolution and collective actions.
+Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups.
+Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers.
+Representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
+
+Within communities, leaders may emerge whenever groups of individuals form.
+The role of leaders in ritual contexts, such as in initiations, and in providing social services, including facilitating marriages and organizing social activities, are also relatively common.
+
+Increased sedentarization and defensibility of resources is often associated with both social stratification and an institutionalized managerial role of community leaders.
+Among horticulturalists and agriculturalists managerial leadership can promote for more efficient use of shared resources such as water reserves and cooperative labor.
+Pastoralists and coastal populations can face similar pressures surrounding grazing lands and fishing access.
+In the wake of economic intensification, the function of leadership becomes increasing associated with the distribution of resources.
+
+Defensibility of resources is also often associated with increased territoriality and increased pressures of inter-group conflicts.
+Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territory are often forced to manage military forces and develop offensive and defensive strategies in warfare, more so than mobile populations.
+
 ## The qualities of leaders across cultures and contexts
-
-## The functions of leadership across cultures and contexts
 
@@
 
+# Study aims
+
+Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity.
+The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and context and the functions they serve is difficult to capture.
+Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others.
+Here, we seek provide the first broad near comprehensive view of the qualities and functions of leaders across from a representative cross-cultural sample.
+
 # Methods
+
+## Ethnographic sample and coding
+
+We evaluate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which CITE "tested" three theoretical models against the ethnographic record.
+Whereas CITE, documented supporting evidence in respect to influential theoretical models and offered direction for novel theoretical synthesis, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in.
+
+Using the same database of 1212 ethnographic paragraphs extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database.
+The leadershipdata was designed to capture a wide sample of the ethnographic record of leadership.
+Two authors coded each paragraph for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leader or followers may receive.
+Cost and benefits were operationalized prior to coding following our reading of the evolutionary leadership literature.
+See Table X.
+The operationalized qualities and functions (See Table X) were developed from the ethnographic text records in the database.
+First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
+Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, **creating new variables and operationalizations for additional qualities and functions encountered**.
+After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Tables XX.
+ZG then followed the same process of coding the entire sample of 1212 text records for the X operationalized variables.
+DISCUSS IRR.
+ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository.
+For additional details on ethnographic sampling and creating of the leadershipdata package see CITE.
+
+## Statistical analysis
 

### 2019-09-25 Variables support plots.R

@@ -67,7 +67,7 @@ cult.sum = function(df) {
 
 
 
-# Fit models --------------------------------------------------------------
+# Qualities & Functions Fit models --------------------------------------------------------------
 
 Model=character(0)
 Variable=character(0)
@@ -133,17 +133,17 @@ for (m in models) {
 d_melt = data.frame(Model=Model, Variable=Variable, Type=Type, value=value, y_se=y_se, y_negse=y_negse, stringsAsFactors = F)
 
 var_names <- c(
-  "function_bestow.mate" = "Bestow mate",                        
-  "function_political.appointments" = "Polistical appointments",             
+  "function_bestow.mate" = "Bestow mates",                        
+  "function_political.appointments" = "Political appointments",             
   "functions_construction.infrastructure"       = "Construction/infastructure",
-  "functions_control.economics"                 = "Constrol economics",
+  "functions_control.economics"                 = "Control economics",
   "functions_council.member"                    = "Council member",
    "functions_group.determination"              = "Group determiniation",
    "functions_hospitality"                       = "Hospitality",
    "functions_military.command"                  = "Military command",
    "functions_new.settlement"                    =  "Movement/migration",
    "functions_prosocial.investment"              = "Prosocial investment",
-   "functions_provide.counsel"                   = "Provide counsel",
+   "functions_provide.counsel"                   = "Provide counsel/direction",
    "functions_punishment"                        = "Punishment",
    "functions_serve.leader"                      = "Serve a leader",
    "functions_strategic.planning" = "Strategic planning",
@@ -153,21 +153,21 @@ var_names <- c(
    "functions_control.immigration"  = "Control immigration",
    "functions_distribute.resources" = "Distribute resources",
    "functions_group.representative" = "Group repsenetative",
-   "functions_medicinal"           = "Medicinal",
+   "functions_medicinal"           = "Medicinal functions",
    "functions_moral.authority"     = "Moral authority",
    "functions_policymaking"        =  "Policy making",
    "functions_protection"          = "Protection",
[…]

### 2019-09-25 leader_contexts_analyses.R

@@ -108,6 +108,8 @@ docID_cultureID<-documents[,c("d_ID","d_culture")]
 docID_cultureID$doc_ID<-docID_cultureID$d_ID
 docID_cultureID<-docID_cultureID[,c("d_culture","doc_ID")]
 
+textID_docID$doc_ID<-as.character(textID_docID$doc_ID)
+
 text_doc_cultureIDs<-left_join(textID_docID,docID_cultureID, by="doc_ID")
 
 leader_text2<-left_join(leader_text2, text_doc_cultureIDs, by="cs_textrec_ID")
@@ -137,8 +139,9 @@ pca_data_qualities[quality_vars==-1]<-0
 pca_data_qualities<-pca_data_qualities[rowSums(pca_data_qualities[quality_vars])>0,]
 
 # Add subsistence category and other vars in pca_data_qualities
-pca_data_qualities <- left_join(pca_data_qualities, text_doc_cultureIDs)
-pca_data_qualities <-left_join(pca_data_qualities, leader_cult)
+pca_data_qualities <- left_join(pca_data_qualities, text_doc_cultureIDs, by = "cs_textrec_ID")
+pca_data_qualities$c_name<-pca_data_qualities$d_culture
+pca_data_qualities <-left_join(pca_data_qualities, leader_cult[,c("subsistence","c_name")], by = "c_name")
 pca_data_qualities$c_culture_code<-pca_data_qualities$d_culture
 pca_data_qualities<-left_join(pca_data_qualities, leader_cult)
 pca_data_qualities<-left_join(pca_data_qualities, leader_text, by="cs_textrec_ID")
@@ -155,7 +158,7 @@ logpca_cv = cv.lpca(pca_data_qualities[quality_vars], ks = 6, ms = 1:10)
 plot(logpca_cv)
 
 
-logpca_model = logisticPCA(pca_data_qualities[quality_vars], k = 6, m = which.min(logpca_cv))
+logpca_model = logisticPCA(pca_data_qualities[quality_vars], k = 6, m = which.min(logpca_cv), main_effects = T)
 clogpca_model = convexLogisticPCA(pca_data_qualities[quality_vars], k = 6, m = which.min(logpca_cv))
 
 #Plots
@@ -184,7 +187,7 @@ plot(logpca_model, type = "scores") +
 
 
 plot(clogpca_model, type = "scores") + 
-  geom_point(aes(colour=pca_data_qualities$subsistence.x)) + 
[…]

### 2019-09-25 leader_contexts_analyses.R
There are some issues in merging dataframes for PCA with variable names. Need to sort this out...

@@ -141,7 +141,7 @@ pca_data_qualities<-pca_data_qualities[rowSums(pca_data_qualities[quality_vars])
 # Add subsistence category and other vars in pca_data_qualities
 pca_data_qualities <- left_join(pca_data_qualities, text_doc_cultureIDs, by = "cs_textrec_ID")
 pca_data_qualities$c_name<-pca_data_qualities$d_culture
-pca_data_qualities <-left_join(pca_data_qualities, leader_cult[,c("subsistence","c_name")], by = "c_name")
+pca_data_qualities <-left_join(pca_data_qualities, leader_cult, by = "c_name")
 pca_data_qualities$c_culture_code<-pca_data_qualities$d_culture
 pca_data_qualities<-left_join(pca_data_qualities, leader_cult)
 pca_data_qualities<-left_join(pca_data_qualities, leader_text, by="cs_textrec_ID")
@@ -169,7 +169,7 @@ plot(logsvd_model, type = "trace")
 
 
 plot(logsvd_model, type = "scores")+ 
-  geom_point(aes(colour=pca_data_qualities$subsistence.x)) + 
+  geom_point(aes(colour=pca_data_qualities$subsistence)) + 
   ggtitle("Exponential Family PCA") +
   scale_colour_brewer(palette = "Set1")
 

### 2019-09-25 leader_contexts_analyses.R
Added a "by" statement in a left_join

@@ -143,7 +143,7 @@ pca_data_qualities <- left_join(pca_data_qualities, text_doc_cultureIDs, by = "c
 pca_data_qualities$c_name<-pca_data_qualities$d_culture
 pca_data_qualities <-left_join(pca_data_qualities, leader_cult, by = "c_name")
 pca_data_qualities$c_culture_code<-pca_data_qualities$d_culture
-pca_data_qualities<-left_join(pca_data_qualities, leader_cult)
+pca_data_qualities<-left_join(pca_data_qualities, leader_cult, by = "c_culture_code")
 pca_data_qualities<-left_join(pca_data_qualities, leader_text, by="cs_textrec_ID")
 pca_data_qualities$demo_sex[pca_data_qualities$demo_sex=="-1"]="unkown"
 pca_data_qualities$demo_sex[is.na(pca_data_qualities$demo_sex)==TRUE]="unkown"

### 2019-09-25 leader_contexts_analyses.R
Link to logisticPCA package

@@ -2,6 +2,7 @@
 #Eventually this will be added to the leadershipdata package, or probably it's own package. 
 # Good stack overflow on PCA with binary data
 # https://stats.stackexchange.com/questions/16331/doing-principal-component-analysis-or-factor-analysis-on-binary-data/16335#16335
+# https://cran.r-project.org/web/packages/logisticPCA/vignettes/logisticPCA.html
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)

### 2019-10-02 leadership_across_cultures_contexts.Rmd
Adding to introduction text.

@@
 
-<!-- Reference a historical example of a short term leader implimenting change with lasting effects, good and/or bad. -->
+We first briefly review current theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership for both leaders and followers across human societies, including WEIRD and non-WEIRD populations.
+We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
+
+<!-- Reference a historical example of a short term leader implementing change with lasting effects, good and/or bad. -->
 
@@
 
+Leadership as a phenomena is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence.
+Also, adopting the broad definition of leadership given above, parents can be defined as leaders of nuclear families.
+Strong community leadership, however, is common cross-culturally.
+
 The emergence of leadership and the functions leaders serve across cultures is closely tied to social and political structure.
+
 Among ethnographically described small-scale societies which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
-Leaders in such contexts often draw on persuasive capacities and leverage community-based respect, their greater knowledge and social support to express their influence.
+Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, their greater knowledge and social support to express their influence.
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of conflict resolution and collective actions.
@@
 
+In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies with specific responsibilities and functions associated with their position.
+The structure of these organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as a figure head embodying the strategic vision of the organization.
+They are not often involved in resolving immediate conflicts, such as disputes between individuals or nested bodies, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
+
 ## The qualities of leaders across cultures and contexts
+
+Despite eminse variation, there do seem to be some commonalities among the qualities of community leaders.
+Across human societies, when strong community leaders exist, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence.
+Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have increased influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations.
+
+Amont small-scale socities with relative egalitarianism and local political autonomy, coomunity leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support.
+Physically, leaders across multople contexts tend to be taller and more physically formidable.
+Such physical embodiments might equip individuals for more eefective leadership in the context of inter-group conflicts as well as in the context of intra-group conflict resolution and in management of collective actions.
 
@@
 Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others.
-Here, we seek provide the first broad near comprehensive view of the qualities and functions of leaders across from a representative cross-cultural sample.
+Here, we provide the first broad and near comprehensive view of the qualities and functions of leaders across from a representative cross-cultural sample.
+We also aim to understand how leader qualities relate to their functions and how these relationships vary across social contexts and levels of cultural complexity.
+We hope these empirical results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundation component of human social behavior and social organization.
 

### 2019-10-08 leadership_across_cultures_contexts.Rmd
Added PCA work in analyses.
Working on introduction.

@@
 
+The Global Leadership and Organizational Behavior Effectivensss (GLOBE) project is a large-scale international research program aimed at understanding leadership effectiveness and culturla variation across large organizations.
+The GLOBE project has produced multiple cross-national datasets on leadership and hundreds of publications.
+For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied of industries across 24 countries.
+The goals of the GLOBE CEO Study were to address (1) how natioanl culture influences expectations for hte behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning adn low-functionsing CEOs, adn (4) the importance of CEO behaviors aligning with societal expectations.
+Through path analysis the study concluded that culturla values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
+That is, leaders moderate their behavior through culturally learned values and leadership-specific values.
+
+The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership beaviors." Among the global leadership behaviors, the GLOBE CEO STudy found that charimatic leadership beahvior is consistently the most important behavioral trait regarding firm performance and dedication of top managemnet team (TMT) members, followed by team-oriented behavior and human-oritented behavior.
+Participative leadership was moderately related to TMT dedication, however was not related to firm performance.
+The most successful CEOs are those who are rated as more visonary, performance oriented, and decisve.
+Furthermore, results strongly suggest the fit between a CEOs beahavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
+
 ## The qualities of leaders across cultures and contexts
@@
 
-Amont small-scale socities with relative egalitarianism and local political autonomy, coomunity leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support.
-Physically, leaders across multople contexts tend to be taller and more physically formidable.
-Such physical embodiments might equip individuals for more eefective leadership in the context of inter-group conflicts as well as in the context of intra-group conflict resolution and in management of collective actions.
+Among small-scale socities with relative egalitarianism and local political autonomy coomunity leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support.
+Physically, leaders across multiple contexts tend to be taller and more physically formidable.
+Such physical embodiments might equip individuals for more eefective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
+Charisma and oratory skills are common properties of leaders, but are espeically important in the context of egalitarianism where individual leaders must rely on persuassion and more subtle forms of influence to convince community members to come to agreement.
+
+## The benefits of leadership across cultures and contexts
+
+The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context are able to promote group policies which either align with their individual interests or which are not especially costly for themselves.
+
+Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating oppurtunities and therefore, given the aformentioned male-biase in top leadership positions, these leaders also tend to have greater reproductive success.
+
+Instituionalized leadership positions are also typically associated with increased social status.
+
+Leaders and high-status individuals are also more likely to be the recipeints of social support and material benefits in times of need.
 
@@
 
-## The benefits of leadership across cultures and contexts
+The unique costs associated with leadership positions have recieved less attention, relative to the benefits of leadership.
 
-The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.'
+Group responsibilities necessarily impose oppurtunity costs on individual priorities and agendas.
+
+In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
 
@@
 
-Using the same database of 1212 ethnographic paragraphs extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database.
+Using the same database of 1212 ethnographic paragraphs (termed text-records) extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database.
 The leadershipdata was designed to capture a wide sample of the ethnographic record of leadership.
@@
 
+We analyze data from our researcher-coded operationalized variables at the text record level and at the culture level.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables.
+
+<!-- At the text record level, hierarchical cluster analysis was then used to identify natural groupings of all our coded variables.
+If variables from a model clustered together, and separately from the other variables, that would lend support to the theoretical cohesiveness of that model.
+If, on the other hand, variables from different models clustered together, that would suggest theoretical models may need revision or that clusters had identified components of a novel theoretical model.
+-->
+
+To reduce the dimensionality of the `r length(quality_vars)` quality variables and the
+
 # Results
+
+The geographic distribution of the cultures in our sample is displayed in Fig.
+1, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
+The 1212 text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
+The vast majority of text records pertained to male leadership (89.8%, 898 text records) and very few pertained to female leadership (2.5%, 25 text records).In7 instances (0.7%)textrecordspertain tobothmales and females,and in69 text records (6.9%) sex was unknown.
+
+NEED TO DISCUSS 1187 and split records.
+
+## Variable distributions
+
+For each of the X coded variables, we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
+See FIG.
+
+```{r, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8}
+plot.variable.support
+```
+
+Costs and benefits.
+
+```{r, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8}
+plot.variable.support_costs_benefits
+```
 

### 2019-10-08 leader_contexts_analyses.R

@@ -5,6 +5,7 @@
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)
+load("leader_text2.rda")
 
 # Load libraries ----------------------------------------------------------
 library(tidyverse)
@@ -13,7 +14,7 @@ library(dendextend)
 library(logisticPCA)
 library(tibble)
 library(gridExtra)
-library(patchwork)
+#library(patchwork)
 
 
 # Create functions --------------------------------------------------------
@@ -128,375 +129,6 @@ by_culture = leader_text2 %>%
                                     follower_benefit_vars, follower_cost_vars))) %>% 
   summarise_each(list(mean=mean))
 
-
-
-
-
-# PCA Qualities ---------------------------------------------------------
-# Logistic PCA on leader qualities
-
-# Create data frames without all 0 rows
-
-#Create dataframe of variables for logistic PCA of qualities
-pca_data_qualities<-leader_text2[,c(quality_vars, "cs_textrec_ID", "group.structure2")]
-#Remove -1s for now
-pca_data_qualities[quality_vars==-1]<-0
-#Remove rows with all 0s
-pca_data_qualities<-pca_data_qualities[rowSums(pca_data_qualities[quality_vars])>0,]
-
-# Add subsistence category and other vars in pca_data_qualities
-pca_data_qualities <- left_join(pca_data_qualities, text_doc_cultureIDs, by = "cs_textrec_ID")
-
-# Rename culture name in leader_cult data
-leader_cult$d_culture<-leader_cult$c_culture_code
-leader_cult<-de_factor(leader_cult)
-#Subset culture vars of interest
-culture_vars<-leader_cult[,c("d_culture","subsistence","c_cultural_complexity")]
-
-pca_data_qualities <-left_join(pca_data_qualities, culture_vars, by = "d_culture")
-
-#Don't think we need this...
-#pca_data_qualities<-left_join(pca_data_qualities, leader_text, by="cs_textrec_ID")
-
-#Set components
-k=3
-#Fit the SVD with k = 6. 6 is the minimum for 80% of the variance
-logsvd_model = logisticSVD(pca_data_qualities[quality_vars], k = k)
-logsvd_model
-
-#Cross validate optimal m
[…]

### 2019-10-09 leader_contexts_analyses.R
Minor tweaks to model fitting args, and ggplot ellipses

@@ -151,16 +151,17 @@ pca_data_qualities$demo_sex[is.na(pca_data_qualities$demo_sex)==TRUE]="unkown"
 
 
 #Fit the SVD with k = 6. 6 is the minimum for 80% of the variance
-logsvd_model = logisticSVD(pca_data_qualities[quality_vars], k = 6)
+k = 3
+logsvd_model = logisticSVD(pca_data_qualities[quality_vars], k = k)
 logsvd_model
 
 #Cross validate optimal m
-logpca_cv = cv.lpca(pca_data_qualities[quality_vars], ks = 6, ms = 1:10)
+logpca_cv = cv.lpca(pca_data_qualities[quality_vars], ks = k, ms = 1:10)
 plot(logpca_cv)
 
 
-logpca_model = logisticPCA(pca_data_qualities[quality_vars], k = 6, m = which.min(logpca_cv), main_effects = T)
-clogpca_model = convexLogisticPCA(pca_data_qualities[quality_vars], k = 6, m = which.min(logpca_cv))
+logpca_model = logisticPCA(pca_data_qualities[quality_vars], k = k, m = which.min(logpca_cv), main_effects = T)
+clogpca_model = convexLogisticPCA(pca_data_qualities[quality_vars], k = k, m = which.min(logpca_cv))
 
 #Plots
 
@@ -171,12 +172,14 @@ plot(logsvd_model, type = "trace")
 
 plot(logsvd_model, type = "scores")+ 
   geom_point(aes(colour=pca_data_qualities$subsistence)) + 
+  stat_ellipse(aes(colour=pca_data_qualities$subsistence)) +
   ggtitle("Exponential Family PCA") +
   scale_colour_brewer(palette = "Set1")
 
 # Indicate group structure of log PCA model
 plot(logpca_model, type = "scores") + 
   geom_point(aes(colour=pca_data_qualities$group.structure2)) + 
+  stat_ellipse(aes(colour=pca_data_qualities$group.structure2)) + 
   ggtitle("Logistic PCA") +
   scale_colour_brewer(palette = "Set1")
 

### 2019-10-25 leadership_across_cultures_contexts.Rmd
Fixed errors on analyses script with repeated PCA analyses.
Adding CSV of SCCS and eRHAF (OWC) IDs.

Minor changes in Markdown file.

@@
 
-To reduce the dimensionality of the `r length(quality_vars)` quality variables and the
+To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` we relied on logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package CITE.
 
@@
 The 1212 text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
-The vast majority of text records pertained to male leadership (89.8%, 898 text records) and very few pertained to female leadership (2.5%, 25 text records).In7 instances (0.7%)textrecordspertain tobothmales and females,and in69 text records (6.9%) sex was unknown.
+The vast majority of text records pertained to male leadership (89.8%, 898 text records) and very few pertained to female leadership (2.5%, 25 text records).
+In 7 instances (0.7%)textrecordspertain tobothmales and females,and in 69 text records (6.9%) sex was unknown.
 

### 2019-10-25 leader_contexts_analyses.R

@@ -6,7 +6,7 @@
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)
-# load("leader_text2.rda")
+load("leader_text2.rda")
 
 # Load libraries ----------------------------------------------------------
 library(tidyverse)
@@ -18,6 +18,7 @@ library(gridExtra)
 library(car)
 library(visreg)
 library(effects)
+library(lme4)
 #library(patchwork)
 
 
@@ -43,6 +44,8 @@ leader_text2$group.structure2[leader_text2$group.structure.coded=="kin group"]="
 leader_text2$group.structure2[leader_text2$group.structure.coded=="local group"]="social group"
 leader_text2$group.structure2[leader_text2$group.structure.coded=="performance group"]="social group"
 
+leader_text2$group.structure2[leader_text2$group.structure.coded=="multiple domains"]="other"
+leader_text2$group.structure2[leader_text2$group.structure.coded=="unkown"]="other"
 
 
 # Create named lists of variables by group --------------------------------
@@ -135,153 +138,31 @@ by_culture = leader_text2 %>%
 
 
 
+# Culturel level variable manipulations -----------------------------------
+
+# Rename culture name in leader_cult data
+leader_cult$d_culture<-leader_cult$c_culture_code
+leader_cult<-de_factor(leader_cult)
+#Subset culture vars of interest
+culture_vars<-leader_cult[,c("d_culture","subsistence","c_cultural_complexity", "settlement_fixity", "pop_density","com_size")]
+
+# Add more SCCS variables
+load("sccs.RData")
+sccs<-data.frame(sccs)
+sccs<-sccs[,c("SCCS.","V1648")]
+sccs_ehraf_ID<-read.csv("sccs_ehraf_IDs.csv")
+sccs_ehraf_ID$c_culture_code<-sccs_ehraf_ID$OWC
+
+sccs_vars<-left_join(sccs, sccs_ehraf_ID, by = "SCCS.")
+sccs_vars<-sccs_vars[,c("c_culture_code","V1648")]
+sccs_vars<-distinct(sccs_vars)
+sccs_vars<-sccs_vars[!is.na(sccs_vars$c_culture_code)==T,]
 
 
-# PCA ---------------------------------------------------------
-# Logistic PCA on leader qualities
 
-# Create data frames without all 0 rows
-
[…]

### 2019-10-25 leader_contexts_analyses.R
Minor edits checking SCCS and eHRAF overlap.

@@ -153,6 +153,11 @@ sccs<-sccs[,c("SCCS.","V1648")]
 sccs_ehraf_ID<-read.csv("sccs_ehraf_IDs.csv")
 sccs_ehraf_ID$c_culture_code<-sccs_ehraf_ID$OWC
 
+sccs_ehraf_our_sample<-left_join(leader_cult, sccs_ehraf_ID)
+sccs_ehraf_our_sample<-sccs_ehraf_our_sample[,c("c_name","c_culture_code",
+                                                "SCCS.")]
+sccs_ehraf_our_sample<-sccs_ehraf_our_sample[!is.na(sccs_ehraf_our_sample$SCCS.)==T,]
+
 sccs_vars<-left_join(sccs, sccs_ehraf_ID, by = "SCCS.")
 sccs_vars<-sccs_vars[,c("c_culture_code","V1648")]
 sccs_vars<-distinct(sccs_vars)

### 2019-10-25 leadership_across_cultures_contexts.Rmd
Working on results

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=F)
 load("Leader2.Rdata")
@@
 1, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
-The 1212 text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
-The vast majority of text records pertained to male leadership (89.8%, 898 text records) and very few pertained to female leadership (2.5%, 25 text records).
-In 7 instances (0.7%)textrecordspertain tobothmales and females,and in 69 text records (6.9%) sex was unknown.
+The `r nrow(leader_text2)` text records from `r length(unique(documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
+The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
 
@@
 For each of the X coded variables, we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
-See FIG.
+See Figure \@ref(fig:varsupport).
 
-```{r, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8}
+```{r varsupport, fig.width=15, fig.height=8, fig.cap="Support for each coded variable. Blue: percent of text records with evidence of support. Red: percent of cultures with evidence of support."}
 plot.variable.support
@@
 
-Costs and benefits.
+Costs and benefits. See Figure \@ref(fig:varsupportcb).
 
-```{r, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8}
+```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8}
 plot.variable.support_costs_benefits
 ```
+
+## Principle components analysis
 

### 2019-10-25 leader_contexts_analyses.R

@@ -6,7 +6,7 @@
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)
-load("leader_text2.rda")
+# load("leader_text2.rda")
 
 # Load libraries ----------------------------------------------------------
 library(tidyverse)
@@ -29,6 +29,11 @@ df %>% dplyr::mutate_if(is.factor, as.character) -> df
 }
 
 # Recode variables --------------------------------------------------------
+
+# Add sex
+leader_text2 <- left_join(leader_text2, leader_text_original[c("cs_textrec_ID", "demo_sex")])
+male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
+female_leader_pct <- signif(100*sum(leader_text2$demo_sex=='female', na.rm=T)/nrow(leader_text2), 2)
 
 # Collapse group structure types
 
@@ -149,8 +154,8 @@ culture_vars<-leader_cult[,c("d_culture","subsistence","c_cultural_complexity", 
 # Add more SCCS variables
 load("sccs.RData")
 sccs<-data.frame(sccs)
-sccs<-sccs[,c("SCCS.","V1648")]
-sccs_ehraf_ID<-read.csv("sccs_ehraf_IDs.csv")
+sccs<-sccs[,c("SCCS.","V1648")] # ID, Frequency of warfare
+sccs_ehraf_ID<-distinct(read.csv("sccs_ehraf_IDs.csv")) # Some dupes in csv file
 sccs_ehraf_ID$c_culture_code<-sccs_ehraf_ID$OWC
 
 sccs_ehraf_our_sample<-left_join(leader_cult, sccs_ehraf_ID)
@@ -158,16 +163,14 @@ sccs_ehraf_our_sample<-sccs_ehraf_our_sample[,c("c_name","c_culture_code",
                                                 "SCCS.")]
 sccs_ehraf_our_sample<-sccs_ehraf_our_sample[!is.na(sccs_ehraf_our_sample$SCCS.)==T,]
 
-sccs_vars<-left_join(sccs, sccs_ehraf_ID, by = "SCCS.")
+sccs_vars<-left_join(sccs, sccs_ehraf_ID[c('SCCS.', 'c_culture_code')], by = "SCCS.")
+sccs_vars<-na.omit(sccs_vars) # Some eHRAF cultures do not map to SCCS cultures
 sccs_vars<-sccs_vars[,c("c_culture_code","V1648")]
-sccs_vars<-distinct(sccs_vars)
[…]

### 2019-10-26 leadership_across_cultures_contexts.Rmd
Working on MS.
Added rstanarm models to analyses script.

@@
 
-The Global Leadership and Organizational Behavior Effectivensss (GLOBE) project is a large-scale international research program aimed at understanding leadership effectiveness and culturla variation across large organizations.
-The GLOBE project has produced multiple cross-national datasets on leadership and hundreds of publications.
+The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leadership effectiveness and cultural variation across large organizations.
+The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications.
 For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied of industries across 24 countries.
-The goals of the GLOBE CEO Study were to address (1) how natioanl culture influences expectations for hte behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning adn low-functionsing CEOs, adn (4) the importance of CEO behaviors aligning with societal expectations.
-Through path analysis the study concluded that culturla values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
+The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
+Through path analysis the study concluded that cultural values indirectly predict CEO leadership behavior through culturally endorsed leadership expectations.
 That is, leaders moderate their behavior through culturally learned values and leadership-specific values.
 
-The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership beaviors." Among the global leadership behaviors, the GLOBE CEO STudy found that charimatic leadership beahvior is consistently the most important behavioral trait regarding firm performance and dedication of top managemnet team (TMT) members, followed by team-oriented behavior and human-oritented behavior.
+The GLOBE CEO Study defines six "global leadership behaviors" developed from 21 "primary leadership behaviors." Among the global leadership behaviors, the GLOBE CEO Study found that charismatic leadership behavior is consistently the most important behavioral trait regarding firm performance and dedication of top management team (TMT) members, followed by team-oriented behavior and human-oriented behavior.
 Participative leadership was moderately related to TMT dedication, however was not related to firm performance.
-The most successful CEOs are those who are rated as more visonary, performance oriented, and decisve.
-Furthermore, results strongly suggest the fit between a CEOs beahavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
+The most successful CEOs are those who are rated as more visionary, performance oriented, and decisive.
+Furthermore, results strongly suggest the fit between a CEOs behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
 
@@
 
-Despite eminse variation, there do seem to be some commonalities among the qualities of community leaders.
+Despite immense variation, there do seem to be some commonalities among the qualities of community leaders.
 Across human societies, when strong community leaders exist, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence.
@@
 
-Among small-scale socities with relative egalitarianism and local political autonomy coomunity leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support.
+Among small-scale societies with relative egalitarianism and local political autonomy community leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support.
 Physically, leaders across multiple contexts tend to be taller and more physically formidable.
-Such physical embodiments might equip individuals for more eefective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
-Charisma and oratory skills are common properties of leaders, but are espeically important in the context of egalitarianism where individual leaders must rely on persuassion and more subtle forms of influence to convince community members to come to agreement.
+Such physical embodiment might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
+Charisma and oratory skills are common properties of leaders, but are especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement.
 
@@
 
-Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating oppurtunities and therefore, given the aformentioned male-biase in top leadership positions, these leaders also tend to have greater reproductive success.
+Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
 
-Instituionalized leadership positions are also typically associated with increased social status.
+Institutionalized leadership positions are also typically associated with increased social status.
 
-Leaders and high-status individuals are also more likely to be the recipeints of social support and material benefits in times of need.
+Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need.
 
@@
 
-The unique costs associated with leadership positions have recieved less attention, relative to the benefits of leadership.
+The unique costs associated with leadership positions have received less attention, relative to the benefits of leadership.
 
-Group responsibilities necessarily impose oppurtunity costs on individual priorities and agendas.
+Group responsibilities necessarily impose opportunity costs on individual priorities and agendas.
 
@@
 
-We evaluate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which CITE "tested" three theoretical models against the ethnographic record.
-Whereas CITE, documented supporting evidence in respect to influential theoretical models and offered direction for novel theoretical synthesis, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in.
+We evaluate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
+Whereas @garfield_evolutionary_2019, documented supporting evidence in reference to influential theoretical models, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not tricly theoretically driven.
 
-Using the same database of 1212 ethnographic paragraphs (termed text-records) extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database.
-The leadershipdata was designed to capture a wide sample of the ethnographic record of leadership.
-Two authors coded each paragraph for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leader or followers may receive.
-Cost and benefits were operationalized prior to coding following our reading of the evolutionary leadership literature.
-See Table X.
+Here we draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database [@naroll_proposed_1967].
+The leadershipdata database was designed to capture a wide sample of the ethnographic record of leadership.
+For the current study, two authors coded each paragraph for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive.
+
 The operationalized qualities and functions (See Table X) were developed from the ethnographic text records in the database.
 First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, **creating new variables and operationalizations for additional qualities and functions encountered**.
-After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Tables XX.
+Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, creating new variables and operationalizations for additional qualities and functions encountered.
+After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Tables X.
 ZG then followed the same process of coding the entire sample of 1212 text records for the X operationalized variables.
@@
 ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository.
+
+Coding operationalizations of the costs and benefits of leadership were eveloped *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
+For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status/reputation, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, other resources, and other costs or benefits not captured by these categories.
+See Table X.
+
 For additional details on ethnographic sampling and creating of the leadershipdata package see CITE.
@@
 
-We analyze data from our researcher-coded operationalized variables at the text record level and at the culture level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables.
+We analyze data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to illustrate the dimensions of leadership (e.g., the qualities, functions, costs and benefits) which are most frequent cross-culturally, in the ethnographic record.
 
@@
 
-To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` we relied on logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package CITE.
+To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` we relied on logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package [@logisticPCA].
+We specify a $k = 3$ rank solution *a priori* and set $m$ (the value to approximate the saturated model) using cross-validation.
+Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitiave measure of leader qualities and leader functions for additional exploratory analyses.
 
@@
 1, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
-The `r nrow(leader_text2)` text records from `r length(unique(documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
+The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
 The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
 
-NEED TO DISCUSS 1187 and split records.
+<!-- NEED TO DISCUSS 1187 and split records.  -->
 
@@
 
-For each of the X coded variables, we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
+For each of the `r length(quality_vars) + length(function_vars) + length(leader_cost_vars) + length(leader_benefit_vars) + length(follower_benefit_vars) + length(follower_cost_vars)` coded variables (see Table X), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
 See Figure \@ref(fig:varsupport).
 
-```{r varsupport, fig.width=15, fig.height=8, fig.cap="Support for each coded variable. Blue: percent of text records with evidence of support. Red: percent of cultures with evidence of support."}
+```{r varsupport, fig.width=15, fig.height=8, fig.cap="Support for each coded variable.
+Blue circles: percent of text records with evidence of support.
+Red triangles: percent of cultures with evidence of support."}
 plot.variable.support
@@
 
-Costs and benefits. See Figure \@ref(fig:varsupportcb).
+The most common *functions* leaders serve, which are documented in over 70% of cultures, are providing counsel or direction in follower's affairs, conflict resolution, and organizing various cooperative activities.
+Leaders also commonly provide social functions, serve as a representative in inter-group interactions, and provide context-specific leadership such as military command and in providing ritual services.
+In about half of the cultures sampled here leaders punish other group members, control economic systems, and are responsible for policy development.
+Notable infrequent functions identified here include controlling group movement and migrations.
 
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8}
+The most common *qualities* of leaders are being of high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent.
+Also, in about 70% of cultures leaders maintained coercive authority over followers.
+Strategic nepotism, generosity, and having a favorable personality are also common leader qualities, documented in over 50% of cultures.
+Notable infrequent qualities of leaders in these data include charisma and fairness.
+
+Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
+Recieving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
+The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
+Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services.
+Mating and fitness benefits are relatively rare for followers.
+Followers most frequently encur costs in the form of resources lost and in providing social services.
+For both leaders and followers, individuals stand to lose what they are able to gain.
+The clear differences lie in leaders' opportunities for increased social status and matting opportunities.
+See Figure \@ref(fig:varsupportcb).
+
+```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8, fig.cap="Support for each coded variable.
+Blue circles: percent of text records with evidence of support.
+Red triangles: percent of cultures with evidence of support."}
 plot.variable.support_costs_benefits
@@
 
-## Principle components analysis
+## Principle components analyses
+
+A three-rank logistic PCA on the `r length(quality_vars)` coded leader quality variables with $m$ set at `r logpca_model_qualities$m` via cross-validation explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance and converged following `r logpca_model_qualities$iters` iterations.
+We draw on the first of the three leader quality components for interpretation and additional analyses and identify this component as a *Prestige vs.
+Dominance* measure.
+Low values on Leader qualities PC 1 (Prestige vs.
+Dominance) are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Oratory skill* and *Innovative*.
+High values are strongly associated with *Polygnous* and *Coercive authority* and moderately associated with *Many children* and *High quality spouse*.
+Therefore, the "dominance" side of the *Prestige vs.
+Dominance* component is also closely liked to reproduction.
+Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component.
+See Figure \@ref(fig:qualitiesPCplot).
+
+```{r qualitiesPCplot, fig.width=9, fig.height=6, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first of three components."}
+qualities_component1_plot
+```
+
+Similarly, a three-rank logistic PCA on the `r length(function_vars)` coded leader function variables with $m$ set at `r logpca_model_functions$m` via cross-validation explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance and converged following `r logpca_model_functions$iters` iterations.
+The first Leader functions component distinguishes between interpersonal mediation functions and group organizational and social functions, we define as a *Mediation vs.
+Organization* measures.
+Low values on Leader functions PC 1 (Mediation vs.
+Organization) are strongly associated with the variables *Resolve conflict* and *Provide counsel/direction* and moderately associated with *Political appointments* and *Punishment*, whereas high values are strongly associated with *Social functions* and *Organize cooperation* and moderately associated with *Ritual functions*.
+See Figure \@ref(fig:functionsPCplot).
+
+```{r functionsPCplot, fig.width=9, fig.height=6, fig.cap="Logistic PCA loadings of coded leader function variables on the first of three components."}
+functions_component1_plot
+```
+
+## Exploring population-level variation
+
+To explore potential population variation in the *Prestige vs.
+Dominance* leader qualities component and the *Mediation vs.
+Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture.
+
+We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “huntergatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
+We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
+We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, levels of male aggression, and social stratification.
+Fifty of these cultures are the same as, or very similar to, cultures in the PSF.
 

### 2019-10-26 leader_contexts_analyses.R

@@ -6,7 +6,8 @@
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)
-# load("leader_text2.rda")
+
+load("leader_text2.rda")
 
 # Load libraries ----------------------------------------------------------
 library(tidyverse)
@@ -665,7 +666,7 @@ qualities_component1_plot <-
   ggalt::geom_lollipop(horizontal = T, size = 1) +
   scale_color_gradient2(low = 'red', mid = 'white', 'high' = 'blue') +
   theme_bw(15) +
-  labs(title = "Leader qualities PC 1 (Prestige-Dominance)", x = "\nLoading", y = "")
+  labs(title = "Leader qualities PC 1 (Prestige vs. Dominance)", x = "\nLoading", y = "")
 
 qualities_component1_plot
 
@@ -711,18 +712,22 @@ pca_data_functions <- left_join(pca_data_functions, text_doc_cultureIDs, by = "c
 
 pca_data_functions <-left_join(pca_data_functions, culture_vars, by = "d_culture")
 
+# Fix names
+pca_data_functions2 <- pca_data_functions[function_vars]
+names(pca_data_functions2) <- var_names[names(pca_data_functions2)]
+
 
 #Fit the SVD 
-logsvd_model_functions = logisticSVD(pca_data_functions[function_vars], k = k)
+logsvd_model_functions = logisticSVD(pca_data_functions2, k = k)
 logsvd_model_functions
 
 #Cross validate optimal m
-logpca_cv_function = cv.lpca(pca_data_functions[function_vars], ks = k, ms = 1:10)
+logpca_cv_function = cv.lpca(pca_data_functions2, ks = k, ms = 1:10)
 plot(logpca_cv_function)
 
 
-logpca_model_functions = logisticPCA(pca_data_functions[function_vars], k = k, m = which.min(logpca_cv_function), main_effects = T)
-clogpca_model_functions = convexLogisticPCA(pca_data_functions[function_vars], k = k, m = which.min(logpca_cv_function))
+logpca_model_functions = logisticPCA(pca_data_functions2, k = k, m = which.min(logpca_cv_function), main_effects = T)
+clogpca_model_functions = convexLogisticPCA(pca_data_functions2, k = k, m = which.min(logpca_cv_function))
 
 #Plots
 
[…]

### 2019-10-28 leadership_across_cultures_contexts.Rmd
Working on MS text.
Exploring Bayesian models.

@@
 Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced to a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., CITE], accompanied by a growing body of empirical results [e.g., CITE].
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010 and @kaplan_evolutionary_2009], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
 This approach as been informative.
-Drawing on the study of leadership across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015].
+Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019].
 Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains.
 Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated.
+Current outstanding questions for evolutionary leaderhip scholars include the degree to which the nature of leadership is shaped by population-level variation and the costs and benefits leaders and followers encur across distinct leadership systems.
 
-Leaders are individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
+Evolutionay scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
 Leadership roles are also often associated with specific rights and responsibilities.
-Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the challenges specific groups face.
-Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor certain types of leaders, however there can be substantial variation in the qualities of leaders within these groups as well as and between them.
+Not surprisingly, the functions leaders serve then are often closely linked to group contexts and the salient challenges groups face.
+Certain types of groups, such as kin groups, economic groups, or political groups, may tend to favor leaders with specific qualities, however there can be substantial variation in the traits of leaders within these groups as well as between them.
 
@@
 Also, adopting the broad definition of leadership given above, parents can be defined as leaders of nuclear families.
-Strong community leadership, however, is common cross-culturally.
+Strong community leadership, however, is common cross-culturally [@garfield_review_2019; @lewis_leaders_1974].
 
@@
 
-The geographic distribution of the cultures in our sample is displayed in Fig.
-1, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
+The geographic distribution of the cultures in our sample is displayed in Figure X, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
 The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
@@
 
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=8, fig.cap="Support for each coded variable.
+```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable.
 Blue circles: percent of text records with evidence of support.
@@
 
-```{r qualitiesPCplot, fig.width=9, fig.height=6, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first of three components."}
+```{r qualitiesPCplot, fig.width=9, fig.height=7, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first of three components."}
 qualities_component1_plot
@@
 
-```{r functionsPCplot, fig.width=9, fig.height=6, fig.cap="Logistic PCA loadings of coded leader function variables on the first of three components."}
+```{r functionsPCplot, fig.width=9, fig.height=7, fig.cap="Logistic PCA loadings of coded leader function variables on the first of three components."}
 functions_component1_plot
@@
 Dominance* leader qualities component and the *Mediation vs.
-Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture.
+Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture with these componets as outcome measures.
 
+For predictor variables we incorporate additional population-level measures.
 We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “huntergatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
 We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
-We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, levels of male aggression, and social stratification.
+
+We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification.
 Fifty of these cultures are the same as, or very similar to, cultures in the PSF.
+Specifically, we use the SCCS measures of community size, population density, and cultural complexity.
+Using these measures as covariates allows us to determine the degree to which, an aggregate measure of cultural complexity, mode of subsistence, or measures of community size and population dispersal most strongly impact the qualities and functions of leaders.
+We also include our researcher coded measure of group structure (See Table X) as a predictor.
+
+```{r, fig.height=20, fig.width=9}
+qc_bayes_model_post_plot
+```
 

### 2019-10-28 leader_contexts_analyses.R

@@ -21,6 +21,24 @@ library(visreg)
 library(effects)
 library(lme4)
 #library(patchwork)
+library(dplyr)
+library(tidyr)
+library(tidybayes)
+library(ggplot2)
+library(cowplot)
+library(brms)
+library(ggstance)
+library(rstanarm)
+library(ggpubr)
+library(bayesplot)
+
+library(magrittr)
+library(forcats)
+library(purrr)
+library(modelr)
+library(ggridges)
+library(RColorBrewer)
+library(gganimate)
 
 
 # Create functions --------------------------------------------------------
@@ -437,7 +455,8 @@ plot.variable.support = ggplot(d_melt, aes(value, Variable, xmin=y_negse, xmax=y
   theme_bw() +
   theme(strip.text.y = element_text(angle=0))+
   scale_shape_manual(name="", values=c(17,16), labels=c('Cultures', 'Text records'))+
-  scale_fill_manual(name="", values=c("red", "blue"), labels=c('Cultures', 'Text records'))
+  scale_fill_manual(name="", values=c("red", "blue"), labels=c('Cultures', 'Text records')) +
+  labs(x="\nValue",y="")
 plot.variable.support
 
 
@@ -568,7 +587,8 @@ plot.variable.support_costs_benefits = ggplot(d_melt_cb, aes(value, Variable, xm
   theme_bw() +
   theme(strip.text.y = element_text(angle=0))+
   scale_shape_manual(name="", values=c(17,16), labels=c('Cultures', 'Text records'))+
-  scale_fill_manual(name="", values=c("red", "blue"), labels=c('Cultures', 'Text records'))
+  scale_fill_manual(name="", values=c("red", "blue"), labels=c('Cultures', 'Text records'))+
+  labs(x="\nValue",y="")
 plot.variable.support_costs_benefits
 
 
@@ -1083,7 +1103,7 @@ leader_text2$pop_density <-
   ordered(
     leader_text2$pop_density,
     levels = c(
-      "≤ 1 person / 1-5 sq. mile", 
+      #"≤ 1 person / 1-5 sq. mile", 
       "1-25 persons / sq. mile", 
       "26-100 persons / sq. mile",
       "101-500 persons / sq. mile",
@@ -1093,13 +1113,22 @@ leader_text2$pop_density <-
 
 leader_text2<-left_join(leader_text2, sccs_vars, by=c("d_culture"="c_culture_code"))
 
[…]

### 2019-10-28 leader_contexts_analyses.R
Minor edits to posterior plot.

@@ -1248,8 +1248,9 @@ qc_m_post_df_plot_long<-gather(qc_m_post_df_plot, variable, value)
 theme_set(theme_tidybayes())
 qc_bayes_model_post_plot <- qc_m_post_df_plot_long %>% 
   ggplot(aes(x=value, y=variable))+
-  geom_halfeyeh()+
+  geom_halfeyeh(trim = TRUE)+
   vline_0() +
+  xlim(-10,10) +
   scale_y_discrete(labels=rev(c("Subsistence:Pastoralists",
                             "Subsistence:Mixed",
                             "Subsistence:Hunter-gatherers",

### 2019-10-28 leader_contexts_analyses.R
Removed redundent packages; tweaked lmer vars

@@ -24,7 +24,6 @@ library(lme4)
 library(dplyr)
 library(tidyr)
 library(tidybayes)
-library(ggplot2)
 library(cowplot)
 library(brms)
 library(ggstance)
@@ -33,8 +32,6 @@ library(ggpubr)
 library(bayesplot)
 
 library(magrittr)
-library(forcats)
-library(purrr)
 library(modelr)
 library(ggridges)
 library(RColorBrewer)
@@ -1127,8 +1124,8 @@ qc_m2 <- lmer(
     #functions_component1 +
     subsistence +
     c_cultural_complexity +
-    pop_density2 +
-    com_size2 +
+    pop_density +
+    com_size +
     group.structure2 +
     # warfare_freq +
     (1|d_culture/doc_ID),

### 2019-11-02 leader_contexts_analyses.R
Reformated one line. Fixed possible bug in another

@@ -147,15 +147,20 @@ text_doc_cultureIDs<-left_join(textID_docID,docID_cultureID, by="doc_ID")
 
 leader_text2<-left_join(leader_text2, text_doc_cultureIDs, by="cs_textrec_ID")
 
-
-by_culture = leader_text2 %>% 
+by_culture <- leader_text2 %>% 
+  dplyr::select(
+    d_culture,
+    one_of(
+      c(quality_vars,
+        function_vars,
+        leader_benefit_vars,
+        leader_cost_vars,
+        follower_benefit_vars,
+        follower_cost_vars)
+      )
+    ) %>%
   group_by(d_culture) %>%
-  dplyr::select(d_culture, one_of(c(quality_vars, function_vars,
-                                    leader_benefit_vars, leader_cost_vars,
-                                    follower_benefit_vars, follower_cost_vars))) %>%
-  summarise_each(funs(mean))
-
-
+  summarise_each(lst(mean))
 
 # Culturel level variable manipulations -----------------------------------
 
@@ -240,7 +245,12 @@ authorID_textID<-leader_text_original[,c("author_ID","cs_textrec_ID")]
 all_data <- left_join(leader_text2, authorID_textID)
 
 # Remove negative values from vars
-all_data[all_data==-1]<-0
+# all_data[all_data==-1]<-0 # Some vars are factors, so this code won't work
+
+all_data <- all_data %>% 
+  mutate_if(
+    is.numeric, function(x) ifelse(x < 0, 0, x)
+  )
 
 #Groups of variables to model support
 model_vars<-c(quality_vars, function_vars)

### 2019-11-02 leader_contexts_analyses.R
Need to cross validate both m and k
Tweaked plotting code

@@ -607,7 +607,13 @@ plot.variable.support_costs_benefits
 #Create dataframe of variables for logistic PCA of qualities
 pca_data_qualities<-leader_text2[,c(quality_vars, "cs_textrec_ID", "group.structure2")]
 #Remove -1s for now
-pca_data_qualities[quality_vars==-1]<-0
+# pca_data_qualities[quality_vars==-1]<-0 # This code makes no sense
+
+pca_data_qualities <- pca_data_qualities %>% 
+  mutate_if(
+    is.numeric, function(x) ifelse(x < 0, 0, x) # Remove -1's
+  )
+
 #Remove rows with all 0s
 pca_data_qualities<-pca_data_qualities[rowSums(pca_data_qualities[quality_vars])>0,]
 
@@ -632,16 +638,16 @@ logsvd_model_qualities = logisticSVD(pca_data_qualities2, k = k)
 logsvd_model_qualities
 
 #Cross validate optimal m
-logpca_cv_qualities = cv.lpca(pca_data_qualities2, ks = k, ms = 1:10)
-plot(logpca_cv_qualities)
+# logpca_cv_qualities = cv.lpca(pca_data_qualities2, ks = k, ms = 1:10)
+# plot(logpca_cv_qualities)
 
 # Need to cross validate both k and m
-# cvlpca <- cv.lpca(pca_data_qualities2, ks = 1:20, ms = 5:10)
+# This takes a long time
+qual_cvlpca <- cv.lpca(pca_data_qualities2, ks = 1:20, ms = 5:10)
+plot(qual_cvlpca)
 # optimal values seem to be
-# k = 6
-# m = 9 or 10
-k = 6
-m = 10
+k = 10
+m = 12
 logpca_model_qualities = logisticPCA(pca_data_qualities2, k = k, m = m, main_effects = T)
 clogpca_model_qualities = convexLogisticPCA(pca_data_qualities2, k = k, m = which.min(logpca_cv_qualities))
 
@@ -689,28 +695,28 @@ ggplot(qual_loadings, aes(Loading, variable)) +
 
 # Plot individual components
 ## Component 1
-qual_loadings_X1<-qual_loadings[qual_loadings$Component=="X1",]
-X1_sort<-qual_loadings_X1$variable[order(qual_loadings_X1$Loading)]
-qual_loadings_X1$variable<-factor(qual_loadings_X1$variable, levels =  X1_sort)
+# qual_loadings_X1<-qual_loadings[qual_loadings$Component=="X1",]
[…]

### 2019-11-02 leader_contexts_analyses.R
Added pvclust analysis. Added png to gitignore

@@ -13,6 +13,7 @@ library(leadershipdata)
 library(tidyverse)
 library(forcats)
 library(NMF)
+library(pvclust)
 library(dendextend)
 library(logisticPCA)
 library(gridExtra)
@@ -630,6 +631,10 @@ pca_data_qualities <-left_join(pca_data_qualities, leader_cult, by = "d_culture"
 # Fix names
 pca_data_qualities2 <- pca_data_qualities[quality_vars]
 names(pca_data_qualities2) <- var_names[names(pca_data_qualities2)]
+
+# Cluster anaysis
+
+m <- pvclust(pca_data_qualities2, method.hclust = 'ward', method.dist = 'correlation')
 
 #Set components
 k=3

### 2019-11-03 leader_contexts_analyses.R
pvclust for functions too.

@@ -634,7 +634,9 @@ names(pca_data_qualities2) <- var_names[names(pca_data_qualities2)]
 
 # Cluster anaysis
 
-m <- pvclust(pca_data_qualities2, method.hclust = 'ward', method.dist = 'correlation')
+m <- pvclust(pca_data_qualities2, method.hclust = 'ward', method.dist = 'correlation', nboot = 2000)
+plot(m)
+pvrect(m)
 
 #Set components
 k=3
@@ -771,6 +773,10 @@ pca_data_functions <-left_join(pca_data_functions, culture_vars, by = "d_culture
 pca_data_functions2 <- pca_data_functions[function_vars]
 names(pca_data_functions2) <- var_names[names(pca_data_functions2)]
 
+# pvclust
+m <- pvclust(pca_data_functions2, method.hclust = 'ward', method.dist = 'correlation', nboot = 2000)
+plot(m)
+pvrect(m, alpha = 0.9)
 
 #Fit the SVD 
 logsvd_model_functions = logisticSVD(pca_data_functions2, k = k)

### 2019-11-03 leader_contexts_analyses.R
Fixed pop density recode.

@@ -1100,7 +1100,8 @@ leader_text2$pop_density <-
   ordered(
     leader_text2$pop_density,
     levels = c(
-      #"≤ 1 person / 1-5 sq. mile", 
+      "1 or less person / 1-5 sq. mile", 
+      "1-5 persons / sq. mile",
       "1-25 persons / sq. mile", 
       "26-100 persons / sq. mile",
       "101-500 persons / sq. mile",
@@ -1110,13 +1111,13 @@ leader_text2$pop_density <-
 
 leader_text2<-left_join(leader_text2, sccs_vars, by=c("d_culture"="c_culture_code"))
 
-#Set com size2 contrast to cubic function
-leader_text2$com_size2<-leader_text2$com_size
-contrasts(leader_text2$com_size2, 1) <- contr.poly(5)[,3]
-
-# Set pop density2 contrast to quadratic function
-leader_text2$pop_density2<-leader_text2$pop_density
-contrasts(leader_text2$pop_density2, 1) <- contr.poly(4)[,2]
+# #Set com size2 contrast to cubic function
+# leader_text2$com_size2<-leader_text2$com_size
+# contrasts(leader_text2$com_size2, 1) <- contr.poly(5)[,3]
+# 
+# # Set pop density2 contrast to quadratic function
+# leader_text2$pop_density2<-leader_text2$pop_density
+# contrasts(leader_text2$pop_density2, 1) <- contr.poly(4)[,2]
 
 
 qc_m2 <- lmer(
@@ -1124,10 +1125,10 @@ qc_m2 <- lmer(
     #functions_component1 +
     subsistence +
     c_cultural_complexity +
-    pop_density +
+    #pop_density +
     com_size +
     group.structure2 +
-    # warfare_freq +
+    #warfare_freq +
     (1|d_culture/doc_ID),
   data=leader_text2
   )
@@ -1136,6 +1137,7 @@ Anova(qc_m2)
 AIC(qc_m2)
 vif(qc_m2)
 visreg(qc_m2)
+
 plot(allEffects(qc_m2))
 # visreg(qc_m2, by = 'c_cultural_complexity', xvar = 'pop_density')
 # visreg(qc_m2, by = 'group.structure2', xvar = 'warfare_freq')
@@ -1169,8 +1171,8 @@ fc_m2 <- lmer(
     #qualities_component1 +
     subsistence +
     c_cultural_complexity +
-    pop_density +
[…]

### 2019-11-03 leadership_across_cultures_contexts.Rmd

@@
 We evaluate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
-Whereas @garfield_evolutionary_2019, documented supporting evidence in reference to influential theoretical models, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not tricly theoretically driven.
+Whereas @garfield_evolutionary_2019, documented supporting evidence in reference to influential theoretical models, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
 
@@
 
-Coding operationalizations of the costs and benefits of leadership were eveloped *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
+Coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
 For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status/reputation, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, other resources, and other costs or benefits not captured by these categories.

### 2019-11-03 leader_contexts_analyses.R
Re-organized left_joins with leader_text2 to prevent dupicate joins.

@@ -7,7 +7,7 @@
 # Load data library -------------------------------------------------------
 library(leadershipdata)
 
-# load("leader_text2.rda")
+load("leader_text2.rda")
 
 # Load libraries ----------------------------------------------------------
 library(tidyverse)
@@ -22,6 +22,7 @@ library(visreg)
 library(effects)
 library(lme4)
 library(patchwork)
+library(pvclust)
 # library(tidybayes)
 # library(cowplot)
 # library(brms)
@@ -171,6 +172,8 @@ leader_cult<-de_factor(leader_cult)
 #Subset culture vars of interest
 culture_vars<-leader_cult[,c("d_culture","subsistence","c_cultural_complexity", "settlement_fixity", "pop_density","com_size")]
 
+leader_text2<-left_join(leader_text2, culture_vars)
+
 # Add more SCCS variables
 load("sccs.RData")
 sccs<-data.frame(sccs)
@@ -187,11 +190,14 @@ sccs_vars<-left_join(sccs, sccs_ehraf_ID[c('SCCS.', 'c_culture_code')], by = "SC
 sccs_vars<-na.omit(sccs_vars) # Some eHRAF cultures do not map to SCCS cultures
 sccs_vars<-sccs_vars[,c("c_culture_code","V1648")]
 
+
+
 # Convert warfare levels to interval var
 sccs_vars$warfare_freq <- as.numeric(sccs_vars$V1648)
 sccs_vars$warfare_freq <- ifelse(sccs_vars$warfare_freq %in% c(1,20), NA, sccs_vars$warfare_freq)
 sccs_vars$warfare_freq <- 19 - sccs_vars$warfare_freq
 
+leader_text2<-left_join(leader_text2, sccs_vars, by=c("d_culture"="c_culture_code"))
 
 # Heatmaps -----------------------------------------------------------------
 heatmap_data<-leader_text2[,c(quality_vars, "group.structure2")]
@@ -640,7 +646,7 @@ pvrect(m)
 
 #Set components
 k=3
-#Fit the SVD with k = 6. 6 is the minimum for 80% of the variance
+
 logsvd_model_qualities = logisticSVD(pca_data_qualities2, k = k)
 logsvd_model_qualities
 
@@ -1048,7 +1054,7 @@ ggplot(leader_text2, aes(qualities_component1, functions_component1))+
        y="Functions component: Mediation - Organization\n")
 
 # Cultural complexity
[…]

### 2019-11-25 analyses2.R
Fixed bug in initialcompute. Reorganizing in analysis2.R

@@ -5,6 +5,7 @@
 # https://cran.r-project.org/web/packages/logisticPCA/vignettes/logisticPCA.html
 
 library(conflicted)
+conflict_prefer("which", "base")
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)
@@ -36,42 +37,7 @@ library(RColorBrewer)
 
 # Then load computed objects
 
-load("Leader2.Rdata") 
-
-# Heatmaps -----------------------------------------------------------------
-heatmap_data<-leader_text2[,c(quality_vars, "group.structure2")]
-
-#Temporary, -1 and 1 coudl 0 out
-heatmap_data<-heatmap_data[rowSums(heatmap_data[,c(quality_vars)]) > 0, ]
-
-# Rowv  <-
-#   t(as.matrix(heatmap_data)) %>%
-#   dist %>% # Cluster rows (variables) by Spearman rank correlation
-#   hclust(method = 'ward.D') %>%
-#   as.dendrogram %>%
-#   #rotate(order= colnames(data[vars])[order(data[vars][1,])]) %>%
-#   set("branches_k_color", k = 4)
-#
-# Colv  <- heatmap_data %>%
-#   dist %>% # Cluster columns (participants) by Euclidean metric
-#   hclust(method = 'ward.D') %>%
-#   as.dendrogram %>%
-#   #rotate(order = rownames(data)[order(-data$Respect)]) %>%
-#   set("branches_k_color", k = 3)
-
-aheatmap(t(as.matrix(heatmap_data[c(quality_vars)])),
-         width = 15, height = 10,
-         #Rowv  = Rowv,
-         #Colv = Colv,
-         distfun = "euclidean",
-         hclustfun = "ward",
-         cellheight = 5,
-         annCol = list(
-           Group = heatmap_data$group.structure2),
-         # annColors = list(
-         #   ),
-         treeheight = 50,
-         filename = 'heatmap_qualities.pdf')
+load("Leader2.Rdata")
 
 # PCA Qualities ---------------------------------------------------------
 
@@ -81,17 +47,18 @@ plot(m_lpca_qualk2, type = 'scores')
 logisticPCA_loadings_plot(m_lpca_qualk2, data = pca_data_qualities2)
 
 # Add PC scores to df
-pca_data_qualities$PC1k2 <- m_lpca_qualk2$PCs[,1]
[…]

### 2019-11-25 initialcompute.R

@@ -1,3 +1,8 @@
+
+# This script takes about 90 minutes to run
+# Saves all objects at the end, which are 
+# required for the analysis2.R script
+
 library(conflicted)
 conflict_prefer("which", "base")
 
@@ -631,13 +636,13 @@ which.min(m_lpca_funk2cv)
 
 # Qualities and functions
 
-logpca_cv_qfk2 = cv.lpca(pca_data_FQ[c(function_vars, quality_vars)], ks = 1:20, ms = 5:15)
-plot(logpca_cv_qfk2)
-x <- apply(logpca_cv_qfk2, MARGIN = 1, which.min)
-plot(1:20, logpca_cv_qfk2[cbind(1:20, x)], type='l') # elbows at ??
+logpca_cv_qf = cv.lpca(pca_data_FQ[c(function_vars, quality_vars)], ks = 1:20, ms = 5:15)
+plot(logpca_cv_qf)
+x <- apply(logpca_cv_qf, MARGIN = 1, which.min)
+plot(1:20, logpca_cv_qf[cbind(1:20, x)], type='l') # elbows at 5, 16
 
-# kqf <- ??
-# mqf <- ??
+kqf <- 16
+mqf <- 10
 
 # For k=2 only
 logpca_cv_qfk2 = cv.lpca(pca_data_FQ[c(function_vars, quality_vars)], ks = 2, ms = 1:10)

### 2019-11-25 analyses2.R
Moved plotting code to analysis2.R

@@ -39,6 +39,28 @@ library(RColorBrewer)
 
 load("Leader2.Rdata")
 
+
+# Functions ---------------------------------------------------------------
+
+# loadings plot
+logisticPCA_loadings_plot <- function(m, data){
+  df <- data.frame(m$U)
+  df$variable <- names(data)
+  p1 <-
+    ggplot(df, aes(X1, fct_reorder(variable, X1), colour=X1)) +
+    ggalt::geom_lollipop(horizontal = T, size = 1, show.legend = FALSE) +
+    scale_color_gradient2(low = 'red', mid = 'white', 'high' = 'blue', name = 'Loading') +
+    theme_bw(15) +
+    labs(title = "PC 1", x = "\nLoading", y = "")
+  p2<-
+    ggplot(df, aes(X2, fct_reorder(variable, X2), colour=X2)) +
+    ggalt::geom_lollipop(horizontal = T, size = 1, show.legend = FALSE) +
+    scale_color_gradient2(low = 'red', mid = 'white', 'high' = 'blue', name = 'Loading') +
+    theme_bw(15) +
+    labs(title = "PC 2", x = "\nLoading", y = "")
+  p1 + p2
+}
+
 # PCA Qualities ---------------------------------------------------------
 
 # For two PCs only; use optimal m
@@ -563,3 +585,53 @@ aheatmap(t(as.matrix(heatmap_data[,c(quality_vars)])),
          #   ),
          treeheight = 50,
          filename = 'heatmap_qualities.pdf')
+
+
+# Variable support plots --------------------------------------------------
+
+percent<-c("0%","10%","20%","30%","40%","50%","60%","70%","80%","90%","100%")
+# Plot code
+plot.variable.support = ggplot(d_melt, aes(value, Variable, xmin=y_negse, xmax=y_se, colour=Type, shape=Type)) + 
+  geom_errorbarh() + 
+  geom_point() +
+  scale_x_continuous(breaks=seq(0,1,.1), labels=percent, limits=c(0,1)) +
+  scale_colour_discrete(name='', labels=c('Cultures', 'Text records')) +
+  #labs(x='\nPercent', y='') +
+  facet_grid(Model~., scales = "free_y", space='free') +
+  facet_wrap(~Model, scales = "free_y") +
+  theme_bw() +
+  theme(strip.text.y = element_text(angle=0))+
[…]

### 2019-11-25 initialcompute.R

@@ -43,25 +43,6 @@ neg1to0 <- function(df){
     )
 }
 
-# loadings plot
-logisticPCA_loadings_plot <- function(m, data){
-  df <- data.frame(m$U)
-  df$variable <- names(data)
-  p1 <-
-    ggplot(df, aes(X1, fct_reorder(variable, X1), colour=X1)) +
-    ggalt::geom_lollipop(horizontal = T, size = 1, show.legend = FALSE) +
-    scale_color_gradient2(low = 'red', mid = 'white', 'high' = 'blue', name = 'Loading') +
-    theme_bw(15) +
-    labs(title = "PC 1", x = "\nLoading", y = "")
-  p2<-
-    ggplot(df, aes(X2, fct_reorder(variable, X2), colour=X2)) +
-    ggalt::geom_lollipop(horizontal = T, size = 1, show.legend = FALSE) +
-    scale_color_gradient2(low = 'red', mid = 'white', 'high' = 'blue', name = 'Loading') +
-    theme_bw(15) +
-    labs(title = "PC 2", x = "\nLoading", y = "")
-  p1 + p2
-}
-
 # Compute values ----------------------------------------------------------
 
 male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
@@ -137,7 +118,7 @@ sccs_vars$warfare_freq <- 19 - sccs_vars$warfare_freq
 
 leader_text2<-left_join(leader_text2, sccs_vars, by=c("d_culture"="c_culture_code"))
 
-# Variable support plots --------------------------------------------------
+# Variable support plot data prep --------------------------------------------------
 
 # Tabulate evidence for & against at the extract level (d1) and the culture level (d_ev2)
 
@@ -354,25 +335,6 @@ d_melt$Variable <- var_names[d_melt$Variable]
 the_levels <- c(d_melt$Variable[d_melt$Type=='Cultures'][order(d_melt$value[d_melt$Type=='Cultures'])])
 d_melt$Variable = factor(d_melt$Variable, levels=the_levels)
 
-percent<-c("0%","10%","20%","30%","40%","50%","60%","70%","80%","90%","100%")
-# Plot code
-plot.variable.support = ggplot(d_melt, aes(value, Variable, xmin=y_negse, xmax=y_se, colour=Type, shape=Type)) + 
-  geom_errorbarh() + 
[…]

### 2019-11-25 analyses2.R
Relabeled variable support plot

@@ -589,38 +589,77 @@ aheatmap(t(as.matrix(heatmap_data[,c(quality_vars)])),
 
 # Variable support plots --------------------------------------------------
 
-percent<-c("0%","10%","20%","30%","40%","50%","60%","70%","80%","90%","100%")
-# Plot code
+# Functions & qualities
+
 plot.variable.support = ggplot(d_melt, aes(value, Variable, xmin=y_negse, xmax=y_se, colour=Type, shape=Type)) + 
   geom_errorbarh() + 
   geom_point() +
-  scale_x_continuous(breaks=seq(0,1,.1), labels=percent, limits=c(0,1)) +
+  scale_x_continuous(breaks=seq(0,1,.1), labels=scales::percent, limits=c(0,1)) +
   scale_colour_discrete(name='', labels=c('Cultures', 'Text records')) +
-  #labs(x='\nPercent', y='') +
-  facet_grid(Model~., scales = "free_y", space='free') +
-  facet_wrap(~Model, scales = "free_y") +
-  theme_bw() +
+  facet_wrap(~Model, ncol = 1, scales = "free_y") +
   theme(strip.text.y = element_text(angle=0))+
   scale_shape_manual(name="", values=c(17,16), labels=c('Cultures', 'Text records'))+
   scale_fill_manual(name="", values=c("red", "blue"), labels=c('Cultures', 'Text records')) +
-  labs(x="\nValue",y="")
+  labs(x="",y="") +
+  theme_bw(15) 
 plot.variable.support
 
-percent<-c("0%","10%","20%","30%","40%","50%","60%","70%","80%","90%","100%")
-# Plot code
+# Leader/follower benefits and costs
+
+format_label <- function(lbl){
+  relabel <- c(
+    leader.benefits_fitness = 'Fitness',
+    leader.benefits_mating = 'Mating',
+    leader.benefits_other = 'Other',
+    leader.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
+    leader.benefits_resource_food = 'Food',
+    leader.benefits_resource_other = 'Other resource',
+    leader.benefits_social.services = 'Social services',
+    leader.benefits_social.status.reputation = 'Increased status',
+    leader.benefits_territory = 'Territory',
+    leader.costs_fitness.costs = 'Fitness cost',
+    leader.costs_increased.risk.harm.conflict = 'Increased risk of harm',
+    leader.costs_other = 'Other cost',
+    leader.costs_resource_food.cost = 'Food cost',
[…]

### 2019-11-25 leadership_across_cultures_contexts.Rmd
Progress...

@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=F)
-load("Leader2.Rdata")
+source('analyses2.R')
 library(pander)
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure X, with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
 The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
 The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
+
+```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis.
+Symbol shapes indicate mode of subsistence and symbol size indicates the number of documents available for that culture."}
+library(maps)
+world_map <- map_data("world")
+p_map <- ggplot(world_map, aes(x = long, y = lat)) +
+  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
+  geom_point(
+    data = leader_cult, 
+    aes(
+      c_longitude, 
+      c_latitude, 
+      colour = subsistence, 
+      shape = subsistence, 
+      size = documents
+    )
+  ) +
+  coord_fixed() +
+  labs(x='', y='', shape = 'Subsistence', colour = 'Subsistence', size = 'Documents') +
+  guides(shape = guide_legend(override.aes = list(size = 4))) +
+  theme_bw(15)
+p_map
+# ggsave(filename = 'Figures/map.pdf', p_map)
+```
 
@@
 
-For each of the `r length(quality_vars) + length(function_vars) + length(leader_cost_vars) + length(leader_benefit_vars) + length(follower_benefit_vars) + length(follower_cost_vars)` coded variables (see Table X), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Table X), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
 See Figure \@ref(fig:varsupport).
 
-```{r varsupport, fig.width=15, fig.height=8, fig.cap="Support for each coded variable.
-Blue circles: percent of text records with evidence of support.
-Red triangles: percent of cultures with evidence of support."}
+```{r varsupport, fig.width=15, fig.height=15, fig.cap="Support for each coded variable of leader functions (top) and qualities (bottom).
+Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
 plot.variable.support
@@
 
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable.
-Blue circles: percent of text records with evidence of support.
-Red triangles: percent of cultures with evidence of support."}
+```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
+Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
 plot.variable.support_costs_benefits
+```
+
+## Cluster analyses
+
+```{r qualcluster, fig.cap="Cluster analysis of leader quality variables.
+Distances are $1 - cor$.
+Ward agglomeration method.
+p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
+plot(m_pvclust_qual)
+pvrect(m_pvclust_qual, alpha = 0.9)
+```
+
+```{r funcluster, fig.cap="Cluster analysis of leader function variables.
+Distances are $1 - cor$.
+Ward agglomeration method.
+p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
+plot(m_pvclust_fun)
+pvrect(m_pvclust_fun, alpha = 0.9)
 ```
@@
 
-A three-rank logistic PCA on the `r length(quality_vars)` coded leader quality variables with $m$ set at `r logpca_model_qualities$m` via cross-validation explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance and converged following `r logpca_model_qualities$iters` iterations.
+A three-rank logistic PCA on the `r length(quality_vars)` coded leader quality variables with $m$ set at `r logpca_model_qualities$m` via cross-validation explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
 We draw on the first of the three leader quality components for interpretation and additional analyses and identify this component as a *Prestige vs.
@@
 
-```{r qualitiesPCplot, fig.width=9, fig.height=7, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first of three components."}
-qualities_component1_plot
+```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components."}
+logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2)
 ```
 
-Similarly, a three-rank logistic PCA on the `r length(function_vars)` coded leader function variables with $m$ set at `r logpca_model_functions$m` via cross-validation explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance and converged following `r logpca_model_functions$iters` iterations.
+Similarly, a three-rank logistic PCA on the `r length(function_vars)` coded leader function variables with $m$ set at `r logpca_model_functions$m` via cross-validation explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
 The first Leader functions component distinguishes between interpersonal mediation functions and group organizational and social functions, we define as a *Mediation vs.
@@
 
-```{r functionsPCplot, fig.width=9, fig.height=7, fig.cap="Logistic PCA loadings of coded leader function variables on the first of three components."}
-functions_component1_plot
+```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components."}
+logisticPCA_loadings_plot(logpca_model_functions, data = pca_data_functions2)
 ```
@@
 ```{r, fig.height=20, fig.width=9}
-qc_bayes_model_post_plot
+# qc_bayes_model_post_plot
 ```

### 2019-11-25 analyses2.R

@@ -29,7 +29,6 @@ library(modelr)
 library(ggridges)
 library(RColorBrewer)
 
-
 # Load precomputed objects ------------------------------------------------
 
 # First run initialcompute.R, which takes ~90 minutes on my machine,
@@ -38,7 +37,6 @@ library(RColorBrewer)
 # Then load computed objects
 
 load("Leader2.Rdata")
-
 
 # Functions ---------------------------------------------------------------
 
@@ -60,6 +58,93 @@ logisticPCA_loadings_plot <- function(m, data){
     labs(title = "PC 2", x = "\nLoading", y = "")
   p1 + p2
 }
+
+# Variable support plots --------------------------------------------------
+
+# Functions & qualities
+
+plot.variable.support = ggplot(d_melt, aes(value, Variable, xmin=y_negse, xmax=y_se, colour=Type, shape=Type)) + 
+  geom_errorbarh() + 
+  geom_point() +
+  scale_x_continuous(breaks=seq(0,1,.1), labels=scales::percent, limits=c(0,1)) +
+  scale_colour_discrete(name='', labels=c('Cultures', 'Text records')) +
+  facet_grid(Model~., scales = "free_y", space = "free_y") +
+  theme(strip.text.y = element_text(angle=0)) +
+  scale_shape_manual(name="", values=c(17,16), labels=c('Cultures', 'Text records'))+
+  scale_fill_manual(name="", values=c("red", "blue"), labels=c('Cultures', 'Text records')) +
+  labs(x="",y="") +
+  theme_bw(15) +
+  theme(strip.text.y = element_text(angle=0))
+plot.variable.support
+
+# Leader/follower benefits and costs
+
+format_label <- function(lbl){
+  relabel <- c(
+    leader.benefits_fitness = 'Fitness',
+    leader.benefits_mating = 'Mating',
+    leader.benefits_other = 'Other',
+    leader.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
+    leader.benefits_resource_food = 'Food',
+    leader.benefits_resource_other = 'Other resource',
+    leader.benefits_social.services = 'Social services',
+    leader.benefits_social.status.reputation = 'Increased status',
+    leader.benefits_territory = 'Territory',
+    leader.costs_fitness.costs = 'Fitness cost',
[…]

### 2019-11-26 leadership_across_cultures_contexts.Rmd
Minor reorganization of script file, minor tweaks to ms.

@@
 ---
-title: "Leadership across cultures and contexts"
-author: "Zachary Garfield, Kristen Syme, and Edward Hagen"
+title: "Leader functions and qualities across 59 cultures: an exploratory analysis"
+author: "Zachary J. Garfield, Kristen L. Syme, and Edward H. Hagen"
 date: "4/25/2019"
@@
 
-We analyze data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to illustrate the dimensions of leadership (e.g., the qualities, functions, costs and benefits) which are most frequent cross-culturally, in the ethnographic record.
+We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most frequently noted in the ethnographic record.
 
@@
 
-To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` we relied on logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package [@logisticPCA].
-We specify a $k = 3$ rank solution *a priori* and set $m$ (the value to approximate the saturated model) using cross-validation.
+To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` function variables, all of which take binary values (0, 1), we used logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package [@logisticPCA].
+Prior to analysis we removed all rows that had no evidence for any variable.
+We optimized rank, $k$, and $m$ (the value to approximate the saturated model) using cross-validation, per recommendation of the package authors.
 Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitiave measure of leader qualities and leader functions for additional exploratory analyses.
@@
 
-A three-rank logistic PCA on the `r length(quality_vars)` coded leader quality variables with $m$ set at `r logpca_model_qualities$m` via cross-validation explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
+A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_qualities$U)[2]`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
 We draw on the first of the three leader quality components for interpretation and additional analyses and identify this component as a *Prestige vs.
@@
 
-Similarly, a three-rank logistic PCA on the `r length(function_vars)` coded leader function variables with $m$ set at `r logpca_model_functions$m` via cross-validation explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
+Similarly, A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
 The first Leader functions component distinguishes between interpersonal mediation functions and group organizational and social functions, we define as a *Mediation vs.

### 2019-11-26 analyses2.R

@@ -38,6 +38,8 @@ library(RColorBrewer)
 
 load("Leader2.Rdata")
 
+#+ fig.height=15, fig.width=15
+
 # Functions ---------------------------------------------------------------
 
 # loadings plot
@@ -58,6 +60,28 @@ logisticPCA_loadings_plot <- function(m, data){
     labs(title = "PC 2", x = "\nLoading", y = "")
   p1 + p2
 }
+
+
+# Recode variables --------------------------------------------------------
+
+leader_text2$com_size <-
+  ordered(
+    leader_text2$com_size,
+    levels = c("< 99", "100-199", "200-399", "400-999", "> 1,000")
+  )
+
+leader_text2$pop_density <-
+  ordered(
+    leader_text2$pop_density,
+    levels = c(
+      "1 or less person / 1-5 sq. mile",
+      "1-5 persons / sq. mile",
+      "1-25 persons / sq. mile",
+      "26-100 persons / sq. mile",
+      "101-500 persons / sq. mile",
+      "over 500 persons / sq. mile"
+    )
+  )
 
 # Variable support plots --------------------------------------------------
 
@@ -138,13 +162,13 @@ plot.variable.support_costs_benefits
 # Cluster plots -----------------------------------------------------------
 
 plot(m_pvclust_qual)
-pvrect(m_pvclust_qual)
+pvrect(m_pvclust_qual, alpha = 0.9)
 
 plot(m_pvclust_fun)
 pvrect(m_pvclust_fun, alpha = 0.9)
 
 plot(qual_func_clust)
-pvrect(qual_func_clust)
+pvrect(qual_func_clust, alpha = 0.9)
 
 # PCA Qualities ---------------------------------------------------------
 
@@ -186,6 +210,7 @@ plot(logpca_model_qualities, type = "scores") +
 
 m_lpca_funk2 <- logisticPCA(pca_data_functions2, k = 2, m = which.min(m_lpca_funk2cv), main_effects = T)
 plot(m_lpca_funk2, type = 'score')
+logisticPCA_loadings_plot(m_lpca_funk2, data = pca_data_functions2)
 
 pca_data_functions$fPC1k2 <- m_lpca_funk2$PCs[,1]
 pca_data_functions$fPC2k2 <- m_lpca_funk2$PCs[,2]
@@ -334,28 +359,7 @@ plot(logpca_model_qf, type = "scores") +
 # # plot(qc_m)
[…]

### 2019-11-26 analyses2.R
Moved compute values to analysis2.R

@@ -680,3 +680,14 @@ plot(logpca_model_qf, type = "scores") +
 # NMF ---------------------------------------------------------------------
 # library(NMF)
 # m_nmf <- nmf(t(pca_data_qualities2), rank = 2:10)
+
+# Compute values ----------------------------------------------------------
+
+male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
+female_leader_pct <- signif(100*sum(leader_text2$demo_sex=='female', na.rm=T)/nrow(leader_text2), 2)
+
+intelltxts <- sum(leader_text2$qualities.knowlageable.intellect)
+polytxts <- sum(leader_text2$qualities.polygynous)
+statustxts <- sum(leader_text2$qualities_high.status)
+intellpolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities.knowlageable.intellect)
+statuspolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities_high.status)

### 2019-11-26 initialcompute.R

@@ -43,11 +43,6 @@ neg1to0 <- function(df){
     )
 }
 
-# Compute values ----------------------------------------------------------
-
-male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
-female_leader_pct <- signif(100*sum(leader_text2$demo_sex=='female', na.rm=T)/nrow(leader_text2), 2)
-
 # Create vectors of variable names by type --------------------------------
 
 lt2vars <- names(leader_text2)
@@ -309,7 +304,7 @@ var_names <- c(
   "qualities.feared"                = "Feared",
   "qualities.humble"                = "Humility",
   "qualities.innovative"            = "Innovative",
-  "qualities.knowlageable.intellect" = "Knowledgeable/intelligence",
+  "qualities.knowlageable.intellect" = "Knowledgeable/intelligent",
   "qualities.oratory.skill"         = "Oratory skill",
   "qualities.polygynous"            = "Polygnous",
   "qualities.social.contacts"       = "Social contacts",

### 2019-11-26 leadership_across_cultures_contexts.Rmd
Moved sections around
Latest stuff

@@
 title: "Leader functions and qualities across 59 cultures: an exploratory analysis"
-author: "Zachary J. Garfield, Kristen L. Syme, and Edward H. Hagen"
+author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
 date: "4/25/2019"
@@
 
+## Distribution of text records by group type and sex
+
+The type of group to which a leader text referred was coded into seven categories.
+Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups; it also indicates the proportion of male and female leaders of each type of group.
+
+```{r groupsex, fig.width=15, fig.cap="The distribution of text records by group type and leader sex.
+Areas of the bars are proportional to the number of text records in that category.
+There were no female leaders in any military or religious group."}
+plot_group_sex
+```
+
 ## Variable distributions
@@
 Notable infrequent qualities of leaders in these data include charisma and fairness.
-
-Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
-Recieving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
-The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
-Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services.
-Mating and fitness benefits are relatively rare for followers.
-Followers most frequently encur costs in the form of resources lost and in providing social services.
-For both leaders and followers, individuals stand to lose what they are able to gain.
-The clear differences lie in leaders' opportunities for increased social status and matting opportunities.
-See Figure \@ref(fig:varsupportcb).
-
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
-Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
-Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
-plot.variable.support_costs_benefits
-```
 
@@
 
+## Benefits and costs of leadership and followership
+
+Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
+Recieving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
+The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
+Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services.
+Mating and fitness benefits are relatively rare for followers.
+Followers most frequently encur costs in the form of resources lost and in providing social services.
+For both leaders and followers, individuals stand to lose what they are able to gain.
+The clear differences lie in leaders' opportunities for increased social status and matting opportunities.
+See Figure \@ref(fig:varsupportcb).
+
+```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
+Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
+plot.variable.support_costs_benefits
+```
+
 ## Exploring population-level variation

### 2019-11-26 analyses2.R

@@ -4,8 +4,8 @@
 # https://stats.stackexchange.com/questions/16331/doing-principal-component-analysis-or-factor-analysis-on-binary-data/16335#16335
 # https://cran.r-project.org/web/packages/logisticPCA/vignettes/logisticPCA.html
 
-library(conflicted)
-conflict_prefer("which", "base")
+# library(conflicted)
+# conflict_prefer("which", "base")
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)
@@ -28,6 +28,7 @@ library(magrittr)
 library(modelr)
 library(ggridges)
 library(RColorBrewer)
+library(ggmosaic)
 
 # Load precomputed objects ------------------------------------------------
 
@@ -105,39 +106,40 @@ plot.variable.support
 
 format_label <- function(lbl){
   relabel <- c(
-    leader.benefits_fitness = 'Fitness',
+    leader.benefits_fitness = 'Inclusive fitness',
     leader.benefits_mating = 'Mating',
-    leader.benefits_other = 'Other',
+    leader.benefits_other = 'Misc. non-material benefit',
     leader.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
     leader.benefits_resource_food = 'Food',
-    leader.benefits_resource_other = 'Other resource',
+    leader.benefits_resource_other = 'Material resources',
     leader.benefits_social.services = 'Social services',
     leader.benefits_social.status.reputation = 'Increased status',
     leader.benefits_territory = 'Territory',
-    leader.costs_fitness.costs = 'Fitness cost',
+    follower.benefits_fitness = 'Inclusive fitness',
+    follower.benefits_mating = 'Mating',
+    follower.benefits_other = 'Misc. non-material benefit',
+    follower.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
+    follower.benefits_resource_food = 'Food',
+    follower.benefits_resource_other = 'Material resources',
+    follower.benefits_social.services = 'Social services',
+    follower.benefits_social.status.reputation = 'Increased status',
[…]

### 2019-11-26 initialcompute.R

@@ -285,7 +285,7 @@ var_names <- c(
   "qualities_generous"                 = "Generosity",
   "qualities.age"                      = "Age",
   "qualities.attractive"              = "Attractive",
-  "qualities.coercive.authority"      = "Coercive authority",
+  "qualities.coercive.authority"      = "Authority",
   "qualities.culturally.progressive"  = "Culturally progressive",
   "qualities.favorable.personality"   = "Favorable personality",
   "qualities.honest"                  = "Honesty",
@@ -520,12 +520,12 @@ qual_func_vars <- leader_text2 %>%
   select(-functions_context) %>% 
   select(-contains("component"))
 
-qual_func_clust <- pvclust(
-  qual_func_vars, 
-  method.hclust = 'ward', 
-  method.dist = 'correlation', 
-  nboot = 10000,
-  parallel = T)
+# qual_func_clust <- pvclust(
+#   qual_func_vars, 
+#   method.hclust = 'ward', 
+#   method.dist = 'correlation', 
+#   nboot = 10000,
+#   parallel = T)
 
 # Cross-validation for logisticPCA ----------------------------------------
 
@@ -538,10 +538,10 @@ which.min(qual_cvlpca[9,]) # optimal? k=9, m=11
 
 # Plot all minima
 x <- apply(qual_cvlpca, MARGIN = 1, which.min)
-plot(1:20, qual_cvlpca[cbind(1:20, x)], type='l') # elbows at 9 & 14
+plot(1:20, qual_cvlpca[cbind(1:20, x)], type='l') # elbows at 8 & 13
 
-kq <- 9
-mq <- 11
+kq <- 8
+mq <- 12
 
 # Assuming k=2, cross validate for optimal m
 qual_cvlpcak2 <- cv.lpca(pca_data_qualities2, ks = 2, ms = 1:10)
@@ -555,12 +555,11 @@ fun_cvlpca = cv.lpca(pca_data_functions2, ks = 1:20, ms = 5:15)
 plot(fun_cvlpca)
 
 x <- apply(fun_cvlpca, MARGIN = 1, which.min)
-plot(1:20, fun_cvlpca[cbind(1:20, x)], type='l') # elbows at 5, 10, 15
[…]

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Code to output pvclust to pdf. Insert pngs into doc.

@@
 
-```{r groupsex, fig.width=15, fig.cap="The distribution of text records by group type and leader sex.
+```{r groupsex, fig.width=15, fig.height=15, fig.cap="The distribution of text records by group type and leader sex.
 Areas of the bars are proportional to the number of text records in that category.
@@
 p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
-plot(m_pvclust_qual)
-pvrect(m_pvclust_qual, alpha = 0.9)
+# plot(m_pvclust_qual)
+# pvrect(m_pvclust_qual, alpha = 0.9)
+knitr::include_graphics("Figures/m_pvclust_qual.final.png")
 ```
@@
 p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
-plot(m_pvclust_fun)
-pvrect(m_pvclust_fun, alpha = 0.9)
+# plot(m_pvclust_fun)
+# pvrect(m_pvclust_fun, alpha = 0.9)
+knitr::include_graphics("Figures/m_pvclust_fun.final.png")
 ```

### 2019-11-27 analyses2.R

@@ -163,11 +163,15 @@ plot.variable.support_costs_benefits
 
 # Cluster plots -----------------------------------------------------------
 
+pdf(file = "Figures/m_pvclust_qual.pdf", width=12, height=8)
 plot(m_pvclust_qual)
 pvrect(m_pvclust_qual, alpha = 0.9)
+dev.off()
 
+pdf(file = "Figures/m_pvclust_fun.pdf", width=12, height=8)
 plot(m_pvclust_fun)
 pvrect(m_pvclust_fun, alpha = 0.9)
+dev.off()
 
 # plot(qual_func_clust)
 # pvrect(qual_func_clust, alpha = 0.9)

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Removed group X sex mosaic plot; ink to info ratio not favorable. Put info in the text instead.

@@
 The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
+There were female leaders in almost all types of groups, however, except military and religious.
+Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
 
@@
 
-## Distribution of text records by group type and sex
+## Distribution of text records by mode of subsistence and group type
 
-The type of group to which a leader text referred was coded into seven categories.
-Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups; it also indicates the proportion of male and female leaders of each type of group.
+The type of group to which a leader text referred was coded into seven categories. Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups.
 
-```{r groupsex, fig.width=15, fig.height=15, fig.cap="The distribution of text records by group type and leader sex.
-Areas of the bars are proportional to the number of text records in that category.
-There were no female leaders in any military or religious group."}
-plot_group_sex
+```{r groupsex, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group type.
+Areas of the bars are proportional to the number of text records in that category."}
+plot_group_subsis #+ plot_group_sex
 ```

### 2019-11-27 analyses2.R

@@ -758,39 +758,54 @@ statuspolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities
 
 # Group structure by sex --------------------------------------------------
 
-xtabs(~group.structure2 + demo_sex, leader_text2)
-
-df_groupsex <- 
+df_groups <- 
   leader_text2 %>% 
   dplyr::select(
     group.structure2,
-    demo_sex
+    demo_sex,
+    subsistence
   ) %>% 
-  dplyr::filter(demo_sex != 'both', demo_sex != 'unknown', group.structure2 != 'other') %>% 
+  dplyr::filter(group.structure2 != 'other') %>% 
   mutate(
-    demo_sex = factor(demo_sex, levels = c('male', 'female'))
+    demo_sex = factor(demo_sex, levels = c('male', 'female')),
+    group =   factor(
+      group.structure2,
+      levels = c(
+        'residential subgroup',
+        'kin group',
+        'economic group',
+        'political group (community)',
+        'political group (supracommunity)',
+        'religious group',
+        'military group'
+      )
+    )
   )
 
-# df_groupsex$demo_sex[df_groupsex$demo_sex != 'both',]
-# df_groupsex$demo_sex <- factor(df_groupsex$demo_sex, levels = c('male', 'female', 'both', 'unknown'))
+plot_group_subsis <-
+  ggplot(df_groups) +
+  geom_mosaic(aes(x = product(group, subsistence), fill = group)) +
+  labs(x="", y="", fill = "Group type") +
+  guides(fill = NULL) +
+  theme_bw(15) #+ theme(legend.position = "none") 
+plot_group_subsis
 
-df_groupsex$group.structure2 <- 
-  factor(
-    df_groupsex$group.structure2,
-    levels = c(
-      'residential subgroup',
-      'kin group',
-      'economic group',
-      'political group (community)',
-      'political group (supracommunity)',
-      'religious group',
-      'military group'
-    )
-    )
+df_group_sex <- 
+  df_groups %>% 
+  dplyr::filter(
+    demo_sex != 'both', demo_sex != 'unknown'
+  )
 
 plot_group_sex <- 
-  ggplot(df_groupsex) + 
-  geom_mosaic(aes(x = product(group.structure2, demo_sex), fill = group.structure2)) +
+  ggplot(df_group_sex) + 
[…]

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Working on introduction. Additions to bib file.

--- 2019-11-26 leadership_across_cultures_contexts.Rmd
@@
 title: "Leader functions and qualities across 59 cultures: an exploratory analysis"
-author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
+author: "Zachary J. Garfield, Kristen L. Syme, and Edward H. Hagen"
 date: "4/25/2019"
@@
 Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda.
-The interactions between the strategies of individual leaders, pressures of followership, and cultural forces and group dynamics likely play an out-sized role in contributing to organizational variation across human populations including social and economic trajectories of populations [@wiessner_risk_1982], and cultural evolutionary processes [@henrich_big_2015].
+The interactions between the strategies of individual leaders, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural evolutionary processes [@henrich_big_2015].
+
 Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies.
 Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010 and @kaplan_evolutionary_2009], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
-This approach as been informative.
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010 and @kaplan_evolutionary_2009; @von_rueden_roots_2014], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
+This integration as been informative.
 Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019].
-Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains.
+Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and at all levels of social organization.
 Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated.
@@
 We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
+Through novel empirical and exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches.
 
@@
 
-Leadership as a phenomena is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence.
-Also, adopting the broad definition of leadership given above, parents can be defined as leaders of nuclear families.
-Strong community leadership, however, is common cross-culturally [@garfield_review_2019; @lewis_leaders_1974].
+Leadership is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence over group behavior and outcomes.
+Additionally, under the broad definition of leadership given above, parents can be defined as leaders of nuclear families.
+Strong community leadership, however, is common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @lewis_leaders_1974].
 
-The emergence of leadership and the functions leaders serve across cultures is closely tied to social and political structure.
-
+The nature of leaderhsip across human societies and social contexts is highly diverse and the emergence of leaders and the functions they serve are closely tied to social and political structure.
 Among ethnographically described small-scale societies which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
-Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, their greater knowledge and social support to express their influence.
+Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect and greater knowledge and social support to achieve or maintain community influence.
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of conflict resolution and collective actions.
@@
 Within communities, leaders may emerge whenever groups of individuals form.
-The role of leaders in ritual contexts, such as in initiations, and in providing social services, including facilitating marriages and organizing social activities, are also relatively common.
+The role of leaders in ritual contexts, such as in initiations, and in providing social services, including facilitating marriages and organizing social activities, are also expected to be common functions leaders serve.
 
-Increased sedentarization and defensibility of resources is often associated with both social stratification and an institutionalized managerial role of community leaders.
+Increased sedentarization and defensibility of resources is often associated with both social stratification and an institutionalized managerial role and stronger authority of community leaders.
 Among horticulturalists and agriculturalists managerial leadership can promote for more efficient use of shared resources such as water reserves and cooperative labor.
-Pastoralists and coastal populations can face similar pressures surrounding grazing lands and fishing access.
+Pastoralists and coastal populations often face similar pressures surrounding grazing lands and fishing access.
 In the wake of economic intensification, the function of leadership becomes increasing associated with the distribution of resources.
 
-Defensibility of resources is also often associated with increased territoriality and increased pressures of inter-group conflicts.
-Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territory are often forced to manage military forces and develop offensive and defensive strategies in warfare, more so than mobile populations.
+Defensibility of resources necessarily creates increased oppurtunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts.
+Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on disperrsed, less defesnible resources.
 
-In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies with specific responsibilities and functions associated with their position.
-The structure of these organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as a figure head embodying the strategic vision of the organization.
+In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies with institutionalized responsibilities and functions associated with their position.
+The structure of these organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization.
 They are not often involved in resolving immediate conflicts, such as disputes between individuals or nested bodies, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
@@
 Despite immense variation, there do seem to be some commonalities among the qualities of community leaders.
-Across human societies, when strong community leaders exist, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence.
+Across human societies, when strong community leaders emerge, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence.
 Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have increased influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations.
@@
 Physically, leaders across multiple contexts tend to be taller and more physically formidable.
-Such physical embodiment might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
+Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
 Charisma and oratory skills are common properties of leaders, but are especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement.
@@
 
-The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context are able to promote group policies which either align with their individual interests or which are not especially costly for themselves.
+The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the oppurtunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners.
 
@@
 
-Institutionalized leadership positions are also typically associated with increased social status.
+Institutionalized leadership positions are also often, if not universally, associated with increased social status.
+The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional.
 
@@
 
-The unique costs associated with leadership positions have received less attention, relative to the benefits of leadership.
+The unique costs associated with leadership positions have received less attention, relative to the benefits of leadership across disciplines.
 
@@
 Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity.
-The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and context and the functions they serve is difficult to capture.
-Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others.
-Here, we provide the first broad and near comprehensive view of the qualities and functions of leaders across from a representative cross-cultural sample.
-We also aim to understand how leader qualities relate to their functions and how these relationships vary across social contexts and levels of cultural complexity.
-We hope these empirical results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundation component of human social behavior and social organization.
+The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions they serve is difficult to capture.
+Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations.
+Here, we provide the first broad and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample.
+We also provide the first comparative empirical view describing the costs and benefits of leadership for both leaders nad followers in an evolutionary biological framework.
+We also aim to understand how the qualities and functions of leaders vary across social contexts and levels of cultural complexity.
+We hope these exploratory results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and social organization.
 
@@
 First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, creating new variables and operationalizations for additional qualities and functions encountered.
+Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
 After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Tables X.
@@
 DISCUSS IRR.
-ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository.
+ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a dataframe named leader_text2).
 
 Coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
-For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status/reputation, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, other resources, and other costs or benefits not captured by these categories.
+For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscelaneous costs or benefits not captured by these categories.
 See Table X.
@@
 
-## Distribution of text records by group type and sex
-
-The type of group to which a leader text referred was coded into seven categories.
-Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups; it also indicates the proportion of male and female leaders of each type of group.
-
-```{r groupsex, fig.width=15, fig.cap="The distribution of text records by group type and leader sex.
-Areas of the bars are proportional to the number of text records in that category.
-There were no female leaders in any military or religious group."}
-plot_group_sex
-```
-
 ## Variable distributions
@@
 Notable infrequent qualities of leaders in these data include charisma and fairness.
+
+Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
+Recieving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
+The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
+Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services.
+Mating and fitness benefits are relatively rare for followers.
+Followers most frequently encur costs in the form of resources lost and in providing social services.
+For both leaders and followers, individuals stand to lose what they are able to gain.
+The clear differences lie in leaders' opportunities for increased social status and matting opportunities.
+See Figure \@ref(fig:varsupportcb).
+
+```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
+Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
+Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
+plot.variable.support_costs_benefits
+```
 
@@
 
-## Benefits and costs of leadership and followership
-
-Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
-Recieving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
-The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
-Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services.
-Mating and fitness benefits are relatively rare for followers.
-Followers most frequently encur costs in the form of resources lost and in providing social services.
-For both leaders and followers, individuals stand to lose what they are able to gain.
-The clear differences lie in leaders' opportunities for increased social status and matting opportunities.
-See Figure \@ref(fig:varsupportcb).
-
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
-Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
-Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
-plot.variable.support_costs_benefits
-```
-
 ## Exploring population-level variation

### 2019-11-27 leadership_across_cultures_contexts.Rmd

@@
 
+## Distribution of text records by mode of subsistence and group type
+
+The type of group to which a leader text referred was coded into seven categories. Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups.
+
+```{r groupsex, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group type.
+Areas of the bars are proportional to the number of text records in that category."}
+plot_group_subsis #+ plot_group_sex
+```
+
 ## Variable distributions
@@
 
-Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
-Recieving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
-The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
-Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services.
-Mating and fitness benefits are relatively rare for followers.
-Followers most frequently encur costs in the form of resources lost and in providing social services.
-For both leaders and followers, individuals stand to lose what they are able to gain.
-The clear differences lie in leaders' opportunities for increased social status and matting opportunities.
-See Figure \@ref(fig:varsupportcb).
-
-```{r varsupportcb, echo=FALSE, warning=FALSE, message=F, fig.width=15, fig.height=11, fig.cap="Support for each coded variable for the costs and benefits of leadership and followership.
-Blue circles: percent of text records with evidence of support (95% CI computed with a binomial mixed effects models with random intercept for author nested within culture).
-Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
-plot.variable.support_costs_benefits
-```
-
 ## Cluster analyses

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Working on MS.
Minor lable edits to analyses files.

@@
 Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010 and @kaplan_evolutionary_2009; @von_rueden_roots_2014], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
 This integration as been informative.
@@
 Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated.
-Current outstanding questions for evolutionary leaderhip scholars include the degree to which the nature of leadership is shaped by population-level variation and the costs and benefits leaders and followers encur across distinct leadership systems.
+Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is shaped by population-level variation and the costs and benefits leaders and followers incur across distinct leadership systems.
 
-Evolutionay scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
+Evolutionary scholars typically define leaders as individuals who maintain disproportionate influence over decision-making within groups [@von_rueden_leadership_2015], whatever the group may be.
 Leadership roles are also often associated with specific rights and responsibilities.
@@
 
-The nature of leaderhsip across human societies and social contexts is highly diverse and the emergence of leaders and the functions they serve are closely tied to social and political structure.
+The nature of leadership across human societies and social contexts is highly diverse and the emergence of leaders and the functions they serve are closely tied to social and political structure.
 Among ethnographically described small-scale societies which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
@@
 
-Defensibility of resources necessarily creates increased oppurtunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts.
-Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on disperrsed, less defesnible resources.
+Defensibility of resources necessarily creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts.
+Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on dispersed, less defensible resources.
 
@@
 Despite immense variation, there do seem to be some commonalities among the qualities of community leaders.
-Across human societies, when strong community leaders emerge, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence.
-Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have increased influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations.
+Across human societies, when strong community leaders emerge, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006].
+Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have increased influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
+Drawing on cross-national data through the GLOBE project including data from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are perceived to be associated with effective leadership.
+In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
+Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].
 
-Among small-scale societies with relative egalitarianism and local political autonomy community leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support.
+Among small-scale societies with relative egalitarianism and local political autonomy community leaders, in addition to being older, respected males, tend to be individuals with large social networks and strong social support [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012].
 Physically, leaders across multiple contexts tend to be taller and more physically formidable.
-Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
-Charisma and oratory skills are common properties of leaders, but are especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement.
+Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014].
+Charisma and oratory skills are common properties of leaders, but are especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement [@grabo_charismatic_2016].
+
+Among societies with institutionalized social stratification, such as chiefdoms and agricultural communities, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004].
+Managerial elites, who maintain influence over community decisions are endowed with a high degree of ideological status associated with the role and kin group [@stanish_evolution_2010].
+In such contexts community leaders often lack coercive authority, but maintain strong influence through ideological systems and ritual or religious status and prowess inaccessible to non-elites [@stanish_evolution_2010].
+The association of formal community leadership and supernatural abilities is likely an important precursor of institutionalized authority and social stratification.
+
+Within state-level societies including non-industrial kingdoms leadership positions are proscribed and institutionalized.
+Ruling classes, parties, or political bodies monopolize social and political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda.
+
+Across discplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [@johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
+Recent evidence from within managerial studies suggests a curvilinear effect in that exceptionally intelligence individuals at high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
 
@@
 
-The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the oppurtunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners.
-
-Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
+The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners.
 
@@
 
+Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
+In a phylogenetic multilevel metaanalysis of 288 reported statistical associations among nonindustrial societies, @von_rueden_mens_2016 found that most measures of reproductive success were positively associated with male social status, independent of subsistence type or status measure.
+
+@von_rueden_leadership_2015 outilne four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin, (2) claim a fee for their services, such a tax or greater share of returns, (3) recieve reciprocal exchange in other currencies from followers who have benefited from their services, or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
+
 ## The costs of leadership across cultures and contexts
 
-The unique costs associated with leadership positions have received less attention, relative to the benefits of leadership across disciplines.
+Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers.
+However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines.
 
-Group responsibilities necessarily impose opportunity costs on individual priorities and agendas.
+Group responsibilities necessarily impose opportunity costs on individual priorities and agendas, given leadership responsibilities require investments of time, energy, and resources that could be invested elsewhere.
+
+Mediating conflicts between individuals can be physically and socially costly.
+Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved.
+Therefore, leaders stand to become the targets of animosity and social ridicule.
+In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression [@boehm_purposive_2008].
+Physically formidable individuals with strong social networks are likely to be able o bear these costs more-so than their counterparts and unsurprisingly larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015].
+With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude.
 
@@
 Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity.
-The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions they serve is difficult to capture.
+The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture.
 Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations.
 Here, we provide the first broad and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample.
-We also provide the first comparative empirical view describing the costs and benefits of leadership for both leaders nad followers in an evolutionary biological framework.
-We also aim to understand how the qualities and functions of leaders vary across social contexts and levels of cultural complexity.
-We hope these exploratory results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and social organization.
+We also provide the first comparative empirical view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
+We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity.
+
+In sum, we aim to empirically and systematically answer the following questions: 1) what are the most frequent qualities of leaders across cultures?; 2) what functions are leaders most commonly serving across cultures?; 3) what are the most salient costs and benefits of leadership for both leaders and followers?; and 4) how does variation across these measures relate to population-level variation?
+
+We hope these exploratory results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
 
@@
 
-The operationalized qualities and functions (See Table X) were developed from the ethnographic text records in the database.
+The operationalized qualities and functions (See Table \@ref(tab:qualfuncvars)) were developed from the ethnographic text records in the database.
 First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
 Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
-After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Tables X.
-ZG then followed the same process of coding the entire sample of 1212 text records for the X operationalized variables.
+After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
+ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables.
 DISCUSS IRR.
-ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a dataframe named leader_text2).
+ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named leader_text2).
+
+Table: (\#tab:qualfuncvars) Operational definitions for coded leader qualities and functions.
+
+```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
+library(readr)
+qual_func_var_tab<-read_csv('func_qual_definitions.csv', na = 'NA')
+library(knitr)
+kable(qual_func_var_tab)
+```
 
 Coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
-For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscelaneous costs or benefits not captured by these categories.
-See Table X.
+For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
+See Table \@ref(tab:costbenefitvars).
+
+Table: (\#tab:costbenefitvars) Operational definitions for coded benefits and costs for leaders and followers.
+
+```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE, }
+library(readr)
+cost_benefit_vars_tab<-read_csv('costs_benefits.csv', na = 'NA')
+library(knitr)
+kable(cost_benefit_vars_tab)
+```
 
@@
 We optimized rank, $k$, and $m$ (the value to approximate the saturated model) using cross-validation, per recommendation of the package authors.
-Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitiave measure of leader qualities and leader functions for additional exploratory analyses.
+Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitative measure of leader qualities and leader functions for additional exploratory analyses.
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Table X), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars) ), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
 See Figure \@ref(fig:varsupport).
@@
 Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
-Recieving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
+Receiving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
 The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
@@
 Mating and fitness benefits are relatively rare for followers.
-Followers most frequently encur costs in the form of resources lost and in providing social services.
+Followers most frequently incur costs in the form of resources lost and in providing social services.
 For both leaders and followers, individuals stand to lose what they are able to gain.
@@
 Dominance* leader qualities component and the *Mediation vs.
-Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture with these componets as outcome measures.
+Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture with these components as outcome measures.
 
@@
 
-We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification.
+<!-- We drew on the Standard Cross-Cultural Sample (SCCS) database [@murdock_standard_1969] for additional population-level measures, which scores 186 cultures on many qualitative or quantitative variables, such as kinship system, dependence on agriculture, and social stratification.
 Fifty of these cultures are the same as, or very similar to, cultures in the PSF.
@@
 We also include our researcher coded measure of group structure (See Table X) as a predictor.
+-->
 

### 2019-11-27 analyses2.R

@@ -106,23 +106,23 @@ plot.variable.support
 
 format_label <- function(lbl){
   relabel <- c(
-    leader.benefits_fitness = 'Inclusive fitness',
-    leader.benefits_mating = 'Mating',
+    leader.benefits_fitness = 'Inclusive fitness benefit',
+    leader.benefits_mating = 'Mating benefit',
     leader.benefits_other = 'Misc. non-material benefit',
     leader.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
-    leader.benefits_resource_food = 'Food',
+    leader.benefits_resource_food = 'Food benefit',
     leader.benefits_resource_other = 'Material resources',
     leader.benefits_social.services = 'Social services',
-    leader.benefits_social.status.reputation = 'Increased status',
+    leader.benefits_social.status.reputation = 'Increased social status',
     leader.benefits_territory = 'Territory',
-    follower.benefits_fitness = 'Inclusive fitness',
-    follower.benefits_mating = 'Mating',
+    follower.benefits_fitness = 'Inclusive fitness benefit',
+    follower.benefits_mating = 'Mating benefit',
     follower.benefits_other = 'Misc. non-material benefit',
     follower.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
-    follower.benefits_resource_food = 'Food',
+    follower.benefits_resource_food = 'Food benefit',
     follower.benefits_resource_other = 'Material resources',
     follower.benefits_social.services = 'Social services',
-    follower.benefits_social.status.reputation = 'Increased status',
+    follower.benefits_social.status.reputation = 'Increased social status',
     follower.benefits_territory = 'Territory',
 
     leader.costs_fitness.costs = 'Inclusive fitness cost',

### 2019-11-27 initialcompute.R

@@ -259,7 +259,7 @@ var_names <- c(
   "functions_construction.infrastructure"       = "Construction/infastructure",
   "functions_control.economics"                 = "Control economics",
   "functions_council.member"                    = "Council member",
-  "functions_group.determination"              = "Group determiniation",
+  "functions_group.determination"              = "Group determiniation/cohesiveness",
   "functions_hospitality"                       = "Hospitality",
   "functions_military.command"                  = "Military command",
   "functions_new.settlement"                    =  "Movement/migration",
@@ -270,7 +270,7 @@ var_names <- c(
   "functions_strategic.planning" = "Strategic planning",
   "function_organize.cooperation"  = "Organize cooperation",
   "function_resolve.conflcit"      = "Resolve conflict",
-  "functions_control.calendar"     = "Control calander",
+  "functions_control.calendar"     = "Control calender",
   "functions_control.immigration"  = "Control immigration",
   "functions_distribute.resources" = "Distribute resources",
   "functions_group.representative" = "Group representative",

### 2019-11-27 leadership_across_cultures_contexts.Rmd
* Fixed bug in bib file. * Added cluster analyses with jaccard distance. * Edited stats section of ms.

--- 2019-11-27 leadership_across_cultures_contexts.Rmd
@@
 Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda.
-The interactions between the strategies of individual leaders, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural evolutionary processes [@henrich_big_2015].
-
+The interactions between the strategies of individual leaders, pressures of followership, and cultural forces and group dynamics likely play an out-sized role in contributing to organizational variation across human populations including social and economic trajectories of populations [@wiessner_risk_1982], and cultural evolutionary processes [@henrich_big_2015].
 Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies.
 Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010 and @kaplan_evolutionary_2009; @von_rueden_roots_2014], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
-This integration as been informative.
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010 and @kaplan_evolutionary_2009], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
+This approach as been informative.
 Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019].
-Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and at all levels of social organization.
+Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains.
 Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated.
@@
 We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
-Through novel empirical and exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches.
 
@@
 
-Leadership is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence over group behavior and outcomes.
-Additionally, under the broad definition of leadership given above, parents can be defined as leaders of nuclear families.
-Strong community leadership, however, is common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @lewis_leaders_1974].
+Leadership as a phenomena is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence.
+Also, adopting the broad definition of leadership given above, parents can be defined as leaders of nuclear families.
+Strong community leadership, however, is common cross-culturally [@garfield_review_2019; @lewis_leaders_1974].
 
-The nature of leaderhsip across human societies and social contexts is highly diverse and the emergence of leaders and the functions they serve are closely tied to social and political structure.
+The emergence of leadership and the functions leaders serve across cultures is closely tied to social and political structure.
+
 Among ethnographically described small-scale societies which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
-Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect and greater knowledge and social support to achieve or maintain community influence.
+Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, their greater knowledge and social support to express their influence.
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of conflict resolution and collective actions.
@@
 Within communities, leaders may emerge whenever groups of individuals form.
-The role of leaders in ritual contexts, such as in initiations, and in providing social services, including facilitating marriages and organizing social activities, are also expected to be common functions leaders serve.
+The role of leaders in ritual contexts, such as in initiations, and in providing social services, including facilitating marriages and organizing social activities, are also relatively common.
 
-Increased sedentarization and defensibility of resources is often associated with both social stratification and an institutionalized managerial role and stronger authority of community leaders.
+Increased sedentarization and defensibility of resources is often associated with both social stratification and an institutionalized managerial role of community leaders.
 Among horticulturalists and agriculturalists managerial leadership can promote for more efficient use of shared resources such as water reserves and cooperative labor.
-Pastoralists and coastal populations often face similar pressures surrounding grazing lands and fishing access.
+Pastoralists and coastal populations can face similar pressures surrounding grazing lands and fishing access.
 In the wake of economic intensification, the function of leadership becomes increasing associated with the distribution of resources.
 
-Defensibility of resources necessarily creates increased oppurtunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts.
-Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on disperrsed, less defesnible resources.
+Defensibility of resources is also often associated with increased territoriality and increased pressures of inter-group conflicts.
+Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territory are often forced to manage military forces and develop offensive and defensive strategies in warfare, more so than mobile populations.
 
-In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies with institutionalized responsibilities and functions associated with their position.
-The structure of these organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization.
+In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies with specific responsibilities and functions associated with their position.
+The structure of these organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as a figure head embodying the strategic vision of the organization.
 They are not often involved in resolving immediate conflicts, such as disputes between individuals or nested bodies, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
@@
 Despite immense variation, there do seem to be some commonalities among the qualities of community leaders.
-Across human societies, when strong community leaders emerge, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence.
+Across human societies, when strong community leaders exist, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence.
 Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have increased influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations.
@@
 Physically, leaders across multiple contexts tend to be taller and more physically formidable.
-Such embodied capital might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
+Such physical embodiment might equip individuals for more effective leadership in the context of inter-group conflicts as well as in the context of intra-group dispute settlement and in management of collective actions.
 Charisma and oratory skills are common properties of leaders, but are especially important in the context of egalitarianism where individual leaders must rely on persuasion and more subtle forms of influence to convince community members to come to agreement.
@@
 
-The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the oppurtunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners.
+The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context are able to promote group policies which either align with their individual interests or which are not especially costly for themselves.
 
@@
 
-Institutionalized leadership positions are also often, if not universally, associated with increased social status.
-The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional.
+Institutionalized leadership positions are also typically associated with increased social status.
 
@@
 
-The unique costs associated with leadership positions have received less attention, relative to the benefits of leadership across disciplines.
+The unique costs associated with leadership positions have received less attention, relative to the benefits of leadership.
 
@@
 Evolutionary theoretical models and empirical studies often seek to integrate findings within a broad framework of human cultural diversity.
-The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions they serve is difficult to capture.
-Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations.
-Here, we provide the first broad and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample.
-We also provide the first comparative empirical view describing the costs and benefits of leadership for both leaders nad followers in an evolutionary biological framework.
-We also aim to understand how the qualities and functions of leaders vary across social contexts and levels of cultural complexity.
-We hope these exploratory results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and social organization.
+The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and context and the functions they serve is difficult to capture.
+Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others.
+Here, we provide the first broad and near comprehensive view of the qualities and functions of leaders across from a representative cross-cultural sample.
+We also aim to understand how leader qualities relate to their functions and how these relationships vary across social contexts and levels of cultural complexity.
+We hope these empirical results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundation component of human social behavior and social organization.
 
@@
 First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, creating new variables and operationalizations for additional qualities and functions encountered.
 After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Tables X.
@@
 DISCUSS IRR.
-ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a dataframe named leader_text2).
+ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository.
 
 Coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
-For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscelaneous costs or benefits not captured by these categories.
+For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status/reputation, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, other resources, and other costs or benefits not captured by these categories.
 See Table X.
@@
 We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most frequently noted in the ethnographic record.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
 
-<!-- At the text record level, hierarchical cluster analysis was then used to identify natural groupings of all our coded variables.
-If variables from a model clustered together, and separately from the other variables, that would lend support to the theoretical cohesiveness of that model.
-If, on the other hand, variables from different models clustered together, that would suggest theoretical models may need revision or that clusters had identified components of a novel theoretical model.
--->
+Our second goal was to discover "features" in our data, e.g., to determine if particular variables tended to covary or cluster together.
+Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
+Or they might suggest new and theoretically important dimensions of leadership.
 
-To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` function variables, all of which take binary values (0, 1), we used logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package [@logisticPCA].
-Prior to analysis we removed all rows that had no evidence for any variable.
-We optimized rank, $k$, and $m$ (the value to approximate the saturated model) using cross-validation, per recommendation of the package authors.
-Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitiave measure of leader qualities and leader functions for additional exploratory analyses.
+There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
+To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package.
+This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].
 
+Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables.
+We optimized rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), using cross-validation, per recommendation of the package authors.
+
+We also performed hierarchical cluster analysis on our quality and function variables.
+We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
+We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
+We used 10,000 bootstrapped samples.
+
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record.
+
+Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
+We used linear mixed effects models from the lme4 package [@xx], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
 # Results
@@
 
-The type of group to which a leader text referred was coded into seven categories. Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups.
+The type of group to which a leader text referred was coded into seven categories.
+Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence.
 

### 2019-11-27 initialcompute.R

@@ -506,11 +506,29 @@ m_pvclust_qual <- pvclust(
   parallel = T
 )
 
+m_pvclust_qual_jaccard <- pvclust(
+  pca_data_qualities2, 
+  method.hclust = 'ward', 
+  method.dist = 'binary', 
+  nboot = 10000,
+  parallel = T
+)
+
 m_pvclust_fun <- 
   pvclust(
     pca_data_functions2, 
     method.hclust = 'ward', 
     method.dist = 'correlation', 
+    nboot = 10000,
+    parallel = T
+  )
+
+library(pvclust)
+m_pvclust_fun_jaccard <- 
+  pvclust(
+    pca_data_functions2, 
+    method.hclust = 'ward', 
+    method.dist = 'binary', 
     nboot = 10000,
     parallel = T
   )

### 2019-11-27 leadership_across_cultures_contexts.Rmd

--- 2019-11-27 leadership_across_cultures_contexts.Rmd
@@
 We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most frequently noted in the ethnographic record.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
 
-<!-- At the text record level, hierarchical cluster analysis was then used to identify natural groupings of all our coded variables.
-If variables from a model clustered together, and separately from the other variables, that would lend support to the theoretical cohesiveness of that model.
-If, on the other hand, variables from different models clustered together, that would suggest theoretical models may need revision or that clusters had identified components of a novel theoretical model.
--->
+Our second goal was to discover "features" in our data, e.g., to determine if particular variables tended to covary or cluster together.
+Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
+Or they might suggest new and theoretically important dimensions of leadership.
 
-To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` function variables, all of which take binary values (0, 1), we used logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package [@logisticPCA].
-Prior to analysis we removed all rows that had no evidence for any variable.
-We optimized rank, $k$, and $m$ (the value to approximate the saturated model) using cross-validation, per recommendation of the package authors.
-Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitative measure of leader qualities and leader functions for additional exploratory analyses.
+There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
+To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package.
+This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].
+
+Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables.
+We optimized rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), using cross-validation, per recommendation of the package authors.
+
+We also performed hierarchical cluster analysis on our quality and function variables.
+We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
+We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
+We used 10,000 bootstrapped samples.
+
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record.
+
+Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
+We used linear mixed effects models from the lme4 package [@xx], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
 

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Adding functions/qual and costs/benefits vars table file.
Adding text in results section.

--- 2019-11-27 leadership_across_cultures_contexts.Rmd
@@
 
-Across discplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [@johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
+Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [@johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
 Recent evidence from within managerial studies suggests a curvilinear effect in that exceptionally intelligence individuals at high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
@@
 
-@von_rueden_leadership_2015 outilne four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin, (2) claim a fee for their services, such a tax or greater share of returns, (3) recieve reciprocal exchange in other currencies from followers who have benefited from their services, or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
+@von_rueden_leadership_2015 outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin, (2) claim a fee for their services, such a tax or greater share of returns, (3) receive reciprocal exchange in other currencies from followers who have benefited from their services, or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
 
@@
 
+To idenfity natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster anlayses using the pvclust package [@suzuki_pvclust:_2015].
+These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary.
+
+All analyses were conducted with R version...
+
 # Results
@@
 
+Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i,.
+v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$.
+Using the pvclust package [@suzuki_pvclust:_2015] we clustered the variables using the matrix of all pairwise distances using the Ward aggolomeration algorithm.
+
+Figure \@ref(fig:qualcluster) displays a dendrogram from the cluster analysis of the `r length(quality_vars)` ledaer quality variables and Figure \@ref(fig:funcluster) displays a dendrogram from the cluster analysis of the `r length(function_vars)`.
+Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.
+We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).
+
 ```{r qualcluster, fig.cap="Cluster analysis of leader quality variables.
@@
 ```
+
+The leader quality cluster analysis (Figure \@ref(fig:qualcluster)) idenfitified two top-level clusters.
+Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text.
+Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (e.g., *Authority*, *Aggressiveness*, *Feared*) and shamanism (e.g., *Supernatural*, *Artistic performance*).
+The second sub-cluster includes qualities of leaders related to material, social, and reproductive success.
 
@@
 
-A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_qualities$U)[2]`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
-We draw on the first of the three leader quality components for interpretation and additional analyses and identify this component as a *Prestige vs.
+A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
+We draw on the first two leader quality components for interpretation and additional analyses.
+We identify the first component (PC1) as a *Prestige vs.
 Dominance* measure.
-Low values on Leader qualities PC 1 (Prestige vs.
-Dominance) are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Oratory skill* and *Innovative*.
-High values are strongly associated with *Polygnous* and *Coercive authority* and moderately associated with *Many children* and *High quality spouse*.
+Low values on Leader qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*.
+High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*.
 Therefore, the "dominance" side of the *Prestige vs.
-Dominance* component is also closely liked to reproduction.
+Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:qualcluster)).
 Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component.
 See Figure \@ref(fig:qualitiesPCplot).
+The second leader qualities component we identify as an *In-group favoritism vs.
+Social Status* measure.
+Low values on the Leader qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*.
+High Values are strongly associated with *High status* and moderately associated with *Fairness* and *Age* and *Decisiveness/decision-making*.
 
@@
 
-Similarly, A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
-The first Leader functions component distinguishes between interpersonal mediation functions and group organizational and social functions, we define as a *Mediation vs.
-Organization* measures.
-Low values on Leader functions PC 1 (Mediation vs.
-Organization) are strongly associated with the variables *Resolve conflict* and *Provide counsel/direction* and moderately associated with *Political appointments* and *Punishment*, whereas high values are strongly associated with *Social functions* and *Organize cooperation* and moderately associated with *Ritual functions*.
+Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
+The first Leader functions component distinguishes between social functions and general prosocial function from other more specific leader functions, which we identify as a *Social functions* component.
+High values on Leader function PC 1 are strongly associated with *Social functions*.
+The Leader functions PC 2 distinguished between organizational and mediation-based functions, which we identify as a *Organization vs.
+Mediation* component.
+Low values on Leader functions PC 2 are strongly associated with the variables *Organize cooperation* and *Political appointments*, whereas high values are strongly associated with *Resolve conflict* and moderately associated with *Provide counsel/direction*.
 See Figure \@ref(fig:functionsPCplot).
@@
 
-Leaders are most commonly receiving benefits in the form of greater access to resources and increased social status.
-Receiving social services and mating benefits are also relatively frequent benefits of leadership cross-culturally.
+Leaders are most commonly receiving benefits in the form of greater access to mateiral resources and increased social status.
+Receiving social services, mating benefits, and food benefits are also relatively frequent benefits of leadership cross-culturally.
 The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
-Followers on the other hand most commonly benefit from leadership systems through receiving resources and social services.
+
+Followers on the other hand also benefit from leadership systems through receiving material resources but otherwise most frequently benefit through social *services* provided.
 Mating and fitness benefits are relatively rare for followers.
-Followers most frequently incur costs in the form of resources lost and in providing social services.
+Followers most frequently incur costs in the form of material resources lost and in providing social services.
 For both leaders and followers, individuals stand to lose what they are able to gain.
@@
 
+## Limitations
+
 # References

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Commenting out jaccard plotting for now. Also not evaluating tables because csv files not available.

--- 2019-11-27 leadership_across_cultures_contexts.Rmd
@@
 
-Table: (\#tab:qualfuncvars) Operational definitions for coded leader qualities and functions.
-
-```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
+```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE, eval=F}
 library(readr)
@@
 library(knitr)
-kable(qual_func_var_tab)
+kable(qual_func_var_tab, caption = "Operational definitions for coded leader qualities and functions.")
 ```
@@
 
-Table: (\#tab:costbenefitvars) Operational definitions for coded benefits and costs for leaders and followers.
-
-```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE, }
+```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE, eval=F}
 library(readr)
@@
 library(knitr)
-kable(cost_benefit_vars_tab)
+kable(cost_benefit_vars_tab, caption = "Operational definitions for coded benefits and costs for leaders and followers.")
 ```
@@
 A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_qualities$U)[2]`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
-We draw on the first of the three leader quality components for interpretation and additional analyses and identify this component as a *Prestige vs.
+We draw on the first two leader quality components for interpretation and additional analyses and identify this component as a *Prestige vs.
 Dominance* measure.
@@
 
-Similarly, A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
+Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
 The first Leader functions component distinguishes between interpersonal mediation functions and group organizational and social functions, we define as a *Mediation vs.

### 2019-11-27 analyses2.R

@@ -168,20 +168,20 @@ plot(m_pvclust_qual)
 pvrect(m_pvclust_qual, alpha = 0.9)
 dev.off()
 
-pdf(file = "Figures/m_pvclust_qual_jaccard.pdf", width=12, height=8)
-plot(m_pvclust_qual_jaccard)
-pvrect(m_pvclust_qual_jaccard, alpha = 0.9)
-dev.off()
+# pdf(file = "Figures/m_pvclust_qual_jaccard.pdf", width=12, height=8)
+# plot(m_pvclust_qual_jaccard)
+# pvrect(m_pvclust_qual_jaccard, alpha = 0.9)
+# dev.off()
 
 pdf(file = "Figures/m_pvclust_fun.pdf", width=12, height=8)
 plot(m_pvclust_fun)
 pvrect(m_pvclust_fun, alpha = 0.9)
 dev.off()
 
-pdf(file = "Figures/m_pvclust_fun_jaccard.pdf", width=12, height=8)
-plot(m_pvclust_fun_jaccard)
-pvrect(m_pvclust_fun_jaccard, alpha = 0.9)
-dev.off()
+# pdf(file = "Figures/m_pvclust_fun_jaccard.pdf", width=12, height=8)
+# plot(m_pvclust_fun_jaccard)
+# pvrect(m_pvclust_fun_jaccard, alpha = 0.9)
+# dev.off()
 
 # plot(qual_func_clust)
 # pvrect(qual_func_clust, alpha = 0.9)

### 2019-11-27 leadership_across_cultures_contexts.Rmd

--- 2019-11-27 leadership_across_cultures_contexts.Rmd
@@
 We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most frequently noted in the ethnographic record.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
 
-<!-- At the text record level, hierarchical cluster analysis was then used to identify natural groupings of all our coded variables.
-If variables from a model clustered together, and separately from the other variables, that would lend support to the theoretical cohesiveness of that model.
-If, on the other hand, variables from different models clustered together, that would suggest theoretical models may need revision or that clusters had identified components of a novel theoretical model.
+Our second goal was to discover "features" in our data, e.g., to determine if particular variables tended to covary or cluster together.
+Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
+Or they might suggest new and theoretically important dimensions of leadership.
+
+There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
+To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package.
+This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].
+
+Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables.
+We optimized rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), using cross-validation, per recommendation of the package authors.
+
+<!-- To idenfity natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster anlayses using the pvclust package [@suzuki_pvclust:_2015].
+These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary.
 -->
 
-To reduce the dimensionality of the `r length(quality_vars)` quality variables and the `r length(function_vars)` function variables, all of which take binary values (0, 1), we used logistic PCA from the Binary Dimensionality Reduction (logisticPCA) package [@logisticPCA].
-Prior to analysis we removed all rows that had no evidence for any variable.
-We optimized rank, $k$, and $m$ (the value to approximate the saturated model) using cross-validation, per recommendation of the package authors.
-Results from logistic PCA stand to (1) reveal which measures (coded variables) distinguish leaders and (2) produce a quantitative measure of leader qualities and leader functions for additional exploratory analyses.
+All analyses were conducted with R version...
 
-To idenfity natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster anlayses using the pvclust package [@suzuki_pvclust:_2015].
-These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary.
+We also performed hierarchical cluster analysis on our quality and function variables.
+We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
+We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
+We used 10,000 bootstrapped samples.
 
-All analyses were conducted with R version...
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record.
+
+Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
+We used linear mixed effects models from the lme4 package [@xx], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
+>>>>>>> 59ccacb71b5cc33ebe98b5790a61cccb17f65052
 

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Also moved table captions into "caption" arg in kable.

@@
 
-```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE, eval=F}
+```{r qualfuncvars, echo=FALSE, message=F, warning=FALSE}
 library(readr)
@@
 
-```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE, eval=F}
+```{r costbenefitvars, echo=FALSE, message=F, warning=FALSE}
 library(readr)

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Removed crap from merge conflict

@@
 Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
-We used linear mixed effects models from the lme4 package [@xx], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
->>>>>>> 59ccacb71b5cc33ebe98b5790a61cccb17f65052
+We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
 

### 2019-11-27 leadership_across_cultures_contexts.Rmd
* Created all_study_vars * computed text stats

@@
 The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
-The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, with a median word count of X, a mean of X, and standard deviation of X.
+The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
+Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.
+
 The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
@@
 
-Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i,.
-v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$.
+Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$.
 Using the pvclust package [@suzuki_pvclust:_2015] we clustered the variables using the matrix of all pairwise distances using the Ward aggolomeration algorithm.
 
-Figure \@ref(fig:qualcluster) displays a dendrogram from the cluster analysis of the `r length(quality_vars)` ledaer quality variables and Figure \@ref(fig:funcluster) displays a dendrogram from the cluster analysis of the `r length(function_vars)`.
+Figure \@ref(fig:qualcluster) displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:funcluster) displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
 Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.

### 2019-11-27 analyses2.R

@@ -12,6 +12,7 @@ library(leadershipdata)
 
 # Load libraries ----------------------------------------------------------
 library(tidyverse)
+library(tidytext)
 library(forcats)
 library(NMF)
 library(pvclust)
@@ -756,6 +757,8 @@ visreg(m_qPC2)
 
 # Compute values ----------------------------------------------------------
 
+final_record_count <- sum(rowSums(leader_text2[all_study_vars])>0)
+
 male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
 female_leader_pct <- signif(100*sum(leader_text2$demo_sex=='female', na.rm=T)/nrow(leader_text2), 2)
 
@@ -765,6 +768,18 @@ statustxts <- sum(leader_text2$qualities_high.status)
 intellpolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities.knowlageable.intellect)
 statuspolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities_high.status)
 
+# text analysis
+# leader_text has 1000 rows
+# need raw texts for all 1212 rows
+
+textstats <- leader_text %>% 
+  dplyr::select(cs_textrec_ID, raw_text) %>% 
+  unnest_tokens(word, raw_text) %>% 
+  # dplyr::filter(is.na(as.numeric(word))) %>% # filters out numbers, some of which are page numbers
+  group_by(cs_textrec_ID) %>% 
+  summarise(count = n()) %>% 
+  summarise(min = min(count), max = max(count), mean = mean(count), median = median(count), sd = sd(count)) %>% 
+  round(1)
 
 # Group structure by sex --------------------------------------------------
 

### 2019-11-27 initialcompute.R

@@ -57,6 +57,8 @@ leader_cost_vars <- lt2vars[str_detect(lt2vars, 'leader.costs')]
 follower_benefit_vars <- lt2vars[str_detect(lt2vars, 'follower.benefits')]
 follower_cost_vars <- lt2vars[str_detect(lt2vars, 'follower.costs')]
 
+all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
+
 # Aggregate at Culture level ----------------------------------------------
 
 # df that links the ids of texts, docs, and cultures

### 2019-11-27 leadership_across_cultures_contexts.Rmd
Changed some wonky bib keys

@@
 Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$.
-Using the pvclust package [@suzuki_pvclust:_2015] we clustered the variables using the matrix of all pairwise distances using the Ward aggolomeration algorithm.
+Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward aggolomeration algorithm.
 

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Minor edits.

@@
 Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables.
-We optimized rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), using cross-validation, per recommendation of the package authors.
+We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors.
 
@@
 -->
-
-All analyses were conducted with R version...
 
@@
 We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
+
+All analyses were conducted with `r R.version.string`.
 
@@
 For predictor variables we incorporate additional population-level measures.
-We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “huntergatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
+We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
 We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

### 2019-11-28 analyses2.R

@@ -85,6 +85,8 @@ leader_text2$pop_density <-
     )
   )
 
+all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
+
 # Variable support plots --------------------------------------------------
 
 # Functions & qualities
@@ -801,8 +803,8 @@ df_groups <-
         'economic group',
         'political group (community)',
         'political group (supracommunity)',
-        'religious group',
-        'military group'
+        'military group',
+        'religious group'
       )
     )
   )

### 2019-11-28 initialcompute.R

@@ -57,8 +57,6 @@ leader_cost_vars <- lt2vars[str_detect(lt2vars, 'leader.costs')]
 follower_benefit_vars <- lt2vars[str_detect(lt2vars, 'follower.benefits')]
 follower_cost_vars <- lt2vars[str_detect(lt2vars, 'follower.costs')]
 
-all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
-
 # Aggregate at Culture level ----------------------------------------------
 
 # df that links the ids of texts, docs, and cultures

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Edits to MS.

@@
 
-Figure \@ref(fig:qualcluster) displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:funcluster) displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
+Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
 Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data.
@@
 
-```{r qualcluster, fig.cap="Cluster analysis of leader quality variables.
+```{r cluster, fig.cap="Cluster analysis of leader quality variables (left) and leader function variables (right).
 Distances are $1 - cor$.
@@
 # pvrect(m_pvclust_qual, alpha = 0.9)
-knitr::include_graphics("Figures/m_pvclust_qual.final.png")
+knitr::include_graphics("Figures/m_pvclust.both.png")
 ```
 
-The leader quality cluster analysis (Figure \@ref(fig:qualcluster)) idenfitified two top-level clusters.
+The leader quality cluster analysis (Figure \@ref(fig:cluster)A) idenfitified two top-level clusters.
 Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text.
@@
 The second sub-cluster includes qualities of leaders related to material, social, and reproductive success.
-
-```{r funcluster, fig.cap="Cluster analysis of leader function variables.
-Distances are $1 - cor$.
-Ward agglomeration method.
-p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
-# plot(m_pvclust_fun)
-# pvrect(m_pvclust_fun, alpha = 0.9)
-knitr::include_graphics("Figures/m_pvclust_fun.final.png")
-```
 
@@
 Therefore, the "dominance" side of the *Prestige vs.
-Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:qualcluster)).
+Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)).
 Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component.

### 2019-11-28 analyses2.R

@@ -12,6 +12,7 @@ library(leadershipdata)
 
 # Load libraries ----------------------------------------------------------
 library(tidyverse)
+library(tidytext)
 library(forcats)
 library(NMF)
 library(pvclust)
@@ -83,6 +84,8 @@ leader_text2$pop_density <-
       "over 500 persons / sq. mile"
     )
   )
+
+all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
 
 # Variable support plots --------------------------------------------------
 
@@ -168,20 +171,20 @@ plot(m_pvclust_qual)
 pvrect(m_pvclust_qual, alpha = 0.9)
 dev.off()
 
-pdf(file = "Figures/m_pvclust_qual_jaccard.pdf", width=12, height=8)
-plot(m_pvclust_qual_jaccard)
-pvrect(m_pvclust_qual_jaccard, alpha = 0.9)
-dev.off()
+# pdf(file = "Figures/m_pvclust_qual_jaccard.pdf", width=12, height=8)
+# plot(m_pvclust_qual_jaccard)
+# pvrect(m_pvclust_qual_jaccard, alpha = 0.9)
+# dev.off()
 
 pdf(file = "Figures/m_pvclust_fun.pdf", width=12, height=8)
 plot(m_pvclust_fun)
 pvrect(m_pvclust_fun, alpha = 0.9)
 dev.off()
 
-pdf(file = "Figures/m_pvclust_fun_jaccard.pdf", width=12, height=8)
-plot(m_pvclust_fun_jaccard)
-pvrect(m_pvclust_fun_jaccard, alpha = 0.9)
-dev.off()
+# pdf(file = "Figures/m_pvclust_fun_jaccard.pdf", width=12, height=8)
+# plot(m_pvclust_fun_jaccard)
+# pvrect(m_pvclust_fun_jaccard, alpha = 0.9)
+# dev.off()
 
 # plot(qual_func_clust)
 # pvrect(qual_func_clust, alpha = 0.9)
@@ -756,6 +759,8 @@ visreg(m_qPC2)
 
 # Compute values ----------------------------------------------------------
 
+final_record_count <- sum(rowSums(leader_text2[all_study_vars])>0)
+
 male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
[…]

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Added EHB csl file
Working on introduction and results text.

@@
 
-In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies with institutionalized responsibilities and functions associated with their position.
-The structure of these organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization.
+In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies.
+Other public service and nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders CITE BASS.
+These positions typically are associated with institutionalized responsibilities and specific functions associated with their role.
+The structure of such organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization.
 They are not often involved in resolving immediate conflicts, such as disputes between individuals or nested bodies, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
 
+Scholars within managerial studies have long discussed what leaders and managers do and have developed multiple models of the functions of leadership.
+Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [@bass_handbook_1990].
+These perspectives focused on the rational, mechanistic aspects of group function and overlooked interpersonal dynamics.
+Later theorists incorporated cognitive and socioemotional components of leadership functions.
+For example, @selznick_leadership_1957 describes the functions of organization leaders is to define the goals of the organization, creating structures for accomplishing group goals, defending institutional integrity, and maintain internal consistency.
+
+@wofford_factor_1970 provides an early empirical view of managerial functions.
+A factor analysis of managerial behaviors from over 130 US employees' responses to 183 questionnaire items identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
+Wofford's analysis provided a foundation for future empirical studies of leadership and management functions.
+He contrasted factor one type management, described as the professional administrator who effectively collaborates and integrates team members, from factors two and five, described as autocratic leaders strongly motivated by personal power and willing to rely on aggressive tactics [@wofford_factor_1970].
+
+More recently, many managerial scholars have adopted a comparative approach to better understand cultural diversity in leadership systems, drawing on Culturally Endorsed Implicit Leadership Theory (CLT), which suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles and behaviors [@moore_literature_2018].
 The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leadership effectiveness and cultural variation across large organizations.
@@
 
-## The benefits of leadership across cultures and contexts
+This theoretical development ultimately produced perspectives emphasizing that organizations become larger, increasing hierarchical, adm more complex leaders...
+(3078 Bass) -- INTELLIGENCE and COGNITIVE COMPLEXITY IN LARGER INSTITUIONS
+
+## The benefits and costs of leadership across cultures and contexts
 
@@
 Institutionalized leadership positions are also often, if not universally, associated with increased social status.
-The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional.
+The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional, but is widely associated with numerous social, material, and reproductive benefits.
 
 Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need.
+Across many industries and organizations leadership positions are associated with increased pay.
+High social status, material wealth, and leadership status are also viewed as sexually attractive, particularly among women [@buss_strategies_2006].
+
+<!-- NEED BETTER CITATIONS ABOVE -->
 
@@
 
-@von_rueden_leadership_2015 outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin, (2) claim a fee for their services, such a tax or greater share of returns, (3) receive reciprocal exchange in other currencies from followers who have benefited from their services, or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
-
-## The costs of leadership across cultures and contexts
+@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
 
@@
 In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
+
+Conceptualizing leadership as a possible solution to problems associated with collective goods [@olson_logic_1965], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leader incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
+In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the imposed costs.
+@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in urban neighborhoods and neighborhood organizations, which can be viewed as self-interested groups which face group-wide problems and necessarily share public benefits.
+The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination).
+@rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization.
+Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position.
+@rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US.
+Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics.
+Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.
 
@@
 
-The most common *functions* leaders serve, which are documented in over 70% of cultures, are providing counsel or direction in follower's affairs, conflict resolution, and organizing various cooperative activities.
-Leaders also commonly provide social functions, serve as a representative in inter-group interactions, and provide context-specific leadership such as military command and in providing ritual services.
-In about half of the cultures sampled here leaders punish other group members, control economic systems, and are responsible for policy development.
+The most common *functions* leaders serve, which are documented in over 70% of cultures, are resolving conflicts, providing various social functions, providing counsel or direction in follower's affairs, and organizing various cooperative activities.
+Leaders also commonly serve as a representative in inter-group interactions, manage the distribution of resources, and provide context-specific leadership such as military command and in providing ritual services.
+In about half of the cultures sampled here leaders punish other group members and manage economic systems.
 Notable infrequent functions identified here include controlling group movement and migrations.
 
-The most common *qualities* of leaders are being of high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent.
-Also, in about 70% of cultures leaders maintained coercive authority over followers.
-Strategic nepotism, generosity, and having a favorable personality are also common leader qualities, documented in over 50% of cultures.
+The most common *qualities* of leaders are being of high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures).
+Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, and maintained strong authority over followers.
 Notable infrequent qualities of leaders in these data include charisma and fairness.
@@
 Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$.
-Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward aggolomeration algorithm.
+Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward agglomeration algorithm.
 
@@
 
-The leader quality cluster analysis (Figure \@ref(fig:cluster)A) idenfitified two top-level clusters.
+The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters.
 Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text.
-Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (e.g., *Authority*, *Aggressiveness*, *Feared*) and shamanism (e.g., *Supernatural*, *Artistic performance*).
-The second sub-cluster includes qualities of leaders related to material, social, and reproductive success.
+Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (i.e., *Authority*, *Aggressiveness*, *Feared*) and shamanism (i.e., *Supernatural*, *Artistic performance*).
+The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygnous*).
+Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).
+
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as an organizational functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
 
@@
 A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
-We draw on the first two leader quality components for interpretation and additional analyses.
+We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)).
 We identify the first component (PC1) as a *Prestige vs.
 Dominance* measure.
-Low values on Leader qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*.
+Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*.
 High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*.
@@
 Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component.
-See Figure \@ref(fig:qualitiesPCplot).
 The second leader qualities component we identify as an *In-group favoritism vs.
 Social Status* measure.
-Low values on the Leader qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*.
-High Values are strongly associated with *High status* and moderately associated with *Fairness* and *Age* and *Decisiveness/decision-making*.
+Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*.
+High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*.
+Therefore, Leader Qualities PC 2 also reflects elements of a presetige-dominance distinction in contrasting aggressiveness with high social status.
 
-```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components."}
+```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components.
+PC 1 is identified as a *Prestige vs.
+Dominance* component and PC 2 is identified as an *In-group favoritism vs.
+Social Status* component."}
 logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2)
@@
 
-Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
-The first Leader functions component distinguishes between social functions and general prosocial function from other more specific leader functions, which we identify as a *Social functions* component.
-High values on Leader function PC 1 are strongly associated with *Social functions*.
-The Leader functions PC 2 distinguished between organizational and mediation-based functions, which we identify as a *Organization vs.
+Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
+The first leader functions component distinguishes between social functions (a very broad and generalized measure) from other more specific leader functions, which we identify as a *Social functions* component.
+Low values on Leader function PC 1 are strongly associated with *Social functions*.
+The Leader functions PC 2 distinguished between organizational and mediation-based functions, which we identify as an *Organization vs.
 Mediation* component.
@@
 
-```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components."}
+```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components.
+PC 1 is identified as a *Social functions* component and PC 2 is identified as an *Organization vs.
+Mediation* component."}
 logisticPCA_loadings_plot(logpca_model_functions, data = pca_data_functions2)
@@
 
-Leaders are most commonly receiving benefits in the form of greater access to mateiral resources and increased social status.
+In these data leaders are most commonly receiving benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
 Receiving social services, mating benefits, and food benefits are also relatively frequent benefits of leadership cross-culturally.
-The costs follow the potential benefits and leaders stand to lose their high social status and greater access to resources.
+The costs follow the potential benefits and leaders stand to lose their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures).
 
-Followers on the other hand also benefit from leadership systems through receiving material resources but otherwise most frequently benefit through social *services* provided.
+Followers on the other hand also benefit from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently benefit through social *services* provided (documented in about 35% of cultures).
 Mating and fitness benefits are relatively rare for followers.
-Followers most frequently incur costs in the form of material resources lost and in providing social services.
+Followers most frequently incur costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
+
 For both leaders and followers, individuals stand to lose what they are able to gain.
@@
 To explore potential population variation in the *Prestige vs.
-Dominance* leader qualities component and the *Mediation vs.
-Organization* leader functions component, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture with these components as outcome measures.
+Dominance* and *In-group favoritism vs.
+Social Status* leader qualities components and the *Social functions* and *Mediation vs.
+Organization* leader functions components, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture with these components as outcome measures.
 

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Minor edit

--- 2019-11-28 leadership_across_cultures_contexts.Rmd
@@
 bibliography: Leader2.bib
+csl: evolution-and-human-behavior.csl
 ---
@@
 Ward agglomeration method.
-p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
+AU p-values (red) computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
 # plot(m_pvclust_qual)

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Working on MS, intro, discussion, results.

--- 2019-11-28 leadership_across_cultures_contexts.Rmd
@@
 
-In sum, we aim to empirically and systematically answer the following questions: 1) what are the most frequent qualities of leaders across cultures?; 2) what functions are leaders most commonly serving across cultures?; 3) what are the most salient costs and benefits of leadership for both leaders and followers?; and 4) how does variation across these measures relate to population-level variation?
+In sum, we aim to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent qualities of leaders and the functions they serve; 2) identify features of the qualities and functions of leaders in reference to theoretical models or novel interpretations; 3) identify the most cross-culturally frequent costs and benefits of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
 
-We hope these exploratory results will facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
+Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
 
@@
 
-<!-- To idenfity natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster anlayses using the pvclust package [@suzuki_pvclust:_2015].
+<!-- To identify natural groupings of variables among the research coded leader quality and leader function measures, we conducted two cluster analyses using the pvclust package [@suzuki_pvclust:_2015].
 These analyses allow us to interpret *post-hoc* (1) the degree to which coded variables relate to established theoretical models, (2) identify potential phenotypic profiles of leaders, and (3) offer empirical and theoretical future directions related to suites of qualities or functions likely to covary.
@@
 Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence.
+We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
+We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
 
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars) ), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars) ), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or philosophy).
 See Figure \@ref(fig:varsupport).
@@
 High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*.
-Therefore, Leader Qualities PC 2 also reflects elements of a presetige-dominance distinction in contrasting aggressiveness with high social status.
+Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.
 
@@
 Social Status* leader qualities components and the *Social functions* and *Mediation vs.
-Organization* leader functions components, we fit a series of hierarchical linear models at the text-level, using nested random effects of document within culture with these components as outcome measures.
+Organization* leader functions components, we fit four hierarchical linear models at the text-level, using nested random effects of document within culture with each of these components as outcome measures.
+For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:groupsex)).
 
-For predictor variables we incorporate additional population-level measures.
+<!-- For predictor variables we incorporate additional population-level measures.
 We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
 We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
+-->
 
@@
 
+Results of the current study represent the broadest and most comprehensive view of the qualities of leaders, (e.g., what types of people occupy leadership roles?), the functions of leaders, (e.g., what do leaders do?), and the costs and benefits of leadership for both leaders and followers, across a diverse and culturally representative sample of the ethnographic record.
+These results suggest universal dimensions of human leadership while also suggesting important variation by group type, subsistence strategy, and sex of the leader.
+
+## The universality of human leadership systems
+
+Across a broad range of cultures and social contexs leaders are nearly universally high status males described as knoweldage or intelligent who are experienced and have a demonstrated history of previous accomplishments or special skills.
+They most frequently resolve conflicts within their respective groups while also providing important social functions, organizign cooperation among followers, and providing counsel or input in followers' affairs.
+
 ## Limitations
 
+This study is based on the ethnographic record.
+Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample.
+Nonetheless, these results are conditional upon the content ethnographers chose to discuss.
+The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results.
+We, however, view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description.
+Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record.
+These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent.
+
 # References

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Commenting out logisticPCA stuff for now.

--- 2019-11-28 leadership_across_cultures_contexts.Rmd
@@
 
-In sum, we aim to empirically and systematically answer the following questions: 1) what are the most frequent qualities of leaders across cultures?; 2) what functions are leaders most commonly serving across cultures?; 3) what are the most salient costs and benefits of leadership for both leaders and followers?; and 4) how does variation across these measures relate to population-level variation?
+In sum, we aim to empirically and systematically answer the following questions: 1) what are the most frequent qualities of leaders across cultures?
+2) What functions are leaders most commonly serving across cultures?; 3) What "features" exist in our data, e.g., do particular variables tended to covary or cluster together?
+Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
+Or they might suggest new and theoretically important dimensions of leadership.
+4) What are the most salient costs and benefits of leadership for both leaders and followers?
+And 5) how does variation across these measures relate to population-level variation?
 
@@
 Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
+We computed confidence intervals for the proportion of text records supporting each variable using generalized linear mixed effects models and random intercepts based on document authors nested within cultures.
+We computed confidence intervals for the proportion of cultures with evidence for each variable using a cluster bootstrap.
 
-Our second goal was to discover "features" in our data, e.g., to determine if particular variables tended to covary or cluster together.
-Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
-Or they might suggest new and theoretically important dimensions of leadership.
+We then explored "features" in our data.
+There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
+We used hierarchical cluster analysis on our quality and function variables.
+We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
+We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
+We used 10,000 bootstrapped samples.
 
-There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
-To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package.
+<!-- To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package.
 This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].
-
 Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables.
 We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors.
+-->
 
@@
 
-We also performed hierarchical cluster analysis on our quality and function variables.
-We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
-We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
-We used 10,000 bootstrapped samples.
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above.
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record.
-
-Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
-We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
+Our fourth goal was to explore associations between the most common leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
+We used linear mixed effects models from the lme4 package [@bates2015], with common qualities and functions as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
 

### 2019-11-28 initialcompute.R

@@ -224,8 +224,8 @@ for (m in models){
     intercept = cfs$coefficients[[1]]
     se = cfs$coefficients[[2]]
     value = c(value, logit.inv(intercept))
-    y_se = c(y_se, logit.inv(intercept + se))
-    y_negse = c(y_negse, logit.inv(intercept - se))           
+    y_se = c(y_se, logit.inv(intercept + 2*se))
+    y_negse = c(y_negse, logit.inv(intercept - 2*se))           
     
   }
 }
@@ -370,8 +370,8 @@ for (m in models){
     intercept = cfs$coefficients[[1]]
     se = cfs$coefficients[[2]]
     value = c(value, logit.inv(intercept))
-    y_se = c(y_se, logit.inv(intercept + se))
-    y_negse = c(y_negse, logit.inv(intercept - se))           
+    y_se = c(y_se, logit.inv(intercept + 2*se))
+    y_negse = c(y_negse, logit.inv(intercept - 2*se))           
     
   }
 }

### 2019-11-30 leadership_across_cultures_contexts.Rmd
Working in entire MS.
Added gg visreg plots in exploratory models. Enhanced group by subsistence plot labels.

Set varaible distribution plot replication to 10,000.

--- 2019-11-28 leadership_across_cultures_contexts.Rmd
@@
 
-We first briefly review current theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership for both leaders and followers across human societies, including WEIRD and non-WEIRD populations.
+We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
 We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
-Through novel empirical and exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches.
+Through novel empirical and exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
 
@@
 
-Leadership is a universal feature of human groups; among all societies there are particular contexts in which one individual maintains greater influence over group behavior and outcomes.
-Additionally, under the broad definition of leadership given above, parents can be defined as leaders of nuclear families.
-Strong community leadership, however, is common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @lewis_leaders_1974].
+Leaders emerge in all human societies.
+That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group and functions to facilitate group goals and provides services to the group.
+Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families.
+Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @lewis_leaders_1974; @von_rueden_leadership_2015].
 
-The nature of leadership across human societies and social contexts is highly diverse and the emergence of leaders and the functions they serve are closely tied to social and political structure.
-Among ethnographically described small-scale societies which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
-Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect and greater knowledge and social support to achieve or maintain community influence.
-The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of conflict resolution and collective actions.
-Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups.
-Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers.
-Representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
-
-Within communities, leaders may emerge whenever groups of individuals form.
+Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are closely tied to group structure.
+Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
+Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions conflict resolution [@glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015].
+Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
+Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
+Across relatively small-scale societies, at the community level, representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of leaders.
+Within communities, however, leaders may emerge whenever individuals for groups.
 The role of leaders in ritual contexts, such as in initiations, and in providing social services, including facilitating marriages and organizing social activities, are also expected to be common functions leaders serve.
 
-Increased sedentarization and defensibility of resources is often associated with both social stratification and an institutionalized managerial role and stronger authority of community leaders.
-Among horticulturalists and agriculturalists managerial leadership can promote for more efficient use of shared resources such as water reserves and cooperative labor.
-Pastoralists and coastal populations often face similar pressures surrounding grazing lands and fishing access.
-In the wake of economic intensification, the function of leadership becomes increasing associated with the distribution of resources.
+Increased sedentarization and defensibility of resources is often associated with both social stratification and institutionalized managerial roles including stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
+Among horticulturalists and agriculturalists managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
+Pastoralists and coastal populations often face similar pressures concerning grazing lands and fishing access [e.g., @widmer_evolution_1988; @stevens_sherpa_1990].
+In the wake of economic intensification, the functions of leadership becomes increasingly associated with the managing economic systems and distribution of resources.
+Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts [@glowacki_evolutionary_2017].
+Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on dispersed, less defensible resources [@lopez_evolutionary_2017].
 
-Defensibility of resources necessarily creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts.
-Community leaders among populations more reliant on domesticated plant foods, livestock, and specific territories are often required to manage military forces and develop both offensive and defensive strategies in warfare, more so than mobile populations that rely on dispersed, less defensible resources.
+In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
+Other public service and nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
+These positions typically are associated with institutionalized responsibilities and specific functions associated with their role.
+The structure of such organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
+They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
 
-In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies.
-Other public service and nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders CITE BASS.
-These positions typically are associated with institutionalized responsibilities and specific functions associated with their role.
-The structure of such organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At the higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization.
-They are not often involved in resolving immediate conflicts, such as disputes between individuals or nested bodies, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
-
-Scholars within managerial studies have long discussed what leaders and managers do and have developed multiple models of the functions of leadership.
+Scholars in managerial studies have long discussed the behaviors of leaders and managers have developed multiple models on the functions of leadership.
 Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [@bass_handbook_1990].
-These perspectives focused on the rational, mechanistic aspects of group function and overlooked interpersonal dynamics.
+These perspectives focused on the rational, mechanistic aspects of group functioning and overlooked interpersonal dynamics.
 Later theorists incorporated cognitive and socioemotional components of leadership functions.
@@
 
+NEED CONCLUSION PARAGRAPH/STATEMENT.
+
 ## The qualities of leaders across cultures and contexts
 
-Despite immense variation, there do seem to be some commonalities among the qualities of community leaders.
-Across human societies, when strong community leaders emerge, they tend to be high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006].
+Across culturally and economically diverse populations there are some broad commonalities in the qualities of community leaders.
+When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
 Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have increased influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
-Drawing on cross-national data through the GLOBE project including data from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are perceived to be associated with effective leadership.
+Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are perceived to be associated with effective leadership.
 In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
-Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].
+
+<!-- Other attributes of charismatic leadership found to be culturally contingent included ambitiousness, self-effacing, enthusiasm, and risk-taking [@den_hartog_culture_1999].  -->
 
@@
 
-Among societies with institutionalized social stratification, such as chiefdoms and agricultural communities, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004].
-Managerial elites, who maintain influence over community decisions are endowed with a high degree of ideological status associated with the role and kin group [@stanish_evolution_2010].
-In such contexts community leaders often lack coercive authority, but maintain strong influence through ideological systems and ritual or religious status and prowess inaccessible to non-elites [@stanish_evolution_2010].
-The association of formal community leadership and supernatural abilities is likely an important precursor of institutionalized authority and social stratification.
+In the context of institutionalized social stratification, such as chiefdoms and many agricultural populations, community leadership is often highly influenced by heredity and formal class structures [@earle_how_1997; @stanish_evolution_2004; @redmond_introduction_1998].
+Managerial elites, who maintain influence over community decisions are endowed with a high degree of social status rooted in ideologies and associated with their roles or kin group [@stanish_evolution_2010].
+In such contexts community leaders often lack strong coercive authority, but maintain substantial influence through ideological systems and ritual or religious prowess inaccessible to non-elites [@stanish_evolution_2010].
+The association of formal community leadership and supernatural abilities is likely an important precursor of institutionalized and coercive authority.
 
 Within state-level societies including non-industrial kingdoms leadership positions are proscribed and institutionalized.
-Ruling classes, parties, or political bodies monopolize social and political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda.
+Ruling classes, parties, or political bodies monopolize social and political influence and maintain control through a variety of mechanisms including military force, ideological systems, and through the control of information and propaganda [@cohen_comparative_1967; @bodley_cultural_2011].
 
 Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [@johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
-Recent evidence from within managerial studies suggests a curvilinear effect in that exceptionally intelligence individuals at high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
+Recent cross-cultural evidence and from a small-scale society implicates strong relationships between perceptions of intelligence and leadership among males [@garfield_evolutionary_2019; @garfield_investigating_2019], when considered in the context of broad interdisciplinary results suggests high intelligence and specialized knowledge are universal traits of political and community leaders.
+Recent evidence from within managerial studies, however, suggests a curvilinear effect in that exceptionally intelligence individuals at high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
 
-This theoretical development ultimately produced perspectives emphasizing that organizations become larger, increasing hierarchical, adm more complex leaders...
-(3078 Bass) -- INTELLIGENCE and COGNITIVE COMPLEXITY IN LARGER INSTITUIONS
+Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002].
+More complex organizations require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view, as well as be equipped to socially manage diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
+
+<!-- COULD GET INTO LEADER VS MANAGER DISTINCTION -->
 
@@
 
-The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners.
+<!-- BENEFITS: DECISIONS IN THEIR INTERST -->
+The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @pietraszewski_evolution_2019; @hagen_leadership_2019].
 
-Institutionalized leadership positions are also often, if not universally, associated with increased social status.
-The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional, but is widely associated with numerous social, material, and reproductive benefits.
+<!-- BENEFITS: HIGH SOCIAL STATUS -->
+Leadership positions, including formalized institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status.
+The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional, but leadership roles through their association with increased social status are widely associated with numerous social, material, and reproductive benefits [@kantner_identifying_2010].
 
-Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need.
-Across many industries and organizations leadership positions are associated with increased pay.
-High social status, material wealth, and leadership status are also viewed as sexually attractive, particularly among women [@buss_strategies_2006].
+<!-- BENEFITS: SOCIAL BENEFITS, SOCIAL NETWORK -->
+Leaders and high-status individuals are also more likely to be the recipients of social support and material benefits in times of need [@gurven_its_2000; @von_rueden_roots_2014; @sugiyama_illness_2004].
+Across many industries and organizations leadership positions are associated with increased financial compensation [@tsui_income_2017].
+High social status, material wealth, intelligence, and leadership status are also viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
 
@@
 
+<!-- BENFITS: MATING -->
 Across diverse populations and organizations, community leaders, political leaders, and organization leaders tend to have greater mating opportunities and therefore, given the aforementioned male-bias in top leadership positions, these leaders also tend to have greater reproductive success.
@@
 
+<!-- BENEFITS: 'MODEL' OF POTENTIAL BENEFIST -->
 @von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
 
+<!-- COSTS: OPPURTUNITY COSTS -->
 Evolutionary models of leadership generally assume the benefits of leadership outweigh the costs, for both leaders and followers.
 However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines.
-
 Group responsibilities necessarily impose opportunity costs on individual priorities and agendas, given leadership responsibilities require investments of time, energy, and resources that could be invested elsewhere.
 
-Mediating conflicts between individuals can be physically and socially costly.
+<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
+Mediating conflicts between individuals can be physically and socially costly [@wiessner_power_2010].
 Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved.
@@
 With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude.
-
 In managing collective action and enacting group-level policies, some behaviors of leaders will upset some followers and leader-follower interactions can be hostile.
 
+<!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
 Conceptualizing leadership as a possible solution to problems associated with collective goods [@olson_logic_1965], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leader incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
@@
 
-The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
-There were female leaders in almost all types of groups, however, except military and religious.
-Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
-
 ```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis.
@@
 
-The type of group to which a leader text referred was coded into seven categories.
+The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups.
 Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence.
@@
 ```
+
+The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
+There were female leaders in almost all types of groups, however, except military and religious.
+Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders.
+
+As Figure \@ref(fig:groupsex) reveals, the distribution of text records by group type varies across subsistence type.
+Across all subsistence groups, the most frequent group type is political groups at the community level, with the exception of agriculturalitsts in which supracommunity political groups are the most frequent group type.
+Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers.
+Leadership within residential groups is most frequently associated hunter gatherers (`r hg_residential_pct`% of the total sample of hunter gatherer text records) and leadership within kin groups is most frequently associated with hunter gatherers (`r hg_kin_pct`% of the total hunter gatherers sample) and horticulturalists (`r hort_kin_pct`% of the total horticulturalists sample).
+
+<!-- SHOULD WE PUT THE ENTIRE TABLE IN SI? -->
 
@@
 The most common *functions* leaders serve, which are documented in over 70% of cultures, are resolving conflicts, providing various social functions, providing counsel or direction in follower's affairs, and organizing various cooperative activities.
-Leaders also commonly serve as a representative in inter-group interactions, manage the distribution of resources, and provide context-specific leadership such as military command and in providing ritual services.
+Leaders also commonly serve as a representatives in inter-group interactions, manage the distribution of resources, and provide context-specific leadership such as military command and in providing ritual services.
 In about half of the cultures sampled here leaders punish other group members and manage economic systems.
-Notable infrequent functions identified here include controlling group movement and migrations.
+Notable infrequent functions include controlling group movement and migrations.
 
 The most common *qualities* of leaders are being of high social status (documented in over 90% of cultures), being experienced or accomplished, and being described as knowledgeable or intelligent (documented in approximately 80% of cultures).
-Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, and maintained strong authority over followers.
+Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained strong authority over followers.
 Notable infrequent qualities of leaders in these data include charisma and fairness.
@@
 
-Results of the current study represent the broadest and most comprehensive view of the qualities of leaders, (e.g., what types of people occupy leadership roles?), the functions of leaders, (e.g., what do leaders do?), and the costs and benefits of leadership for both leaders and followers, across a diverse and culturally representative sample of the ethnographic record.
-These results suggest universal dimensions of human leadership while also suggesting important variation by group type, subsistence strategy, and sex of the leader.
+Results of the current study represent the broadest and most comprehensive view of the qualities of leaders, (e.g., the types of people that occupy leadership roles), the functions of leaders, (e.g., what do leaders do), and the costs and benefits of leadership for both leaders and followers, across a diverse and culturally representative sample of the ethnographic record.
+These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence strategy, and sex of the leader.
 
-## The universality of human leadership systems
+## The universality of human leadership
 
-Across a broad range of cultures and social contexs leaders are nearly universally high status males described as knoweldage or intelligent who are experienced and have a demonstrated history of previous accomplishments or special skills.
-They most frequently resolve conflicts within their respective groups while also providing important social functions, organizign cooperation among followers, and providing counsel or input in followers' affairs.
+Across a broad range of cultures and social contexts leaders are nearly universally high status males described as knowledgable or intelligent who are experienced and have a demonstrated history of previous accomplishments or special skills.
+They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, and providing counsel or input in followers' affairs.
+See Figure \@ref(fig:varsupport).
+
+At all subsistence levels leaders emerge across a broad range of group types, including at lower-order social units such as within residential groups and kin groups; within specialized groups such as religous, economic, or military groups; and in political contexts at the community level and beyond.
+Political leadership at the commmunity level is widespread across all subsistence categories.
+Economic, military, and religious group leadership, although relatively less frequent, are fairly consistently represented across subsistence levels.
+See Figure \@ref(fig:groupsex).
+
+## The cultural dependancy of human leadership
+
+The distribiution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistance strategies, but at variable frequencies.
+Hunter-gatherers and pastoralists, populations more likely to expeirence high residential mobility exhibited greater relative frequencies of leaderhsip wtihin residential groups and kin groups.
+Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community mores o than hunter-gatherers and on par with horticulturalists.
+Both pastoralists and horticulturalists are known for elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaboratored political leadership systems.
+
+Figure \@ref(fig:groupsex) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups.
+One possible interpretation is that as political structures become more instituionalized and economic systems transcend the household, leadership within households and residental groups errodes.
+It is interesting, however, that even among the agricultural populations that are charachterized by leadership systems beyong the community level, leadership within residential and family group persists.
+
+<!-- MAY WANT TO RETHINK INTERPRETATIONS, LOTS OF STORIES HERE -->
+
+## The features of human leadership
 
@@
 These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent.
+The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.
 

### 2019-11-30 analyses2.R

@@ -481,7 +481,17 @@ m_fPC1 <- lmer(
 summary(m_fPC1)
 Anova(m_fPC1)
 AIC(m_fPC1)
-visreg(m_fPC1)
+
+
+m_fPC1_sub_plot <- visreg(m_fPC1, "subsistence", type = "contrast", gg = T) +
+  labs(y = "Functions PC 1: \nSocial functions\n",
+       x = "\nSubsistence")
+m_fPC1_group_plot <- visreg(m_fPC1, "group.structure2", type = "contrast", gg = T) + 
+  labs(y = "Functions PC 1: \nSocial functions\n",
+       x = "\nGroup type")
+
+m_fPC1_sub_plot + m_fPC1_group_plot
+
 
 m_fPC2 <- lmer(
   fPC2 ~
@@ -498,7 +508,15 @@ m_fPC2 <- lmer(
 summary(m_fPC2)
 Anova(m_fPC2)
 AIC(m_fPC2)
-visreg(m_fPC2)
+
+m_fPC2_sub_plot <- visreg(m_fPC2, "subsistence", type = "contrast", gg = T) +
+  labs(y = "Functions PC 2: \nOrganization vs. Mediation\n",
+       x = "\nSubsistence")
+m_fPC2_group_plot <- visreg(m_fPC2, "group.structure2", type = "contrast", gg = T) +
+  labs(y = "Functions PC 2: \nOrganization vs. Mediation\n",
+       x = "\nGroup type")
+
+m_fPC2_sub_plot + m_fPC2_group_plot
 
 # Leader qualities
 
@@ -516,7 +534,15 @@ m_qPC1 <- lmer(
 )
 summary(m_qPC1)
 Anova(m_qPC1)
-visreg(m_qPC1)
+
+m_qPC1_sub_plot <- visreg(m_qPC1, "subsistence", type = "contrast", gg = T) +
+  labs(y = "Qualities PC 1: \nPrestige vs. Dominance\n",
+       x = "\nSubsistence")
+m_qPC1_group_plot <- visreg(m_qPC1, "group.structure2", type = "contrast", gg = T) +
+  labs(y = "Qualities PC 1: \nPrestige vs. Dominance\n",
+       x = "\nGroup type")
+
+m_qPC1_sub_plot + m_qPC1_group_plot
 
 m_qPC2 <- lmer(
   qPC2 ~
@@ -532,7 +558,17 @@ m_qPC2 <- lmer(
 )
 summary(m_qPC2)
 Anova(m_qPC2)
[…]

### 2019-11-30 initialcompute.R

@@ -244,7 +244,7 @@ for (m in models) {
     d.tmp = all_data[, c(model_vars[v], 'd_culture', 'author_ID')]
     val = cult.sum(d.tmp)
     value = c(value, val)
-    replicates = replicate(10, cult.sum(resample(d.tmp, cluster=c('d_culture', 'author_ID'), replace=c(T,F))))
+    replicates = replicate(1000, cult.sum(resample(d.tmp, cluster=c('d_culture', 'author_ID'), replace=c(T,F))))
     q=quantile(replicates, c(0.025, 0.975))
     y_se = c(y_se, q[[2]])
     y_negse = c(y_negse, q[[1]])
@@ -390,7 +390,7 @@ for (m in models) {
     d.tmp = all_data[, c(model_vars[v], 'd_culture', 'author_ID')]
     val = cult.sum(d.tmp)
     value = c(value, val)
-    replicates = replicate(10, cult.sum(resample(d.tmp, cluster=c('d_culture', 'author_ID'), replace=c(T,F))))
+    replicates = replicate(1000, cult.sum(resample(d.tmp, cluster=c('d_culture', 'author_ID'), replace=c(T,F))))
     q=quantile(replicates, c(0.025, 0.975))
     y_se = c(y_se, q[[2]])
     y_negse = c(y_negse, q[[1]])

### 2019-11-30 leadership_across_cultures_contexts.Rmd
Minor edits to exploratory model plots.
Working on MS.

Added visreg exploratory model effects plots to MS to visualize effects. Might remove these plots/analyses prior to submission.

@@
 Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology.
-Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016].
+Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018].
 This integration as been informative.
@@
 
-```{r, fig.height=20, fig.width=9}
-# qc_bayes_model_post_plot
+Functions PC 1 model
+
+```{r funcPCmodel1, fig.height=8, fig.width=20}
+m_fPC1_sub_plot + m_fPC1_group_plot
+```
+
+Functions PC 2 model
+
+```{r funcPCmodel2, fig.height=10, fig.width=20}
+m_fPC2_sub_plot + m_fPC2_group_plot
+```
+
+Qualities PC 1 model
+
+```{r qualPCmodel1, fig.height=10, fig.width=20}
+m_qPC1_sub_plot + m_qPC1_group_plot
+```
+
+Qualities PC 2 model
+
+```{r qualPCmodel2, fig.height=10, fig.width=20}
+m_qPC2_sub_plot + m_qPC2_group_plot
 ```
@@
 
-## The cultural dependancy of human leadership
+## The cultural and context dependancy of human leadership
 
 The distribiution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistance strategies, but at variable frequencies.
-Hunter-gatherers and pastoralists, populations more likely to expeirence high residential mobility exhibited greater relative frequencies of leaderhsip wtihin residential groups and kin groups.
-Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community mores o than hunter-gatherers and on par with horticulturalists.
-Both pastoralists and horticulturalists are known for elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaboratored political leadership systems.
+Hunter-gatherers and pastoralists, populations more likely to expeirence high residential mobility, exhibited greater relative frequencies of leaderhsip wtihin residential groups and kin groups.
+Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
+Pastoral and horticultural populations are often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaboratored political leadership systems.
 
@@
 
+# Conclusion
+
 # References

### 2019-11-30 analyses2.R

@@ -485,7 +485,7 @@ AIC(m_fPC1)
 
 m_fPC1_sub_plot <- visreg(m_fPC1, "subsistence", type = "contrast", gg = T) +
   labs(y = "Functions PC 1: \nSocial functions\n",
-       x = "\nSubsistence")
+       x = "\nSubsistence type")
 m_fPC1_group_plot <- visreg(m_fPC1, "group.structure2", type = "contrast", gg = T) + 
   labs(y = "Functions PC 1: \nSocial functions\n",
        x = "\nGroup type")
@@ -511,7 +511,7 @@ AIC(m_fPC2)
 
 m_fPC2_sub_plot <- visreg(m_fPC2, "subsistence", type = "contrast", gg = T) +
   labs(y = "Functions PC 2: \nOrganization vs. Mediation\n",
-       x = "\nSubsistence")
+       x = "\nSubsistence type")
 m_fPC2_group_plot <- visreg(m_fPC2, "group.structure2", type = "contrast", gg = T) +
   labs(y = "Functions PC 2: \nOrganization vs. Mediation\n",
        x = "\nGroup type")
@@ -537,7 +537,7 @@ Anova(m_qPC1)
 
 m_qPC1_sub_plot <- visreg(m_qPC1, "subsistence", type = "contrast", gg = T) +
   labs(y = "Qualities PC 1: \nPrestige vs. Dominance\n",
-       x = "\nSubsistence")
+       x = "\nSubsistence type")
 m_qPC1_group_plot <- visreg(m_qPC1, "group.structure2", type = "contrast", gg = T) +
   labs(y = "Qualities PC 1: \nPrestige vs. Dominance\n",
        x = "\nGroup type")
@@ -561,7 +561,7 @@ Anova(m_qPC2)
 
 m_qPC2_sub_plot <- visreg(m_qPC2, "subsistence", type = "contrast", gg = T) + 
   labs(y = "Qualities PC 2: \nIn-group favoritism vs. Social Status\n",
-       x = "\nSubsistence")
+       x = "\nSubsistence type")
 m_qPC2_group_plot <- visreg(m_qPC2, "group.structure2", type = "contrast", gg = T) +
   labs(y = "Qualities PC 2: \nIn-group favoritism vs. Social Status\n",
        x = "\nGroup type")

### 2019-12-01 analyses2.R
Added cluster analysis with Euclidean distances. Added some more heatmaps.

@@ -30,7 +30,7 @@ library(modelr)
 library(ggridges)
 library(RColorBrewer)
 library(ggmosaic)
-
+library(proxy)
 # Load precomputed objects ------------------------------------------------
 
 # First run initialcompute.R, which takes ~90 minutes on my machine,
@@ -680,8 +680,61 @@ visreg(m_qPC2)
 # 
 # 
 # 
-# # Heatmaps -----------------------------------------------------------------
-# # Heatmaps -----------------------------------------------------------------
+# Heatmaps -----------------------------------------------------------------
+
+# Heatmap of qualities with Euclidean distance
+
+heatmap(
+  t(as.matrix(pca_data_qualities2)),
+  hclustfun = function(x) hclust(x, method = 'ward.D'),
+  scale = 'none'
+  )
+
+# aheatmap version
+
+aheatmap(
+  t(as.matrix(pca_data_qualities2)),
+  distfun = "euclidean", 
+  hclustfun = "ward",
+  scale = "none",
+  filename = "Figures/heatmap_qualities_euc.pdf"
+)
+
+aheatmap(
+  t(as.matrix(pca_data_qualities2)),
+  Rowv = c('correlation', 'ward'),
+  Colv = c('euclidean', 'ward'),
+  scale = "none",
+  filename = "Figures/heatmap_qualities_euc_cor.pdf"
+)
+
+# # Heatmap of qualities with cor & binary distance
+
+heatmap(
+  t(as.matrix(pca_data_qualities2)),
+  hclustfun = function(x) hclust(x, method = 'ward.D'),
+  distfun = function(x) proxy::dist(x, method = 'binary'),
+  scale = 'none'
+  )
+
+# Heatmap of functions with Euclidean distance
+
+# heatmap(
+#   t(as.matrix(pca_data_functions2)),
+#   hclustfun = function(x) hclust(x, method = 'ward.D'),
+#   scale = 'none',
+#   main = "Functions with Euclidean distance"
+# )
+# 
+# # Heatmap of functions with cor distance
+# heatmap(
+#   t(as.matrix(pca_data_functions2)),
+#   hclustfun = function(x) hclust(x, method = 'ward.D'),
+#   distfun = function(x) proxy::dist(x, method = 'correlation'),
+#   scale = 'none',
+#   main = "Functions with Correlation distance"
+# )
+
 # heatmap_data<-leader_text2[,c(quality_vars, "group.structure2")]
 # 
[…]

### 2019-12-01 initialcompute.R

@@ -514,6 +514,14 @@ m_pvclust_qual_jaccard <- pvclust(
   parallel = T
 )
 
+m_pvclust_qual_euc <- pvclust(
+  pca_data_qualities2, 
+  method.hclust = 'ward', 
+  method.dist = 'euclidean', 
+  nboot = 10000,
+  parallel = T
+)
+
 m_pvclust_fun <- 
   pvclust(
     pca_data_functions2, 
@@ -523,12 +531,20 @@ m_pvclust_fun <-
     parallel = T
   )
 
-library(pvclust)
 m_pvclust_fun_jaccard <- 
   pvclust(
     pca_data_functions2, 
     method.hclust = 'ward', 
     method.dist = 'binary', 
+    nboot = 10000,
+    parallel = T
+  )
+
+m_pvclust_fun_euc <- 
+  pvclust(
+    pca_data_functions2, 
+    method.hclust = 'ward', 
+    method.dist = 'euclidean', 
     nboot = 10000,
     parallel = T
   )

### 2019-12-01 leadership_across_cultures_contexts.Rmd

@@
 
-In sum, we aim to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent qualities of leaders and the functions they serve; 2) identify features of the qualities and functions of leaders in reference to theoretical models or novel interpretations; 3) identify the most cross-culturally frequent costs and benefits of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
+In sum, we aim to answer the following questions: 1) what are the most frequent qualities of leaders across cultures?
+2) What functions are leaders most commonly serving across cultures?
+3) What "features" exist in our data, e.g., do particular variables tended to covary or cluster together?
+Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
+Or they might suggest new and theoretically important dimensions of leadership.
+4) What are the most salient costs and benefits of leadership for both leaders and followers?
+And 5) how does variation across these measures relate to population-level variation?
 

### 2019-12-01 analyses2.R
new code in ggmosaic screws up the plot somehow. Commented out for now.

@@ -872,7 +872,7 @@ textstats <- leader_text %>%
   summarise(min = min(count), max = max(count), mean = mean(count), median = median(count), sd = sd(count)) %>% 
   round(1)
 
-# Group structure by sex --------------------------------------------------
+# Group structure by subsistence --------------------------------------------------
 
 df_groups <- 
   leader_text2 %>% 
@@ -912,28 +912,34 @@ plot_group_subsis <-
   geom_mosaic(aes(x = product(group, subsistence), fill = group)) +
   labs(x="", y="", fill = "Group type") +
   guides(fill = guide_legend(reverse = T)) +
-  theme_bw(15) + 
-  theme(axis.text.x = element_text(size=16),
-        axis.text.y = element_text(size=16)) +
-  scale_y_productlist(labels = c("Residential subgroup",
-                                 "Kin group",
-                                 "Economic group",
-                                 "Political group\n(community)",
-                                 "Political group\n(supracommunity)",
-                                 "Military group",
-                                 "Religious group")) +
-  scale_x_productlist(labels = c("Hunter-gatherers",
-                                 "Pastoralists",
-                                 "Mixed",
-                                 "Horticulturalists",
-                                 "Agriculturalists")) + 
-  scale_fill_discrete(labels = c("Residential subgroup",
-                                  "Kin group",
-                                  "Economic group",
-                                  "Political group (community)",
-                                  "Political group (supracommunity)",
-                                  "Military group",
-                                  "Religious group"))
+  theme_bw(15) 
+
+# Leave off for now. Might not need this.
+# + 
+#   theme(axis.text.x = element_text(size=16),
+#         axis.text.y = element_text(size=16)) +
+
+# This code screws up the plot somehow
+#   scale_y_productlist(labels = c("Residential subgroup",
+#                                  "Kin group",
+#                                  "Economic group",
+#                                  "Political group\n(community)",
+#                                  "Political group\n(supracommunity)",
+#                                  "Military group",
+#                                  "Religious group")) +
+#   scale_x_productlist(labels = c("Hunter-gatherers",
+#                                  "Pastoralists",
+#                                  "Mixed",
+#                                  "Horticulturalists",
+#                                  "Agriculturalists")) + 
[…]

### 2019-12-01 analyses2.R
Consolidate computed values

@@ -846,31 +846,6 @@ heatmap(
 # library(NMF)
 # m_nmf <- nmf(t(pca_data_qualities2), rank = 2:10)
 
-# Compute values ----------------------------------------------------------
-
-final_record_count <- sum(rowSums(leader_text2[all_study_vars])>0)
-
-male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
-female_leader_pct <- signif(100*sum(leader_text2$demo_sex=='female', na.rm=T)/nrow(leader_text2), 2)
-
-intelltxts <- sum(leader_text2$qualities.knowlageable.intellect)
-polytxts <- sum(leader_text2$qualities.polygynous)
-statustxts <- sum(leader_text2$qualities_high.status)
-intellpolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities.knowlageable.intellect)
-statuspolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities_high.status)
-
-# text analysis
-# leader_text has 1000 rows
-# need raw texts for all 1212 rows
-
-textstats <- leader_text %>% 
-  dplyr::select(cs_textrec_ID, raw_text) %>% 
-  unnest_tokens(word, raw_text) %>% 
-  # dplyr::filter(is.na(as.numeric(word))) %>% # filters out numbers, some of which are page numbers
-  group_by(cs_textrec_ID) %>% 
-  summarise(count = n()) %>% 
-  summarise(min = min(count), max = max(count), mean = mean(count), median = median(count), sd = sd(count)) %>% 
-  round(1)
 
 # Group structure by subsistence --------------------------------------------------
 
@@ -970,3 +945,26 @@ hg_residential_pct <- signif(group_sub_tbl['hunter gatherers', 'residential subg
 hg_kin_pct <- signif(group_sub_tbl['hunter gatherers', 'kin group']/sum(leader_text$subsistence == 'hunter gatherers'), 3)*100
 hort_kin_pct <- signif(group_sub_tbl['horticulturalists', 'kin group']/sum(leader_text$subsistence == 'horticulturalists'), 3)*100
 
+final_record_count <- sum(rowSums(leader_text2[all_study_vars])>0)
+
+male_leader_pct <- signif(100*sum(leader_text2$demo_sex=='male', na.rm=T)/nrow(leader_text2), 3)
[…]

### 2019-12-01 leadership_across_cultures_contexts.Rmd
Working on MS, all sections.
Updated variables definition tables.

--- 2019-11-30 leadership_across_cultures_contexts.Rmd
@@
 Across human societies, the influence of leaders stands to substantially shape the nature of group structure and group outcomes.
-Simultaneously, social norms, pressures from competing leaders or rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda.
+Simultaneously, social norms, competing leaders, pressures from rival or cooperative groups, and followers preferences can greatly constrain a leader's agenda.
 The interactions between the strategies of individual leaders, pressures of followership, and group dynamics are likely to strongly contribute to organizational variation across human populations including social and economic trajectories [@wiessner_risk_1982] and cultural evolutionary processes [@henrich_big_2015].
@@
 This integration as been informative.
-Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019].
+Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
 Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and at all levels of social organization.
-Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated.
-Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is shaped by population-level variation and the costs and benefits leaders and followers incur across distinct leadership systems.
+Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have been relatively under-investigated.
+Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is shaped by population-level variation at different levels of social organizations and the costs and benefits leaders and followers incur across distinct leadership systems.
 
@@
 We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
-Through novel empirical and exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
+Through novel exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
 
@@
 Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families.
-Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @lewis_leaders_1974; @von_rueden_leadership_2015].
+Strong community leadership, however, is extremely common cross-culturally including among hunter-gatherers and relatively small-scale societies [@garfield_review_2019; @von_rueden_leadership_2015].
 
-Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are closely tied to group structure.
-Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders tends to emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
+Despite the universality of the phenomenon of leadership, the nature of leadership across populations and across social contexts is highly diverse and the emergence of leaders and the functions they serve are expected to be closely tied to group structure and group goals [@van_vugt_cognitive_2007].
+Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
 Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
-The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions conflict resolution [@glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015].
+The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
 Demands of mediation in within-group conflict resolution, although variable in scale, are probably a universal feature of human groups [@king_origins_2009; @van_vugt_leadership_2008].
@@
 Across relatively small-scale societies, at the community level, representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of leaders.
-Within communities, however, leaders may emerge whenever individuals for groups.
-The role of leaders in ritual contexts, such as in initiations, and in providing social services, including facilitating marriages and organizing social activities, are also expected to be common functions leaders serve.
+Within communities, however, leaders may emerge whenever individuals form groups.
+The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions leaders serve.
 
@@
 Pastoralists and coastal populations often face similar pressures concerning grazing lands and fishing access [e.g., @widmer_evolution_1988; @stevens_sherpa_1990].
-In the wake of economic intensification, the functions of leadership becomes increasingly associated with the managing economic systems and distribution of resources.
+In the wake of economic intensification, the functions of leadership becomes increasingly associated with managing economic systems and distribution of resources.
 Defensibility of critical resources creates increased opportunities for resource management and is often associated with increased territoriality and increased pressures of inter-group conflicts [@glowacki_evolutionary_2017].
@@
 In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
-Other public service and nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
-These positions typically are associated with institutionalized responsibilities and specific functions associated with their role.
+Other public service, nonprofit leaders, and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
+These positions typically are associated with institutionalized responsibilities and specific functions associated with their roles.
 The structure of such organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
@@
 Scholars in managerial studies have long discussed the behaviors of leaders and managers have developed multiple models on the functions of leadership.
-Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [@bass_handbook_1990].
-These perspectives focused on the rational, mechanistic aspects of group functioning and overlooked interpersonal dynamics.
+Classic management theorists [e.g., @davis_learning_1951] emphasized the planning, organizing, and controlling functions of leaders in formal organizational and business settings [see @bass_handbook_1990].
+These perspectives focused on the rational, mechanistic aspects of group functioning and often overlooked interpersonal dynamics.
 Later theorists incorporated cognitive and socioemotional components of leadership functions.
-For example, @selznick_leadership_1957 describes the functions of organization leaders is to define the goals of the organization, creating structures for accomplishing group goals, defending institutional integrity, and maintain internal consistency.
+For example, @selznick_leadership_1957 describes the functions of organization leaders is to define the goals of the organization, create structures for accomplishing group goals, defend institutional integrity, and maintain internal consistency among group members.
 
 @wofford_factor_1970 provides an early empirical view of managerial functions.
-A factor analysis of managerial behaviors from over 130 US employees' responses to 183 questionnaire items identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
+A factor analysis of managerial behaviors from over 130 US employees' survey responses identified five interpretable, orthogonal factors: (1) an *order and group achievement* factor, identifying participative, democratic, and organizational planning function; (2) a *personal enhancement oriented* factor, associated with authority and control, (3) a *personal interaction* factor, associated with interpersonal relationships and employee support, (4) a *security and maintenance* factor, associated with self-preservation, and (5) a *dynamic and achievement oriented* factor, associated with aggressive, self-assured actions.
 Wofford's analysis provided a foundation for future empirical studies of leadership and management functions.
-He contrasted factor one type management, described as the professional administrator who effectively collaborates and integrates team members, from factors two and five, described as autocratic leaders strongly motivated by personal power and willing to rely on aggressive tactics [@wofford_factor_1970].
+He contrasted factor one type management, described as the "professional administrator" who effectively collaborates and integrates team members, from factors two and five, described as "autocratic leaders" strongly motivated by personal power and a willingness to rely on aggressive tactics [@wofford_factor_1970].
 
-More recently, many managerial scholars have adopted a comparative approach to better understand cultural diversity in leadership systems, drawing on Culturally Endorsed Implicit Leadership Theory (CLT), which suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles and behaviors [@moore_literature_2018].
-The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leadership effectiveness and cultural variation across large organizations.
+More recently, many managerial scholars have adopted a comparative approach to better understand cultural diversity in leadership systems, drawing on Culturally Endorsed Implicit Leadership Theory (CLT), which suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018].
+The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations.
 The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications.
@@
 Across culturally and economically diverse populations there are some broad commonalities in the qualities of community leaders.
-When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have differential influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
-Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have increased influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
+When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
+Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
 Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are perceived to be associated with effective leadership.
@@
 
-Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [@johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
-Recent cross-cultural evidence and from a small-scale society implicates strong relationships between perceptions of intelligence and leadership among males [@garfield_evolutionary_2019; @garfield_investigating_2019], when considered in the context of broad interdisciplinary results suggests high intelligence and specialized knowledge are universal traits of political and community leaders.
-Recent evidence from within managerial studies, however, suggests a curvilinear effect in that exceptionally intelligence individuals at high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
+Across disciplines many theoretical models and empirical findings have implicated high intelligence as an important quality positively associated with effective leadership [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009].
+Recent cross-cultural evidence and from a small-scale society implicates strong relationships between perceptions of intelligence and leadership among males [@garfield_evolutionary_2019; @garfield_investigating_2019].
+These results suggests high intelligence and specialized knowledge are universal traits of political and community leaders.
 
@@
 More complex organizations require leaders equipped to process complex information, understand distinct departmentalized bodies, and resolve conflicting points of view, as well as be equipped to socially manage diverse individuals [@bass_handbook_1990; @schneider_organizations_2006; @antonakis_does_2009].
+However, recent evidence suggests a curvilinear effect in that exceptionally intelligence individuals at the high end of the distribution may not be the most effective leaders in contemporary business settings [@antonakis_can_2017].
 
 <!-- COULD GET INTO LEADER VS MANAGER DISTINCTION -->
+
+NEED CONCLUSION PARAGRAPH/STATEMENT.
 
@@
 <!-- BENEFITS: DECISIONS IN THEIR INTERST -->
-The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @pietraszewski_evolution_2019; @hagen_leadership_2019].
-
+The ability to shape group dynamics can provide enormous individual benefits to leaders, even if 'just for a while.' Fundamentally, leaders in any context have the opportunity to promote group-level policies which either align with their individual interests or which are not especially costly for themselves, their kin, or their close social partners [@garfield_evolutionary_2019; @kantner_identifying_2010; @hagen_leadership_2019].
 <!-- BENEFITS: HIGH SOCIAL STATUS -->
 Leadership positions, including formalized institutionalized roles and informal ephemeral roles, are also often if not universally associated with increased social status.
-The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional, but leadership roles through their association with increased social status are widely associated with numerous social, material, and reproductive benefits [@kantner_identifying_2010].
-
+The causal relationship between high social status and community or political leadership is complex and likely to be context-specific and bi-directional, but leadership roles through their association with increased social status are widely associated with numerous social, material, and reproductive benefits [@von_rueden_roots_2014; @von_rueden_multiple_2008; @grammer_symptoms_1996; @cheng_psychology_2014].
 <!-- BENEFITS: SOCIAL BENEFITS, SOCIAL NETWORK -->
@@
 High social status, material wealth, intelligence, and leadership status are also viewed as sexually attractive, particularly among women [@buss_strategies_2006; @stanik_who_2010].
-
 <!-- NEED BETTER CITATIONS ABOVE -->
-
 <!-- BENFITS: MATING -->
@@
 <!-- BENEFITS: 'MODEL' OF POTENTIAL BENEFIST -->
-@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) may gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
+@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
 
@@
 However, the unique costs associated with various leadership positions have received less attention, relative to the benefits of leadership, across disciplines.
-Group responsibilities necessarily impose opportunity costs on individual priorities and agendas, given leadership responsibilities require investments of time, energy, and resources that could be invested elsewhere.
-
+Group responsibilities necessarily impose opportunity costs on individual priorities, given leadership responsibilities require investments of time, energy, and resources that could be invested elsewhere.
 <!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - SSS -->
@@
 Devising and proposing compromises and optimal outcomes can be cognitively challenging and the solutions leaders propose are unlikely to be equally satisfactory for everyone involved.
-Therefore, leaders stand to become the targets of animosity and social ridicule.
-In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression [@boehm_purposive_2008].
-Physically formidable individuals with strong social networks are likely to be able o bear these costs more-so than their counterparts and unsurprisingly larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015].
+Therefore, leaders stand to become the targets of animosity.
+In the absence of state-based systems and institutionalized inequalities, all group members are on relative equal footing and equipped to challenge leaders directly or indirectly using a variety of leveling mechanisms including overt physical aggression or indirect aggression, such as gossip and social ridicule [@boehm_purposive_2008; @hess_gossip_2017].
+Physically formidable individuals with strong social networks are likely to be able to bear these costs more so than their counterparts.
+Larger body size and high social capital are qualities often associated with leaders in small-scale societies [@glowacki_leadership_2015].
 With greater population density and wealth inequality, these risks are expected to concomitantly increase in magnitude.
@@
 <!-- COSTS: INCREASED SOCIAL OR PHYSICAL RISK FROM MEDIATION - LSS -->
-Conceptualizing leadership as a possible solution to problems associated with collective goods [@olson_logic_1965], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leader incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
-In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the imposed costs.
-@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in urban neighborhoods and neighborhood organizations, which can be viewed as self-interested groups which face group-wide problems and necessarily share public benefits.
+Conceptualizing leadership as a possible solution to problems associated with collective goods [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation.
+In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs.
+@rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in urban neighborhoods and neighborhood organizations in WEIRD settings, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits.
 The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination).
@@
 
+NEED CONCLUSTION STATEMENT/PARAGRAPH.
+
 # Study aims
@@
 The universality of human leadership systems and the incredible diversity among the qualities of leaders across cultures and contexts and the functions leaders serve is difficult to capture.
-Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations.
-Here, we provide the first broad and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample.
-We also provide the first comparative empirical view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
+Scholars are then left to haphazardly self-select ethnographic cases or rely on the self-selected reviews of others for generalizations [e.g., @garfield_review_2019; @von_rueden_leadership_2015].
+Here, we seek to provide the first empirical and near comprehensive view of the qualities and functions of leaders across human societies from a representative cross-cultural sample.
+We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
 We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity.
 
-In sum, we aim to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent qualities of leaders and the functions they serve; 2) identify features of the qualities and functions of leaders in reference to theoretical models or novel interpretations; 3) identify the most cross-culturally frequent costs and benefits of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
-
+In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*; 2) identify *features* of the qualities and functions of leaders in reference to theoretical models or novel interpretations; 3) identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
 Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
@@
 We evaluate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
-Whereas @garfield_evolutionary_2019, documented supporting evidence in reference to influential theoretical models, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
+Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
 
 Here we draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database [@naroll_proposed_1967].
-The leadershipdata database was designed to capture a wide sample of the ethnographic record of leadership.
-For the current study, two authors coded each paragraph for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive.
+The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership.
+For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive.
 
@@
 DISCUSS IRR.
-ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named leader_text2).
+ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
 
@@
 
-For additional details on ethnographic sampling and creating of the leadershipdata package see CITE.
+For additional details on ethnographic sampling and creation of the *leadershipdata* package see @garfield_evolutionary_2019.
 
@@
 We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities and functions of leadership that are most and least frequently noted in the ethnographic record.
 
@@
 
-Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
+Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups.
 We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture (for details, see Table S1).
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture.
 The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
@@
 We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
-We renamed the eHRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
+We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.
 
@@
 As Figure \@ref(fig:groupsex) reveals, the distribution of text records by group type varies across subsistence type.
-Across all subsistence groups, the most frequent group type is political groups at the community level, with the exception of agriculturalitsts in which supracommunity political groups are the most frequent group type.
+Across all subsistence types, the most frequent group type is political groups at the community level, with the exception of agriculturalitsts in which supracommunity political groups are the most frequent group type.
 Leadership beyond the community level is relatively rare outside of agricultural populations and is especially rare among hunter-gatherers.
-Leadership within residential groups is most frequently associated hunter gatherers (`r hg_residential_pct`% of the total sample of hunter gatherer text records) and leadership within kin groups is most frequently associated with hunter gatherers (`r hg_kin_pct`% of the total hunter gatherers sample) and horticulturalists (`r hort_kin_pct`% of the total horticulturalists sample).
+Leadership within residential groups is most frequently associated with hunter-gatherers (`r hg_residential_pct`% of the total sample of hunter-gatherer text records) and leadership within kin groups is most frequently associated with hunter-gatherers (`r hg_kin_pct`% of the total hunter-gatherers sample) and horticulturalists (`r hort_kin_pct`% of the total horticulturalists sample).
 
@@
 
-The most common *functions* leaders serve, which are documented in over 70% of cultures, are resolving conflicts, providing various social functions, providing counsel or direction in follower's affairs, and organizing various cooperative activities.
+The most common *functions* leaders serve, which are documented in over 70% of cultures, are resolving conflicts, providing various social functions, organizing various cooperative activities, and providing counsel or direction in follower's affairs.
 Leaders also commonly serve as a representatives in inter-group interactions, manage the distribution of resources, and provide context-specific leadership such as military command and in providing ritual services.
@@
 Also, in about 60% of the cultures sampled here leaders were wealthy, nepotistic, possessed supernatural qualities, were generous, or maintained strong authority over followers.
-Notable infrequent qualities of leaders in these data include charisma and fairness.
+Notable infrequent qualities of leaders in these data include physical formidability and high social contacts (documented in about 35% of cultures) and charisma and fairness (documented in about 20% of cultures).
 
@@
 Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).
-
 The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as an organizational functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
@@
 Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)).
-Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component.
+Measures of physical formidability are also moderately associated with the "dominance" side of this component.
 The second leader qualities component we identify as an *In-group favoritism vs.
@@
 
+The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above.
 In these data leaders are most commonly receiving benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures).
@@
 Social Status* leader qualities components and the *Social functions* and *Mediation vs.
-Organization* leader functions components, we fit four hierarchical linear models at the text-level, using nested random effects of document within culture with each of these components as outcome measures.
+Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures.
 For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:groupsex)).
@@
 
-Across a broad range of cultures and social contexts leaders are nearly universally high status males described as knowledgable or intelligent who are experienced and have a demonstrated history of previous accomplishments or special skills.
+Across a broad range of cultures and social contexts leaders are nearly universally high status males described as knowledgeable or intelligent who are experienced and have a demonstrated history of previous accomplishments or special skills.
 They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, and providing counsel or input in followers' affairs.
@@
 
-At all subsistence levels leaders emerge across a broad range of group types, including at lower-order social units such as within residential groups and kin groups; within specialized groups such as religous, economic, or military groups; and in political contexts at the community level and beyond.
-Political leadership at the commmunity level is widespread across all subsistence categories.
+At all subsistence levels leaders emerge across a broad range of group types, including at lower-order social units such as within residential groups and kin groups; within specialized groups such as religious, economic, or military groups; and in political contexts at the community level and beyond.
+Political leadership at the community level is widespread across all subsistence categories.
 Economic, military, and religious group leadership, although relatively less frequent, are fairly consistently represented across subsistence levels.
@@
 
-## The cultural and context dependancy of human leadership
+## The cultural and context dependency of human leadership
 
-The distribiution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistance strategies, but at variable frequencies.
-Hunter-gatherers and pastoralists, populations more likely to expeirence high residential mobility, exhibited greater relative frequencies of leaderhsip wtihin residential groups and kin groups.
+The distribution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistence strategies, but at variable frequencies.
+Hunter-gatherers and pastoralists, populations more likely to experience high residential mobility, exhibited greater relative frequencies of leadership within residential groups and kin groups.
 Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
-Pastoral and horticultural populations are often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaboratored political leadership systems.
+Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems.
 
 Figure \@ref(fig:groupsex) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups.
-One possible interpretation is that as political structures become more instituionalized and economic systems transcend the household, leadership within households and residental groups errodes.
-It is interesting, however, that even among the agricultural populations that are charachterized by leadership systems beyong the community level, leadership within residential and family group persists.
+One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups erodes.
+It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
 
@@
 ## The features of human leadership
+
+Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001].
+Exploratory analyses reported here suggest multiple leadership phenotypes and associated features.
+
+There is likely a strong connection between shamanism and ritual-based influence and the dominance style of leadership.
+Super natural abilities, being feared, aggressiveness, and strong authority formed a strong cluster of leadership qualities (Figure \@ref(fig:cluster)A, *Dominant/shaman* cluster $AU = 96$).
+Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
+@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n=43$) shamans were associated with providing decision-making services and associated with military power.
+Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
+Widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support dominance-based leadership.
+
+We have also identified a broad suite of positive personal characteristics, including many of the most cross-culturally frequent qualities of leadership, which generally correspond to prestige and charismatic leadership [@van_vugt_dual_2019; @cheng_pride_2010; @grabo_charismatic_2016; @henrich_evolution_2001; @price_evolution_2014; @von_rueden_roots_2014].
+Being knowledgeable or intelligent and highly experienced or accomplished cluster closely with high social status (the three most cross-culturally frequent qualities of leaders).
+Other personal characteristics, such as having a favorable personality, humility, oratory skill, charisma, and interpersonal skills are also associated with this feature of leadership (Figure \@ref(fig:cluster)A, *Positive personal characteristics* cluster $AU = 100$.).
+These features  suggest a strong link between prestige-based leadership, associated with deference towards individuals with special kills, high intelligence, and accomplishments [@cheng_dominance_2019; @van_vugt_dual_2019; @henrich_evolution_2001], and charismatic leadership, associated with oratory skills, high charisma, confidence, and interpersonal skills [@den_hartog_culture_1999; @grabo_charismatic_2016; @howell_role_2005].
+
+The most common benegits associated with leadership are material resources, increased social status, and social services (See Figure \@ref(fig:varsupportcb)).
+Measures of material and social wealth coupled with reproductive success represent a feature of leadership qualities descriptive of benefits of social influence and high social status (Figure \@ref(fig:cluster)A, *Material, social, & reproductive success* cluster $AU = 96$.).
 
@@
 
+These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and across all levels of social organization.
+
 # References

### 2019-12-01 leadership_across_cultures_contexts.Rmd
Fixed bib entry.
Added paragraphy in discussion.

@@
 
+<!-- Qualities features -->
+
 There is likely a strong connection between shamanism and ritual-based influence and the dominance style of leadership.
@@
 
-The most common benegits associated with leadership are material resources, increased social status, and social services (See Figure \@ref(fig:varsupportcb)).
+The most common benefits associated with leadership are material resources, increased social status, and social services (See Figure \@ref(fig:varsupportcb)).
 Measures of material and social wealth coupled with reproductive success represent a feature of leadership qualities descriptive of benefits of social influence and high social status (Figure \@ref(fig:cluster)A, *Material, social, & reproductive success* cluster $AU = 96$.).
+
+<!-- Functions features -->
+
+Results suggest X features of the functions of leaders.
+Leaders may serve various prosocial functions, such as providing subsistence and distributing resources, they may be tasked with providing hospitality services to guests or followers, they function to protect the group via a variety of behaviors, and very commonly are leading in various social functions tied to cultural values and social practices (Figure \@ref(fig:cluster)B, *Prosocial* cluster $AU = 95$).
 

### 2019-12-01 leadership_across_cultures_contexts.Rmd
Corrected spelling

--- 2019-12-01 leadership_across_cultures_contexts.Rmd
@@
 
-Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$.
+Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$.
 Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward agglomeration algorithm.
@@
 Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (i.e., *Authority*, *Aggressiveness*, *Feared*) and shamanism (i.e., *Supernatural*, *Artistic performance*).
-The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygnous*).
+The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygynous*).
 Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).
@@
 Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*.
-High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*.
+High values are strongly associated with *Polygynous* and *Authority* and moderately associated with *High quality spouse*.
 Therefore, the "dominance" side of the *Prestige vs.

### 2019-12-01 initialcompute.R

@@ -306,7 +306,7 @@ var_names <- c(
   "qualities.innovative"            = "Innovative",
   "qualities.knowlageable.intellect" = "Knowledgeable/intelligent",
   "qualities.oratory.skill"         = "Oratory skill",
-  "qualities.polygynous"            = "Polygnous",
+  "qualities.polygynous"            = "Polygynous",
   "qualities.social.contacts"       = "Social contacts",
   "qualities.supernatural"    = "Supernatural",
   "qualities_exp.accomplished"       = "Experienced/accomplished",

### 2019-12-01 leadership_across_cultures_contexts.Rmd

--- 2019-12-01 leadership_across_cultures_contexts.Rmd
@@
 
-In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*; 2) identify *features* of the qualities and functions of leaders in reference to theoretical models or novel interpretations; 3) identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
+In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*.
+2) Identify *features* of the qualities and functions of leaders (e.g., do particular variables tended to covary or cluster together?) in reference to theoretical models or novel interpretations.
+Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
+3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
 Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
@@
 We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level.
-Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities and functions of leadership that are most and least frequently noted in the ethnographic record.
+Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
+We computed confidence intervals for the proportion of text records supporting each variable using generalized linear mixed effects models and random intercepts based on document authors nested within cultures.
+We computed confidence intervals for the proportion of cultures with evidence for each variable using a cluster bootstrap.
 
-Our second goal was to discover "features" in our data, e.g., to determine if particular variables tended to covary or cluster together.
-Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
-Or they might suggest new and theoretically important dimensions of leadership.
+We then explored "features" in our data.
+There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
+We used hierarchical cluster analysis on our quality and function variables.
+We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
+We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
+We used 10,000 bootstrapped samples.
 
-There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
-To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package.
+<!-- To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package.
 This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].
-
 Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables.
 We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors.
+-->
 
@@
 
-We also performed hierarchical cluster analysis on our quality and function variables.
-We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
-We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
-We used 10,000 bootstrapped samples.
+Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above.
 
-Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record.
+<!-- Our fourth goal was to explore associations between the most common leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
+We used linear mixed effects models from the lme4 package [@bates2015], with common qualities and functions as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
+-->
 
@@
 Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (i.e., *Authority*, *Aggressiveness*, *Feared*) and shamanism (i.e., *Supernatural*, *Artistic performance*).
-The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygnous*).
+The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygynous*).
 Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).
+
 The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as an organizational functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
+
+<!-- ## Principle components analyses -->
+
+<!-- A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
+We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)).
+We identify the first component (PC1) as a *Prestige vs.
+Dominance* measure.
+Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*.
+High values are strongly associated with *Polygynous* and *Authority* and moderately associated with *High quality spouse*.
+Therefore, the "dominance" side of the *Prestige vs.
+Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)).
+Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component.
+The second leader qualities component we identify as an *In-group favoritism vs.
+Social Status* measure.
+Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*.
+High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*.
+Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.
+-->
 

### 2019-12-01 analyses2.R
Changed visreg plots and added tmp model template for exploration.

@@ -509,10 +509,10 @@ summary(m_fPC2)
 Anova(m_fPC2)
 AIC(m_fPC2)
 
-m_fPC2_sub_plot <- visreg(m_fPC2, "subsistence", type = "contrast", gg = T) +
+m_fPC2_sub_plot <- visreg(m_fPC2, "subsistence", type = "conditional", gg = T) +
   labs(y = "Functions PC 2: \nOrganization vs. Mediation\n",
        x = "\nSubsistence type")
-m_fPC2_group_plot <- visreg(m_fPC2, "group.structure2", type = "contrast", gg = T) +
+m_fPC2_group_plot <- visreg(m_fPC2, "group.structure2", type = "conditional", gg = T) +
   labs(y = "Functions PC 2: \nOrganization vs. Mediation\n",
        x = "\nGroup type")
 
@@ -568,6 +568,26 @@ m_qPC2_group_plot <- visreg(m_qPC2, "group.structure2", type = "contrast", gg = 
 
 m_qPC2_sub_plot + m_qPC2_group_plot
 
+
+
+# TMP models
+
+tmp <- glmer(
+  function_resolve.conflcit ~
+    #qualities_component1 +
+    subsistence +
+    #c_cultural_complexity +
+    #pop_density +
+    #com_size +
+    group.structure2 +
+    # warfare_freq +
+    (1|d_culture/doc_ID),
+  family = binomial(link = "logit"),
+  data=leader_text2
+)
+summary(tmp)
+Anova(tmp)
+visreg(tmp)
 
 # #visreg(m_fPC1, xvar = 'warfare_freq', by = 'group.structure2')
 # 
@@ -909,4 +929,50 @@ hg_residential_pct <- signif(group_sub_tbl['hunter gatherers', 'residential subg
 
 hg_kin_pct <- signif(group_sub_tbl['hunter gatherers', 'kin group']/sum(leader_text$subsistence == 'hunter gatherers'), 3)*100
 hort_kin_pct <- signif(group_sub_tbl['horticulturalists', 'kin group']/sum(leader_text$subsistence == 'horticulturalists'), 3)*100
+
+
+
+# Violin plots of PCs by group vars ---------------------------------------
+
+
+ggplot(pca_data_qualities, aes(group.structure2, qPC1))+
+  geom_violin()+
+  geom_jitter(height = 0, width = 0.1) +
+  geom_boxplot(width=.15)
+
[…]

### 2019-12-01 initialcompute.R
Added cluster analysis of rows using Euclidean distance.
Changed name of social functions -> misc. social functions

@@ -22,7 +22,6 @@ library(visreg)
 library(effects)
 library(lme4)
 library(patchwork)
-library(pvclust)
 library(magrittr)
 library(modelr)
 library(ggridges)
@@ -280,7 +279,7 @@ var_names <- c(
   "functions_protection"          = "Protection",
   "functions_provide.subsistence" = "Provide subsistence",
   "functions_ritual"              = "Ritual functions",
-  "functions_social.functions" = "Social functions",
+  "functions_social.functions" = "Misc. social functions",
   "qualities_artistic.performance"     = "Artistic performance",
   "qualities_generous"                 = "Generosity",
   "qualities.age"                      = "Age",
@@ -503,6 +502,22 @@ m_pvclust_qual <- pvclust(
   method.hclust = 'ward', 
   method.dist = 'correlation', 
   nboot = 10000,
+  parallel = T
+)
+
+# m_pvclust_qual_texts <- pvclust(
+#   t(pca_data_qualities2), 
+#   method.hclust = 'ward', 
+#   method.dist = 'correlation', 
+#   nboot = 1000,
+#   parallel = T
+# )
+
+m_pvclust_qual_texts_euc <- pvclust(
+  t(pca_data_qualities2),
+  method.hclust = 'ward',
+  method.dist = 'euclidean',
+  nboot = 1000,
   parallel = T
 )
 

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Commeted out readacted sections.

--- 2019-12-01 leadership_across_cultures_contexts.Rmd
@@
 
-## Principle components analyses
+<!-- ## Principle components analyses -->
 
-A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
+<!-- A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
 We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)).
@@
 Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.
+-->
 
-```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components.
+<!-- ```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components.
 PC 1 is identified as a *Prestige vs.
 Dominance* component and PC 2 is identified as an *In-group favoritism vs.
-Social Status* component."}
-logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2)
-```
+Social Status* component."} -->
+<!-- logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2) -->
+<!-- ``` -->
 
-Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
+<!-- Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
 The first leader functions component distinguishes between social functions (a very broad and generalized measure) from other more specific leader functions, which we identify as a *Social functions* component.
@@
 See Figure \@ref(fig:functionsPCplot).
+-->
 
-```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components.
+<!-- ```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components.
 PC 1 is identified as a *Social functions* component and PC 2 is identified as an *Organization vs.
-Mediation* component."}
-logisticPCA_loadings_plot(logpca_model_functions, data = pca_data_functions2)
-```
+Mediation* component."} -->
+<!-- logisticPCA_loadings_plot(logpca_model_functions, data = pca_data_functions2) -->
+<!-- ``` -->
 
@@
 
-## Exploring population-level variation
+<!-- ## Exploring population-level variation -->
 
-To explore potential population variation in the *Prestige vs.
+<!-- To explore potential population variation in the *Prestige vs.
 Dominance* and *In-group favoritism vs.
@@
 For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:groupsex)).
+-->
 
@@
 
-Functions PC 1 model
+<!-- Functions PC 1 model -->
 
-```{r funcPCmodel1, fig.height=8, fig.width=20}
-m_fPC1_sub_plot + m_fPC1_group_plot
-```
+<!-- ```{r funcPCmodel1, fig.height=8, fig.width=20} -->
+<!-- m_fPC1_sub_plot + m_fPC1_group_plot -->
+<!-- ``` -->
 
-Functions PC 2 model
+<!-- Functions PC 2 model -->
 
-```{r funcPCmodel2, fig.height=10, fig.width=20}
-m_fPC2_sub_plot + m_fPC2_group_plot
-```
+<!-- ```{r funcPCmodel2, fig.height=10, fig.width=20} -->
+<!-- m_fPC2_sub_plot + m_fPC2_group_plot -->
+<!-- ``` -->
 
-Qualities PC 1 model
+<!-- Qualities PC 1 model -->
 
-```{r qualPCmodel1, fig.height=10, fig.width=20}
-m_qPC1_sub_plot + m_qPC1_group_plot
-```
+<!-- ```{r qualPCmodel1, fig.height=10, fig.width=20} -->
+<!-- m_qPC1_sub_plot + m_qPC1_group_plot -->
+<!-- ``` -->
 
-Qualities PC 2 model
+<!-- Qualities PC 2 model  -->
 
-```{r qualPCmodel2, fig.height=10, fig.width=20}
-m_qPC2_sub_plot + m_qPC2_group_plot
-```
+<!-- ```{r qualPCmodel2, fig.height=10, fig.width=20} -->
+<!-- m_qPC2_sub_plot + m_qPC2_group_plot -->
+<!-- ``` -->
 

### 2019-12-02 analyses2.R
Some heatmap stuff, and a couple of models.

@@ -264,17 +264,17 @@ plot(logpca_model_functions, type = "scores") +
 
 # PCA Qualities & Functions -----------------------------------------------
 
-# logpca_model_qfk2 = logisticPCA(pca_data_FQ[c(function_vars, quality_vars)], k = 2, m = which.min(logpca_cv_qfk2), main_effects = T)
-# plot(logpca_model_qfk2, type = 'scores')
-# logisticPCA_loadings_plot(logpca_model_qfk2, data = pca_data_FQ[,c(function_vars, quality_vars)])
+logpca_model_qfk2 = logisticPCA(pca_data_FQ[c(function_vars, quality_vars)], k = 2, m = which.min(logpca_cv_qfk2), main_effects = T)
+plot(logpca_model_qfk2, type = 'scores')
+logisticPCA_loadings_plot(logpca_model_qfk2, data = pca_data_FQ[,c(function_vars, quality_vars)])
 # 
 # pca_data_FQ$fqPC1k2 <- logpca_model_qfk2$PCs[,1]
 # pca_data_FQ$fqPC2k2 <- logpca_model_qfk2$PCs[,2]
 # leader_text2 <- left_join(leader_text2, pca_data_FQ[c('cs_textrec_ID', 'fqPC1k2', 'fqPC2k2')])
 # 
-# logpca_model_qf = logisticPCA(pca_data_FQ[c(function_vars, quality_vars)], k = kqf, m = which.min(logpca_cv_qf), main_effects = T)
+logpca_model_qf = logisticPCA(pca_data_FQ[c(function_vars, quality_vars)], k = kqf, m = mqf, main_effects = T)
 # plot(logpca_model_qf, type = "scores")
-# logisticPCA_loadings_plot(logpca_model_qf, data = pca_data_FQ[c(function_vars, quality_vars)])
+logisticPCA_loadings_plot(logpca_model_qf, data = pca_data_FQ[c(function_vars, quality_vars)])
 # 
 # pca_data_FQ$fqPC1 <- logpca_model_qf$PCs[,1]
 # pca_data_FQ$fqPC2 <- logpca_model_qf$PCs[,2]
@@ -375,7 +375,31 @@ plot(logpca_model_functions, type = "scores") +
 # 
 # 
 # # Exploratory models ------------------------------------------------------
-# 
+
+mqPC1 <- lmer(
+  qPC1 ~ 
+    subsistence + 
+    c_cultural_complexity +
[…]

### 2019-12-02 initialcompute.R

@@ -513,13 +513,13 @@ m_pvclust_qual <- pvclust(
 #   parallel = T
 # )
 
-m_pvclust_qual_texts_euc <- pvclust(
-  t(pca_data_qualities2),
-  method.hclust = 'ward',
-  method.dist = 'euclidean',
-  nboot = 1000,
-  parallel = T
-)
+# m_pvclust_qual_texts_euc <- pvclust(
+#   t(pca_data_qualities2),
+#   method.hclust = 'ward',
+#   method.dist = 'euclidean',
+#   nboot = 1000,
+#   parallel = T
+# )
 
 m_pvclust_qual_jaccard <- pvclust(
   pca_data_qualities2, 

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Working on discussion.

--- 2019-12-02 leadership_across_cultures_contexts.Rmd
@@
 
-In sum, we seek to empirically and systematically accomplish the following four goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*; 2) identify *features* of the qualities and functions of leaders in reference to theoretical models or novel interpretations; 3) identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
+In sum, we seek to empirically and systematically accomplish the following three goals: 1) identify the most cross-culturally frequent *qualities of leaders* and the *functions they serve*; 2) identify *features* of the qualities and functions of leaders in reference to theoretical models or novel interpretations; 3) identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers.
 Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
+
+<!-- and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.  -->
 
@@
 
-Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups.
+<!-- Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups.
 We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
+-->
 
@@
 
-<!-- ```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components.
+<!-- ```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader quality variables on the first two components.
 PC 1 is identified as a *Prestige vs.
@@
 These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and across all levels of social organization.
+Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers.
+Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups.
+Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history.
+Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].
+
+Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014].
+Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
+However, the ubiquity of conflict resolution as a function of leaders across social contexts calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits.
+[@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018].
 

### 2019-12-02 leadership_across_cultures_contexts.Rmd

--- 2019-12-02 leadership_across_cultures_contexts.Rmd
@@
 We also provide the first comparative view describing the costs and benefits of leadership for both leaders and followers in an evolutionary biological framework.
-We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity.
+We also aim to explore how the qualities and functions of leaders vary across social contexts and levels of cultural complexity
 
@@
 Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
-3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers; and 4) explore relationships between measures of leader qualities and functions with group structure and population-level measures.
+3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers.
 Such exploratory results stand to facilitate more fine-grained approaches within leadership studies and greater theoretical and empirical work beyond community leadership, while also revealing the importance of a broad, cross-cultural perspective in understanding the psychology and cultural patterning of perhaps the most foundational component of human social behavior and organization.
@@
 
-<!-- Our fourth goal was to explore associations between the most common leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size.
-We used linear mixed effects models from the lme4 package [@bates2015], with common qualities and functions as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
+<!-- Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups.
+We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
 -->
-
-Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups.
-We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
 

### 2019-12-02 analyses2.R
Commenting out unneeded analyses in analyses script.
Additions in bib file.

@@ -192,90 +192,90 @@ dev.off()
 # PCA Qualities ---------------------------------------------------------
 
 # For two PCs only; use optimal m
-m_lpca_qualk2 <- logisticPCA(pca_data_qualities2, k = 2, m = which.min(qual_cvlpcak2), main_effects = T)
-plot(m_lpca_qualk2, type = 'scores')
-logisticPCA_loadings_plot(m_lpca_qualk2, data = pca_data_qualities2)
-
-# Add PC scores to df
-pca_data_qualities$qPC1k2 <- m_lpca_qualk2$PCs[,1]
-pca_data_qualities$qPC2k2 <- m_lpca_qualk2$PCs[,2]
-leader_text2 <- left_join(leader_text2, pca_data_qualities[c('cs_textrec_ID', 'qPC1k2', 'qPC2k2')])
-
-# For optimal k, m determined by cv in initialcompute.R
-kq <- 8
-mq <- 12
-
-logpca_model_qualities = logisticPCA(pca_data_qualities2, k = kq, m = mq, main_effects = T)
-plot(logpca_model_qualities, type = "scores")
-logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2)
-
-pca_data_qualities$qPC1 <- logpca_model_qualities$PCs[,1]
-pca_data_qualities$qPC2 <- logpca_model_qualities$PCs[,2]
-leader_text2 <- left_join(leader_text2, pca_data_qualities[c('cs_textrec_ID', 'qPC1', 'qPC2')])
-
-# Indicate group structure of log PCA model
-plot(logpca_model_qualities, type = "scores") + 
-  geom_point(aes(colour=pca_data_qualities$group.structure2)) + 
-  stat_ellipse(aes(colour=pca_data_qualities$group.structure2)) + 
-  ggtitle("Logistic PCA") +
-  scale_colour_brewer(palette = "Set1")
-
-# Indicate subsistence stype of log PCA model
-plot(logpca_model_qualities, type = "scores") + 
-  geom_point(aes(colour=pca_data_qualities$subsistence)) + 
-  ggtitle("Logistic PCA") +
-  scale_colour_brewer(palette = "Set1")
-
-# PCA Functions -----------------------------------------------------------
-
-# Logistic PCA on leader functions
-
[…]

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Modified cluster dendrograms. Minor edits to texts.

@@
 Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.
-We computed confidence intervals for the proportion of text records supporting each variable using generalized linear mixed effects models and random intercepts based on document authors nested within cultures.
+We computed confidence intervals for the proportion of text records supporting each variable using generalized linear mixed effects models and random intercepts based on document authors nested within cultures (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or theoretical framework).
 We computed confidence intervals for the proportion of cultures with evidence for each variable using a cluster bootstrap.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars) ), we computed the percentage of text records that provided support, estimating standard errors using a binomial generalized linear mixed effects model with a random intercept for author nested within culture (which partially accounts for potential variation due to differences in ethnographers such as gender, nationality, or philosophy).
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars) ), we computed the percentage of text records that provided support, and we computed the percentage of cultures with at least one text record providing support for that variable.
 See Figure \@ref(fig:varsupport).

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Removed Logistic PCA text.

--- 2019-12-02 leadership_across_cultures_contexts.Rmd
@@
 
-<!-- ## Principle components analyses -->
-
-<!-- A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
-We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)).
-We identify the first component (PC1) as a *Prestige vs.
-Dominance* measure.
-Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*.
-High values are strongly associated with *Polygynous* and *Authority* and moderately associated with *High quality spouse*.
-Therefore, the "dominance" side of the *Prestige vs.
-Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)).
-Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component.
-The second leader qualities component we identify as an *In-group favoritism vs.
-Social Status* measure.
-Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*.
-High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*.
-Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.
--->
-
-<!-- ## Principle components analyses -->
-
-<!-- A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance.
-We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)).
-We identify the first component (PC1) as a *Prestige vs.
-Dominance* measure.
-Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*.
-High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*.
-Therefore, the "dominance" side of the *Prestige vs.
-Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)).
-Measures of physical formidability are also moderately associated with the "dominance" side of this component.
-The second leader qualities component we identify as an *In-group favoritism vs.
-Social Status* measure.
-Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*.
-High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*.
-Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.
--->
-
-<!-- ```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader quality variables on the first two components.
-PC 1 is identified as a *Prestige vs.
-Dominance* component and PC 2 is identified as an *In-group favoritism vs.
-Social Status* component."} -->
-<!-- logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2) -->
-<!-- ``` -->
-
-<!-- Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance.
-The first leader functions component distinguishes between social functions (a very broad and generalized measure) from other more specific leader functions, which we identify as a *Social functions* component.
-Low values on Leader function PC 1 are strongly associated with *Social functions*.
-The Leader functions PC 2 distinguished between organizational and mediation-based functions, which we identify as an *Organization vs.
-Mediation* component.
-Low values on Leader functions PC 2 are strongly associated with the variables *Organize cooperation* and *Political appointments*, whereas high values are strongly associated with *Resolve conflict* and moderately associated with *Provide counsel/direction*.
-See Figure \@ref(fig:functionsPCplot).
--->
-
-<!-- ```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components.
-PC 1 is identified as a *Social functions* component and PC 2 is identified as an *Organization vs.
-Mediation* component."} -->
-<!-- logisticPCA_loadings_plot(logpca_model_functions, data = pca_data_functions2) -->
-<!-- ``` -->
-
 ## Benefits and costs of leadership and followership

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Edited cluster analysis section

@@
 We used hierarchical cluster analysis on our quality and function variables.
-We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
+We computed distance between our binary vectors using both $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
 We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
@@
 
-Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$.
-Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward agglomeration algorithm.
-
 Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
@@
 
-The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters.
-Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text.
-Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (i.e., *Authority*, *Aggressiveness*, *Feared*) and shamanism (i.e., *Supernatural*, *Artistic performance*).
-The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygynous*).
-Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).
+The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters, one that comprised mostly variables related to material, social, and reproductive success, and one that comprised mostly variables indicating the relationship between leaders and their groups, which we term *social orientation*.
+Within the latter top-level cluster we identify two interpretable though weakly supported sub-clusters which are annotated with text.
+Within the first top-level cluster, we identify a subcluster mainly comprising variables related to the ability to provide social benefits, and a subcluster mainly comprising variables related to the ability to impose social costs.
 
-The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as an organizational functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
+Within the management cluster, we identified a mediation subcluster (e.g., *Resolve conflict*, *Provide counsel*), and a an organization subcluster (e.g., *Organize cooperation*, *Military command*).
+
+In naming the clusters and subclusters, we gave extra weight to variables that were strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), and therefore not all variables within these clusters were clearly related to the themes we identified.
 

### 2019-12-02 analyses2.R
Editing discussions sections on group types, qualities, and functions

@@ -1026,6 +1026,12 @@ textstats <- leader_text %>%
   summarise(min = min(count), max = max(count), mean = mean(count), median = median(count), sd = sd(count)) %>% 
   round(1)
 
+movement_pct <- signif(d_melt$value[d_melt$Variable == "Movement/migration" & d_melt$Type == "Cultures"], 3)*100
+moral_pct <- signif(d_melt$value[d_melt$Variable == "Moral authority" & d_melt$Type == "Cultures"], 3)*100
+fairness_pct <- signif(d_melt$value[d_melt$Variable == "Fairness" & d_melt$Type == "Cultures"], 3)*100
+feared_pct <- signif(d_melt$value[d_melt$Variable == "Feared" & d_melt$Type == "Cultures"], 3)*100
+attractive_pct <- signif(d_melt$value[d_melt$Variable == "Attractive" & d_melt$Type == "Cultures"], 3)*100
+
 # # Violin plots of PCs by group vars ---------------------------------------
 # 
 # 

### 2019-12-02 leadership_across_cultures_contexts.Rmd

@@
 The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups.
-Figure \@ref(fig:groupsex) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence.
+Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence.
 We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category.
@@
 
-```{r groupsex, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group type.
+```{r grouptype, fig.width=20, fig.height=10, fig.cap="The distribution of text records by subsistence and group type.
 Areas of the bars are proportional to the number of text records in that category."}
@@
 
-As Figure \@ref(fig:groupsex) reveals, the distribution of text records by group type varies across subsistence type.
+As Figure \@ref(fig:grouptype) reveals, the distribution of text records by group type varies across subsistence type.
 Across all subsistence types, the most frequent group type is political groups at the community level, with the exception of agriculturalitsts in which supracommunity political groups are the most frequent group type.
@@
 For both leaders and followers, individuals stand to lose what they are able to gain.
-The clear differences lie in leaders' opportunities for increased social status and matting opportunities.
+The clear differences lie in leaders' opportunities for increased social status and mating opportunities.
 See Figure \@ref(fig:varsupportcb).
@@
 Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures.
-For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:groupsex)).
+For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:grouptype)).
 -->
@@
 
-## The universality of human leadership
-
-Across a broad range of cultures and social contexts leaders are nearly universally high status males described as knowledgeable or intelligent who are experienced and have a demonstrated history of previous accomplishments or special skills.
-They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, and providing counsel or input in followers' affairs.
-See Figure \@ref(fig:varsupport).
-
-At all subsistence levels leaders emerge across a broad range of group types, including at lower-order social units such as within residential groups and kin groups; within specialized groups such as religious, economic, or military groups; and in political contexts at the community level and beyond.
-Political leadership at the community level is widespread across all subsistence categories.
-Economic, military, and religious group leadership, although relatively less frequent, are fairly consistently represented across subsistence levels.
-See Figure \@ref(fig:groupsex).
-
-## The cultural and context dependency of human leadership
+## The group context of human leadership
 
 The distribution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistence strategies, but at variable frequencies.
+See Figure \@ref(fig:grouptype).
+
+### Universal dimensions
+
+In groups of all subsistence types (e.g., hunter-gatherers, pastoralists) leaders emerge across a similar range of group types, including at lower-level social units such as within residential groups and kin groups; within specialized groups such as religious, economic, or military groups; and in political contexts at the community and supracommunity levels.
+A large percentage of text records in all subsistence groups document political leadership at the community level.
+Leadership at the supracommunity level is also common across all subsistence groups save hunter-gatherers, which are largely autonomous.
+Economic, military, and religious group leadership, although relatively less frequent, are fairly consistently represented across subsistence levels.
+
+### Variable dimensions
+
+Leadership in residential subgroups (e.g., hunting party, work group) was relatively common among hunter-gatherers but relatively rare among agriculturists.
+Leadership at the surpacommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
+
 Hunter-gatherers and pastoralists, populations more likely to experience high residential mobility, exhibited greater relative frequencies of leadership within residential groups and kin groups.
@@
 
-Figure \@ref(fig:groupsex) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups.
+Figure \@ref(fig:grouptype) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups.
 One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups erodes.
@@
 
-<!-- MAY WANT TO RETHINK INTERPRETATIONS, LOTS OF STORIES HERE -->
+## Leader functions and qualities
 
-## The features of human leadership
+Our analysis identified several leader functions and qualities that were near cultural universals, and others that were relatively rare. See Figure \@ref(fig:varsupport).
+
+### Universal dimensions
+
+Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent and who are experienced and have a demonstrated history of previous accomplishments or special skills.
+They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
+See Figure \@ref(fig:varsupport).
+
+### Variable dimensions
+
+Several functions and qualities that are theoretically important were relatively rare in the ethnographic record.
+In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures.
+Several theories of leadership also emphasize the moral authority of leaders [e.g., @xx; @yy], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
+The "Dominance" theory of leadership [@xx; @yy] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
+Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders, but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
+
+## Clusters of leader functions and qualities
 
@@
 
+## Synthesis
+
 ## Limitations

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Adding citations in discussion sections.

@@
 In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures.
-Several theories of leadership also emphasize the moral authority of leaders [e.g., @xx; @yy], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
-The "Dominance" theory of leadership [@xx; @yy] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
-Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders, but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
+Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
+The "Dominance" theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019:1; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
+Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @leadership], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
 

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Fixed bib errors.

@@
 Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
-The "Dominance" theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019:1; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
-Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @leadership], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
+The "Dominance" theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
+Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
 

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Changed png to svg.

@@
 # pvrect(m_pvclust_qual, alpha = 0.9)
-knitr::include_graphics("Figures/m_pvclust.both.png")
+knitr::include_graphics("Figures/m_pvclust.both.svg")
 ```

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Final edits

@@
 We used hierarchical cluster analysis on our quality and function variables.
-We computed distance between our binary vectors using both $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm.
-We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
+We computed distance between our binary vectors using $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and agglomerated using the Ward algorithm.
+We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
 We used 10,000 bootstrapped samples.
@@
 
-For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars) ), we computed the percentage of text records that provided support, and we computed the percentage of cultures with at least one text record providing support for that variable.
+For each of the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` coded variables (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)), we computed the percentage of text records that provided support, and we computed the percentage of cultures with at least one text record providing support for that variable.
 See Figure \@ref(fig:varsupport).
@@
 
-The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identify as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
-Within the management cluster, we identified a mediation subcluster (e.g., *Resolve conflict*, *Provide counsel*), and a an organization subcluster (e.g., *Organize cooperation*, *Military command*).
+The leader functions cluster analysis (Figure \@ref(fig:cluster)B) also identified two top-level clusters; the larger of which we identified as a management functions cluster (i.e., *Organize cooperation*, *Military command*, *Resolve conflict*) and the smaller of which we identified as a prosociality cluster (i.e., *Social functions*, *Distribute resources*, *Hospitality*).
+Within the management cluster, we identified a mediation subcluster (e.g., *Resolve conflict*, *Provide counsel*), and an organization subcluster (e.g., *Organize cooperation*, *Military command*).
 
@@
 The costs follow the potential benefits and leaders stand to lose their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures).
+See Figure \@ref(fig:varsupportcb).
 
@@
 Followers most frequently incur costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures).
-
-For both leaders and followers, individuals stand to lose what they are able to gain.
-The clear differences lie in leaders' opportunities for increased social status and mating opportunities.
 See Figure \@ref(fig:varsupportcb).
@@
 
-Results of the current study represent the broadest and most comprehensive view of the qualities of leaders, (e.g., the types of people that occupy leadership roles), the functions of leaders, (e.g., what do leaders do), and the costs and benefits of leadership for both leaders and followers, across a diverse and culturally representative sample of the ethnographic record.
-These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence strategy, and sex of the leader.
+This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record.
+These variables were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
+Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers, across a diverse and culturally representative sample of the ethnographic record.
+These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence stype, and sex of the leader.
 
@@
 See Figure \@ref(fig:varsupport).
+These qualities and functions correspond to those highlighted by several evolutionary theories of leadership [@garfield_review_2019].
+See the *Synthesis* section below.
 
@@
 
-## Clusters of leader functions and qualities
+## Clusters of leader qualities
 
 Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001].
-Exploratory analyses reported here suggest multiple leadership phenotypes and associated features.
+Exploratory cluster analyses reported here suggest multiple leadership phenotypes and associated features.
 
-<!-- Qualities features -->
+We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories.
+Variables were clustered together if they were positively correlated.
+Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster.
+See Figure \@ref(fig:cluster)A.
 
+Most of the leader quality variables were found in the *Social orientation* cluster, which had two smaller and more weakly supported subclusters that we identified as *Ability to provide social benefits* and *Ability to impose social costs*.
+The social benefit subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport), a point to which we shall return.
+The remaining social benefit variables included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which correspond to a charismatic style of leadership that has been identified in industrialized settings [@den_hartog_culture_1999; @grabo_charismatic_2016; @howell_role_2005].
+A few variables did not seem to fit the social benefit theme, however, such as *Xenophobia* and *Drug consumption* (although drug use is often an important social activity).
+
+The subcluster of social cost variables, which included *Physically formidable*, *Aggressiveness*, and *Feared*, align fairly closely with the *Dominance* theory of status and leadership [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013].
+Interestingly, this cluster also contained *Supernatural*.
 There is likely a strong connection between shamanism and ritual-based influence and the dominance style of leadership.
-Super natural abilities, being feared, aggressiveness, and strong authority formed a strong cluster of leadership qualities (Figure \@ref(fig:cluster)A, *Dominant/shaman* cluster $AU = 96$).
 Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].
+
+Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (top cluster, Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
+Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
+
+## Clusters of leader functions
+
+The leader function variables formed two major clusters, one that we identified as "Prosociality" and one that we identified as "Management".
+The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+
+The management subcluster included many more variables, which were grouped into two subclusters that we identified as "Mediation" and "Organization".
+The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991].
+Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@xx; @yy].
+
+The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*.
+This subcluster aligns with evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [@hooper_theory_2010; @van_vugt_cognitive_2007; for review, see @garfield_evolutionary_2019].
+
+## Benefits and costs of leadership and followership
+
+The evolution of leadership and followership is somewhat perplexing.
+Leadership often involves the provisioning of a public good at an individual cost, raising the question of how the individual benefits of leadership outweigh the costs.
+Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest.
+The benefits of followership must also outweigh the costs.
+We therefore coded all benefits and costs of both leadership and followership.
+Leaders were widely reported to receive material benefits and social status, but loss of material resources and social status were also widely reported.
+The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
+Our results cannot determine, however, if the benefits to leaders and followers outweighed their costs.
+See Figure \@ref(fig:varsupportcb).
+
+For both leaders and followers, in summary, individuals stood to lose what they were able to gain.
+The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers.
+See Figure \@ref(fig:varsupportcb).
+
+## Synthesis
+
+In general, our results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership [for review, see @garfield_evolutionary_2019].
+Leaders in the ethnographic record are portrayed as generally prosocial, and serve to resolve conflicts and organize cooperation.
+Leaders also receive material, social, and reproductive benefits.
+There is also a "dominant" style of leadership involving physical formidability, fear, and aggression.
+
+One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership.
+Our social cost cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable.
 @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n=43$) shamans were associated with providing decision-making services and associated with military power.
@@
 
-We have also identified a broad suite of positive personal characteristics, including many of the most cross-culturally frequent qualities of leadership, which generally correspond to prestige and charismatic leadership [@van_vugt_dual_2019; @cheng_pride_2010; @grabo_charismatic_2016; @henrich_evolution_2001; @price_evolution_2014; @von_rueden_roots_2014].
-Being knowledgeable or intelligent and highly experienced or accomplished cluster closely with high social status (the three most cross-culturally frequent qualities of leaders).
-Other personal characteristics, such as having a favorable personality, humility, oratory skill, charisma, and interpersonal skills are also associated with this feature of leadership (Figure \@ref(fig:cluster)A, *Positive personal characteristics* cluster $AU = 100$.).
-These features  suggest a strong link between prestige-based leadership, associated with deference towards individuals with special kills, high intelligence, and accomplishments [@cheng_dominance_2019; @van_vugt_dual_2019; @henrich_evolution_2001], and charismatic leadership, associated with oratory skills, high charisma, confidence, and interpersonal skills [@den_hartog_culture_1999; @grabo_charismatic_2016; @howell_role_2005].
+We also found a link between prestige-based leadership, associated with deference towards individuals with special skills, high intelligence, and accomplishments [@cheng_dominance_2019; @van_vugt_dual_2019; @henrich_evolution_2001], and charismatic leadership, associated with oratory skills, high charisma, confidence, and interpersonal skills [@vanvugtXXX].
+See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster.
+This suggests that these leadership strategies might be closely related, or part of a broader strategy of social influence based on the provisioning of various benefits.
 
-The most common benefits associated with leadership are material resources, increased social status, and social services (See Figure \@ref(fig:varsupportcb)).
-Measures of material and social wealth coupled with reproductive success represent a feature of leadership qualities descriptive of benefits of social influence and high social status (Figure \@ref(fig:cluster)A, *Material, social, & reproductive success* cluster $AU = 96$.).
-
-<!-- Functions features -->
-
-Results suggest X features of the functions of leaders.
-Leaders may serve various prosocial functions, such as providing subsistence and distributing resources, they may be tasked with providing hospitality services to guests or followers, they function to protect the group via a variety of behaviors, and very commonly are leading in various social functions tied to cultural values and social practices (Figure \@ref(fig:cluster)B, *Prosocial* cluster $AU = 95$).
-
-## Synthesis
+Intergenerational transmission; The fact that variables indicating success in these three domains were correlated supports the...
 
@@
 Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample.
-Nonetheless, these results are conditional upon the content ethnographers chose to discuss.
+Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it.
 The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results.
@@
 
+Conflict resolution, decision-making, providing counsel. Dominance. Benefits/costs.
+
 # References

### 2019-12-02 analyses2.R

@@ -1017,7 +1017,7 @@ statuspolytxts <- sum(leader_text2$qualities.polygynous & leader_text2$qualities
 # leader_text has 1000 rows
 # need raw texts for all 1212 rows
 
-textstats <- leader_text %>% 
+textstats <- text_records %>% 
   dplyr::select(cs_textrec_ID, raw_text) %>% 
   unnest_tokens(word, raw_text) %>% 
   # dplyr::filter(is.na(as.numeric(word))) %>% # filters out numbers, some of which are page numbers

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Minor MS changes here and bib edits.

--- 2019-12-02 leadership_across_cultures_contexts.Rmd
@@
 
-NEED CONCLUSION PARAGRAPH/STATEMENT.
+<!-- NEED CONCLUSION PARAGRAPH/STATEMENT.  -->
 
@@
 
-NEED CONCLUSION PARAGRAPH/STATEMENT.
+<!-- NEED CONCLUSION PARAGRAPH/STATEMENT. -->
 
@@
 
-NEED CONCLUSTION STATEMENT/PARAGRAPH.
+<!-- NEED CONCLUSTION STATEMENT/PARAGRAPH.  -->
 
@@
 2) Identify *features* of the qualities and functions of leaders (e.g., do particular variables tended to covary or cluster together?) in reference to theoretical models or novel interpretations.
-Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001].
+Such features might lend support to existing theories; for instance they might correspond to qualities associated with "prestige" or "dominance" [@henrich_evolution_2001] or functions associated with conflict or cooperation [@gavrilets_collective_2015; @glowacki_evolutionary_2017; @hooper_theory_2010].
 3) Identify the most cross-culturally frequent *costs and benefits* of leadership for both leaders and followers.

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Adding ethnographic examples in discussion. Edits in discussion section.

@@
 ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables.
-DISCUSS IRR.
 ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
+
+<!-- DISCUSS IRR. -->
 
@@
 These variables were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
-Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers, across a diverse and culturally representative sample of the ethnographic record.
+Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
 These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence stype, and sex of the leader.
@@
 
-The distribution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistence strategies, but at variable frequencies.
+The distribution of text records by group and subsistence type suggests leaders emerge within similar social domains among populations with varied subsistence strategies, but at variable frequencies.
 See Figure \@ref(fig:grouptype).
@@
 Leadership at the surpacommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
+The only example of supracommunity leadership among hunter-gatherers was from the Adamans in LOCATION, by AUTHOR [@az02-001] (45).
+(reproducing AUTHOR), who discussed the roles of a "head chief" from a permnant residential location and their marginal influence over other, local residential leaders:
 
-Hunter-gatherers and pastoralists, populations more likely to experience high residential mobility, exhibited greater relative frequencies of leadership within residential groups and kin groups.
+>Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief.
+The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e.
+those in authority over each community, consisting of from 20 to 50 individuals, maicla .
+The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited.
+It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers.
+It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating.
+It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved.
+The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other."
+
+Hunter-gatherers and pastoralists, populations more likely to experience high residential mobility [@wild_making_2019; @rigby_persistent_1985; @kelly_foraging_2013], exhibited greater relative frequencies of leadership within residential groups and kin groups.
 Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists.
-Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems.
+Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992].
+For example, among the Somali, a politically complex and socially stratigied agro-pastoralists group, DESCIRBE, AUTHOR [@mo04-021] (p.
+58) notes that, "the Sultan does not lead the army of the tribe in war.
+He names a chief of the armed forces."
 
 Figure \@ref(fig:grouptype) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups.
-One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups erodes.
-It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
+One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
+For example, among rural Korean populations a chairwomen of local organizations maintain imporant leadership positions within the groups na drepresent the group to govermental forces.
+AUTHOR [@aa01-062] (p.
+173) describes that in these local 'clubs', "the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
+In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process.
+She is not a figurehead but rather a middleman in the meeting, functioning, in effect, as a go-between for the group.
+There is no despotic idea in the context of a total consensus in this essentially egalitarian society.
+The chairwoman stands as a middleman between the government and the village.
+Although the government may call her a ‘leader’ in the village, she properly does not occupy this role."
 
@@
 They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
+AUTHOR [@oc06-034] (p.
+236) in describing kin group leaders among the Iban in LOCATION provides example of this suite of traits in leaders:
+
+>"Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society.
+The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111).
+Although usually a man, women tuai rumah are not unknown.
+The tuai rumah must inspire the confidence of his followers and exemplify the qualities of bravery, endurance, and determination that are essential for survival.
+Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement.
+The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom."
+
 See Figure \@ref(fig:varsupport).
-These qualities and functions correspond to those highlighted by several evolutionary theories of leadership [@garfield_review_2019].
+These qualities and functions correspond to those highlighted by several evolutionary theories of leadership which have drawn heavily on select ethnographic cases [@garfield_review_2019].
 See the *Synthesis* section below.
@@
 Several functions and qualities that are theoretically important were relatively rare in the ethnographic record.
-In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures.
+In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
+Among the Blackfoot in the North American plains for example, AUTHOR [@nf06-061] (p.
+411) descirbeds that, "If there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
+He did not order the others to follow, but he knew that they probably would go with him."
+
 Several theories of leadership also emphasize the moral authority of leaders [e.g., @henrich_big_2015; @henrich_evolution_2001; @richerson_tribal_2009; @tietjen_moral_1985; @boggild_evolved_2016], yet leaders were seen as sources of moral authority or "fair" in only a minority of cultures (`r moral_pct`% and `r fairness_pct`% of cultures, respectively).
-The "Dominance" theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
+The *Dominance* theory of leadership [e.g., @chapais_competence_2015; @cheng_dominance_2019; @cheng_pride_2010; @henrich_evolution_2001; @tiger_imperial_1971] emphasizes leaders' ability to instill fear, but in only `r feared_pct`% of cultures were leaders characterized as feared.
 Finally, a large literature on leadership in industrialized societies highlights the attractiveness of leaders [e.g., @sinclair_sexuality_1995; @altemeyer_sexual_1974; @march_leadership_2009], but leaders were characterized as attractive in only `r attractive_pct`% of cultures.
@@
 Leadership and high social status are expected to be associated with particular suites of characteristics, which potentially vary by cultural or social context, personality, or via other inter-individual variation [@cheng_two_2013; @von_rueden_roots_2014; @hogan_personality_2020; @lukaszewski_extraversion_2015; @henrich_evolution_2001].
-Exploratory cluster analyses reported here suggest multiple leadership phenotypes and associated features.
+Exploratory cluster analyses reported here suggest multiple leadership phenotypes.
 
@@
 See Figure \@ref(fig:cluster)A.
+Several of traits often cooccur.
+Among the Trobriands horticulturalists in LOCATION, AUTHOR [@ol06-037] (p.
+47 ) describes relationships between plogynous marriages, increased wealthm, and widespread social contacts:
+
+>Chiefs managed unusually large quantities of resources and it is agreed that the basis of such wealth was polygamy.
+Rank was quite directly expressed in the number of wives.
+The chief of Omarakana had 16 wives in Malinowski's time, while his predecessor Enamakala had 19 in 1893 (Annual Report 1893-4:19).
+In 1950 despite long-established missionary activity, there were still 13 (Powell 1960: 136).
+In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
+An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)...
 
 Most of the leader quality variables were found in the *Social orientation* cluster, which had two smaller and more weakly supported subclusters that we identified as *Ability to provide social benefits* and *Ability to impose social costs*.
-The social benefit subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport), a point to which we shall return.
-The remaining social benefit variables included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which correspond to a charismatic style of leadership that has been identified in industrialized settings [@den_hartog_culture_1999; @grabo_charismatic_2016; @howell_role_2005].
+The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport), a point to which we shall return.
+As brief examples of the very many in the ethngorraphic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them." [fo07-056] (p.
+85); and among the Highland Scots, AUTHOR [es10-003] (p.
+147) notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit heirachy in the context of cultural transitions.
+AUTHOR [@nm09-056] (p.
+183) describes Minnie Kellog (born in 1880), a land-claims activist, writer, and perhaps the most famous female leader among the Iriquous:
+
+>Despite her exceptional gifts -- a brilliant mind, beauty, selfconfidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
+It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
+Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
+Unfortunately, because of her questionable ethies and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
+Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
+
+The remaining social benefit variables included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005].
 A few variables did not seem to fit the social benefit theme, however, such as *Xenophobia* and *Drug consumption* (although drug use is often an important social activity).
@@
 
-Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (top cluster, Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
+Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (top cluster, Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013].
 Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme.
@@
 The leader function variables formed two major clusters, one that we identified as "Prosociality" and one that we identified as "Management".
-The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
+In the absenece of institutionalized leadership roles, which charachterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.
 
@@
 The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991].
-Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@xx; @yy].
+Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@de_cremer_intergroup_2002].
+In the absenece of state-instiutions local leaders are often tasked with managing various economic systems and providing social services.
 
 The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*.
-This subcluster aligns with evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [@hooper_theory_2010; @van_vugt_cognitive_2007; for review, see @garfield_evolutionary_2019].
+This subcluster aligns with evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [@hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
 
@@
 The benefits of followership must also outweigh the costs.
-We therefore coded all benefits and costs of both leadership and followership.
+We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading or following.
 Leaders were widely reported to receive material benefits and social status, but loss of material resources and social status were also widely reported.
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
-Our results cannot determine, however, if the benefits to leaders and followers outweighed their costs.
+Our results cannot determine, however, if the benefits to leaders and followers outweighed their costs, but shed light on the social and biological pressures associated with leadership and social hierarchy.
 See Figure \@ref(fig:varsupportcb).
@@
 
-In general, our results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership [for review, see @garfield_evolutionary_2019].
+In general, our results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theoretical predictions [for review, see @garfield_evolutionary_2019].
 Leaders in the ethnographic record are portrayed as generally prosocial, and serve to resolve conflicts and organize cooperation.
@@
 One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership.
-Our social cost cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable.
+Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable.
 @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n=43$) shamans were associated with providing decision-making services and associated with military power.
@@
 
-We also found a link between prestige-based leadership, associated with deference towards individuals with special skills, high intelligence, and accomplishments [@cheng_dominance_2019; @van_vugt_dual_2019; @henrich_evolution_2001], and charismatic leadership, associated with oratory skills, high charisma, confidence, and interpersonal skills [@vanvugtXXX].
+We also found a link between prestige-based leadership, which is associated with deference towards individuals with special skills, high intelligence, and accomplishments [@cheng_dominance_2019; @van_vugt_dual_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, high charisma, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005].
 See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster.
-This suggests that these leadership strategies might be closely related, or part of a broader strategy of social influence based on the provisioning of various benefits.
+This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits.
 
 Intergenerational transmission; The fact that variables indicating success in these three domains were correlated supports the...
+
+The strong support for material and social benefits assocaited with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequlity [@bowles_emergence_2010; @mattison_evolution_2016].
+Various sources of material wealth and strategic or instituionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporates and pay close attention to group processes and social strategies associated with leadership and followership.
 
@@
 Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
-However, the ubiquity of conflict resolution as a function of leaders across social contexts calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits.
-[@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018].
+However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018].
 
-Conflict resolution, decision-making, providing counsel. Dominance. Benefits/costs.
+decision-making, providing counsel. Dominance. Benefits/costs.
 

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Added more ethnographic examples in discussion section.

@@
 Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
-These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence stype, and sex of the leader.
+These results reveal universal dimensions of human leadership while also suggesting important variation by group type, subsistence type, and sex of the leader.
 
@@
 Leadership at the surpacommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
-The only example of supracommunity leadership among hunter-gatherers was from the Adamans in LOCATION, by AUTHOR [@az02-001] (45).
-(reproducing AUTHOR), who discussed the roles of a "head chief" from a permnant residential location and their marginal influence over other, local residential leaders:
+The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands  by @az02-001: p.
+45, who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders:
 
@@
 Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992].
-For example, among the Somali, a politically complex and socially stratigied agro-pastoralists group, DESCIRBE, AUTHOR [@mo04-021] (p.
+For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 (p.
 58) notes that, "the Sultan does not lead the army of the tribe in war.
@@
 One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
-For example, among rural Korean populations a chairwomen of local organizations maintain imporant leadership positions within the groups na drepresent the group to govermental forces.
-AUTHOR [@aa01-062] (p.
+For example, among rural Korean populations a chairwomen of local organizations maintain important leadership positions within the groups and represent the group to governmental forces.
+@aa01-062 (p.
 173) describes that in these local 'clubs', "the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
@@
 They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
-AUTHOR [@oc06-034] (p.
-236) in describing kin group leaders among the Iban in LOCATION provides example of this suite of traits in leaders:
+@oc06-034 (p.
+236) in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders:
 
@@
 In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
-Among the Blackfoot in the North American plains for example, AUTHOR [@nf06-061] (p.
-411) descirbeds that, "If there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
+Among the Blackfoot in the North American plains for example, @nf06-061 (p.
+411) describes that, "If there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
 He did not order the others to follow, but he knew that they probably would go with him."
@@
 See Figure \@ref(fig:cluster)A.
-Several of traits often cooccur.
-Among the Trobriands horticulturalists in LOCATION, AUTHOR [@ol06-037] (p.
-47 ) describes relationships between plogynous marriages, increased wealthm, and widespread social contacts:
+Several of traits often co-occur.
+Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 (p.
+47 ) describes relationships between polygnyous marriages, increased wealth, and widespread social contacts:
 
@@
 The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport), a point to which we shall return.
-As brief examples of the very many in the ethngorraphic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them." [fo07-056] (p.
-85); and among the Highland Scots, AUTHOR [es10-003] (p.
-147) notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit heirachy in the context of cultural transitions.
-AUTHOR [@nm09-056] (p.
-183) describes Minnie Kellog (born in 1880), a land-claims activist, writer, and perhaps the most famous female leader among the Iriquous:
+As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them." [fo07-056] (p.
+85); and among the Highland Scots, es10-003 (p.
+147) notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
+@nm09-056 (p.
+183) describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
 
@@
 The prosocial cluster included *Distribute resources*, *Protection*, and *Prosocial investment*, which are functions that are commonly highlighted in the anthropological literature on leadership in nonindustrial societies [@sahlins_poor_1963; @johnson_organizational_1982; @service_origins_1975; @boehm_egalitarian_1993; @boehm_hierarchy_1999; @johnson_evolution_1987].
-In the absenece of institutionalized leadership roles, which charachterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.
+In the absence of institutionalized leadership roles, which characterize many of the relatively small-scale populations in this sample, leaders often maintain social influence via continued community support and the demands of followers can shape the functions of leaders to meet local needs.
 
-The management subcluster included many more variables, which were grouped into two subclusters that we identified as "Mediation" and "Organization".
-The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991].
+The management cluster included many more variables, which were grouped into two subclusters that we identified as "Mediation" and "Organization".
+The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
+Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 (p.
+185) describes the nature of conflict resolution processes in the absence of strong authority:
+
+>The elder settles numerous disputes, both within and without his hunting band.
+Such disputes mainly concern marital disagreements, Schebesta says.
+The elders often settle controversies by delivering speeches during the evening, leading frequently to noisy debates between the elders of different bands or lineages.
+Outside the lineage context, the elder is not in any way different from the others; there are none of the trappings of villager chieftainship.
+He delivers his decisions without hesitancy and is obeyed in the same way within his lineage.
+His power is limited, however, by custom and tradition as well as by context.
+He may neither ignore nor alter traditional practices.
+
 Other variables in this subcluster indicated more generic group strategizing, such as *Strategic planning* and *Group representative* [@de_cremer_intergroup_2002].
-In the absenece of state-instiutions local leaders are often tasked with managing various economic systems and providing social services.
+In the absence of state-institutions local leaders are often tasked with managing various economic systems and providing social services.
 
 The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*.
-This subcluster aligns with evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [@hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
+This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
+In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overlly conflict with individual or household-level goals, such as among the Tukano of the Braziallian Amazon, where @sq19-018 (p.
+127) notes:
+
+>Wanano chieftainship entails control over three major aspects of village life: fishing resources, labor, and ritual.
+A mahsa wami must hold sway over all three areas to lead effectively.
+Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor, Control, then, implies the obligation to redistribute.
+The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict.
+I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.
 
@@
 
-Intergenerational transmission; The fact that variables indicating success in these three domains were correlated supports the...
-
-The strong support for material and social benefits assocaited with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequlity [@bowles_emergence_2010; @mattison_evolution_2016].
-Various sources of material wealth and strategic or instituionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporates and pay close attention to group processes and social strategies associated with leadership and followership.
+The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
+Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporates and pay close attention to group processes and social strategies associated with leadership and followership.
 

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Additional minor edits.

@@
 It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved.
-The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other."
+The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.
 
@@
 
->"Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society.
+>Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society.
 The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111).
@@
 Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement.
-The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom."
+The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom.
 
@@
 A mahsa wami must hold sway over all three areas to lead effectively.
-Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor, Control, then, implies the obligation to redistribute.
+Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor.
+Control, then, implies the obligation to redistribute.
 The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict.
@@
 Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable.
-@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n=43$) shamans were associated with providing decision-making services and associated with military power.
+@winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services and associated with military power.
 Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)).
@@
 
-We also found a link between prestige-based leadership, which is associated with deference towards individuals with special skills, high intelligence, and accomplishments [@cheng_dominance_2019; @van_vugt_dual_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, high charisma, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005].
+We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005].
 See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster.

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Other edits in discussion sections, etc.

@@
 @oc06-034 (p.
-236) in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders:
+236) in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):
 
@@
 Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 (p.
-47 ) describes relationships between polygnyous marriages, increased wealth, and widespread social contacts:
+47 ) describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
 
@@
 This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
-In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overlly conflict with individual or household-level goals, such as among the Tukano of the Braziallian Amazon, where @sq19-018 (p.
+In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 (p.
 127) notes:
@@
 The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
-Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporates and pay close attention to group processes and social strategies associated with leadership and followership.
+Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.
 
@@
 We, however, view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description.
-Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record.
+Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].
 These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent.
 The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.
+
+Another limitation lies in our ability to interpret the absence of evidence.
+We are confident that the cross-culturally frequent measures discussed here and strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns in human behavior and culture.
+We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record.
+Nonetheless, results reported here provide a broad, comprehensive, and cross-cultural framework for future empirical and theoretical work to draw on.
 

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Several very minor edits.

@@
 Pastoral and horticultural populations often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaborated political leadership systems [@ember_warfare_1992].
-For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 (p.
-58) notes that, "the Sultan does not lead the army of the tribe in war.
+For example, among the Somali, a politically complex and socially stratified agro-pastoralists group distributed across the Horn of Africa, @mo04-021 [p.
+58] notes that, "the Sultan does not lead the army of the tribe in war.
 He names a chief of the armed forces."
@@
 Figure \@ref(fig:grouptype) is consistent with a negative relationships between political leadership at the community leadership or above and leadership in kin groups or residential groups.
-One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
+One possible interpretation is that as political structures become more institutionalized and economic systems transcend the household, leadership within households and residential groups diminishes.
+It is interesting, however, that even among the agricultural populations that are characterized by leadership systems beyond the community level, leadership within residential and family group persists.
 For example, among rural Korean populations a chairwomen of local organizations maintain important leadership positions within the groups and represent the group to governmental forces.
-@aa01-062 (p.
-173) describes that in these local 'clubs', "the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
+@aa01-062 [p.
+173] describes that in these local 'clubs', "the procedure of discussion and decision-making is not based on an idea of democracy but rather on that of unanimity.
 In this context, the ‘leader’ does not have to be a person of marked influence in the decision-making process.
@@
 They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources.
-@oc06-034 (p.
-236) in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):
+@oc06-034 [p.
+236] in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):
 
@@
 In the literature on leadership among non-human animals, for example, leaders commonly initiate group movement [@couzin_effective_2005], and "movement" was one of four leadership dimensions that @smith_leadership_2016 identified as comparable among humans and non-human animals, yet movement or migration was identified as a leader function in only `r movement_pct`% of cultures, of which the vast majority (71%) were hunter-gatherers.
-Among the Blackfoot in the North American plains for example, @nf06-061 (p.
-411) describes that, "If there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
+Among the Blackfoot in the North American plains for example, @nf06-061 [p.
+411] describes that, "If there was a dispute as to whether the camp should move north or south, the chief might present his arguments, gain support, and then say that he was going south.
 He did not order the others to follow, but he knew that they probably would go with him."
@@
 Several of traits often co-occur.
-Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 (p.
-47 ) describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
+Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p.
+47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:
 
@@
 In traditional Trobriand society it seems that all leaders of village clusters of guyau status became polygamists.
-An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)...
+An emerging leader might also expect to be given wives by traditional allies outside the cluster while he might demand them from traditional rivals of subordinate subclans (Powell 1960:136)....
 
 Most of the leader quality variables were found in the *Social orientation* cluster, which had two smaller and more weakly supported subclusters that we identified as *Ability to provide social benefits* and *Ability to impose social costs*.
-The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport), a point to which we shall return.
-As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them." [fo07-056] (p.
-85); and among the Highland Scots, es10-003 (p.
-147) notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
-@nm09-056 (p.
-183) describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
+The social benefits subcluster contained the *Knowledgeable/intelligent* and *Experienced/accomplished* variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and these variables are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return.
+As brief examples of the very many in the ethnographic record, among the Azande, "though the most knowledgeable witch-doctors tend to become leaders, there is no institutionalized organization or leadership among them" [@fo07-056, p.
+85]; and among the Highland Scots, @es10-003 [p.
+147] notes, "on the whole, their long ties with place and people, coupled with education, made them useful leaders of a community in which leaders were needed to adapt life to a changing economy." Such exceptionally talented leaders, however, also face unique challenges, particularly within more egalitarian societies which emphasizes personal autonomy and resit hierarchy in the context of cultural transitions.
+@nm09-056 [p.
+183] describes Laura Cornelius Kellogg (also known as Minnie and Wynnogene, 1880-1947), a land-claims activist, writer, and perhaps the most famous female leader among the Iroquois:
 
->Despite her exceptional gifts -- a brilliant mind, beauty, selfconfidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
+>Despite her exceptional gifts -- a brilliant mind, beauty, self confidence, unusual oratorical abilities, and her educational attainments -- Kellogg is also the most controversial Iroquois leader of the twentieth century.
 It is clear from her many bizarre involvements that she misused her prodigious talents and/or was incapable of carrying out all the massive designs she had for her people's betterment.
 Although acknowledged today as a major force and brilliant person, she is accused by many Iroquois elders of swindling them out of hundreds of thousands of dollars in her abortive efforts to bring their land claims to fruition and of creating debilitating factionalism that impeded tribal development for decades.
-Unfortunately, because of her questionable ethies and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
+Unfortunately, because of her questionable ethics and her inability to carry out what she espoused, Kellogg is blamed today for all that went wrong in Iroquois history in the interwar period.
 Consequently, her life story had the feel of a Greek tragedy: she wanted to use her extraordinary abilities to help her people but ended up accused by them of being a common outlaw.
@@
 The mediation subcluster included variables such as *Resolve conflict* (the most widely supported function in our sample) and *Provide counsel* (also widely supported), which again are functions that are commonly noted in the anthropological literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991], and concerning conflict resolution specifically, widespread among nonhuman animals as well [@smith_leadership_2016; @smith_obstacles_2018; @wilson_sociobiology:_1980].
-Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 (p.
-185) describes the nature of conflict resolution processes in the absence of strong authority:
+Among the Mbuti, hunter-gatherers in the Congo Basin known for strong cultural values of personal autonomy and egalitarianism, @fo04-001 [p.
+185] describes the nature of conflict resolution processes in the absence of strong authority:
 
@@
 This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019].
-In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 (p.
-127) notes:
+In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p.
+127] notes:
 
@@
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
-Our results cannot determine, however, if the benefits to leaders and followers outweighed their costs, but shed light on the social and biological pressures associated with leadership and social hierarchy.
-See Figure \@ref(fig:varsupportcb).
-
-For both leaders and followers, in summary, individuals stood to lose what they were able to gain.
+Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy.
+For both leaders and followers, individuals stood to lose what they were able to gain.
 The major difference between leaders and followers lay in leaders' opportunities for increased social status and mating opportunities, relative to followers.
@@
 The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016].
-Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.
+Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." These results suggest anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.
 
@@
 The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results.
-We, however, view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description.
+We nevertheless view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description.
 Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Shortened title and moved synthesis to concluding remarks

@@
 ---
-title: "Leader functions and qualities across 59 cultures: an exploratory analysis"
+title: "Leadership across 59 traditional cultures: an exploratory analysis"
 author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
@@
 
-## Synthesis
+## Limitations
 
-In general, our results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theoretical predictions [for review, see @garfield_evolutionary_2019].
+This study is based on the ethnographic record.
+Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample.
+Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it.
+The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results.
+We nevertheless view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description.
+Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].
+These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent.
+The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.
+
+Another limitation lies in our ability to interpret the absence of evidence.
+We are confident that the cross-culturally frequent measures discussed here and strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns in human behavior and culture.
+We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record.
+Nonetheless, results reported here provide a broad, comprehensive, and cross-cultural framework for future empirical and theoretical work to draw on.
+
+# Concluding remarks
+
+Our results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theoretical predictions [for review, see @garfield_evolutionary_2019].
 Leaders in the ethnographic record are portrayed as generally prosocial, and serve to resolve conflicts and organize cooperation.
@@
 
-## Limitations
-
-This study is based on the ethnographic record.
-Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample.
-Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it.
-The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results.
-We nevertheless view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description.
-Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988].
-These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent.
-The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.
-
-Another limitation lies in our ability to interpret the absence of evidence.
-We are confident that the cross-culturally frequent measures discussed here and strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns in human behavior and culture.
-We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record.
-Nonetheless, results reported here provide a broad, comprehensive, and cross-cultural framework for future empirical and theoretical work to draw on.
-
-# Conclusion
-
 These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and across all levels of social organization.

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Edits to first paragraph of conclusion

@@
 
-Our results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theoretical predictions [for review, see @garfield_evolutionary_2019].
-Leaders in the ethnographic record are portrayed as generally prosocial, and serve to resolve conflicts and organize cooperation.
-Leaders also receive material, social, and reproductive benefits.
+Leaders in the ethnographic record are portrayed as generally prosocial, serve to resolve conflicts and organize cooperation, and receive material, social, and reproductive benefits.
 There is also a "dominant" style of leadership involving physical formidability, fear, and aggression.
+These results conform to the characteristics of leaders that are emphasized in the literature on the evolution of leadership and provide broad cross-cultural support for several influential theories, such as the prestige-dominance model, the collective action model, the service-for-prestige model, and the computational services model [for review, see @garfield_evolutionary_2019].
 

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Added ref to computational services model to final paragraph

@@
 Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group.
-However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018].
-
-decision-making, providing counsel. Dominance. Benefits/costs.
+However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019].
 

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Provided coding example

@@
 
+We illustrate our coding with an example from Andamanese hunter-gatherers [@az02-001, p. 45]:
+
+>Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief.
+The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e.
+those in authority over each community, consisting of from 20 to 50 individuals, maicla.
+The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited.
+It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers.
+It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating.
+It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved.
+The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.
+
+This text record was coded has having evidence for the functions *Resolve conflict*, *Organize cooperation*, *Movement/migration*, *Provide subsistence*, and *Misc.
+social functions*.
+For qualities, there was evidence for *High status*.
+Although authority is mentioned, we did not code evidence for this variable because the chief "possesses no power to punish or enforce obedience to his wishes," contrary to our operationalization.
 For additional details on ethnographic sampling and creation of the *leadershipdata* package see @garfield_evolutionary_2019.
@@
 Leadership at the surpacommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists.
-The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands  by @az02-001: p.
-45, who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders:
-
->Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief.
-The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e.
-those in authority over each community, consisting of from 20 to 50 individuals, maicla .
-The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited.
-It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers.
-It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating.
-It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved.
-The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.
+The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands by [@az02-001, p.
+45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).
 

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Adj. figure size

@@
   theme_bw(15)
+ggsave(filename = 'Figures/map.pdf', p_map)
 p_map
-# ggsave(filename = 'Figures/map.pdf', p_map)
 ```
@@
 Areas of the bars are proportional to the number of text records in that category."}
+ggsave(file="Figures/group_subsis.pdf", plot_group_subsis, width=16, height=6)
 plot_group_subsis #+ plot_group_sex
@@
 Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
+ggsave(file="Figures/variablesupport.pdf", plot.variable.support)
 plot.variable.support
@@
 Red triangles: percent of cultures with evidence of support (95% CI computed with a cluster bootstrap)."}
+ggsave(file="Figures/variablecostsbenefits.pdf", plot.variable.support_costs_benefits)
 plot.variable.support_costs_benefits

### 2019-12-07 leadership_across_cultures_contexts.Rmd
Final changes before submission.

@@
 output: 
-  bookdown::html_document2
+  bookdown::word_document2
 bibliography: Leader2.bib
@@
 The primary benefits leaders offer to followers are resolving conflicts, organizing cooperation, providing social functions, leading in inter-group conflicts, providing ritual services, and distributing resources.
-The primary benefits leaders receive from followers include non-food resources, increased social status, and social services from followers.
-The most frequent costs of leadership for leaders include loss of social status and resources, and increased risk of harm or conflict.
+The primary benefits leaders receive from followers include material resources, increased social status, and social services from followers.
+The most frequent costs of leadership for leaders include loss of social status, resources or social services and increased risk of harm or conflict.
 For followers, the most frequent costs include loss of resources, costs in providing social services, and increased risk of harm or conflict.
 
-Leadership in WEIRD settings often differs substantially from non-WEIRD settings.
+Leadership in WEIRD settings often differs substantially from non-WEIRD settings and dominant theories in managerial studies have not previously had broad empirical data on leadership, management, and social organization across diverse populations.
 In WEIRD settings influence often extends into multiple domains, organizations are more hierarchical, and interactions between leaders and followers are less frequently face-to-face.
@@
 The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications.
-For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied of industries across 24 countries.
+For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied of industries across 24 countries [@house2013strategic].
 The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
@@
 The association of formal community leadership and supernatural abilities is likely an important precursor of institutionalized and coercive authority.
-
 Within state-level societies including non-industrial kingdoms leadership positions are proscribed and institutionalized.
@@
 These results suggests high intelligence and specialized knowledge are universal traits of political and community leaders.
-
 Within managerial studies scholars often suggest the cognitive demands of leadership escalate with increasing organizational complexity [@zaccaro_organizational_2002].

### 2020-03-20 Response to reviewers.Rmd
Adding response to reviewers template.

---
title: "Response to reviewers"
author: "Zach Garfield"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Response to editor and reviewer comments

### Editor comments

**Statistical reporting and further analysis: The GLMM should be reported in full somewhere. This could go in the supplemental materials. However, can we learn more from it? Do the random effects matter for the ethnographers? What about the national origin effects for the ethnographers? We recommend taking seriously the reflexive analytical techniques of putting the ethnographer in the analysis (see work by Robert Aunger working in the Congo Basin). Does the ethnographer matter? If so, how much? Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population? This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**

**Could further analyses be done to obtain more information on the nature of the variation within the data?**

**1.	The cluster analysis is nice, but it opens up the option to combine items within both qualities and functions to create a smaller number of higher-level categories or codes that, arguably, are better measured. The issue here is that some things that don’t appear frequently might rise in frequency when they are combined with other similar items. Suppose you use a higher-level category that contains 4 codes and you code the society as ‘1’ if any of the codes appears and zero otherwise. In our view, you need not go crazy with this approach, but look for any tightly bound clusters and consider pursuing those. If you are able to do this, the presentation could be more compact and easier to digest for the reader.**

**2.	Why not take either some of your original codes or some of these higher-level codes (from the above cluster analysis) and regress them on societal-level features related to subsistence type (foragers, pastoralist, etc. or simply forager vs. non-forager if that’s appropriate), levels of jurisdictional hierarchy (or other measures of societal complexity), population size, and potentially ecological variables. Your sample is not big, but you might be able to get some juice out of this. If interesting findings do emerge, you should consider adding or otherwise altering your visualizations.**

**3.	Be clearer about how you are accounting for research effort in each society. For example, are you controlling for the total number of words or pages for each population? This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**

**4.	Did you control for observation year? This might be too highly correlated with ethnographer identity, but worth checking**

**5.	Phylogenetic controls. While this is the probability sample, worries persist about non-independence (see R1). There are a variety of approaches to mitigate (and study!) this. Is there any clumping among related groups that cannot be explained by other variables? Do the patterns hold when you either control for linguistic distance or cluster in some way (continental fixed effects at a minimum)?**

**6.	Analyze the degree to which ‘researcher efforts’ (say measured by total words on that society and/or number of paragraphs on leadership) predict the appearance of a code. For example, what if the total ethnographic pages strongly predict the the appearance of “high status” as a code. Suppose the 10% of populations who don’t mention “high status” hardly wrote anything about leadership. On the other hand, if for other traits, there’s no relationship between researcher effort and appearance, then a case can be made that absence of evidence could very well be evidence of absence.**

**We recommend making liberal use of supplemental materials. Some of the analyses may not yield findings of sufficient interest to warrant placement in the main text, but some readers may want to have a look at them, so put them in the supplemental and briefly mention them in the main.**

**Returning to the WEIRD Problem: in your discussion, please circle back to the question of what research might be missing by focusing on leadership in Western or industrialized societies. Say more on how leaders in small-scale societies are the same and different—and perhaps you will be able to explain some of that variation with variables like jurisdictional hierarchy or subsistence type or mobility measures. As mentioned by R1, this likely involves returning to the GLOBE data and putting it together with your findings more directly.**

**Status vs. Leadership: R2 asks about status vs. leadership. Given recent debates about a “leadership psychology” as distinct and separable from status psychologies (like prestige and dominance, van Vugt), it might be worth expanding a bit on what these data say about the connection between status and leadership.**

**Expanding the dataset: the natural next step with this project is to use this small sample of hand coded categories to create an automated approach for all of the eHRAF, perhaps based on key words. The idea is to develop a set of keywords or other patterns that reliably predict when a paragraph should be coded in a particular way. The method can be developed and benchmarked using the dataset you’ve created and then simply applied in automated fashion to the entire eHRAF corpus. We offer this as a suggestion, but realize that it’s a lot to ask in this context.**

**Finally, the reviewers made a number of valuable suggestions, not enumerated above,  that should help you sharpen you presentation and discussion. Please attend to those.**

## Reviewer 1 comments: "Leadership across 59 traditional cultures: an exploratory analysis" Evolution & Human Behavior

**This manuscript presents a descriptive analysis of the HRAF to provide information about leadership in traditional societies. Overall, I enjoyed reading the paper and thought it was a useful resource. It provides a comprehensive analysis of leadership across a wide swath of traditional cultures. It reports on the types of functions leaders serve, the qualities possessed by leaders, and the costs and benefits of leadership experienced by both leaders and followers. The paper will likely be of interest to anyone working in the area of leadership, social hierarchy, or group behavior. I provide some comments below aimed at helping the authors increase the potential accessibility and impact of the work.**

**The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines. Neither the background section nor the Discussion section focuses much on contrasting WEIRD and traditional societies. The paper appears to miss out on a useful opportunity to provide some commentary on this issue. Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**

**The paper suggests that one of the most important qualities and benefits of being a leader involves being of "high status." That's a little dissatisfying. What's the difference between being high status and being a leader? I realize the authors provided brief operational definitions of the two constructions, but might they simply be two ways of describing the same thing (having a high degree of influence)? Perhaps the authors could do just a bit more to differentiate them.**

**Some of the benefits of leadership also appear as costs. Greater reflection and explanation on this aspect of the data seems warranted. I understand that leaders can enjoy certain benefits (access to resources) but then if they lose their status as leader they might lose those benefits (if someone falls from grace, they no longer get a larger piece of the pie). Do the findings reflect anything more complex than that? Their discussion of benefits of costs was brief and a little disappointing. The paper might benefit from enhancing that discussion.**

**Some readers might appreciate a bit more detail about how the HRAF database was created, including how the ethnographic records were identified and selected and whom and when they are from.**

**I don't work with ethnographic data, so perhaps this question is naïve, but: What conclusions can we draw about the actual frequency of leadership functions and traits from the frequency with which they are mentioned in the ethnographic record? There is an implicit assumption that the two are closely aligned. Has such an alignment been confirmed? Is that even testable?**

**What does it mean for someone to be a leader at the "supracommunity" level? I'm not familiar with the term, so had difficulty contextualizing the relevant findings. And how exactly do the authors define "community" versus "supracommunity"?**

## Reviewer 2 comments

**The authors present results of a Herculean task of mining the ethnographic record to estimate frequency of and covariation between functions, qualities, and costs/benefits of leadership. While the study itself is exploratory, it will be very useful to future work testing models of the evolution of leadership and social inequality in general. I have a number of suggestions and comments that may improve the paper, none highly critical:**

**The abstract could be tightened up in terms of description of particular costs and benefits to leaders and followers. And a brief justification in the abstract for why the study compares costs and benefits would be useful- e.g. understanding evolution of leader or follower strategies requires understanding the (fitness-relevant) costs and benefits to them.**

**Lines 88-89 suggest that the paper's ethnographic analysis will include leadership at all levels of social organization. But does the theory and analysis reported in the paper really extend to all levels of social organization- such as child play groups or households? It seems the focus is still community-level cooperation among adults. Perhaps clarify.**

**In the "Functions of leadership" section, I would expand on your claim that leadership emerges when there is demand for a "managerial presence". You might describe collective action or coordination problems here. That might better motivate the subsequent discussion of different functions of leadership in small-scale societies, as well as discussion of the conditions (e.g. demographic change, wealth access) that increase institutionalization or authority of leaders. Glowacki and von Rueden (2015) suggest that inter-family conflict resolution is where leader institutionalization often first arises, and Roscoe (2009) makes a similar argument concerning leadership in highland New Guinea, where the institutionalized display of material wealth organized by Big Men is rewarded more than ad hoc leadership during war.**

**The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs. transformational leadership and their sub-components being one of the most popular. So what should a reader make of this literature, with respect to evolution of leadership in small-scale societies? Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies? I'd be more explicit on this in the introduction. For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**

**There is little mention of gender in the Qualities of leadership section, even though this might be first attribute an alien might recognize as covarying with community-level leadership in human societies. Even in the more egalitarian societies, men often wield more overt political influence during camp or band-wide political discussions. Without getting into arguments about patriarchy and its origins, it might at least be useful to cite your own work on Chabu and work on the Tsimane (von Rueden et al. 2018) that finds similar traits predict leadership for men and women, implying its not gender per se that gives men a leadership advantage in politics.**

**How often are shamans also the most influential political figure (given your suggestion in line 317)? My sense from various ethnographies is that this was not the majority of cases (though often the case), whether in more egalitarian societies or chiefdoms. I could be wrong.**

**Even though this paper is more exploratory, in the last section before the methods, can you venture any predictions about leader function-leader quality covariation? Or about how reported costs and benefits of leadership and followership may vary depending on subsistence type?**

**Many of the variables coded as qualities of leaders may also fall under leader costs and benefits, e.g. polygyny, physical health, wealthy. Were variables like these analyzed in both cases?**

**Can you control for phylogeny (e.g. linguistic affinity). See Minocher et al. 2019 in Evolution and Human Behavior. It may be that covariation you detect between qualities and leader functions is driven by shared cultural history more than direct causal relationship between qualities and leader functions.**

**Can you report text records pertaining to male vs. female leadership by subsistence type (lines 966-982)?**

**While your analysis strategy was not designed to test particular predictions, I'd be very interested to see you report how the authority variable covaries with different leader functions in different subsistence types, which has relevance to institutionalization of leadership and evolution of social inequality.**

**Give more explanation of prestige theory and computational services theory (lines 1279-1280), including what different predictions they make.**

**Followers in many cases may be making best of limited options, particularly in societies with steeper hierarchy. So I'm not sure benefits of followership must outweigh the costs (1414-1415) in same way as for leadership.**

### 2020-04-02 Response to reviewers.Rmd
Starting to work on response to reviewers.
Adding cost v benefits code.

@@
 
-## Reviewer 1 comments: "Leadership across 59 traditional cultures: an exploratory analysis" Evolution & Human Behavior
+## Reviewer 1 comments
 
@@
 
+We greatly appreciate the reviewers detailed comments and very useful suggestions.
+
 **The paper was framed partially as providing an analysis of differences between traditional societies and WEIRD societies but, unless I missed it, there was not much analysis along these lines.
@@
 Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**
+We agree our initial submisson did not sufficently accomplish this goal.
+We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.
+
+NEED TO ADD TO DISCUSSION & ANALYSES?
 
@@
 Perhaps the authors could do just a bit more to differentiate them.**
+
+<!-- IDEA: Perform some textual analysis on teh 10% of text records in which there was not evidence leaders were high status.
+How does these records compare RE gender, group structure, or other qualities (maybe), or the most frequent functions, i.e., try to charachterize non-high status leaders.
+-->
 
@@
 
+<!-- IDEA: Represent the benefits and costs of leadership within group type.
+Would need to standardize the estimates, such our culture scores in first study, but it could be group structure benefit score, etc.
+Then we can roughly compare how costs and benefits comapre across group types.
+-->
+
 **Some readers might appreciate a bit more detail about how the HRAF database was created, including how the ethnographic records were identified and selected and whom and when they are from.**
+
+We added a few additional details in the methods section on the eHRAF and ethngraphic sample and now more clearly direct readers to the original pubication from which the database was built, which includes additional details.
+
+We also include additional details in the results and supplimentary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.
 
@@
 
+The reviewer brings up an important point.
+For one, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables.
+Despite this low threshold for acceptign supporting evidence, the ethnogrtaphic record (sampled here) does not provide support for every variable from every culture.
+At this culture-level, we feel more confident our data provide a estiamte of cross-cultural frequency, albeit an inherently noisy estimation.
+Also, the entire field of ethnology (the branch of anthropology focused on comparing charachteristics of populations across the globe), is based on developing comparable and often quantitaive measures from primary ethnogrpahic information.
+For example, the Ethnographic Atlas, which provides hundreds of coded quantitaive and qualitive variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operatiopnalized variables here.
+Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations.
+
+At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic.
+However, fundamentally these results provide a systematic view of the ethnographic record.
+From these results field researchers and theoreticians can draw inference from a more comprehensive view of human culturla diversity rather than haphazardly self-selecting a few ethngoraphic cases studies.
+
+We now discuss these limitations and points more clearly in the limitations setion. NEED TO DO.
+
 **What does it mean for someone to be a leader at the "supracommunity" level?
@@
 And how exactly do the authors define "community" versus "supracommunity"?**
+
+In the methods section we now include a description of all levels of our *group structure" coding to clearly explain this coding scheme.
+We define supracommunity leaders when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders.
 

### 2020-04-02 leadership_across_cultures_contexts.Rmd
Starting to work on response to reviewers.
Adding cost v benefits code.

@@
 ---
-title: "Leadership across 59 traditional cultures: an exploratory analysis"
+title: "Leadership across cultures and contexts: an exploratory analysis"
 author: "Zachary H. Garfield, Kristen L. Syme, and Edward H. Hagen"
@@
 output: 
-  bookdown::word_document2
+  bookdown::html_document2
 bibliography: Leader2.bib
@@
 
-We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations.
+We first briefly review theoretical, ethnographic, and empirical perspectives on the qualities of leaders, the functions they serve, and the costs and benefits of leadership across human societies, including WEIRD and non-WEIRD populations [i.e., @henrich_weirdest_2010].
 We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
@@
 
-<!-- NEED CONCLUSION PARAGRAPH/STATEMENT.  -->
+<!-- ADDED CONCLUSION PARAGRAPH/STATEMENT.  -->
+In contrasting the functions leaders serve in smaller-scale, non-industrial, and more traditional societies (i.e., non-WEIRD), from larger-scale, (post)industrial contexts (i.e., WEIRD), community leaders in non-WEIRD settings are more likely to have domain-specific functions based on role-specialization [see @smith_leadership_2016].
+The hunt leader does not necessarily have authority over military raids and the lineage head does not necessarily have authority over ritual ceremonies.
+In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple domains.
+In the United States, despite a designed seperation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
+Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage inter-group social dynamics.
+<!-- Could add ethnographic examples for SSS and LSS  -->
 
@@
 Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
-Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are perceived to be associated with effective leadership.
+Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
 In particular, inducing motivation, having foresight, being encouraging and communicative, trustworthiness, and building self-confidence among followers are suggested to be universal attributes of outstanding leadership in managerial contexts.
@@
 <!-- NEED CONCLUSION PARAGRAPH/STATEMENT. -->
+Much of the theoretical leadership literature across disciplines suggests relative uniformity in many qualities of community leaders.
+Intelligence, charisma, effective decision-making capabilities, high social status and strong social networks are likely to be universal qualities of political and community leaders.
+In more socially stratified, politically complex, and in increasingly WEIRD contexts, heredity, ascribed social statuses, and affiliations are more likely to be critical qualities in achieving and/or maintaining leadership positions.
+We also envision a curvilinear relationship between leadership status and supernatural abilities and qualities.
+Community leaders among small-scale hunter-gatherers may be seen as experts in the use of ritual and supernatural "tools" though they typically do not have a monopoly in their use.
+In chiefdoms and and in early prehistoric state societies political leaders are often seen as divine and have special supernatural qualities commoners do not.
+In contemporary WEIRD societies there is a trend of secularization among political leaders, generally, although religious and spiritual leaders do have substantial influence over social and moral norms.
 
@@
 
-We evaluate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
-Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record speak for itself in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
+We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
+The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at teh paragraph level following a the Outline of Cultural Materials (OCM) coding scheme.
+The eHRAF is a subset of the larger HRAF and is continually growing in scope.
 
-Here we draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the electronic Human Relations Area Files (eHRAF) database [@naroll_proposed_1967].
-The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership.
+Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
+We draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
+The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers.
 For the current study, two authors coded each text record for presence of `r length(quality_vars)` distinct qualities of leaders, `r length(function_vars)` distinct functions leaders serve, and `r length(leader_benefit_vars)` benefits and `r length(leader_cost_vars)` costs that leaders or followers may receive.
+See @garfield_evolutionary_2019 for additional details on the ethnographic sample and methodology.
 
@@
 After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
-ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables.
+ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
@@
 # pvrect(m_pvclust_qual, alpha = 0.9)
-knitr::include_graphics("Figures/m_pvclust.both.svg")
+knitr::include_graphics("Figures/m_pvclust.both.png")
 ```

### 2020-04-02 analyses2.R

@@ -991,7 +991,44 @@ plot_group_sex <-
 plot_group_sex
 
 
-# Compute values ----------------------------------------------------------
+# Costs and benefits by group structure type ------------------------------
+se <- function(x) sd(x)/sqrt(length(x))
+
+# try to tidy data to: Group structure, Benefit type, Cost type, Status type (leader, follower), Mean, SE
+
+# Group by group structure and calculate mean and SE
+cb_grp <- leader_text2 %>% 
+  group_by(group.structure2) %>% 
+  select(contains(c("benefits", "costs"))) %>%
+  summarise_each(funs(mean, se)) %>% 
+  pivot_longer(cols=contains(c("benefits", "costs"))) # couldn't figure out how to pivot_longer on three vars
+
+# label means and SE to split the data frame
+cb_grp$value_type <- gsub("^.*\\_","",cb_grp$name)
+
+#Create dataframe of means and fix names
+cb_grp_means <- data.frame(split(cb_grp, cb_grp$value_type)[1])
+cb_grp_means$name <- str_sub(cb_grp_means$mean.name, end=-6)
+cb_grp_means$group.structure2 <- cb_grp_means$mean.group.structure2
+cb_grp_means <- cb_grp_means[,c("mean.value","name","group.structure2")]
+
+#Create dataframe of SEs and fix names
+cb_grp_se <- data.frame(split(cb_grp, cb_grp$value_type)[2])
+cb_grp_se$name <- str_sub(cb_grp_se$se.name, end=-4)
+cb_grp_se$group.structure2 <- cb_grp_se$se.group.structure2
+cb_grp_se <- cb_grp_se[,c("se.value","name","group.structure2")]
+
+#Merge mean and SE data back together               
+cb_grp_all <- full_join(cb_grp_means, cb_grp_se, by = c("name", "group.structure2"))
+
+# Create status and type variables
+cb_grp_all$status <- str_extract(cb_grp_all$name, ".+?(?=\\.)")
+cb_grp_all$type <- str_match(cb_grp_all$name, "\\.(.*?)_")
+
+
+
+
+# Compute values -----------------------------------------------------------
 
 group_sex_tbl <- xtabs(~demo_sex+group.structure2, leader_text2)
 female_residential_pct <- signif(group_sex_tbl['female', 'residential subgroup']/sum(leader_text$demo_sex == 'female'), 3)*100

### 2020-04-02 leader_contexts_analyses.R

@@ -24,6 +24,7 @@ library(effects)
 library(lme4)
 library(patchwork)
 library(pvclust)
+library(readxl)
 # library(tidybayes)
 # library(cowplot)
 # library(brms)
@@ -1231,6 +1232,38 @@ aheatmap(t(as.matrix(heatmap_data[,c(quality_vars)])),
          treeheight = 50,
          filename = 'heatmap_qualities.pdf')
 
+
+# GLOBE data analysis -----------------------------------------------------
+
+globe_leader <- read_excel("GLOBE-Phase-2-Aggregated-Leadership-Data.xls")
+
+# Remove dimebnsion aggregates
+globe_leader2 <- globe_leader %>% 
+  select(., -contains("Dimension"))
+
+# Cluster analysis
+globe_leader_clust <- pvclust(
+  globe_leader2[,c(3:23)], 
+  method.hclust = 'ward', 
+  method.dist = 'euclidean', 
+  nboot = 10000,
+  parallel = T
+)
+plot(globe_leader_clust)
+pvrect(globe_leader_clust)
+
+globe_societal <- read_excel("GLOBE-Phase-2-Aggregated-Societal-Culture-Data.xls")
+
+# Cluster analysis
+globe_societal_clust <- pvclust(
+  globe_societal[,c(3:20)], 
+  method.hclust = 'ward', 
+  method.dist = 'correlation', 
+  nboot = 10000,
+  parallel = T
+)
+plot(globe_societal_clust)
+pvrect(globe_societal_clust)
 
 
 # Save RData envrionment --------------------------------------------------

### 2020-04-02 leadership_across_cultures_contexts.Rmd
Added analysis of female co-authors and female leadership

@@
 
+## Effects of document meta-data on leadership patterns
+
+```{r tabauthorgender}
+unique_authors <- unique(authorship[c('author_ID', 'author_gender')])
+tab_author_gender <- table(unique_authors$author_gender)
+mm_coauthorSUM <- summary(mm_coauthor)
+mm_coauthorP <- mm_coauthorSUM$coefficients['female_coauthorTRUE', 'Pr(>|z|)']
+```
+
+We explored whether document features, such as date of publication and gender of authors, influenced reporting of leader functions and qualities.
+There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine; some of these authors were co-authors.
+We then tested if text records with a female author or co-author were more likely to discuss female leaders.
+Of the `r sum(tab_coauthor['FALSE',])` text records with no female co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
+Of the `r sum(tab_coauthor['TRUE',])` text records with female co-author, on the other hand, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
+However, after accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
+
 # Discussion

### 2020-04-02 analyses2.R

@@ -1111,3 +1111,64 @@ attractive_pct <- signif(d_melt$value[d_melt$Variable == "Attractive" & d_melt$T
 #   geom_violin()+
 #   geom_jitter(height = 0, width = 0.1) +
 #   geom_boxplot(width=.15)
+
+
+# Female leaders by ethnographer gender -----------------------------------
+
+# leader_text2$demo_sex
+# authorship
+# text_records
+
+female_coauthor <- function(cs_textrec_ID){
+  document_ID <- text_records$document_d_ID[text_records$cs_textrec_ID == cs_textrec_ID]
+  author_genders <- authorship$author_gender[authorship$document_ID == document_ID]
+  'female' %in% author_genders
+}
+
+leader_text2$female_coauthor <- sapply(leader_text2$cs_textrec_ID, female_coauthor)
+
+leader_text3 <- 
+  leader_text2 %>% 
+  dplyr::select(
+    cs_textrec_ID,
+    demo_sex,
+    female_coauthor
+  ) %>% 
+  mutate(
+    female_leader_present = case_when(
+      demo_sex == 'unknown' ~ 'unknown',
+      demo_sex == 'male' ~ 'no',
+      TRUE ~ 'yes'
+    ),
+    female_leader_present2 = female_leader_present == 'yes'
+  ) %>% 
+  dplyr::filter(
+    demo_sex != 'unknown'
+  ) %>% 
+  left_join(
+    text_records[c("document_d_ID", "cs_textrec_ID")]
+  ) 
+
+# Assuming each record is independent
+tab_coauthor <- xtabs(~female_coauthor + female_leader_present, leader_text3)
+summary(tab_coauthor)
+# plot(tab_coauthor)
+
+m_coauthor <- glm(
+  female_leader_present2 ~
+    female_coauthor,
+  family = binomial,
+  data = leader_text3
+)
+
+# Hierarchical model
+
+mm_coauthor <- glmer(
+  female_leader_present2 ~
+    female_coauthor +
+    (1|document_d_ID),
+  family = binomial,
+  data = leader_text3
+)
+mm_coauthorOR <- exp(fixef(mm_coauthor))[[2]]
+

### 2020-04-02 SI analyses.R
Removed unnecessary load

@@ -1,6 +1,5 @@
-### Suplimentary analyses
+### Supplementary analyses
 
-load("Leader2.Rdata")
 source("analyses2.R")
 
 library(ggalt)

### 2020-04-03 Response to reviewers.Rmd
Adding text to main MS.
Added stan_glmer models.

@@
 
-We greatly appreciate the reviewers detailed comments and very useful suggestions.
+We greatly appreciate the detailed comments and very useful suggestions.
 
@@
 For one, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables.
-Despite this low threshold for acceptign supporting evidence, the ethnogrtaphic record (sampled here) does not provide support for every variable from every culture.
-At this culture-level, we feel more confident our data provide a estiamte of cross-cultural frequency, albeit an inherently noisy estimation.
-Also, the entire field of ethnology (the branch of anthropology focused on comparing charachteristics of populations across the globe), is based on developing comparable and often quantitaive measures from primary ethnogrpahic information.
+Despite this low threshold for accepting supporting evidence, the ethnogrtaphic record (sampled here) does not provide support for every variable from every culture.
+At this culture-level we feel more confident our data provide a estiamte of cross-cultural frequency, albeit an inherently noisy, "course-grained" estimation.
+Also, the entire field of ethnology (the branch of anthropology focused on comparing charachteristics of populations across the globe) is based on developing comparable and often quantitaive measures from primary ethnogrpahic information.
 For example, the Ethnographic Atlas, which provides hundreds of coded quantitaive and qualitive variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operatiopnalized variables here.
 Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations.
-
 At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic.
-However, fundamentally these results provide a systematic view of the ethnographic record.
-From these results field researchers and theoreticians can draw inference from a more comprehensive view of human culturla diversity rather than haphazardly self-selecting a few ethngoraphic cases studies.
+However, fundamentally these results provide a systematic view of the ethnographic record on these measures.
+From our results field researchers and theoreticians can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting a few ethngoraphic cases studies.
 
@@
 I have a number of suggestions and comments that may improve the paper, none highly critical:**
+
+We greatly appreciate the detailed comments and very useful suggestions.
 
@@
 
+We now discuss sex-differences in community and political leadership, and refernce in greater deatils (albeit breif) the universal male bias in community leadership in the introuction and cite these works.
+
 **How often are shamans also the most influential political figure (given your suggestion in line 317)?

### 2020-04-03 leadership_across_cultures_contexts.Rmd
Adding text to main MS.
Added stan_glmer models.

@@
 
-Across culturally and economically diverse populations there are some broad commonalities in the qualities of community leaders.
+There are some broad commonalities in the qualities of community leaders across culturally and economically diverse populations.
 When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
-Elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
+Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse non-human social species [@smith_obstacles_2018; @tiger_female_1978].
+@low_sex_1992 found in a systematic cross-cultural sample that in only seven of 93 societies political leaders of both sexes were present and in only two societies were female political leaders identified as equally powerful as their male counterparts.
+Even among these rare cases, powerful female political leaders are still infrequent within the group.
+Women were more likely to hold influential leadership positions within kin groups [@low_sex_1992].
+More recently, empirical studies have suggested sex-differences in political influence and leadership across human societies including among more egalitarian cultural contexts are more likely driven by ecological and social structural factors, rather than sex (or gender identity) per se, including social capital and access to same-sex allies [@yanca_female_2004] and/or sex differences in the sexual division of labor [@von_rueden_sex_2018].
+Ethnographic fieldwork comparing leaders and non-leaders in two distinct small-scale societies has revealed sex is not a strong predictor of community influence or elected leadership positions, after accounting for a broad range of demographic, phenotypic, cognitive, social, and other traits derived from the evolutionary theoretical literature [e.g., @garfield_investigating_2019; @von_rueden_sex_2018].
+
+Across cultural contexts elders, respected individuals, and highly intelligent or knowledgeable individuals tend to have greater influence over community decisions across cultures, including among the most egalitarian hunter-gatherers, stratified chiefdoms, and contemporary state-level organizations [@silverman_how_1978; @van_vugt_cognitive_2007; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990].
 Drawing on cross-national data through the GLOBE project from over 15,000 middle managers across 60 countries, @den_hartog_culture_1999 conclude that elements of charismatic and transformational leadership -- leadership styles when leaders express an inclusive vision of future progress, lead through inspiration, and develop initiatives in collaboration -- are universally endorsed across nations and are qualities perceived to be associated with effective leadership.
@@
 We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
-The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at teh paragraph level following a the Outline of Cultural Materials (OCM) coding scheme.
+The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following a the Outline of Cultural Materials (OCM) coding scheme.
 The eHRAF is a subset of the larger HRAF and is continually growing in scope.
@@
 After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
-ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phratries, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
@@
 Of the `r sum(tab_coauthor['FALSE',])` text records with no female co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
-Of the `r sum(tab_coauthor['TRUE',])` text records with female co-author, on the other hand, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
+Of the `r sum(tab_coauthor['TRUE',])` text records with a female co-author, on the other hand, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
 However, after accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).

### 2020-04-03 analyses2.R

@@ -1172,3 +1172,80 @@ mm_coauthor <- glmer(
 )
 mm_coauthorOR <- exp(fixef(mm_coauthor))[[2]]
 
+library(rstanarm)
+library(bayesplot)
+options(mc.cores = parallel::detectCores())
+
+# stan_mm_coauthor <- stan_glmer(
+#   female_leader_present2 ~
+#     female_coauthor +
+#     (1|document_d_ID),
+#   family = binomial(link = "logit"),
+#   data = leader_text3,
+#   prior_intercept = normal(0,1),
+#   prior = normal(0,1), 
+#   chains = 4, 
+#   iter = 40000)
+# 
+# summary(stan_mm_coauthor, pars = "female_coauthorTRUE")
+# posterior_stan_mm_coauthor <- as.matrix(stan_mm_coauthor)
+# mcmc_areas(posterior_stan_mm_coauthor,
+#            pars = c("female_coauthorTRUE"),
+#            prob = 0.95)
+
+# Female leaders by publication year --------------------------------------
+
+leader_text4 <- left_join(leader_text3, documents, by = c("document_d_ID" = "d_ID"))
+
+# mm_pubyear <- glmer(
+#   female_leader_present2 ~
+#     `d_publication date` +
+#     (1|document_d_ID),
+#   family = binomial,
+#   data = leader_text4
+# )
+# mm_pubyearOR <- exp(fixef(mm_pubyear))[[2]]
+# 
+# stan_mm_pubyear <- stan_glmer(
+#   female_leader_present2 ~
+#     `d_publication date`  +
+#     (1|document_d_ID),
+#   family = binomial(link = "logit"),
+#   data = leader_text4,
+#   prior_intercept = normal(0,1),
+#   prior = normal(0,1), 
+#   chains = 4, 
+#   iter = 40000)
+# 
+# summary(stan_mm_pubyear, pars = "`d_publication date`")
+# posterior_stan_mm_coauthor <- as.matrix(stan_mm_pubyear)
+# mcmc_areas(posterior_stan_mm_coauthor,
+#            pars = c("`d_publication date`"),
+#            prob = 0.95)
+
+## Both gender and publication year in one model
+
+# ***********************
+#    USE THIS MODEL????
+# ***********************
+
+# stan_mm_pubyear_gender <- stan_glmer(
+#   female_leader_present2 ~
+#     female_coauthor +
[…]

### 2020-04-03 analyses2.R
added space.

@@ -1226,7 +1226,7 @@ leader_text4 <- left_join(leader_text3, documents, by = c("document_d_ID" = "d_I
 ## Both gender and publication year in one model
 
 # ***********************
-#    USE THIS MODEL????
+#  USE THIS MODEL????
 # ***********************
 
 # stan_mm_pubyear_gender <- stan_glmer(

### 2020-04-03 analyses2.R
Added idea in response to reviewers doc.
Commenting out and changing prior in stan author mode.

@@ -1232,16 +1232,16 @@ leader_text4 <- left_join(leader_text3, documents, by = c("document_d_ID" = "d_I
 # stan_mm_pubyear_gender <- stan_glmer(
 #   female_leader_present2 ~
 #     female_coauthor +
-#     (1|`d_publication date`)  +
+#     #(1|`d_publication date`)  +
 #     (1|document_d_ID),
 #   family = binomial(link = "logit"),
 #   data = leader_text4,
-#   prior_intercept = normal(0,1),
-#   prior = normal(0,1), 
-#   chains = 4, 
-#   iter = 40000)
+#   prior_intercept = student_t(1,0,1),
+#   prior = normal(0,1),
+#   chains = 4,
+#   iter = 20000)
 # 
-# summary(stan_mm_pubyear_gender, pars = c("female_coauthorTRUE"))
+# summary(stan_mm_pubyear_gender, pars = c("female_coauthorTRUE", "(Intercept)"))
 # posterior_stan_mm_coauthor_pubyear <- as.matrix(stan_mm_pubyear_gender)
 # mcmc_areas(posterior_stan_mm_coauthor_pubyear,
 #            pars = c("female_coauthorTRUE"),

### 2020-04-04 leadership_across_cultures_contexts.Rmd
Added inline r code for number of text records, and also for number of pages of text for the PSF cultures in the eHRAF

--- 2020-04-02 leadership_across_cultures_contexts.Rmd
@@
 Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
-We draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
+We draw on the same database of `r nrow(leader_text2)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
+The latter comprises over `r trunc(sum(dfwc$`N Pages eHRAF`)/1000)`K pages of digitized text.
+
 The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers.
@@
 First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+Then, KS coded all `r nrow(leader_text2)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
 After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
-ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG then followed the same process of coding the entire sample of `r nrow(leader_text2)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
@@
 
+## Effects of document meta-data on leadership patterns
+
+```{r tabauthorgender}
+unique_authors <- unique(authorship[c('author_ID', 'author_gender')])
+tab_author_gender <- table(unique_authors$author_gender)
+mm_coauthorSUM <- summary(mm_coauthor)
+mm_coauthorP <- mm_coauthorSUM$coefficients['female_coauthorTRUE', 'Pr(>|z|)']
+```
+
+We explored whether document features, such as date of publication and gender of authors, influenced reporting of leader functions and qualities.
+There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine, some of which were co-authors.
+We then tested if text records with a female author or co-author were more likely to discuss female leaders.
+Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
+Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
+After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
+
 # Discussion

### 2020-04-04 leadership_across_cultures_contexts.Rmd
Added https to doi. More detail on PSF. Add info on correlation between pages and text records, incl. scatter plot.

@@
 
-We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
+We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (https://doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
 The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at teh paragraph level following a the Outline of Cultural Materials (OCM) coding scheme.
@@
 We draw on the same database of `r nrow(leader_text2)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
-The latter comprises over `r trunc(sum(dfwc$`N Pages eHRAF`)/1000)`K pages of digitized text.
+The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethographic coverage from each of 60 culture regions.
+It comprises over `r trunc(sum(dfwc['N Pages eHRAF'])/1000)`K pages of digitized text.
 
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture.
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture.
 The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
 Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.
+
+```{r corpus}
+cor_pages_tr <- cor.test(~number_leader_records + `N Pages eHRAF`, dfwc)
+```
+
+There was a weak positive relationship between the number of leader text records we retrieved from each culture and the total number of pages of ethnography for each culture in eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
+A loess regression suggested that the number of records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure \@ref(fig:eHRAFpages)).
 
@@
 
+# Supplementary information
+
+```{r eHRAFpages, fig.height=6, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF.
+Each dot is one culture.
+Line fit by loess regression."}
+plot_pages_tr
+```
+
 # References

### 2020-04-06 Response to reviewers.Rmd
Added idea in response to reviewers doc.
Commenting out and changing prior in stan author mode.
Working on response to reviewers and MS text.
Looked at texts with no evidence for status by subsistence in mosaic plot.

Adjusted var names in both script files to match updated data package.

@@
 Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**
-We agree our initial submisson did not sufficently accomplish this goal.
+We agree our initial submisson did not sufficiently accomplish this goal.
 We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.
@@
 How does these records compare RE gender, group structure, or other qualities (maybe), or the most frequent functions, i.e., try to charachterize non-high status leaders.
+
+OR, better idea (?), use LASSO technique from paper one, to identify words predictive of high status present or not
 -->
@@
 
-We added a few additional details in the methods section on the eHRAF and ethngraphic sample and now more clearly direct readers to the original pubication from which the database was built, which includes additional details.
+We added a few additional details in the methods section on the eHRAF and ethnographic sample and now more clearly direct readers to the original publication from which the database was built, which includes additional details.
 
-We also include additional details in the results and supplimentary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.
+We also include additional details in the results and supplementary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.
 
@@
 For one, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables.
-Despite this low threshold for accepting supporting evidence, the ethnogrtaphic record (sampled here) does not provide support for every variable from every culture.
-At this culture-level we feel more confident our data provide a estiamte of cross-cultural frequency, albeit an inherently noisy, "course-grained" estimation.
-Also, the entire field of ethnology (the branch of anthropology focused on comparing charachteristics of populations across the globe) is based on developing comparable and often quantitaive measures from primary ethnogrpahic information.
-For example, the Ethnographic Atlas, which provides hundreds of coded quantitaive and qualitive variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operatiopnalized variables here.
+Despite this low threshold for accepting supporting evidence, the ethnographic record (sampled here) does not provide support for every variable from every culture.
+At this culture-level we feel more confident our data provide a estimate of cross-cultural frequency, albeit an inherently noisy, "course-grained" estimation.
+Also, the entire field of ethnology (the branch of anthropology focused on comparing characteristics of populations across the globe) is based on developing comparable and often quantitative measures from primary ethnographic information.
+For example, the Ethnographic Atlas, which provides hundreds of coded quantitative and qualitative variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operationalized variables here.
 Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations.
@@
 However, fundamentally these results provide a systematic view of the ethnographic record on these measures.
-From our results field researchers and theoreticians can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting a few ethngoraphic cases studies.
+From our results field researchers and theoreticians can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting a few ethnographic cases studies.
 
-We now discuss these limitations and points more clearly in the limitations setion. NEED TO DO.
+We now discuss these limitations and points more clearly in the limitations section. NEED TO DO.
 
@@
 
+We now more clearly motivate our discussion of the functions of leaders and link discussions to these sources and ideas.
+
 **The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs.
@@
 Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies?
-I'd be more explicit on this in the introduction.
-For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**
+I'd be more explicit on this in the introduction.**
+
+We now include additional discussion of managerial frameworks of leader behavior and highlight recent approaches that are more conducive to coupling with evolutionary and anthropological/psychological approaches.
+
+<!-- ADDED SOME DISCUSSION IN INTRO. NEED TO FINISH -->
+
+**For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**
 
@@
 
+The reviewer brings up an important, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benifits.
+This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethngoraphic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.
+
+Concerning the quality of *polygynous* and the benefit of *mating*, note that the *mating* benefits includes polygyny, but also includes other mating benefits, such as having a higher quality spouse (but only one).
+Indeed every text recorded coded as providing evidence for the quality of *polygynous* also provides evidence for the leader *mating* benefit, however the inverse is not the case.
+Therefore these variables are in fact distinct.
+
+Another aspect of potential overlap in the data is the quality *wealthy* and the benefit *material resources*.
+Here, note the operationalization of the quality *wealthy* suggests a relative comparison to followers, or a qualifying statement (we have slightly adjusted the phrasing in our table of operationalization).
+The benefit measures however is focused on the transactional nature of resource transfer or 'payment'.
+
+We now discuss these disctintions in more detail in the methods and limitations sections.
+
 **Can you control for phylogeny (e.g.

### 2020-04-06 leadership_across_cultures_contexts.Rmd
Working on response to reviewers and MS text.
Looked at texts with no evidence for status by subsistence in mosaic plot.

Adjusted var names in both script files to match updated data package.

--- 2020-04-03 leadership_across_cultures_contexts.Rmd
@@
 Among many of the ethnographically described small-scale societies, which generally lack multi-level, institutionalized social and political structures, leadership at the community level is typically informal and ephemeral and leaders emerge contextually when demands of coordination or conflict resolution require a managerial presence [@von_rueden_leadership_2015].
-Leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+In the absence of institutionalized organizational structure leaders often take on essential roles in group coordination and collective action, including facilitating conflict resolution within or between autonomous households or kin groups, managing individual investments to group-wide projects, and administering sanctions to individuals who do not appropriately contribute to collective actions, comply with group policies, or seriously violate social norms.
+@glowacki_leadership_2015 draw on empirical, ethnographic, and theoretical models of leader emergence and suggest in small-scale societies institutionalized leadership roles are likely to initially emerge to resolve conflicts between families.
+@roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military fromidability to rival groups and facilitating within-group conflict resolution and cooperation.
+In addition to military prowess, 'big-men' are characterized by, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009][p.104].
+
+Given specific demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
@@
 
-Increased sedentarization and defensibility of resources is often associated with both social stratification and institutionalized managerial roles including stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
+Increased sedentarization and defensibility of resources is often associated with both social stratification and broader institutionalized managerial roles including stronger authority of community leaders and leadership beyond the community level [@johnson_evolution_1987; @johnson_organizational_1982; @kaplan_evolutionary_2009; @ames_hierarchies_1985; @panchanathan_indirect_2004].
 Among horticulturalists and agriculturalists managerial leadership can promote more efficient use of shared resources such as water reserves and cooperative labor [e.g., @von_rueden_leadership_2014].
@@
 In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
-Other public service, nonprofit leaders, and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
+Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
 These positions typically are associated with institutionalized responsibilities and specific functions associated with their roles.
@@
 
-More recently, many managerial scholars have adopted a comparative approach to better understand cultural diversity in leadership systems, drawing on Culturally Endorsed Implicit Leadership Theory (CLT), which suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018].
+More recently, many managerial scholars have adopted integrative and comparative approaches to better understand cultural diversity in leadership systems.
+@behrendt_integrative_2017 draw on empirical results from the managerial literature and suggest the behaviors of leaders across managerial contexts are classifiable as either *task-oriented*, including enhancing understanding, strengthening motivation and facilitating implementation, or *relation-oriented*, including fostering coordination, promoting cooperation and mobilizing resources.
+This framework is intended to provide a more fine-grained and externally valid view of leadership functions, compared to traditional models of transactional leadership, which tend to oversimplify dimensions of leader-follower dynamics.
+More integrated and holistic models have also allowed managerial scholars to develop comparative theoretical frameworks for leader functioning.
+Culturally Endorsed Implicit Leadership Theory (CLT), suggests cultural variation in social norms has the potential to impact follower preferences for leadership styles [@moore_literature_2018].
 The Global Leadership and Organizational Behavior Effectiveness (GLOBE) project is a large-scale international research program aimed at understanding leader effectiveness and cultural variation across large organizations.
@@
 Furthermore, results strongly suggest the fit between a CEOs behavioral profile and the expectations of leadership within a society strongly predict TMT dedication and firm performance.
+By higher-order features of leadership behavior and situating leadership functions within more detailed classification frameworks, insights from managerial studies can more effectively be coupled with approaches from within psychology and anthropology.
 
@@
 In WEIRD settings however, community, organizational, and political leaders often have decision-making authority across multiple domains.
-In the United States, despite a designed seperation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
+In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
 Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage inter-group social dynamics.
@@
 We estimate the cross-cultural frequency in the functions and qualities of leaders and the costs and benefits of leadership for both leaders and followers using an ethnographic database previously built and published (doi.org/10.5281/zenodo.2541999), from which @garfield_evolutionary_2019 "tested" three theoretical models against the ethnographic record.
-The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following a the Outline of Cultural Materials (OCM) coding scheme.
+The electronic Human Relations Area Files (eHRAF) is a digital collection of ethnographies, subject-coded at the paragraph level following the Outline of Cultural Materials (OCM) coding scheme.
 The eHRAF is a subset of the larger HRAF and is continually growing in scope.
@@
 Coding operationalizations of the costs and benefits of leadership were developed *a priori* based on our reading of the evolutionary theoretical literature from anthropology and biology.
-For both followers and leaders, we code if there is evidence of costs or benefits in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
+For both followers and leaders, we code if there is evidence of costs paid or benefits received in the following categories: social status, reduced or increased risk of harm or conflict, social services, territory, mating, fitness, food resources, material resources, and other miscellaneous costs or benefits not captured by these categories.
 See Table \@ref(tab:costbenefitvars).
+There are multiple instances of conceptual overlap between many of our researcher-coded variables, in particular between various qualities and related costs or benefits.
+The operationalization of costs and benefits generaly capture a wider range of evidence than related qualities and focus more on evidence of transaction rather than description.
+We return to this point in the limitations section.
 

### 2020-04-06 analyses2.R

@@ -136,8 +136,8 @@ format_label <- function(lbl){
     leader.costs_resources_other.cost = 'Loss of material resources',
     leader.costs_social.status = 'Reduced social status',
     leader.costs_territory.cost = 'Loss of territory',
-    leader.costs.mating.cost = 'Mating cost',
-    leader.costs.social.services = 'Loss of social services',
+    leader.costs_mating.cost = 'Mating cost',
+    leader.costs_social.services = 'Loss of social services',
     follower.costs_fitness = 'Inclusive fitness cost',
     follower.costs_increased.risk.harm.conflict = 'Increased risk of harm',
     follower.costs_mating = 'Mating cost',
@@ -991,6 +991,48 @@ plot_group_sex <-
   theme_bw(15) + theme(axis.text.y=element_blank())
 plot_group_sex
 
+## Groups by subsistence for high status texts only
+df_groups_status <- 
+  leader_text2[leader_text2$qualities_high.status==0,] %>% 
+  dplyr::select(
+    group.structure2,
+    demo_sex,
+    subsistence
+  ) %>% 
+  dplyr::filter(group.structure2 != 'other') %>% 
+  mutate(
+    demo_sex = factor(demo_sex, levels = c('male', 'female')),
+    group =   factor(
+      group.structure2,
+      levels = c(
+        'residential subgroup',
+        'kin group',
+        'economic group',
+        'religious group',
+        'military group',
+        'political group (community)',
+        'political group (supracommunity)'
+      )
+    ),
+    subsistence = factor(
+      subsistence,
+      levels = c("hunter gatherers",
+                 "pastoralists",
+                 "mixed",
+                 "horticulturalists",
+                 "agriculturalists"
+      )
+    )
+  )
+
+plot_group_subsis_status <-
+  ggplot(df_groups_status) +
+  geom_mosaic(aes(x = product(group, subsistence), fill = group)) +
+  labs(x="", y="", fill = "Group type") +
+  guides(fill = guide_legend(reverse = T)) +
+  theme_bw(15) 
+plot_group_subsis_status
+
 
 # Costs and benefits by group structure type ------------------------------
 se <- function(x) sd(x)/sqrt(length(x))
[…]

### 2020-04-06 initialcompute.R

@@ -282,46 +282,46 @@ var_names <- c(
   "functions_social.functions" = "Misc. social functions",
   "qualities_artistic.performance"     = "Artistic performance",
   "qualities_generous"                 = "Generosity",
-  "qualities.age"                      = "Age",
-  "qualities.attractive"              = "Attractive",
-  "qualities.coercive.authority"      = "Authority",
-  "qualities.culturally.progressive"  = "Culturally progressive",
-  "qualities.favorable.personality"   = "Favorable personality",
-  "qualities.honest"                  = "Honesty",
-  "qualities.ingroup.member"          = "Ingroup member",
-  "qualities.killer"                  = "Killer",
-  "qualities.many.children"           = "Many children",
-  "qualities.physically.strong"       = "Physically formidable",
-  "qualities.prosocial"               = "Prosocial",
-  "qualities.strategic.planner"       = "Strategic planner",
+  "qualities_age"                      = "Age",
+  "qualities_attractive"              = "Attractive",
+  "qualities_coercive.authority"      = "Authority",
+  "qualities_culturally.progressive"  = "Culturally progressive",
+  "qualities_favorable.personality"   = "Favorable personality",
+  "qualities_honest"                  = "Honesty",
+  "qualities_ingroup.member"          = "Ingroup member",
+  "qualities_killer"                  = "Killer",
+  "qualities_many.children"           = "Many children",
+  "qualities_physically.strong"       = "Physically formidable",
+  "qualities_prosocial"               = "Prosocial",
+  "qualities_strategic.planner"       = "Strategic planner",
   "qualities_drug.consumption"     = "Drug consumption",
   "qualities_high.status"            = "High status",
-  "qualities.aggressive"             = "Aggressiveness",
-  "qualities.bravery"               = "Bravery",
-  "qualities.confident"             = "Confidence",
-  "qualities.decisive"              = "Decisiveness/decision-making",
-  "qualities.feared"                = "Feared",
-  "qualities.humble"                = "Humility",
-  "qualities.innovative"            = "Innovative",
-  "qualities.knowlageable.intellect" = "Knowledgeable/intelligent",
-  "qualities.oratory.skill"         = "Oratory skill",
-  "qualities.polygynous"            = "Polygynous",
-  "qualities.social.contacts"       = "Social contacts",
-  "qualities.supernatural"    = "Supernatural",
[…]

### 2020-04-07 analyses2.R
Added min values in text analyses fit

@@ -1455,6 +1455,8 @@ model_words <- function(pred_df, model_score_var, lam = 'lambda.min', title){
   
   if (lam == 'mid'){
     lmda <- m_cv$lambda.min + (m_cv$lambda.1se - m_cv$lambda.min)/2
+  } else if (lam == 'min'){
+    lmda <- m_cv$lambda.min
   } else {
     lmda = lam
   }
@@ -1488,8 +1490,8 @@ model_words <- function(pred_df, model_score_var, lam = 'lambda.min', title){
   return(plot)
 }
 
-highstatus_plot <- model_words(dtm, 'qualities_high.status', lam = 0.012, title = 'Qualities: High status')
+highstatus_plot <- model_words(dtm, 'qualities_high.status', lam = "min", title = 'Qualities: High status')
 highstatus_plot
 
-ben_highstatus_plot <- model_words(dtm, 'leader.benefits_social.status.reputation', lam = 0.01, title = 'Benefits: Status, reputation')
+ben_highstatus_plot <- model_words(dtm, 'leader.benefits_social.status.reputation', lam = "min", title = 'Benefits: Status, reputation')
 ben_highstatus_plot

### 2020-04-07 Response to reviewers.Rmd
Working on reviewer comments and MS.
Added more stop words in text analyses.

@@
 However, fundamentally these results provide a systematic view of the ethnographic record on these measures.
-From our results field researchers and theoreticians can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting a few ethnographic cases studies.
+Our intention is that from our results field researchers and theoreticians interested in these topics can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting and relying on only a handful of ethnographic cases studies.
 
@@
 Perhaps clarify.**
+
+This statement in the introduction is intended to apply across disciplines focused on leadership dynamics and not specifically apply to the methods of the current study.
+The reviewer brings up a well-recieved point, however.
+Any work in developmental psychology on social hierarchy adn leadership would not be subjec to this critique.
+We have not rephrased this claim to better fit with our objectives and outcomes here.
 
@@
 So I'm not sure benefits of followership must outweigh the costs (1414-1415) in same way as for leadership.**
+
+We completely agree.
+This statement was intneded to be in reference to a priori assumptions on the evolution and emergence of leader-follower dynamics, similar to discourse on the conundrum of the evolution of cooperation.
+Though, clearly, in particular circumstances an individual will cooperate when doing so incurs a cost just as followers in particular circumstances will follow when doing so incurs a cost.
+We assume that over evolutionary history and for leader-follower dynamics to be selected for, both leaders and followers would have to recieve benefits on average, over time.
+
+We have revised this statement. CHECK. NEED TO DO.

### 2020-04-07 leadership_across_cultures_contexts.Rmd
Working on reviewer comments and MS.
Added more stop words in text analyses.

@@
 Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019; @van_vugt_selected_2010].
-Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and at all levels of social organization.
+Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and levels of social organization, such as among kin groups, economic groups, and sub-groupings of individuals within residential communities.
 Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have been relatively under-investigated.
-Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is shaped by population-level variation at different levels of social organizations and the costs and benefits leaders and followers incur across distinct leadership systems.
+Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is associated with population-level variation at different levels of social organizations and the costs and benefits leaders and followers incur across distinct leadership systems.
 
@@
 @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military fromidability to rival groups and facilitating within-group conflict resolution and cooperation.
-In addition to military prowess, 'big-men' are characterized by, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009][p.104].
+In addition to military prowess, 'big-men' are characterized by, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
+104].
 
@@
 When strong community leaders emerge, they are most typically high status males who embody the ideal qualities most respected by the group and necessary for successful leadership in the domains in which they have influence [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019].
+There has been increasing interdisciplinary discussion and consensus on the closely related topics of social status, power, authority, rank, prestige, and leadership [e.g., @blader_whats_2014; @cheng_dominance_2019; @kantner_identifying_2010; @von_rueden_roots_2014].
+We draw on discussions by @blader_whats_2014 and @von_rueden_roots_2014 and coneptualize social status as a relative indicator of an individuals social value by the group and their access to contested resources within the group.
+The relationship between leadership (i.e., social influence) and status (i.e., social value) is expected to be colinear [@von_rueden_leadership_2015; @van_vugt_genes_2020], however, if there is causality between the two it is likely to be complex, context-specific, and driven by underlying correlates.
+
 Substantial evidence suggests there is a near universal male-bias in community, political, or group-level leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse non-human social species [@smith_obstacles_2018; @tiger_female_1978].
@@
 
+To suppliment results from the aformentioend methods, we also performed textual analyses...OTHER METHODS. NEED TO DO.
+
 <!-- Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size, and effects of group type, such as political groups versus kin groups.
@@
 <!-- ``` -->
+
+## Disentangleing the relationship between leadership and high status
+
+Figure \@ref(fig:varsupport) identified *high status* as the most frequent leader quality in the ethnographic record, documented in over 30% of text records and in over 90% of cultures.
+This result is neither surprising nor particularly informative.
+However, these data present an oppurtunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
+To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all “informative” words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words).
+In this data matrix the columns are all the unique word stems (terms), rows were the text records (documents), and the values in each cell were the frequencies of each word stem in each text record.
+The coding of *high status* was joined with this term-document matrix.
+This process produced a matrix with `r dim(dtm)[1]` rows with one for each text-record and `r dim(dtm)[2]` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).
+
+We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$).
+Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting.
+The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005].
+We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality.
+Because this outomce is counts of text recrods, we used Poisson errors.
+We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance.
+We then plotted all the non-zero coefficients (Figure \@ref(fig:status-text).
+The most predictive words then epitomize the semantic content of the text records which provide evidence of high status as a quality of leaders.
+
+```{r status-text, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso Poisson regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record.
+Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010].
+Blue values are positive predictors; red values are negative predictors."}
+highstatus_plot
+```
+
+Figure \@ref(fig:status-text) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficents.
+The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:3]`*, and *`r rev(highstatus_plot$data$Word)[4]`*, respectively.
+The words, *`r rev(highstatus_plot$data$Word)[5:7]`* are moderately strong predictors and notable words with low predictive value (including positive and negative predictors) include *authority*, *influence*, *king*, *chief*, and *headman*.
+The word *follow* is the strongest negative predictor and *company* and *economic* are marginal negative predictors.
 
@@
 Followership, on the other hand, involves relinquishing individual decision-making to a leader who might or might not act in the follower's interest.
-The benefits of followership must also outweigh the costs.
+The benefits of followership (and by extension, group living) must also outweigh the costs for leadership to evolve and be maintained.
 We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading or following.

### 2020-04-07 analyses2.R

@@ -1397,6 +1397,9 @@ sd_wordcount <- sd(x)
 library(dplyr)
 data("stop_words")
 stop_words <- rbind(stop_words, list(word='page', lexicon='garfield'))
+stop_words <- rbind(stop_words, list(word='ijaaj', lexicon='garfield'))
+stop_words <- rbind(stop_words, list(word='tadoelako', lexicon='garfield'))
+stop_words <- rbind(stop_words, list(word='zande', lexicon='garfield'))
 
 words2 <-
   words %>%
@@ -1490,8 +1493,8 @@ model_words <- function(pred_df, model_score_var, lam = 'lambda.min', title){
   return(plot)
 }
 
-highstatus_plot <- model_words(dtm, 'qualities_high.status', lam = "min", title = 'Qualities: High status')
+highstatus_plot <- model_words(dtm, 'qualities_high.status', lam = "min", title = 'Leader quality: High status')
 highstatus_plot
 
-ben_highstatus_plot <- model_words(dtm, 'leader.benefits_social.status.reputation', lam = "min", title = 'Benefits: Status, reputation')
-ben_highstatus_plot
+# ben_highstatus_plot <- model_words(dtm, 'leader.benefits_social.status.reputation', lam = "min", title = 'Benefits: Status, reputation')
+# ben_highstatus_plot

### 2020-04-07 Response to reviewers.Rmd
Added description in response to reviewers document.

@@
 
-In the methods section we now include a description of all levels of our *group structure" coding to clearly explain this coding scheme.
-We define supracommunity leaders when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders.
+In the methods section we now include a description of all levels of our *group structure* coding to clearly explain this coding scheme.
+We coded supracommunity leaders in cases when the "group" involves leadership beyond the community level and across multiple residential communities, such as regional political leaders, chiefs in complex chiefdoms, kings, and state-level leaders.
+This termonology was vaugely inspiried by Murdock and Provost's (1971) Political Integration Scale which codes the political integration at the society-level as either automonous local communities, or having one, two, or three levels of juridictional hiuerarchy "above the community"; i.e., if political authority is exclusive to a local village headman, the group would be coded as "autonomous" but if the village and headman are under authority of a chief from a seperate community, the group would be coded as "one level above community", and so forth.
 

### 2020-04-07 initialcompute.R
Commented out some code to speed up script

@@ -88,11 +88,6 @@ leader_cult$d_culture<-leader_cult$c_culture_code
 #Subset culture vars of interest
 culture_vars<-leader_cult[c("d_culture","subsistence","c_cultural_complexity", "settlement_fixity", "pop_density","com_size")]
 
-leader_text2 <-
-  leader_text2 %>% 
-  left_join(text_doc_cultureID) %>% 
-  left_join(culture_vars)
-
 # Add more SCCS variables
 load("sccs.RData")
 sccs<-data.frame(sccs)
@@ -110,7 +105,14 @@ sccs_vars$warfare_freq <- as.numeric(sccs_vars$V1648)
 sccs_vars$warfare_freq <- ifelse(sccs_vars$warfare_freq %in% c(1,20), NA, sccs_vars$warfare_freq)
 sccs_vars$warfare_freq <- 19 - sccs_vars$warfare_freq
 
-leader_text2<-left_join(leader_text2, sccs_vars, by=c("d_culture"="c_culture_code"))
+culture_vars <-
+  culture_vars %>% 
+  left_join(sccs_vars[c('c_culture_code', 'warfare_freq')], by = c('d_culture' = 'c_culture_code'))
+
+leader_text2 <-
+  leader_text2 %>% 
+  left_join(text_doc_cultureID) %>% 
+  left_join(culture_vars)
 
 # Variable support plot data prep --------------------------------------------------
 
@@ -521,21 +523,21 @@ m_pvclust_qual <- pvclust(
 #   parallel = T
 # )
 
-m_pvclust_qual_jaccard <- pvclust(
-  pca_data_qualities2, 
-  method.hclust = 'ward', 
-  method.dist = 'binary', 
-  nboot = 10000,
-  parallel = T
-)
+# m_pvclust_qual_jaccard <- pvclust(
+#   pca_data_qualities2, 
+#   method.hclust = 'ward', 
+#   method.dist = 'binary', 
+#   nboot = 10000,
+#   parallel = T
+# )
 
-m_pvclust_qual_euc <- pvclust(
-  pca_data_qualities2, 
-  method.hclust = 'ward', 
-  method.dist = 'euclidean', 
-  nboot = 10000,
-  parallel = T
-)
+# m_pvclust_qual_euc <- pvclust(
[…]

### 2020-04-07 Response to reviewers.Rmd
Added response on gender of ethnographer

--- 2020-04-06 Response to reviewers.Rmd
@@
 Does the ethnographer matter?
-If so, how much?
-Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population?
+If so, how much?**
+
+**Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population?
 This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**
+
+We determined the gender of almost all authors of all documents included in our study.
+We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership.
+Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
+After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25).
+We now report this in our results section.
 
@@
 
-We greatly appreciate the detailed comments and very useful suggestions.
+We greatly appreciate the reviewers detailed comments and very useful suggestions.
 
@@
 Also, I realize the paper is decidedly exploratory/descriptive, but did the authors have any hypotheses about differences between WEIRD and traditional societies?**
-We agree our initial submisson did not sufficiently accomplish this goal.
+
+We agree our initial submisson did not sufficently accomplish this goal.
 We have now added summarizing, concluding statements on the qualities and functions of leaders in WEIRD and non-WEIRD settings in the introduction.
@@
 How does these records compare RE gender, group structure, or other qualities (maybe), or the most frequent functions, i.e., try to charachterize non-high status leaders.
-
-OR, better idea (?), use LASSO technique from paper one, to identify words predictive of high status present or not
 -->
@@
 
-We added a few additional details in the methods section on the eHRAF and ethnographic sample and now more clearly direct readers to the original publication from which the database was built, which includes additional details.
+We added a few additional details in the methods section on the eHRAF and ethngraphic sample and now more clearly direct readers to the original pubication from which the database was built, which includes additional details.
 
-We also include additional details in the results and supplementary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.
+We also include additional details in the results and supplimentary information on the corpus of ethnographic documents which provide the sample of ethnographic paragraphs. NEED TO DO.
 
@@
 For one, this is partly our motivation for representing the percentage of cultures which provide at least one text record supporting each of the operationalized variables.
-Despite this low threshold for accepting supporting evidence, the ethnographic record (sampled here) does not provide support for every variable from every culture.
-At this culture-level we feel more confident our data provide a estimate of cross-cultural frequency, albeit an inherently noisy, "course-grained" estimation.
-Also, the entire field of ethnology (the branch of anthropology focused on comparing characteristics of populations across the globe) is based on developing comparable and often quantitative measures from primary ethnographic information.
-For example, the Ethnographic Atlas, which provides hundreds of coded quantitative and qualitative variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operationalized variables here.
+Despite this low threshold for acceptign supporting evidence, the ethnogrtaphic record (sampled here) does not provide support for every variable from every culture.
+At this culture-level, we feel more confident our data provide a estiamte of cross-cultural frequency, albeit an inherently noisy estimation.
+Also, the entire field of ethnology (the branch of anthropology focused on comparing charachteristics of populations across the globe), is based on developing comparable and often quantitaive measures from primary ethnogrpahic information.
+For example, the Ethnographic Atlas, which provides hundreds of coded quantitaive and qualitive variables on thousands of cultures and the Standard Cross-Cultural Sample, which provides thousands of coded quantitative and qualitative variables on 186 cultures, were developed in essentially the same way we have developed our operatiopnalized variables here.
 Developing and contrasting coded measures from primary ethnographic documents is not without limitations, but does represent (perhaps) the only method to systematically and empirically compare measures across a large sample of cultures/populations.
+
 At the text-record level drawing strong conclusions about the actual frequency of traits is slightly more problematic.
-However, fundamentally these results provide a systematic view of the ethnographic record on these measures.
-From our results field researchers and theoreticians can draw inference from a more comprehensive view of human cultural diversity rather than haphazardly self-selecting a few ethnographic cases studies.
+However, fundamentally these results provide a systematic view of the ethnographic record.
+From these results field researchers and theoreticians can draw inference from a more comprehensive view of human culturla diversity rather than haphazardly self-selecting a few ethngoraphic cases studies.
 
-We now discuss these limitations and points more clearly in the limitations section. NEED TO DO.
+We now discuss these limitations and points more clearly in the limitations setion. NEED TO DO.
 
@@
 I have a number of suggestions and comments that may improve the paper, none highly critical:**
-
-We greatly appreciate the detailed comments and very useful suggestions.
 
@@
 
-We now more clearly motivate our discussion of the functions of leaders and link discussions to these sources and ideas.
-
 **The Wofford study (lines 214-233) and GLOBE study reminds me of the vast array of descriptive models in leadership/managerial studies- transactional vs.
@@
 Do bureaucracies, formal legal systems, formal education-based skills, and large inequalities in remuneration limit universality of descriptive models from leadership/managerial studies?
-I'd be more explicit on this in the introduction.**
-
-We now include additional discussion of managerial frameworks of leader behavior and highlight recent approaches that are more conducive to coupling with evolutionary and anthropological/psychological approaches.
-
-<!-- ADDED SOME DISCUSSION IN INTRO. NEED TO FINISH -->
-
-**For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**
+I'd be more explicit on this in the introduction.
+For the discussion section: how does your own descriptive modeling of leadership in traditional societies agree or not with Wofford or other models from large-scale societies?**
 
@@
 2018) that finds similar traits predict leadership for men and women, implying its not gender per se that gives men a leadership advantage in politics.**
-
-We now discuss sex-differences in community and political leadership, and refernce in greater deatils (albeit breif) the universal male bias in community leadership in the introuction and cite these works.
 
@@
 **Many of the variables coded as qualities of leaders may also fall under leader costs and benefits, e.g. polygyny, physical health, wealthy. Were variables like these analyzed in both cases?**
-
-The reviewer brings up an important, i.e., the conceptual and operational overlap in many study variables, and specifically between various qualities and benifits.
-This is partly a consequence of our methodology, given the coding of qualities was purely driven by the ethngoraphic record whereas our coding of costs and benefits was sketched out a priori, based on evolutionary theory.
-
-Concerning the quality of *polygynous* and the benefit of *mating*, note that the *mating* benefits includes polygyny, but also includes other mating benefits, such as having a higher quality spouse (but only one).
-Indeed every text recorded coded as providing evidence for the quality of *polygynous* also provides evidence for the leader *mating* benefit, however the inverse is not the case.
-Therefore these variables are in fact distinct.
-
-Another aspect of potential overlap in the data is the quality *wealthy* and the benefit *material resources*.
-Here, note the operationalization of the quality *wealthy* suggests a relative comparison to followers, or a qualifying statement (we have slightly adjusted the phrasing in our table of operationalization).
-The benefit measures however is focused on the transactional nature of resource transfer or 'payment'.
-
-We now discuss these disctintions in more detail in the methods and limitations sections.
 

### 2020-04-07 analyses2.R
New code for research effort and new cluster vars

@@ -14,6 +14,7 @@ library(leadershipdata)
 library(tidyverse)
 library(tidytext)
 library(forcats)
+library(stringr)
 library(NMF)
 library(pvclust)
 library(dendextend)
@@ -23,14 +24,21 @@ library(car)
 library(visreg)
 library(effects)
 library(lme4)
+library(glmmTMB)
 library(patchwork)
-library(pvclust)
 library(magrittr)
 library(modelr)
 library(ggridges)
 library(RColorBrewer)
 library(ggmosaic)
 library(proxy)
+library(readxl)
+library(broom)
+library(broom.mixed)
+library(ggrepel)
+library(viridis)
+# library(mgcv)
+
 # Load precomputed objects ------------------------------------------------
 
 # First run initialcompute.R, which takes ~90 minutes on my machine,
@@ -66,25 +74,6 @@ logisticPCA_loadings_plot <- function(m, data){
 
 # Recode variables --------------------------------------------------------
 
-leader_text2$com_size <-
-  ordered(
-    leader_text2$com_size,
-    levels = c("< 99", "100-199", "200-399", "400-999", "> 1,000")
-  )
-
-leader_text2$pop_density <-
-  ordered(
-    leader_text2$pop_density,
-    levels = c(
-      "1 or less person / 1-5 sq. mile",
-      "1-5 persons / sq. mile",
-      "1-25 persons / sq. mile",
-      "26-100 persons / sq. mile",
-      "101-500 persons / sq. mile",
-      "over 500 persons / sq. mile"
-    )
-  )
-
 all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
 
 # Variable support plots --------------------------------------------------
@@ -107,47 +96,48 @@ plot.variable.support
 
 # Leader/follower benefits and costs
 
-format_label <- function(lbl){
-  relabel <- c(
-    leader.benefits_fitness = 'Inclusive fitness benefit',
-    leader.benefits_mating = 'Mating benefit',
-    leader.benefits_other = 'Misc. non-material benefit',
-    leader.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
-    leader.benefits_resource_food = 'Food benefit',
-    leader.benefits_resource_other = 'Material resources',
[…]

### 2020-04-09 analyses2.R
Commented out stop words.

@@ -1397,9 +1397,9 @@ sd_wordcount <- sd(x)
 library(dplyr)
 data("stop_words")
 stop_words <- rbind(stop_words, list(word='page', lexicon='garfield'))
-stop_words <- rbind(stop_words, list(word='ijaaj', lexicon='garfield'))
-stop_words <- rbind(stop_words, list(word='tadoelako', lexicon='garfield'))
-stop_words <- rbind(stop_words, list(word='zande', lexicon='garfield'))
+#stop_words <- rbind(stop_words, list(word='ijaaj', lexicon='garfield'))
+#stop_words <- rbind(stop_words, list(word='tadoelako', lexicon='garfield'))
+#stop_words <- rbind(stop_words, list(word='zande', lexicon='garfield'))
 
 words2 <-
   words %>%

### 2020-04-09 Response to reviewers.Rmd
Changes to all docs. Using margins package. More responses to reviewers. Removed vars with v. little evidence. Supplementary info.

@@
 
+Although we determined the gender of almost all ethnographers (see below), we were not able to confirm the national origins of most.
+
+We now briefly describe the random effects for ethnographers. [need to do]
+
 **Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population?
@@
 
+In the original version of the ms we did not describe our numerous efforts at dimension reduction, including standard PCA, various flavors of logistic PCA, non-negative matrix factorization (NMF), and various flavors of cluster analysis, and we regressed the resulting components against our various culture-level variables.
+These efforts did not yield any convincing "stucture" in our data or results beyond what we reported.
+We now briefly describe these efforts so as to better contexualize the results that we do report.
+
+We now additionally create variables based on the clusters we did find, and investigate their relationship to culture-level variables.
+As before, we find little convincing evidence of meaningful patterns, which we now briefly report in the SI.
+
 **2.	Why not take either some of your original codes or some of these higher-level codes (from the above cluster analysis) and regress them on societal-level features related to subsistence type (foragers, pastoralist, etc.
@@
 
-**3.	Be clearer about how you are accounting for research effort in each society.
-For example, are you controlling for the total number of words or pages for each population?
-This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**
+See above.
+
+**3.	Be clearer about how you are accounting for research effort in each society. For example, are you controlling for the total number of words or pages for each population?**
+
+Total number of pages of ethnography for each culture. Results now reported.
+
+**This seems very important since we want to make sure that the variation observed isn’t due to some populations getting more or less total study effort or writing effort.**
+
+We investigated three possible sources of bias: total number of pages of ethnography, year of publication of each document, and the presence or absence of a female author/co-author for each document.
+We found evidence for bias in reporting of evidence for 9 of 105 [get new value] variables based on year of publication only.
+There was no convincing evidene of bias in any variable based on the total number of pages of ethnography or ethographer gender.
+Results now reported.
 
 **4.	Did you control for observation year? This might be too highly correlated with ethnographer identity, but worth checking**
+
+Yes. See above.
 
@@
 
+We agree that worries persist about the putative independence of the PSF. We now investigate continental effects for each variable. We find xxx, yyy, zzz.
+
 **6.	Analyze the degree to which ‘researcher efforts’ (say measured by total words on that society and/or number of paragraphs on leadership) predict the appearance of a code.
@@
 
+As discussed above, we investigated the possible effects of total number of pages of ethnography for each culture on each variable. As we now report, we found no significant effects.
+
 **We recommend making liberal use of supplemental materials.
 Some of the analyses may not yield findings of sufficient interest to warrant placement in the main text, but some readers may want to have a look at them, so put them in the supplemental and briefly mention them in the main.**
+
+We have expanded our supplemental materials, xx, yy, zz.
 
@@
 As mentioned by R1, this likely involves returning to the GLOBE data and putting it together with your findings more directly.**
+
+We agree we need to better situate our study within the theme of the special issue.
+We interpret "Beyond WEIRD" in the following way: The original WEIRD paper was a needed wake-up call for psychology.
+However, we now worry that many psychologists are operationalizing WEIRD as a dichotomy: WEIRD and non-WEIRD.
+This is obviously a mistake, as there is enormous variation among the non-WEIRD.
+Our data set focuses primarily on traditional societies, and thus speaks to variation among the non-WEIRD, a point we now emphasize.
+At the same time, as Edward Said famously argued, anthropologists tend to make the opposite mistake: exoticizing the "Other".
+We hope we have struck the right balance in identifying important common themes and patterns across most of the populations in our data, while doing justice to their differences.
 
@@
 
+This is a good idea. We now include a short section on our "status" variable, and what distinguishes leadership records with evidence for social status from those without such evidence.
+
 **Expanding the dataset: the natural next step with this project is to use this small sample of hand coded categories to create an automated approach for all of the eHRAF, perhaps based on key words.
@@
 We offer this as a suggestion, but realize that it’s a lot to ask in this context.**
+
+This is a good suggestion for future research, but goes beyond remit of the current paper.
 

### 2020-04-09 leadership_across_cultures_contexts.Rmd
Changes to all docs. Using margins package. More responses to reviewers. Removed vars with v. little evidence. Supplementary info.

--- 2020-04-04 leadership_across_cultures_contexts.Rmd
@@
 The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture.
-The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
+Our search yielded `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures.
+The number of text records per culture ranged from `r min(dfwc$number_leader_records, na.rm=T)` to `r max(dfwc$number_leader_records, na.rm=T)`, with a median of `r median(dfwc$number_leader_records, na.rm=T)`, a mean of `r signif(mean(dfwc$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(dfwc$number_leader_records, na.rm=T), 3)`.
+The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
 Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.
@@
 
+## Possible sources of bias
+
+```{r tabauthorgender}
+unique_authors <- unique(authorship[c('author_ID', 'author_gender')])
+tab_author_gender <- table(unique_authors$author_gender)
+mm_coauthorSUM <- summary(mm_coauthor)
+mm_coauthorP <- mm_coauthorSUM$coefficients['female_coauthorTRUE', 'Pr(>|z|)']
+```
+
+We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
+Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)`.
+There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
+Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author.
+Total pages of ethnography on each culture in the PSF ranged from `r min(dfwc[['N Pages eHRAF']])` to `r as.integer(max(dfwc[['N Pages eHRAF']]))`, with a median of `r median(dfwc[['N Pages eHRAF']])`.
+
+We fit `r length(all_study_vars)` logistic mixed effects models with each binary study variable as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures.
+We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995].
+After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models.
+More recent publications were more likely to report evidence for Oratory skill, Group representative, Group determination/cohesiveness, and Knowledgeable/intelligent, and less likely to report evidence for leaders imposing costs (Punishment, Killer) or receiving benefits (Social services, Food, and Reduced risk of harm).
+See Figure \@ref(fig:pubdate).
+See supplementary information for details.
+
+We then tested if text records with a female author or co-author were more likely to discuss female leaders.
+Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
+Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
+After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
+
 ## Variable distributions
@@
 
-```{r cluster, fig.cap="Cluster analysis of leader quality variables (left) and leader function variables (right).
-Distances are $1 - cor$.
+```{r cluster, fig.cap="Cluster analysis of leader quality variables (A) and leader function variables (B).
+Distances were $1 - cor$.
 Ward agglomeration method.
@@
 <!-- ``` -->
-
-## Effects of document meta-data on leadership patterns
-
-```{r tabauthorgender}
-unique_authors <- unique(authorship[c('author_ID', 'author_gender')])
-tab_author_gender <- table(unique_authors$author_gender)
-mm_coauthorSUM <- summary(mm_coauthor)
-mm_coauthorP <- mm_coauthorSUM$coefficients['female_coauthorTRUE', 'Pr(>|z|)']
-```
-
-We explored whether document features, such as date of publication and gender of authors, influenced reporting of leader functions and qualities.
-There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine, some of which were co-authors.
-We then tested if text records with a female author or co-author were more likely to discuss female leaders.
-Of the `r sum(tab_coauthor['FALSE',])` text records with no female author or co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
-Of the `r sum(tab_coauthor['TRUE',])` text records with a female author or co-author, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
-After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
 
@@
 
+## Possible ethnographic or publication biases
+
 ## Limitations
@@
 
+We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` variables: publication date, female vs.
+male-only authorship, and variation in the total number of pages of ethnography for each culture.
+For each binary study variable, we fit a logistic mixed effects model with that variable as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures.
+We fit two sets of models.
+The first univariate set fit separate models for each study variable and each bias variable (total number of models = 105x3 = 315).
+The second multivariate set included all bias variables in each model (total number of models = 105).
+All p-values were corrected for a FDR of 5%.
+After correction, the coefficients of female authorship and pages of ethnography were not significant in any model.
+Seven variables had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model.
+See Figure \@ref(fig:pubdate).
+
+```{r pubdate, fig.cap="The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs.
+Multivariate values are  controlling for female authorship and number of pages of ethonography for each culture.
+All models had random intercepts for ethnographers nested within cultures.
+Models were fit for each of the 105 variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant.
+Publication date was centered and standardized by 1 SD.
+Each dot is one coefficient (converted from log odds to odds ratios).
+Bars are 95% CI."}
+plot_pubdate
+```
+
+## Feature models
+
+```{r features, fig.width=10, fig.cap="Marginal effects of Poission mixed effects regression models of Feature variables by mode of subsistence.
+Only models whose subsistence coefficient had a corrected p-value < 0.05.
+Base level: agriculturalists.
+See text for details."}
+
+x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))
+
+```
+
 # References

### 2020-04-09 analyses2.R

@@ -37,6 +37,8 @@ library(broom)
 library(broom.mixed)
 library(ggrepel)
 library(viridis)
+library(ggcorrplot)
+library(margins)
 # library(mgcv)
 
 # Load precomputed objects ------------------------------------------------
@@ -123,8 +125,8 @@ cost_benefit_dict <- c(
   leader.costs_resources_other.cost = 'Loss of material resources',
   leader.costs_social.status = 'Reduced social status',
   leader.costs_territory.cost = 'Loss of territory',
-  leader.costs.mating.cost = 'Mating cost',
-  leader.costs.social.services = 'Loss of social services',
+  leader.costs_mating.cost = 'Mating cost',
+  leader.costs_social.services = 'Loss of social services',
   follower.costs_fitness = 'Inclusive fitness cost',
   follower.costs_increased.risk.harm.conflict = 'Increased risk of harm',
   follower.costs_mating = 'Mating cost',
@@ -1238,13 +1240,13 @@ df_fun <-
     Variable = fct_reorder(Variable, Intercept)
   )
 
-ggplot(df_fun, aes(Intercept, Variable)) + 
-  geom_point() +
-  geom_segment(aes(x = lower, xend = upper, y = Variable, yend = Variable), alpha=0.5) +
-  # geom_errorbarh(aes(xmin = lower, xmax = upper), alpha = 0.5) + 
-  scale_x_log10() +
-  labs(x = '\nRate of evidence for per 1 SD of ethnographic pages', y = "") +
-  theme_minimal(15)
+# ggplot(df_fun, aes(Intercept, Variable)) + 
+#   geom_point() +
+#   geom_segment(aes(x = lower, xend = upper, y = Variable, yend = Variable), alpha=0.5) +
+#   # geom_errorbarh(aes(xmin = lower, xmax = upper), alpha = 0.5) + 
+#   scale_x_log10() +
+#   labs(x = '\nRate of evidence for per 1 SD of ethnographic pages', y = "") +
+#   theme_minimal(15)
 
 # Nice labels for all vars
 x <- str_to_title(str_split_n(names(cost_benefit_dict), '\\.', n=1))
@@ -1395,40 +1397,70 @@ df_pages_uni <-
 
 # Use clusters from pvclust to create higher-order vars
 
-clust_qual_vars <- 
+## These are out-of-date after some data cleaning
+#
[…]

### 2020-04-09 initialcompute.R

@@ -42,6 +42,17 @@ neg1to0 <- function(df){
     )
 }
 
+
+# Remove variables with little evidence -----------------------------------
+
+evidence <-
+  leader_text2[-c(1:3)] %>%
+  select_if(is.numeric) %>%
+  map_dbl(~sum(.x, na.rm = T))
+
+low_evidence <- names(evidence[evidence<4])
+leader_text2 <- dplyr::select(leader_text2, -any_of(low_evidence))
+
 # Create vectors of variable names by type --------------------------------
 
 lt2vars <- names(leader_text2)
@@ -284,46 +295,50 @@ var_names <- c(
   "functions_social.functions" = "Misc. social functions",
   "qualities_artistic.performance"     = "Artistic performance",
   "qualities_generous"                 = "Generosity",
-  "qualities.age"                      = "Age",
-  "qualities.attractive"              = "Attractive",
-  "qualities.coercive.authority"      = "Authority",
-  "qualities.culturally.progressive"  = "Culturally progressive",
-  "qualities.favorable.personality"   = "Favorable personality",
-  "qualities.honest"                  = "Honesty",
-  "qualities.ingroup.member"          = "Ingroup member",
-  "qualities.killer"                  = "Killer",
-  "qualities.many.children"           = "Many children",
-  "qualities.physically.strong"       = "Physically formidable",
-  "qualities.prosocial"               = "Prosocial",
-  "qualities.strategic.planner"       = "Strategic planner",
+  "qualities_age"                      = "Age",
+  "qualities_attractive"              = "Attractive",
+  "qualities_coercive.authority"      = "Coercive authority",
+  "qualities_culturally.progressive"  = "Culturally progressive",
+  "qualities_favorable.personality"   = "Favorable personality",
+  "qualities_honest"                  = "Honesty",
+  "qualities_ingroup.member"          = "Ingroup member",
+  "qualities_killer"                  = "Killer",
+  "qualities_many.children"           = "Many children",
+  "qualities_physically.strong"       = "Physically formidable",
+  "qualities_prosocial"               = "Prosocial",
+  "qualities_strategic.planner"       = "Strategic planner",
   "qualities_drug.consumption"     = "Drug consumption",
   "qualities_high.status"            = "High status",
-  "qualities.aggressive"             = "Aggressiveness",
-  "qualities.bravery"               = "Bravery",
-  "qualities.confident"             = "Confidence",
[…]

### 2020-04-09 Response to reviewers.Rmd
Resolving merge conflicts across 2 .R and 2 .Rmd files. No issues.

--- 2020-04-07 Response to reviewers.Rmd
@@
 Does the ethnographer matter?
-If so, how much?
-Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population?
+If so, how much?**
+
+Although we determined the gender of almost all ethnographers (see below), we were not able to confirm the national origins of most.
+
+We now briefly describe the random effects for ethnographers. [need to do]
+
+**Further, and we realize there may not be sufficient power for this, but does the gender of the ethnographer predict ‘seeing’ female leaders in a population?
 This can (potentially) speak quantitatively to some of the qualitative discussion about limitations and biases in the approach.**
+
+We determined the gender of almost all authors of all documents included in our study.
+We found that of the 884 text records with no female author or co-author, 21 (2.4%) reported female leadership.
+Of the 225 text records with a female author or co-author, 20 (8.9%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
+After accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, however, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = 3.5), it was not statistically significant (p = 0.25).
+We now report this in our results section.
 

### 2020-04-09 leadership_across_cultures_contexts.Rmd
Resolving merge conflicts across 2 .R and 2 .Rmd files. No issues.

--- 2020-04-07 leadership_across_cultures_contexts.Rmd
@@
 Whereas @garfield_evolutionary_2019, documented supporting evidence for measures of leadership in reference to influential theoretical models, our goals here are to let the ethnographic record "speak for itself" in understanding variation in the dimensions of leadership we are interested in, i.e., our methods and analyses are exploratory and not strictly theoretically driven.
-We draw on the same database of 1212 ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
+We draw on the same database of `r nrow(leader_text2)` ethnographic paragraphs (termed text records) extracted from the 60-culture Probability Sample Files (PSF) of the eHRAF database [@naroll_proposed_1967].
+The PSF aims to mitigate Galton's problem by randomly sampling of one culture with high quality ethographic coverage from each of 60 culture regions.
+It comprises over `r trunc(sum(dfwc['N Pages eHRAF'])/1000)`K pages of digitized text.
+
 The *leadershipdata* database was designed to capture a wide sample of the ethnographic record of leadership across social contexts and was built using an advanced search including 16 OCM codes related to leadership seven key words describing leaders or followers.
@@
 First ZG read a subset of text records and developed initial variable lists and operationalizations of the qualities and functions in those text records.
-Then, KS coded all 1212 text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
+Then, KS coded all `r nrow(leader_text2)` text records (in a random order) for the operationalized qualities, functions, cost, and benefits, while also creating new variables and operationalizations for additional qualities and functions encountered.
 After coding half of the random sample KS and ZG revised the operational definitions to produce the final list of variables in Table \@ref(tab:qualfuncvars).
-ZG then followed the same process of coding the entire sample of 1212 text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phratries, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
+ZG then followed the same process of coding the entire sample of `r nrow(leader_text2)` text records for the `r length(c(quality_vars, function_vars, leader_cost_vars, leader_benefit_vars, follower_benefit_vars, follower_cost_vars))` operationalized variables, as well as a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflcit; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level fo the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders.
 ZG and KS then discussed every instance of disagreement between their respective coding to produce a consensus matrix, which is used in the analyses here and added to the published data repository (a data frame named *leader_text2*).
@@
 
-The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provide ethnographic texts for that culture.
-The `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
+The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture.
+Our search yielded `r nrow(leader_text2)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(leader_text2$d_culture))` cultures.
+The number of text records per culture ranged from `r min(dfwc$number_leader_records, na.rm=T)` to `r max(dfwc$number_leader_records, na.rm=T)`, with a median of `r median(dfwc$number_leader_records, na.rm=T)`, a mean of `r signif(mean(dfwc$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(dfwc$number_leader_records, na.rm=T), 3)`.
+The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`.
 Of the `r nrow(leader_text2)` text records, `r final_record_count` had evidence for at least one variable in the study.
+
+```{r corpus}
+cor_pages_tr <- cor.test(~number_leader_records + `N Pages eHRAF`, dfwc)
+```
+
+There was a weak positive relationship between the number of leader text records we retrieved from each culture and the total number of pages of ethnography for each culture in eHRAF (r = `r signif(cor_pages_tr$estimate, 2)`, p = `r signif(cor_pages_tr$p.value, 2)`).
+A loess regression suggested that the number of records initially increased with the total number of pages of ethnography, plateauing at about 35 records around 4500 pages (see Figure \@ref(fig:eHRAFpages)).
 
@@
 
+# Supplementary information
+
+```{r eHRAFpages, fig.height=6, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF.
+Each dot is one culture.
+Line fit by loess regression."}
+plot_pages_tr
+```
+
+We explored three possible sources of bias in discovery of evidence for each of our `r length(all_study_vars)` variables: publication date, female vs.
+male-only authorship, and variation in the total number of pages of ethnography for each culture.
+For each binary study variable, we fit a logistic mixed effects model with that variable as the outcome, the bias variables as fixed effects, and random intercepts for ethnographers nested within cultures.
+We fit two sets of models.
+The first univariate set fit separate models for each study variable and each bias variable (total number of models = 105x3 = 315).
+The second multivariate set included all bias variables in each model (total number of models = 105).
+All p-values were corrected for a FDR of 5%.
+After correction, the coefficients of female authorship and pages of ethnography were not significant in any model.
+Seven variables had significant associations with year of publication in both the univariate or multivariate sets (with nearly identical coefficients), but Group determination/cohesiveness was only significant in the univariate model, and Leader social service benefits was only significant in the multivariate model.
+See Figure \@ref(fig:pubdate).
+
+```{r pubdate, fig.cap="The effect of publication date on the probability of finding evidence for leader qualities, functions, benefits, and costs.
+Multivariate values are  controlling for female authorship and number of pages of ethonography for each culture.
+All models had random intercepts for ethnographers nested within cultures.
+Models were fit for each of the 105 variables as outcomes, and after adjusing p-values for a false discovery rate of 5%, the effect of publication date on these outcome variables remained significant.
+Publication date was centered and standardized by 1 SD.
+Each dot is one coefficient (converted from log odds to odds ratios).
+Bars are 95% CI."}
+plot_pubdate
+```
+
+## Feature models
+
+```{r features, fig.width=10, fig.cap="Marginal effects of Poission mixed effects regression models of Feature variables by mode of subsistence.
+Only models whose subsistence coefficient had a corrected p-value < 0.05.
+Base level: agriculturalists.
+See text for details."}
+
+x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))
+
+```
+
 # References

### 2020-04-09 analyses2.R

@@ -32,7 +32,6 @@ library(ggridges)
 library(RColorBrewer)
 library(ggmosaic)
 library(proxy)
-<<<<<<< HEAD
 library(readxl)
 library(broom)
 library(broom.mixed)
@@ -41,10 +40,8 @@ library(viridis)
 library(ggcorrplot)
 library(margins)
 # library(mgcv)
-
-=======
 library(stringr)
->>>>>>> befd3dc200c893ed112a9f99a9a4ebb37aff18fc
+
 # Load precomputed objects ------------------------------------------------
 
 # First run initialcompute.R, which takes ~90 minutes on my machine,
@@ -143,9 +140,9 @@ cost_benefit_dict <- c(
 )
 
 format_label <- function(lbl){
-<<<<<<< HEAD
+
   return(cost_benefit_dict[lbl])
-=======
+
   relabel <- c(
     leader.benefits_fitness = 'Inclusive fitness benefit',
     leader.benefits_mating = 'Mating benefit',
@@ -186,7 +183,7 @@ format_label <- function(lbl){
     follower.costs_territory = 'Loss of territory'
   )
   return(relabel[lbl])
->>>>>>> befd3dc200c893ed112a9f99a9a4ebb37aff18fc
+
 }
 
 plot.variable.support_costs_benefits = ggplot(d_melt_cb, aes(value, Variable, xmin=y_negse, xmax=y_se, colour=Type, shape=Type)) + 
@@ -1255,7 +1252,7 @@ mm_coauthor <- glmer(
 )
 mm_coauthorOR <- exp(fixef(mm_coauthor))[[2]]
 
-<<<<<<< HEAD
+
 # Research effort ---------------------------------------------------------
 
 # number of text records for each culture
@@ -1933,4 +1930,4 @@ highstatus_plot
 
 # ben_highstatus_plot <- model_words(dtm, 'leader.benefits_social.status.reputation', lam = "min", title = 'Benefits: Status, reputation')
 # ben_highstatus_plot
->>>>>>> befd3dc200c893ed112a9f99a9a4ebb37aff18fc
+

### 2020-04-09 initialcompute.R

@@ -297,11 +297,7 @@ var_names <- c(
   "qualities_generous"                 = "Generosity",
   "qualities_age"                      = "Age",
   "qualities_attractive"              = "Attractive",
-<<<<<<< HEAD
   "qualities_coercive.authority"      = "Coercive authority",
-=======
-  "qualities_coercive.authority"      = "Authority",
->>>>>>> befd3dc200c893ed112a9f99a9a4ebb37aff18fc
   "qualities_culturally.progressive"  = "Culturally progressive",
   "qualities_favorable.personality"   = "Favorable personality",
   "qualities_honest"                  = "Honesty",
@@ -338,15 +334,11 @@ var_names <- c(
   "qualities_physical.health"        = "Physical health",
   "qualities_proper.behavior"        = "Proper behavior",
   "qualities_strategic.nepotism"     = "Strategic nepotism",
-<<<<<<< HEAD
   "qualities_xenophobic"   = "Xenophobia",
   "qualities_antihonest" = "Dishonest",
   "qualities_antifairness" = "Unfair",
   "qualities_antidrug.consumption" = "No drug consumption",
   "qualities_anticoercive.authority" = "No coercive authority"
-=======
-  "qualities_xenophobic"   = "Xenophobia"
->>>>>>> befd3dc200c893ed112a9f99a9a4ebb37aff18fc
 )
 
 d_melt$Variable <- var_names[d_melt$Variable]

### 2020-04-10 analyses2.R
Deleted residual merge conflict marker

@@ -1646,7 +1646,7 @@ feature_discoveries <-
 # Looking for negative values ---------------------------------------------
 
 x <- map_int(all_data2[all_study_vars], ~ sum(.x < 0))
-=======
+
 library(rstanarm)
 library(bayesplot)
 options(mc.cores = parallel::detectCores())

### 2020-04-12 leadership_across_cultures_contexts.Rmd
Mostly work on the exploratory analyses

@@
 There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
-We used hierarchical cluster analysis on our quality and function variables.
+We used several methods to explore lower dimensional representations of these data that might reveal informative structure, including logistic PCA, non-negative matrix factorization (NMF), and agglomerative hierarchical cluster analysis.
+Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
+NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$.
+NMF estimates $W$ and $H$ by minimizing a measure of distance (typically the Frobenius norm or the Kullback-Leibler divergence) between $X$ and $WH$ [@gaujoux2010].
+Agglomerative hierarchical cluster starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
 We computed distance between our binary vectors using $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and agglomerated using the Ward algorithm.
@@
 
-## Cluster analyses
+## When do leaders also have high status?
 
+```{r highstatuscalc}
+statusFracTxt <- d_melt$value[d_melt$Variable == 'High status' & d_melt$Type == 'Extracts']
+statusPercTxt <- signif(100*statusFracTxt, 3)
+
+statusFracCult <- d_melt$value[d_melt$Variable == 'High status' & d_melt$Type == 'Cultures']
+statusPercCult <- signif(100*statusFracCult, 3)
+
+```
+
+*High status* was the most frequent leader quality in the ethnographic record, with `r statusPercTxt`% of text records and `r statusPercCult`% cultures containing evidence that leaders were of high social status.
+To explore factors that distinguished these records from those with no evidence for status, we first created a document-term matrix (DTM) of all "informative" words in our corpus of texts and the frequency with which they occured in each text record (see SI for details).
+We then fit an elasticnet logistic regression model (with lasso penalty) of high status (0/1) as a function of the frequencies of all `r ncol(leader_dtm)-1` words.
+Words that were strong positive predictors epitomized the semantic content of the text records which provide evidence of high status as a quality of leaders.
+These included *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
+See Figure \@ref(fig:highstatusplot).
+
+## Dimension reduction
+
+Our qualities and functions data comprised sparse $`r nrow(leader_text)` \times `r length(quality_vars)`$ and $`r nrow(leader_text)` \times `r length(function_vars)`$ matrices of binary data, respectively.
+
+Most data reduction methods did not find strong evidence of structure in the data.
+Here we report the cluster analysis of variables that used the correlation distance metric ($1 - cor$), and the Ward agglomeration algorithm (for results from the other methods, see the SI).
 Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables.
@@
 <!-- ``` -->
-
-## Disentangleing the relationship between leadership and high status
-
-Figure \@ref(fig:varsupport) identified *high status* as the most frequent leader quality in the ethnographic record, documented in over 30% of text records and in over 90% of cultures.
-This result is neither surprising nor particularly informative.
-However, these data present an oppurtunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
-To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all “informative” words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words).
-In this data matrix the columns are all the unique word stems (terms), rows were the text records (documents), and the values in each cell were the frequencies of each word stem in each text record.
-The coding of *high status* was joined with this term-document matrix.
-This process produced a matrix with `r dim(dtm)[1]` rows with one for each text-record and `r dim(dtm)[2]` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).
-
-We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$).
-Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting.
-The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005].
-We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality.
-Because this outomce is counts of text recrods, we used Poisson errors.
-We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance.
-We then plotted all the non-zero coefficients (Figure \@ref(fig:status-text).
-The most predictive words then epitomize the semantic content of the text records which provide evidence of high status as a quality of leaders.
-
-```{r status-text, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso Poisson regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record.
-Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010].
-Blue values are positive predictors; red values are negative predictors."}
-highstatus_plot
-```
-
-Figure \@ref(fig:status-text) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficents.
-The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:3]`*, and *`r rev(highstatus_plot$data$Word)[4]`*, respectively.
-The words, *`r rev(highstatus_plot$data$Word)[5:7]`* are moderately strong predictors and notable words with low predictive value (including positive and negative predictors) include *authority*, *influence*, *king*, *chief*, and *headman*.
-The word *follow* is the strongest negative predictor and *company* and *economic* are marginal negative predictors.
-
-## Effects of document meta-data on leadership patterns
-
-```{r tabauthorgender}
-unique_authors <- unique(authorship[c('author_ID', 'author_gender')])
-tab_author_gender <- table(unique_authors$author_gender)
-mm_coauthorSUM <- summary(mm_coauthor)
-mm_coauthorP <- mm_coauthorSUM$coefficients['female_coauthorTRUE', 'Pr(>|z|)']
-```
-
-We explored whether document features, such as date of publication and gender of authors, influenced reporting of leader functions and qualities.
-There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine; some of these authors were co-authors.
-We then tested if text records with a female author or co-author were more likely to discuss female leaders.
-Of the `r sum(tab_coauthor['FALSE',])` text records with no female co-author, `r tab_coauthor['FALSE','yes']` (`r 100*signif(tab_coauthor['FALSE','yes']/sum(tab_coauthor['FALSE',]),2)`%) reported female leadership.
-Of the `r sum(tab_coauthor['TRUE',])` text records with a female co-author, on the other hand, `r tab_coauthor['TRUE','yes']` (`r 100*signif(tab_coauthor['TRUE','yes']/sum(tab_coauthor['TRUE',]),2)`%) reported female leadership, suggesting that female ethnographers were more likely to report female leaders.
-However, after accounting for the non-independence of text records (i.e., multiple records from the same document) using a logistic mixed effects regression model, we found that although there was a positive effect of co-author gender on reporting female leadership (OR = `r signif(mm_coauthorOR, 2)`), it was not statistically significant (p = `r signif(mm_coauthorP, 2)`).
 
@@
 
+## Disentangleing the relationship between leadership and high status
+
+These data present an oppurtunity to explore the relationship between high status and leadership, as represented in ethnographic texts.
+To do so, we draw on the term-document matrix of our ethnographic sample which was developed by computing the frequencies of all "informative" words in each text record after removing all punctuation and English stop words (uninformative words such as the, is, at), stemming the remaining words (reducing inflections such as plurals and past tenses to the root, or stem, words).
+In this data matrix the columns are all the unique word stems (terms), rows were the text records (documents), and the values in each cell were the frequencies of each word stem in each text record.
+The coding of *high status* was joined with this term-document matrix.
+This process produced a matrix with `r nrow(leader_dtm)` rows with one for each text-record and `r ncol(leader_dtm)` columns with a column for each unique term and a column for the researcher-coded *high status* value ($0$ or $1$).
+
+We used lasso regression as a variable selection technique, which is especially useful when the number of predictors is much greater than the number of cases (i.e., $p > > n$).
+Lasso regression has a penalty or shrinkage term, $\lambda$: larger values of $\lambda$ retain more variables in the model, at the risk of over-fitting, whereas smaller values of $\lambda$ decrease the number of predictors retained in the model, at the risk of under-fitting.
+The optimal value of $\lambda$ can be estimated with cross-validation [@tibshirani_regression_1996; @zou_regularization_2005].
+We used the glmnet package [@friedman_regularization_2010] to fit a lasso regression model to determine which words best predicted evidence for high status as a leader quality.
+We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance.
+We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot).
+
+```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record.
+Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010].
+Blue values are positive predictors; red values are negative predictors.
+X-axis is on a log 10 scale."}
+highstatus_plot + scale_x_log10()
+```
+
+Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficents.
+The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:3]`*, and *`r rev(highstatus_plot$data$Word)[4]`*, respectively.
+The words, *`r rev(highstatus_plot$data$Word)[5:7]`* are moderately strong predictors.
+
 ## Feature models

### 2020-04-12 analyses2.R

@@ -41,6 +41,7 @@ library(ggcorrplot)
 library(margins)
 # library(mgcv)
 library(stringr)
+library(glmnet)
 
 # Load precomputed objects ------------------------------------------------
 
@@ -54,7 +55,7 @@ load("Leader2.Rdata")
 #+ fig.height=15, fig.width=15
 
 # Functions ---------------------------------------------------------------
-
+se <- function(x) sd(x)/sqrt(length(x))
 
 # Recode variables --------------------------------------------------------
 
@@ -218,10 +219,10 @@ dev.off()
 # pca_data_qualities$qPC2k2 <- m_lpca_qualk2$PCs[,2]
 # leader_text2 <- left_join(leader_text2, pca_data_qualities[c('cs_textrec_ID', 'qPC1k2', 'qPC2k2')])
 # 
-# # For optimal k, m determined by cv in initialcompute.R
-# kq <- 8
-# mq <- 12
-# 
+# For optimal k, m determined by cv in initialcompute.R
+kq <- 8
+mq <- 12
+
 # logpca_model_qualities = logisticPCA(pca_data_qualities2, k = kq, m = mq, main_effects = T)
 # plot(logpca_model_qualities, type = "scores")
 # logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2)
@@ -785,16 +786,6 @@ dev.off()
 # 
 
 
-# NMF ---------------------------------------------------------------------
-
-# library(NMF)
-# m_nmf <- nmf(t(pca_data_qualities2), rank = 2:15)
-# m_nmf5 <- nmf(t(pca_data_qualities2), rank = 5)
-# m_nmfrandom <- nmf(randomize(t(pca_data_qualities2)), rank=2:15)
-
-# pdf(file = 'consensusmap.pdf', width = 20)
-# consensusmap(m_nmf)
-# dev.off()
 
 # Group structure by subsistence --------------------------------------------------
 
@@ -928,7 +919,6 @@ plot_group_subsis_status
 
 
 # Costs and benefits by group structure type ------------------------------
-se <- function(x) sd(x)/sqrt(length(x))
 
 # try to tidy data to: Group structure, Benefit type, Cost type, Status type (leader, follower), Mean, SE
 
[…]

### 2020-04-12 initialcompute.R

@@ -611,6 +611,31 @@ qual_func_vars <- leader_text2 %>%
 #   nboot = 10000,
 #   parallel = T)
 
+# Cluster all study vars
+
+df_all <-
+  leader_text2 %>% 
+  dplyr::select(all_of(all_study_vars)) %>% 
+  dplyr::filter(rowSums(.) > 0)
+
+all_clust <- pvclust(
+  df_all,
+  method.hclust = 'ward',
+  method.dist = 'binary',
+  nboot = 1000,
+  parallel = T
+)
+
+# PAM
+library(cluster)
+library(factoextra)
+
+nbclust <- fviz_nbclust(t(df_all), pam, method='wss') # 4 or 5 silhouette
+m_pam <- pam(df_all, k = 4)
+
+nbclust <- fviz_nbclust(t(df_all), kmeans, method='silhouette') # 2, 5
+m_kmeans <- kmeans(df_all, centers = 2)
+
 # Cross-validation for logisticPCA ----------------------------------------
 
 # This takes a long time, so putting all code in one section
@@ -619,7 +644,7 @@ if(F){
   # Leader qualities
   qual_cvlpca <- cv.lpca(pca_data_qualities2, ks = 1:20, ms = 5:15)
   plot(qual_cvlpca)
-  which.min(qual_cvlpca[9,]) # optimal? k=9, m=11
+  which.min(qual_cvlpca[12,]) # optimal? k=9, m=11
   
   # Plot all minima
   x <- apply(qual_cvlpca, MARGIN = 1, which.min)

### 2020-04-12 logisticPCA.R

@@ -1,0 +1,10 @@
+library(logisticPCA)
+
+# This code requires pre-computed objects from initialcompute.R. It's not standalone.
+
+load("Leader2.Rdata")
+
+qual_cvlpca <- cv.lpca(pca_data_qualities2, ks = 1:20, ms = 5:15)
+fun_cvlpca = cv.lpca(pca_data_functions2, ks = 1:20, ms = 5:15)
+
+save(qual_cvlpca, fun_cvlpca, file = 'logisticPCA.RData')

### 2020-04-12 text.R

@@ -1,0 +1,43 @@
+library(tidytext)
+library(topicmodels)
+library(ldatuning)
+
+terms <-
+  leader_words %>% 
+  group_by(cs_textrec_ID, lemma) %>% 
+  summarise(value = n())
+
+dtm <- cast_dtm(terms, 'cs_textrec_ID', 'lemma', 'value')
+
+lda_topicsnum <- FindTopicsNumber(
+  dtm,
+  topics = 2:15,
+  metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
+  method = "Gibbs",
+  # control = list(seed = 77),
+  mc.cores = 2L,
+  verbose = TRUE
+)
+
+FindTopicsNumber_plot(lda_topicsnum)
+
+m_lda4 <- LDA(dtm, 8, method = "Gibbs")
+
+top_terms <- 
+  m_lda4 %>%
+  tidy %>% 
+  dplyr::filter(! term %in% c('chief', 'leader')) %>% 
+  group_by(topic) %>%
+  top_n(10, beta) %>%
+  ungroup() %>%
+  arrange(topic, -beta)
+
+# top_terms
+
+top_terms %>%
+  mutate(term = reorder_within(term, beta, topic)) %>%
+  ggplot(aes(beta, term)) +
+  geom_bar(stat = "identity") +
+  scale_y_reordered() +
+  facet_wrap(~ topic, scales = "free_y")
+ 

### 2020-04-12 leadership_across_cultures_contexts.Rmd
Minor edit. Changed a subheading.

@@
 
-We then explored "features" in our data.
-There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
-We used several methods to explore lower dimensional representations of these data that might reveal informative structure, including logistic PCA, non-negative matrix factorization (NMF), and agglomerative hierarchical cluster analysis.
+We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
 Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
@@
 Agglomerative hierarchical cluster starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
-We computed distance between our binary vectors using $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and agglomerated using the Ward algorithm.
+We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm.
 We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
@@
 
+The analyses computing levels of evidence for each variable used the entire set of text records.
+Analyses investigating lower dimensional representations used a slightly smaller set that omitted records with no evidence for any variable (i.e., all zeros).
 All analyses were conducted with `r R.version.string`.
@@
 
-The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with some records not sex-specific.
+The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific.
 There were female leaders in almost all types of groups, however, except military and religious.
@@
 
-## Variable distributions
+## Evidence for leader functions and qualities
 

### 2020-04-15 initialcompute.R
Added all_study_vars list to initial compute script -- it wasn't there but was needed.

@@ -85,6 +85,8 @@ leader_cost_vars <- lt2vars[str_detect(lt2vars, 'leader.costs')]
 follower_benefit_vars <- lt2vars[str_detect(lt2vars, 'follower.benefits')]
 follower_cost_vars <- lt2vars[str_detect(lt2vars, 'follower.costs')]
 
+all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
+
 # Aggregate at Culture level ----------------------------------------------
 
 # df that links the ids of texts, docs, and cultures

### 2020-04-15 analyses2.R
Removed all_study_vars list, moved to initial compute.
Added document years plots from SI script file (deleted that file).

@@ -4,8 +4,10 @@
 # https://stats.stackexchange.com/questions/16331/doing-principal-component-analysis-or-factor-analysis-on-binary-data/16335#16335
 # https://cran.r-project.org/web/packages/logisticPCA/vignettes/logisticPCA.html
 
-# library(conflicted)
+library(conflicted)
 # conflict_prefer("which", "base")
+conflict_prefer("as.matrix", "base")
+conflict_prefer("filter", "dplyr")
 
 # Load data library -------------------------------------------------------
 library(leadershipdata)
@@ -42,6 +44,7 @@ library(margins)
 # library(mgcv)
 library(stringr)
 library(glmnet)
+library(ggalt)
 
 # Load precomputed objects ------------------------------------------------
 
@@ -57,9 +60,6 @@ load("Leader2.Rdata")
 # Functions ---------------------------------------------------------------
 se <- function(x) sd(x)/sqrt(length(x))
 
-# Recode variables --------------------------------------------------------
-
-all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
 
 # Variable support plots --------------------------------------------------
 
@@ -1123,6 +1123,46 @@ plot_pages_tr <-
   labs(x = '\nTotal number of pages of ethnography in the eHRAF', y = 'Number of text records on leadership\n') +
   theme_bw(15)
 plot_pages_tr
+
+
+# Document years plots ----------------------------------------------------
+
+documents$d_publication_date <- as.numeric(documents$d_publication_date)
+
+doc_year_hist <- ggplot(documents, aes(d_publication_date)) + 
+  geom_histogram(binwidth = 5) +
+  scale_x_continuous(breaks = seq(1860,2000,10)) +
+  geom_vline(xintercept = mean(documents$d_publication_date, na.rm=T),
+             linetype="dotted", 
+             color = "grey", size=.5)+
+  labs(x="\nDocument publication years (mean year 1960)", y="Count\n")
+
+## Dumbbell plot of fieldwork time frames
+
+documents$d_field_date_start <- as.numeric(documents$d_field_date_start)
+documents$d_field_date_end <- as.numeric(documents$d_field_date_end)
+
+documents2<-left_join(documents, leader_cult) %>% 
+  filter(d_field_date_start>1869)
+
+doc_fielddates_plot<- ggplot(documents2[!is.na(documents2$d_field_date_start)==T,], 
[…]

### 2020-04-15 analyses2.R
Minor changes to document year plots.

@@ -1135,7 +1135,7 @@ doc_year_hist <- ggplot(documents, aes(d_publication_date)) +
   geom_vline(xintercept = mean(documents$d_publication_date, na.rm=T),
              linetype="dotted", 
              color = "grey", size=.5)+
-  labs(x="\nDocument publication years (mean year 1960)", y="Count\n")
+  labs(x="\nDocument publication years", y="Count\n")
 
 ## Dumbbell plot of fieldwork time frames
 
@@ -1150,7 +1150,7 @@ doc_fielddates_plot<- ggplot(documents2[!is.na(documents2$d_field_date_start)==T
   geom_dumbbell() +
   facet_grid(region~., scales = "free_y", space = "free_y")+ 
   scale_x_continuous(breaks = seq(1870,2000,10), minor_breaks = seq(1870,2000,10)) +
-  labs(x="\nDocument field work start and end dates", y="")+
+  labs(x="\nSpan of earliest and latest field work year for each document", y="")+
   geom_vline(xintercept = mean(documents2$d_field_date_start, na.rm=T),
              linetype="dotted", 
              color = "black", size=.5)+

### 2020-04-15 leadership_across_cultures_contexts.Rmd
Minor edits across manuscript.

--- 2020-04-12 leadership_across_cultures_contexts.Rmd
@@
 We then outline the aims and methods of the current cross-cultural study, before presenting results and interpretations.
-Through novel exploratory analyses encompassing the full range of cultural diversity and social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
+Through novel exploratory analyses encompassing the full range of cultural diversity and diverse social contexts, we aim to advance leadership studies across disciplines and provide researchers with a foundation for more detailed approaches to human leadership across cultures and contexts.
 
@@
 Leaders emerge in all human societies.
-That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group and functions to facilitate group goals and provides services to the group.
+That is, within every population there are particular contexts in which one individual maintains a disproportionate level of influence in the group, functions to facilitate group goals, and provides services to the group.
 Adopting the broad definition of leadership given above, parents can also be defined as leaders of nuclear families.
@@
 @roscoe_social_2009 draws on the ethnography of highland New Guinea in describing the modular structure of human societies and suggests the functions of community leaders in small-scale societies are structured by the joint demands of signaling military fromidability to rival groups and facilitating within-group conflict resolution and cooperation.
-In addition to military prowess, 'big-men' are characterized by, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
+For "big-men" in these societies, in addition to military prowess, "talents for oratory, mediation, and organization were instrumental in reducing their internal conflict and enhancing the ability of their members to coordinate their actions as a group" [@roscoe_social_2009, p.
 104].
 
-Given specific demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
+Given demands of community organization and conflict resolution, leadership tends to be context-specific and leaders in such contexts often draw on persuasive capacities and leverage community-based respect, specialized skills or knowledge, and social support to achieve or maintain community influence [@kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993].
 The theoretical and ethnographic literature suggests these demands most commonly arise in the contexts of collective actions and in conflict resolution [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019].
@@
 Demands of between-group interactions, including both conflict and cooperation are also very common across small-scale societies, including hunter-gatherers [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].
-Across relatively small-scale societies, at the community level, representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of leaders.
-Within communities, however, leaders may emerge whenever individuals form groups.
-The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions leaders serve.
+Across relatively small-scale societies representing the group and managing both internal and external relationships and conflicts are expected to be the most salient functions of community leaders.
+Within communities, however, leaders may emerge whenever individuals congregate.
+The role of leaders in ritual contexts, such as in initiations [@singh_cultural_2017], and in providing social services, including facilitating marriages [@walker_evolutionary_2011] and organizing social activities [@wiessner_food_1996], are also expected to be common functions carried out by a leader.
 
@@
 In WEIRD societies, community leaders are primarily heads of governmental and corporate bodies [@bass_handbook_1990].
-Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
-These positions typically are associated with institutionalized responsibilities and specific functions associated with their roles.
-The structure of such organizations are generally organized hierarchically with nested bodies and a formal internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
+Other public service, nonprofit leaders and managers include law enforcement and police officers, college presidents and school principles, and grassroots community leaders, to name a few [@martiskainen_role_2017; @azzam_community_2003; @grace_school_2005].
+These positions typically are associated with institutionalized responsibilities and specific functions.
+The structure of such organizations are generally hierarchical with nested bodies and an internal 'chain of command.' At higher levels of large-scale organizations, leaders primarily serve as representatives in inter-group interactions, facilitate the creation and adoption of policies, and serve as figure heads embodying the strategic vision of the organization [@keohane_thinking_2010].
 They are not typically directly involved in resolving interpersonal conflicts between individuals, however, they are often the final end-point in a decision making hierarchy and the results of a single decision of these leaders can trickle down through industry or government with global ramifications.
@@
 The GLOBE project has produced multiple cross-national data sets on leadership and hundreds of publications.
-For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied of industries across 24 countries [@house2013strategic].
+For example, *The GLOBE Study of CEO Leadership Behavior and Effectiveness in 24 Countries* draws on data from over 1,000 CEOs and over 5,000 senior executives from corporations in varied industries across 24 countries [@house2013strategic].
 The goals of the GLOBE CEO Study were to address (1) how national culture influences expectations for the behaviors of leaders, (2) which behaviors of CEOs are typically associated with organizational success, (3) differences between high-functioning and low-functioning CEOs, and (4) the importance of CEO behaviors aligning with societal expectations.
@@
 In the United States, despite a designed separation of power across branches of government, presidents, governors, and mayors oversee governmental bodies and have substantial influence in economics, social and legal policies, and international and inter-regional relations.
-Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage inter-group social dynamics.
+Community leadership functions in increasingly WEIRD contexts will likely be increasingly focused on economics and inter-group relations, whereas in less WEIRD contexts community leaders will primarily function to resolve interpersonal conflicts and manage intra-group social dynamics.
 <!-- Could add ethnographic examples for SSS and LSS  -->
@@
 
-The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups.
+The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups (see Methods section).
 Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence.
@@
 We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` variables in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)).
-Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)`.
+Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure \@ref(fig:doc-years)).
 There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors).
@@
 After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models.
-More recent publications were more likely to report evidence for Oratory skill, Group representative, Group determination/cohesiveness, and Knowledgeable/intelligent, and less likely to report evidence for leaders imposing costs (Punishment, Killer) or receiving benefits (Social services, Food, and Reduced risk of harm).
+More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*).
 See Figure \@ref(fig:pubdate).
@@
 
-Our qualities and functions data comprised sparse $`r nrow(leader_text)` \times `r length(quality_vars)`$ and $`r nrow(leader_text)` \times `r length(function_vars)`$ matrices of binary data, respectively.
+Our qualities and functions data comprised sparse $`r nrow(leader_text2)` \times `r length(quality_vars)`$ and $`r nrow(leader_text2)` \times `r length(function_vars)`$ matrices of binary data, respectively.
 
@@
 This exploratory study extracted the qualities, functions, benefits, and costs of leadership from the ethnographic record.
-These variables were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
+The variables on leader qualities and functions were *not* determined prior to the study but were instead identified and coded during several passes through `r nrow(text_records)` texts on leadership obtained from the eHRAF (see Methods section).
 Our results therefore represent the broadest and most comprehensive view of the qualities of leaders (i.e., the types of people that occupy leadership roles), the functions of leaders (i.e., what leaders do), and the costs and benefits of leadership for both leaders and followers across a diverse and culturally representative sample of the ethnographic record.
@@
 See Figure \@ref(fig:cluster)A.
-Several of traits often co-occur.
+Several of these traits often co-occur.
 Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p.
@@
 We therefore coded all benefits and costs of both leadership and followership to better understand which benefits and costs are most commonly associated with leading or following.
-Leaders were widely reported to receive material benefits and social status, but loss of material resources and social status were also widely reported.
+Leaders were widely reported to receive material benefits, social status, and mating benefits but loss of material resources and social status were also widely reported.
 The results were similar for followers: followers benefited from increased material and social resources, but also often lost material and social resources.
+Followers were not reported to recieve mating benefits, relative to leaders or other types benefits.
 Our results cannot determine if the benefits to leaders and followers outweighed their costs, but they do shed light on the social and biological pressures associated with leadership and social hierarchy.
@@
 
-```{r eHRAFpages, fig.height=6, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF.
+## Ethnographic sample
+
+```{r}
+doc_1950_perc <- signif(length(documents$d_publication_date[documents$d_publication_date>1949])/length(documents$d_publication_date),2)*100
+```
+
+The majority of ethnographic documents sampled here (`r doc_1950_perc`%) were published after 1950 and the mean publication year is `r round(mean(documents$d_publication_date),)` (SD = `r round(sd(documents$d_publication_date),2)`).
+Figure \@ref(fig:doc-years) plots the publication years of all documents which contributed text records to the ethnographic sample.
+
+```{r doc-years, fig.height=6, fig.cap="Historgram of eHRAF document publiction year. Vertical line shows the mean publication year."}
+doc_year_hist
+```
+
+The eHRAF provides information on the time frame of field work associated with each document.
+For some documents the span of associated fieldwork is given for the exact months and years the ethnographer was in the field.
+For most documents one or more spans of years are given.
+Figure \@ref(fig:field-years) plots the earliest year and the latest year for which field work occurred for each document.
+The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred.
+We omitted data for three documents (two from North America and one from Insular Pacific) with field work date ranges prior to 1870.
+For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.
+
+```{r field-years, fig.height=8, fig.width=9, fig.cap="Dumbbell plot of eHRAF document field work years. Vertical lines shows the  mean earliest and latest years of field work."}
+doc_fielddates_plot
+```
+
+```{r eHRAFpages, fig.height=8, fig.cap="Scatterplot of the number of leadership text records versus the total number of pages of ethnography in the eHRAF.
 Each dot is one culture.

### 2020-04-15 leadership_across_cultures_contexts.Rmd
Lots of edits, esp to text analysis

--- 2020-04-12 leadership_across_cultures_contexts.Rmd
@@
 Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
+
 NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$.
-NMF estimates $W$ and $H$ by minimizing a measure of distance (typically the Frobenius norm or the Kullback-Leibler divergence) between $X$ and $WH$ [@gaujoux2010].
-Agglomerative hierarchical cluster starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
+NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].
+
+Agglomerative hierarchical cluster analysis starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
 We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm.
@@
 
-Our qualities and functions data comprised sparse $`r nrow(leader_text)` \times `r length(quality_vars)`$ and $`r nrow(leader_text)` \times `r length(function_vars)`$ matrices of binary data, respectively.
+```{r}
+num_removed <- sum(rowSums(leader_text2[all_study_vars]) == 0)
+```
+
+To investigate features in our data, we first removed `r num_removed` text records with no evidence for any variable.
+
+Our qualities and functions data comprised sparse $`r nrow(leader_text2)` \times `r length(quality_vars)`$ and $`r nrow(leader_text2)` \times `r length(function_vars)`$ matrices of binary data, respectively.
 
@@
 
+## Ethnographer and Culture random effects
+
+```{r FunctionRE, fig.cap="Ethnographer and Culture random effects for function variables."}
+plot_fun_auth_cult_re
+```
+
+```{r QualityRE, fig.height=8, fig.cap="Ethnographer and Culture random effects for quality variables."}
+plot_qual_auth_cult_re
+```
+
 ## Disentangleing the relationship between leadership and high status
@@
 We used tenfold cross-validation to choose the $\lambda$ parameter that minimized deviance.
-We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot).
+We then plotted all the non-zero coefficients (Figure \@ref(fig:highstatusplot)).
 
@@
 
+## Dimension reduction
+
 ## Feature models
 
-```{r features, fig.width=10, fig.cap="Marginal effects of Poission mixed effects regression models of Feature variables by mode of subsistence.
+```{r features, fig.width=8, fig.cap="Esimtated marginal means [@lenth2020] of Poission mixed effects regression models of Feature variables by mode of subsistence.
 Only models whose subsistence coefficient had a corrected p-value < 0.05.
-Base level: agriculturalists.
+Plotted on the response scale.
 See text for details."}
 
-x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))
+# x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))
+
+plot_feature_models
 

### 2020-04-15 NMF.R

@@ -21,10 +21,52 @@ save(m_nmf, m_nmfrandom, file = 'NMF.RData')
 
 plot(m_nmf, m_nmfrandom)
 
-m_nmf3 <- nmf(t(df_all), rank = 3)
+library(NMF) # To run job from selection
+m_nmf3 <- nmf(t(df_all), rank = 3, nrun = 100)
 basismap(m_nmf3, filename = 'Figures/basismap3.pdf', hclustfun = 'ward', distfun='correlation')
 coefmap(m_nmf3, hclustfun = 'ward', filename = 'Figures/coefmap3.pdf')
 
-m_nmf4 <- nmf(t(df_all), rank = 4)
+m_nmf4 <- nmf(t(df_all), rank = 4, nrun = 100)
 basismap(m_nmf4, filename = 'Figures/basismap4.pdf', hclustfun = 'ward', distfun='correlation')
-coefmap(m_nmf4, hclustfun = 'ward', filename = 'Figures/coefmap4.pdf')
+coefmap(m_nmf4, hclustfun = 'ward', distfun='correlation', filename = 'Figures/coefmap4.pdf')
+
+save(m_nmf, m_nmfrandom, m_nmf3, m_nmf4, file = 'NMF.RData')
+
+# plotting bases
+
+df_basis_long <-
+  m_nmf4 %>% 
+  basis %>% 
+  round(2) %>% # Set low values to 0
+  as.data.frame %>% 
+  mutate(
+    Variable = rownames(.)
+  ) %>% 
+  gather(key = Component, value = 'Loading', -Variable) %>% 
+  dplyr::filter(Loading > 0)
+
+ggplot(df_basis_long, aes(Loading, reorder_within(Variable, Loading, Component))) +
+  geom_point() +
+  facet_wrap(~Component, scales = 'free_y')
+
+# Analyzing coefs
+
+df_coef_long <-
+  m_nmf4 %>% 
+  coef %>% 
+  round(2) %>% # Set low values to 0
+  t %>% 
+  as.data.frame %>%
+  rowwise() %>% 
+  mutate(
+    max = max(c(V1, V2, V3, V4)),
+    quartile3 = quantile(c(V1, V2, V3, V4))[4]
+  ) %>% 
+  ungroup %>% 
+  summarise_at(
+    1:4,
+    ~ mean(. > quartile3) # Approx. fraction of text records for each topic
+  )
+  
+  
+  

### 2020-04-15 analyses2.R

@@ -42,6 +42,9 @@ library(margins)
 # library(mgcv)
 library(stringr)
 library(glmnet)
+library(treemapify)
+library(emmeans)
+library(hagenutils)
 
 # Load precomputed objects ------------------------------------------------
 
@@ -71,7 +74,6 @@ plot.variable.support = ggplot(d_melt, aes(value, Variable, xmin=y_negse, xmax=y
   scale_x_continuous(breaks=seq(0,1,.1), labels=scales::percent, limits=c(0,1)) +
   scale_colour_discrete(name='', labels=c('Cultures', 'Text records')) +
   facet_grid(Model~., scales = "free_y", space = "free_y") +
-  theme(strip.text.y = element_text(angle=0)) +
   scale_shape_manual(name="", values=c(17,16), labels=c('Cultures', 'Text records'))+
   scale_fill_manual(name="", values=c("red", "blue"), labels=c('Cultures', 'Text records')) +
   labs(x="",y="") +
@@ -799,7 +801,7 @@ df_groups <-
   dplyr::filter(group.structure2 != 'other') %>% 
   mutate(
     demo_sex = factor(demo_sex, levels = c('male', 'female')),
-    group =   factor(
+    group = factor(
       group.structure2,
       levels = c(
         'residential subgroup',
@@ -830,6 +832,40 @@ plot_group_subsis <-
   guides(fill = guide_legend(reverse = T)) +
   theme_minimal(15) 
 plot_group_subsis
+
+# High status by subsistence groups
+
+# Simple proportions
+x <- table(leader_text2$group.structure2, leader_text2$qualities_high.status)
+x <- prop.table(x, margin = 1)[,2]
+plot_status_group <- ggdotchart(x) + scale_x_continuous(limits = c(0, 0.5))
+
+# logistic model
+
+m_status_group <-
+  glmer(
+    qualities_high.status ~
+      group.structure2 +
+      (1|d_culture/author_ID),
+    family = binomial,
+    data = all_data
+  )
+
+emm_status_group <-
+  emmeans(m_status_group, 'group.structure2', type = 'response') %>% 
+  summary %>% 
+  mutate(
+    group.structure2 = fct_reorder(group.structure2, prob)
+  )
+
+plot_emm_status_group <-
[…]

### 2020-04-15 text.R

@@ -40,4 +40,88 @@ top_terms %>%
   geom_bar(stat = "identity") +
   scale_y_reordered() +
   facet_wrap(~ topic, scales = "free_y")
- 
+
+# Word treemaps
+
+df_word_freq <-
+  leader_words %>% 
+  group_by(lemma) %>% 
+  summarise(Frequency = n()) %>% 
+  arrange(-Frequency) %>% 
+  slice_head(n=50)
+
+plot_word_freq <- 
+  ggplot(df_word_freq, aes(area=Frequency, label = lemma)) +
+  geom_treemap(fill='white') +
+  geom_treemap_text(colour = 'black')
+# plot_word_freq
+
+# By subsistence type
+df_word_freq_subsis <-
+  leader_words %>% 
+  left_join(leader_text2[c('cs_textrec_ID', 'subsistence')]) %>% 
+  group_by(subsistence, lemma) %>% 
+  summarise(Frequency = n()) %>% 
+  arrange(-Frequency) %>% 
+  slice_head(n=200)
+
+plot_word_freq_subsis <- 
+  ggplot(df_word_freq_subsis, aes(area=Frequency, label = lemma)) +
+  geom_treemap(fill='white') +
+  geom_treemap_text(colour = 'black') +
+  facet_wrap(~subsistence)
+# plot_word_freq_subsis
+
+# NLP ecosystem
+# http://www.bnosac.be/images/bnosac/blog/NLP-R-ecosystem.png
+
+# Biterm
+# http://www.bnosac.be/index.php/blog/98-biterm-topic-modelling-for-short-texts
+
+library(udpipe)
+library(stopwords)
+library(BTM)
+library(ggraph)
+library(textplot)
+
+text_records2 <-
+  text_records %>% 
+  dplyr::select(document_d_ID, raw_text) %>% 
+  rename(
+    doc_id = document_d_ID,
+    text = raw_text
+  ) %>% 
+  mutate(
+    text = str_to_lower(text),
+    text = str_replace(text, "’", ""),
+    text = str_replace(text, "'", ""),
+    text = str_replace(text, "page", ""),
+    text = str_replace(text, "\\d+", "")
+  )
+
+anno <- udpipe(text_records2, "english", trace = 10)
+
+biterms <-
+  anno %>% 
+  group_by(doc_id) %>% 
+  nest() %>% 
+  mutate(
+    cooc = map(
+      data,
+      ~ cooccurrence(
+        x = .x$lemma,
[…]

### 2020-04-15 leadership_across_cultures_contexts.Rmd

--- 2020-04-15 leadership_across_cultures_contexts.Rmd
@@
 
-We then explored "features" in our data.
-There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques.
-We used several methods to explore lower dimensional representations of these data that might reveal informative structure, including logistic PCA, non-negative matrix factorization (NMF), and agglomerative hierarchical cluster analysis.
+We then used several methods to explore lower dimensional representations of these data that might reveal informative "features." These included logistic PCA, non-negative matrix factorization (NMF), and hierarchical cluster analysis.
 Logistic PCA generalizes Pearson's original formulation of PCA to find a projection of the natural parameters of the saturated model that minimizes the Bernoulli deviance [@logisticPCA].
+
 NMF [@lee1999] finds a low rank approximation of the original matrix, $X \approx WH$, where $X$ is an $n \times p$ matrix, and $W$ and $H$ are $n \times k$ and $k \times p$ matrices, respectively; all elements of $X$, $W$ and $H$ are $\geq 0$; and, typically, $k << min(n,p)$.
-NMF estimates $W$ and $H$ by minimizing a measure of distance (typically the Frobenius norm or the Kullback-Leibler divergence) between $X$ and $WH$ [@gaujoux2010].
-Agglomerative hierarchical cluster starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
-We computed distance between our binary vectors using $1-cor(x, y)$ (so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$), and agglomerated using the Ward algorithm.
+NMF estimates $W$ and $H$ by minimizing a measure of distance between $X$ and $WH$ [we used the default "brunet" algorithm, which uses the Kullback-Leibler divergence; @gaujoux2010].
+
+Agglomerative hierarchical cluster analysis starts with all elements in their own cluster, and then uses a distance metric to combine the two closest clusters into a single cluster; the next two closest clusters are then combined, and so forth, until all elements belong to a single cluster.
+We computed distance between our binary vectors using $1-cor(x, y)$, and agglomerated using the Ward algorithm.
 We assessed the robustness of the clusters using the *pvclust* package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples.
@@
 ## Dimension reduction
+
+```{r}
+num_removed <- sum(rowSums(leader_text2[all_study_vars]) == 0)
+```
+
+To investigate features in our data, we first removed `r num_removed` text records with no evidence for any variable.
 

### 2020-04-15 leadership_across_cultures_contexts.Rmd
Minor edits in ms.

--- 2020-04-15 leadership_across_cultures_contexts.Rmd
@@
 <!-- BENEFITS: 'MODEL' OF POTENTIAL BENEFIST -->
-@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
+@von_rueden_leadership_2015 review a large body of theory and ethnography and outline four possible mechanisms by which leaders in small-scale societies might offset the potential costs associated with various leadership roles; leaders may, (1) facilitate collective actions that yield net benefits to themselves or their close kin; (2) claim a fee for their services, such as a tax or greater share of returns; (3) receive reciprocal exchange in other currencies from followers who have benefited from their services; or (4) gain other social or reproductive benefits by signaling their high quality as a mate and social partner.
 

### 2020-04-16 leadership_across_cultures_contexts.Rmd
Many edits

--- 2020-04-15 leadership_across_cultures_contexts.Rmd
@@
 library(pander)
+library(patchwork)
 ```
@@
 
+We then explored relationships between total leader benefits and costs, summed across all leader benefit and cost variables, respectively, and levels of social and political organization.
+In general, leaders at the highest supracommunity level had the most evidence for receiving benefits and the least evidence for suffering costs.
+See Figure \@ref(fig:leadertotalbenefitscosts).
+
+```{r leadertotalbenefitscosts, fig.cap="Estimated marginal means of A.
+leader total benefits, and B.
+total costs, as a function of group structure, based on a Poisson mixed effects regression model with random intercepts for ethnographer nested within culture.
+Values are on the response scale."}
+p1 <- 
+  emm_plot(emmeans(m_ldrtotben, type = 'response', specs = 'group.structure2')) +
+  labs(title = "Total leader benefits", x = "", y = "")
+
+p2 <- emm_plot(emmeans(m_ldrtotcost, type = 'response', specs = 'group.structure2')) +
+  labs(title = "Total leader costs", x = "\nRate", y = "")
+
+p1 + p2 + plot_layout(ncol = 1) + plot_annotation(tag_levels = 'A')
+
+```
+
 <!-- ## Exploring population-level variation -->
@@
 
-```{r doc-years, fig.height=6, fig.cap="Historgram of eHRAF document publiction year. Vertical line shows the mean publication year."}
+```{r doc-years, fig.height=6, fig.cap="Historgram of eHRAF document publiction year. Vertical line shows the median publication year."}
 doc_year_hist
@@
 
-```{r field-years, fig.height=8, fig.width=9, fig.cap="Dumbbell plot of eHRAF document field work years. Vertical lines shows the  mean earliest and latest years of field work."}
+```{r field-years, fig.height=8, fig.width=9, fig.cap="Span of fieldwork years of study documents. To improve clarity, one document with very early fieldwork from 1791-1804 was omitted."}
 doc_fielddates_plot
@@
 Figure \@ref(fig:highstatusplot) depicts the `r length(highstatus_plot$data$Word)` words with non-zero (mostly positive) coefficents.
-The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:3]`*, and *`r rev(highstatus_plot$data$Word)[4]`*, respectively.
-The words, *`r rev(highstatus_plot$data$Word)[5:7]`* are moderately strong predictors.
+The words most predictive of evidence for high status leaders are *`r rev(highstatus_plot$data$Word)[1:4]`*, and *`r rev(highstatus_plot$data$Word)[5]`*.
+
+We then investigated if the proportion of text records with evidence for high status differed among the different levels of social and political organization.
+There were modest differences.
+Somewhat counterintuitively, the proportion of text records indicating high status was lowest at the highest level of political organization, perhaps because it would have been superfluous to note that kings and chiefs were high status.
+See Figure \@ref(fig:groupstatus).
+
+```{r groupstatus, fig.cap="Estimated marginal means of a mixed effects logistic regression model of evidence for high status (0/1) as a function of group type, with random effects for ethnographer nested within culture."}
+plot_emm_status_group
+```
 

### 2020-04-16 analyses2.R

@@ -58,11 +58,24 @@ library(ggalt)
 
 load("Leader2.Rdata")
 
+# This is also needed in initialcompute.R, so this is redundant
+all_study_vars <- c(function_vars, quality_vars, leader_benefit_vars, leader_cost_vars, follower_benefit_vars, follower_cost_vars)
+
+
 #+ fig.height=15, fig.width=15
 
 # Functions ---------------------------------------------------------------
 se <- function(x) sd(x)/sqrt(length(x))
 
+emm_plot <- function(em){
+  emsum <- summary(em)
+  names(emsum)[1:2] <- c('var', 'estimate')
+  emsum[[1]] <- fct_reorder(emsum[[1]], emsum[[2]])
+  ggplot(emsum, aes(estimate, var, xmin = asymp.LCL, xmax = asymp.UCL)) +
+    geom_errorbarh(height = 0, lwd=2.5, alpha = 0.3) +
+    geom_point() +
+    theme_minimal(15)
+}
 
 # Variable support plots --------------------------------------------------
 
@@ -836,7 +849,7 @@ plot_group_subsis
 # High status by subsistence groups
 
 # Simple proportions
-x <- table(leader_text2$group.structure2, leader_text2$qualities_high.status)
+x <- table(leader_text2$group.structure2, leader_text2$qualities_HighStatus)
 x <- prop.table(x, margin = 1)[,2]
 plot_status_group <- ggdotchart(x) + scale_x_continuous(limits = c(0, 0.5))
 
@@ -844,7 +857,7 @@ plot_status_group <- ggdotchart(x) + scale_x_continuous(limits = c(0, 0.5))
 
 m_status_group <-
   glmer(
-    qualities_high.status ~
+    qualities_HighStatus ~
       group.structure2 +
       (1|d_culture/author_ID),
     family = binomial,
@@ -860,12 +873,15 @@ emm_status_group <-
 
 plot_emm_status_group <-
   ggplot(emm_status_group, aes(prob, group.structure2, xmin = asymp.LCL, xmax = asymp.UCL)) +
-  geom_errorbarh(height = 0) +
+  geom_errorbarh(height = 0, lwd=2.5, alpha = 0.3) +
   geom_point() +
   scale_x_continuous(limits = c(0, 0.7)) +
   labs(title="Proportion of evidence that leaders are high status", x = '\nProportion', y = "") +
-  theme_minimal()
+  theme_minimal(15)
 plot_emm_status_group
+
+
+
 
[…]

### 2020-04-16 initialcompute.R

@@ -77,7 +77,7 @@ lt2vars <- names(leader_text2)
 
 # This includes "functions_context" which is not included in original vector
 function_vars <- lt2vars[str_detect(lt2vars, 'function')]
-function_vars <- function_vars[function_vars != "functions_context"]
+function_vars <- function_vars[function_vars != "functions_Context"]
 
 quality_vars <- lt2vars[str_detect(lt2vars, 'qualities')]
 leader_benefit_vars <- lt2vars[str_detect(lt2vars, 'leader.benefits')]

### 2020-04-17 analyses2.R
Adjusting scripts to match naming convention.

@@ -877,7 +877,7 @@ plot_group_sex
 
 ## Groups by subsistence for high status texts only
 df_groups_status <- 
-  leader_text2[leader_text2$qualities_high.status==0,] %>% 
+  leader_text2[leader_text2$qualities_HighStatus==0,] %>% 
   dplyr::select(
     group.structure2,
     demo_sex,

### 2020-04-17 initialcompute.R

@@ -77,7 +77,7 @@ lt2vars <- names(leader_text2)
 
 # This includes "functions_context" which is not included in original vector
 function_vars <- lt2vars[str_detect(lt2vars, 'function')]
-function_vars <- function_vars[function_vars != "functions_context"]
+function_vars <- function_vars[function_vars != "functions_Context"]
 
 quality_vars <- lt2vars[str_detect(lt2vars, 'qualities')]
 leader_benefit_vars <- lt2vars[str_detect(lt2vars, 'leader.benefits')]
@@ -603,7 +603,7 @@ m_pvclust_fun <-
 
 qual_func_vars <- leader_text2 %>% 
   select(matches("functions|qualities")) %>% 
-  select(-functions_context) %>% 
+  select(-functions_Context) %>% 
   select(-contains("component"))
 
 # qual_func_clust <- pvclust(

### 2020-04-17 initialcompute.R
updated var names list.

@@ -286,79 +286,79 @@ for (m in models) {
 d_melt = data.frame(Model=Model, Variable=Variable, Type=Type, value=value, y_se=y_se, y_negse=y_negse, stringsAsFactors = F)
 
 var_names <- c(
-  "function_bestow.mate" = "Bestow mates",                        
-  "function_political.appointments" = "Political appointments",             
-  "functions_construction.infrastructure"       = "Construction/infastructure",
-  "functions_control.economics"                 = "Control economics",
-  "functions_council.member"                    = "Council member",
-  "functions_group.determination"              = "Group determination/cohesiveness",
-  "functions_hospitality"                       = "Hospitality",
-  "functions_military.command"                  = "Military command",
-  "functions_new.settlement"                    =  "Movement/migration",
-  "functions_prosocial.investment"              = "Prosocial investment",
-  "functions_provide.counsel"                   = "Provide counsel/direction",
-  "functions_punishment"                        = "Punishment",
-  "functions_serve.leader"                      = "Serve a leader",
-  "functions_strategic.planning" = "Strategic planning",
-  "function_organize.cooperation"  = "Organize cooperation",
-  "function_resolve.conflcit"      = "Resolve conflict",
-  "functions_control.calendar"     = "Control calendar",
-  "functions_control.immigration"  = "Control immigration",
-  "functions_distribute.resources" = "Distribute resources",
-  "functions_group.representative" = "Group representative",
-  "functions_medicinal"           = "Medicinal functions",
-  "functions_moral.authority"     = "Moral authority",
-  "functions_policymaking"        =  "Policy making",
-  "functions_protection"          = "Protection",
-  "functions_provide.subsistence" = "Provide subsistence",
-  "functions_ritual"              = "Ritual functions",
-  "functions_social.functions" = "Misc. social functions",
-  "qualities_artistic.performance"     = "Artistic performance",
-  "qualities_generous"                 = "Generosity",
-  "qualities_age"                      = "Age",
-  "qualities_attractive"              = "Attractive",
-  "qualities_coercive.authority"      = "Coercive authority",
-  "qualities_culturally.progressive"  = "Culturally progressive",
-  "qualities_favorable.personality"   = "Favorable personality",
[…]

### 2020-04-17 analyses2.R
updating cost benefit dict names.

@@ -97,43 +97,43 @@ plot.variable.support
 # Leader/follower benefits and costs
 
 cost_benefit_dict <- c(
-  leader.benefits_fitness = 'Inclusive fitness benefit',
-  leader.benefits_mating = 'Mating benefit',
-  leader.benefits_other = 'Misc. non-material benefit',
-  leader.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
-  leader.benefits_resource_food = 'Food benefit',
-  leader.benefits_resource_other = 'Material resources',
-  leader.benefits_social.services = 'Social services',
-  leader.benefits_social.status.reputation = 'Increased social status',
-  leader.benefits_territory = 'Territory',
-  follower.benefits_fitness = 'Inclusive fitness benefit',
-  follower.benefits_mating = 'Mating benefit',
-  follower.benefits_other = 'Misc. non-material benefit',
-  follower.benefits_reduced.risk.harm.conflict = 'Reduced risk of harm',
-  follower.benefits_resource_food = 'Food benefit',
-  follower.benefits_resource_other = 'Material resources',
-  follower.benefits_social.services = 'Social services',
-  follower.benefits_social.status.reputation = 'Increased social status',
-  follower.benefits_territory = 'Territory',
+  leader.benefits_Fitness = 'Inclusive fitness benefit',
+  leader.benefits_Mating = 'Mating benefit',
+  leader.benefits_Other = 'Misc. non-material benefit',
+  leader.benefits_RiskHarmConflict = 'Reduced risk of harm',
+  leader.benefits_FesourceFood = 'Food benefit',
+  leader.benefits_ResourceOther = 'Material resources',
+  leader.benefits_SocialServices = 'Social services',
+  leader.benefits_SocialStatusReputation = 'Increased social status',
+  leader.benefits_Territory = 'Territory',
+  follower.benefits_Fitness = 'Inclusive fitness benefit',
+  follower.benefits_Mating = 'Mating benefit',
+  follower.benefits_Other = 'Misc. non-material benefit',
+  follower.benefits_RiskHarmConflict = 'Reduced risk of harm',
+  follower.benefits_ResourceFood = 'Food benefit',
+  follower.benefits_ResourceOther = 'Material resources',
+  follower.benefits_SocialServices = 'Social services',
+  follower.benefits_SocialStatusReputation = 'Increased social status',
+  follower.benefits_Territory = 'Territory',
   
[…]

### 2020-04-17 analyses2.R
Fixed names in costs benefits dict.

@@ -48,6 +48,7 @@ library(treemapify)
 library(emmeans)
 library(hagenutils)
 library(ggalt)
+library(ggpubr)
 
 # Load precomputed objects ------------------------------------------------
 
@@ -101,7 +102,7 @@ cost_benefit_dict <- c(
   leader.benefits_Mating = 'Mating benefit',
   leader.benefits_Other = 'Misc. non-material benefit',
   leader.benefits_RiskHarmConflict = 'Reduced risk of harm',
-  leader.benefits_FesourceFood = 'Food benefit',
+  leader.benefits_ResourceFood = 'Food benefit',
   leader.benefits_ResourceOther = 'Material resources',
   leader.benefits_SocialServices = 'Social services',
   leader.benefits_SocialStatusReputation = 'Increased social status',
@@ -116,12 +117,12 @@ cost_benefit_dict <- c(
   follower.benefits_SocialStatusReputation = 'Increased social status',
   follower.benefits_Territory = 'Territory',
   
-  leader.costs_FitnessCosts = 'Inclusive fitness cost',
+  leader.costs_Fitness = 'Inclusive fitness cost',
   leader.costs_RiskHarmConflict = 'Increased risk of harm',
   leader.costs_Other = 'Misc. non-material cost',
-  leader.costs_ResourceFoodCost = 'Food cost',
-  leader.costs_ResourcesOther = 'Loss of material resources',
-  leader.costs_SocialStatus = 'Reduced social status',
+  leader.costs_ResourceFood = 'Food cost',
+  leader.costs_ResourceOther = 'Loss of material resources',
+  leader.costs_SocialStatusReputation = 'Reduced social status',
   leader.costs_Territory = 'Loss of territory',
   leader.costs_Mating = 'Mating cost',
   leader.costs_SocialServices = 'Loss of social services',
@@ -132,7 +133,7 @@ cost_benefit_dict <- c(
   follower.costs_ResourceFood = 'Food cost',
   follower.costs_ResourceOther = 'Loss of material resources',
   follower.costs_SocialServices = 'Loss of social services',
-  follower.costs_SocialStatus = 'Reduced social status',
+  follower.costs_SocialStatusReputation = 'Reduced social status',
   follower.costs_Territory = 'Loss of territory'
 )
 
@@ -141,43 +142,43 @@ format_label <- function(lbl){
   return(cost_benefit_dict[lbl])
 
   relabel <- c(
[…]

### 2020-04-17 initialcompute.R
fixing variable name list.

@@ -283,11 +283,11 @@ for (m in models) {
   }
 }
 
-d_melt = data.frame(Model=Model, Variable=Variable, Type=Type, value=value, y_se=y_se, y_negse=y_negse, stringsAsFactors = F)
+  d_melt = data.frame(Model=Model, Variable=Variable, Type=Type, value=value, y_se=y_se, y_negse=y_negse, stringsAsFactors = F)
 
 var_names <- c(
-  "function_BestowMate" = "Bestow mates",                        
-  "function_PoliticalAppointments" = "Political appointments",             
+  "functions_BestowMate" = "Bestow mates",                        
+  "functions_PoliticalAppointments" = "Political appointments",             
   "functions_ConstructionInfrastructure"       = "Construction/infastructure",
   "functions_ControlEconomics"                 = "Control economics",
   "functions_CouncilMember"                    = "Council member",
@@ -300,8 +300,8 @@ var_names <- c(
   "functions_Punishment"                        = "Punishment",
   "functions_ServeLeader"                      = "Serve a leader",
   "functions_StrategicPlanning" = "Strategic planning",
-  "function_OrganizeCooperation"  = "Organize cooperation",
-  "function_ResolveConflcit"      = "Resolve conflict",
+  "functions_OrganizeCooperation"  = "Organize cooperation",
+  "functions_ResolveConflcit"      = "Resolve conflict",
   "functions_ControlCalendar"     = "Control calendar",
   "functions_ControlImmigration"  = "Control immigration",
   "functions_DistributeResources" = "Distribute resources",
@@ -349,11 +349,11 @@ var_names <- c(
   "qualities_Fairness"               = "Fairness",
   "qualities_HighQualitySpouse"    = "High-quality spouse",
   "qualities_Industriousness"        = "Industriousness",
-  "qualities_Interpersonal.skills"   = "Interpersonal skills",
+  "qualities_InterpersonalSkills"   = "Interpersonal skills",
   "qualities_Loyalty"                = "Loyalty",
   "qualities_PhysicalHealth"        = "Physical health",
   "qualities_ProperBehavior"        = "Proper behavior",
-  "qualities_Strategic.Nepotism"     = "Strategic nepotism",
+  "qualities_StrategicNepotism"     = "Strategic nepotism",
   "qualities_Xenophobic"   = "Xenophobia",
   "qualities_AntiHonest" = "Dishonest",
   "qualities_AntiFairness" = "Unfair",

### 2020-04-17 initialcompute.R
Moved var_names to data package

@@ -284,83 +284,6 @@ for (m in models) {
 }
 
 d_melt = data.frame(Model=Model, Variable=Variable, Type=Type, value=value, y_se=y_se, y_negse=y_negse, stringsAsFactors = F)
-
-var_names <- c(
-  "function_BestowMate" = "Bestow mates",                        
-  "function_PoliticalAppointments" = "Political appointments",             
-  "functions_ConstructionInfrastructure"       = "Construction/infastructure",
-  "functions_ControlEconomics"                 = "Control economics",
-  "functions_CouncilMember"                    = "Council member",
-  "functions_GroupDetermination"              = "Group determination/cohesiveness",
-  "functions_Hospitality"                       = "Hospitality",
-  "functions_MilitaryCommand"                  = "Military command",
-  "functions_NewSettlement"                    =  "Movement/migration",
-  "functions_ProsocialInvestment"              = "Prosocial investment",
-  "functions_ProvideCounsel"                   = "Provide counsel/direction",
-  "functions_Punishment"                        = "Punishment",
-  "functions_ServeLeader"                      = "Serve a leader",
-  "functions_StrategicPlanning" = "Strategic planning",
-  "function_OrganizeCooperation"  = "Organize cooperation",
-  "function_ResolveConflcit"      = "Resolve conflict",
-  "functions_ControlCalendar"     = "Control calendar",
-  "functions_ControlImmigration"  = "Control immigration",
-  "functions_DistributeResources" = "Distribute resources",
-  "functions_GroupRepresentative" = "Group representative",
-  "functions_Medicinal"           = "Medicinal functions",
-  "functions_MoralAuthority"     = "Moral authority",
-  "functions_Policymaking"        =  "Policy making",
-  "functions_Protection"          = "Protection",
-  "functions_ProvideSubsistence" = "Provide subsistence",
-  "functions_Ritual"              = "Ritual functions",
-  "functions_SocialFunctions" = "Misc. social functions",
-  "qualities_ArtisticPerformance"     = "Artistic performance",
-  "qualities_Generous"                 = "Generosity",
-  "qualities_Age"                      = "Age",
-  "qualities_Attractive"              = "Attractive",
-  "qualities_CoerciveAuthority"      = "Coercive authority",
[…]

### 2020-04-17 analyses2.R
Minor bug fixes

@@ -48,7 +48,6 @@ library(treemapify)
 library(emmeans)
 library(hagenutils)
 library(ggalt)
-library(ggpubr)
 
 # Load precomputed objects ------------------------------------------------
 
@@ -1814,7 +1813,7 @@ df_allvars_subsistence <-
       )
     ),
     Anova = map(Model, Anova),
-    p_value = map_dbl(Anova, 'Pr(>Chisq)'),
+    p_value = map_dbl(Anova, ~pluck(., 3, 1)),
     adj_pvalue = p.adjust(p_value, method = 'BH')
   )
 
@@ -1888,7 +1887,7 @@ model_words <- function(var, lam = 'lambda.min', exponentiate = T, title){
 }
 
 
-highstatus_plot <- model_words('qualities_high.status', lam = "1se", title = 'Leader quality: High status')
+highstatus_plot <- model_words('qualities_HighStatus', lam = "1se", title = 'Leader quality: High status')
 # highstatus_plot + scale_x_log10()
 
 # ben_highstatus_plot <- model_words('leader.benefits_social.status.reputation', lam = "mid", title = 'Benefits: Status, reputation')

### 2020-04-17 leadership_across_cultures_contexts.Rmd
Added mosaic plot of subsistence X region
Fixed feature X subsistence models. Added a couple of plots.

@@
 
+```{r subsisttree, fig.cap="Composition of the study source data, organized by mode of subsistence and culture.
+Size the rectangles, in descending order, is proportional to the number of text records per mode of subsistence, cultures of each mode, and per document from each culture."}
+plot_cult_docs_subsis
+```
+
 ```{r}
@@
 ```
+
+## Continental region is confounded with mode of subsistence
+
+```{r regionsubsistence, fig.cap=""}
+ggplot(all_data2b) + geom_mosaic(aes(x = product(subsistence, region), fill = subsistence)) + scale_fill_viridis(discrete = T) + theme_minimal(15)
+```
+
+## Bias
 
@@
 
+### Correlation matrix of features
+
+```{r featurecor, fig.cap="Correlation matrix of Feature variables. Numbers and colors indicate Pearson correlation coefficients."}
+plot_features_cor
+```
+
 # References

### 2020-04-17 analyses2.R

@@ -1568,16 +1568,33 @@ clust_fun_vars <- map2_df(fun_branches, names(fun_branches), branch2df)
 clust_vars <- bind_rows(clust_fun_vars, clust_qual_vars)
 features <- unique(clust_vars$Feature)
 names(features) <- features
+features2 <- features
+names(features2) <- paste0(features, 'Failure')
+
 
 feature_var <- function(feature){
   alldatavars <- clust_vars$Variable[clust_vars$Feature == feature]
   rowSums(all_data2[alldatavars])
 }
 
+failures <- function(feature){
+  max <- length(clust_vars$Feature[clust_vars$Feature == feature])
+  successes <- all_data3[[feature]]
+  max - successes
+}
+
 all_data3 <-
   all_data2 %>% 
   dplyr::select(demo_sex:female_coauthor) %>% 
-  bind_cols(map_dfc(features, feature_var))
+  bind_cols(
+    map_dfc(features, feature_var)
+    ) 
+
+all_data3 <- 
+  all_data3 %>% 
+  bind_cols(
+    map_dfc(features2, failures)
+  )
 
 df_culture_sum <-
   all_data3 %>%
@@ -1595,21 +1612,23 @@ heatmap(
   col = viridis(256)
   )
 
-ggcorrplot(cor(df_culture_sum[-c(1, 8)]), hc.order = T, hc.method = 'ward.D', lab=T)
+plot_features_cor <- 
+  ggcorrplot(cor(df_culture_sum[-c(1, 8)]), hc.order = T, hc.method = 'ward.D', lab=T)
 
 # Number of cultures without missing values
 map_int(culture_vars, ~ sum(!is.na(.)))
 
 # Univariate model of each feature by subsistence type
+failure_vars <- names(features2)
 feature_sub_models <-
   tibble(
     Feature = features,
-    Model = map(
-      features, ~ glmer(
-        all_data3[[.x]] ~
+    Model = map2(
+      features, failure_vars, ~ glmer(
+        cbind(all_data3[[.x]], all_data3[[.y]]) ~
           subsistence +
           (1|d_culture/doc_ID),
-        family = poisson,
+        family = binomial,
         data = all_data3
       )),
     Anova = map(Model, Anova),
@@ -1626,12 +1645,12 @@ sub_models_sig <-
 plot_feature_models <-
   ggplot(
     sub_models_sig, 
[…]

### 2020-04-18 analyses2.R
Added hagenutil prefix

@@ -852,7 +852,7 @@ plot_group_subsis
 # Simple proportions
 x <- table(leader_text2$group.structure2, leader_text2$qualities_HighStatus)
 x <- prop.table(x, margin = 1)[,2]
-plot_status_group <- ggdotchart(x) + scale_x_continuous(limits = c(0, 0.5))
+plot_status_group <- hagenutils::ggdotchart(x) + scale_x_continuous(limits = c(0, 0.5))
 
 # logistic model
 

### 2020-04-20 leadership_across_cultures_contexts.Rmd
Minor chnage in MS.

@@
 The eHRAF provides information on the time frame of field work associated with each document.
-For some documents the span of associated fieldwork is given for the exact months and years the ethnographer was in the field.
+For some documents the span of associated fieldwork is given for the exact months and years the ethnographer conducted fieldwork.
 For most documents one or more spans of years are given.
@@
 The time spans plotted do not depict the full span of fieldwork necessarily, but simply illustrate the time span during which fieldwork occurred.
-We omitted data for three documents (two from North America and one from Insular Pacific) with field work date ranges prior to 1870.
+We omitted data points from  one document from North America with very early field work prior to 1805 for clarity.
 For all documents, the average earliest year in which field work occurred was `r signif(mean(documents$d_field_date_start, na.rm=T),4)` and the average latest years in which field work occurred was `r signif(mean(documents$d_field_date_end, na.rm=T),4)`.
